### 2016-11-15 report/survey.tex
Added skeleton for survey paper.

\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{hyperref}

\usepackage{fancyhdr, setspace}
\renewcommand{\headrulewidth}{0pt}
\pagestyle{fancy}
\setlength{\headheight}{15pt} % Removes warnings. Set to 30pt if 2 lines on the header.
\lhead{}
\rhead{}
\usepackage{totcount}
\regtotcounter{page}
\cfoot{\ifnum\totvalue{page} > 1 \thepage \else\fi}

\usepackage{enumitem}
\setlist{parsep=0pt, listparindent=0.5cm}

\usepackage{titlesec}
\titleformat*{\section}{\large\bfseries}

\begin{document}
\begin{center}
{ \huge
	Simplifying Parallel Graph Processing: \\
}
{ \Large
	Survey of Existing Platforms \\
}
Sam Pollard (\href{mailto:spollard@cs.uoregon.edu}{spollard@cs.uoregon.edu}), University of Oregon \\
\today
\end{center}

This is a survey of existing graph analytics frameworks.

\end{document}

### 2016-11-15 get-graphalytics.sh
Got powergraph working.

@@ -347,14 +347,12 @@ install_PowerGraph()
 	fi
 	PLATFORM_DIR="$BASE_DIR/graphalytics-platforms-$platform"
 	cd "$PLATFORM_DIR"
-	if [ "$NUM_CORES" -gt 64 ]; then
-		echo "Using 64 of $NUM_THREADS available threads due to GraphLab limitations."
-		export GRAPHLAB_THREADS_PER_WORKER=64
-	else
-		export GRAPHLAB_THREADS_PER_WORKER=$NUM_CORES
-	fi
-	# Unlike in GraphX, OpenG, you need to have the configuration in the Platform directory
-	cp -r "$GA_DIR/config" "$PLATFORM_DIR/config"
+ 	if [ "$NUM_CORES" -gt 64 ]; then
+ 		echo "Using 64 of $NUM_THREADS available threads due to GraphLab limitations."
+ 		export GRAPHLAB_THREADS_PER_WORKER=64
+ 	else
+ 		export GRAPHLAB_THREADS_PER_WORKER=$NUM_CORES
+ 	fi
 	echo "powergraph.home = $POWERGRAPH_DIR" > "$PLATFORM_DIR/config/powergraph.properties"
 	echo "powergraph.num-threads = $GRAPHLAB_THREADS_PER_WORKER" >> "$PLATFORM_DIR/config/powergraph.properties"
 	#echo "powergraph.command = mpirun -np $NUM_CORES %s %s" >> "$PLATFORM_DIR/config/powergraph.properties" # Distributed Memory

### 2016-11-15 get-graphalytics.sh
Ran benchmark for powergraph and openg. Will compare and report results.

@@ -402,8 +402,8 @@ install_graphalytics
 #download_datasets
 
 ### Run the OpenG benchmark
-#install_OpenG
-#run_benchmark
+install_OpenG
+run_benchmark
 
 ### Run the GraphX benchmark
 #check_hadoop_dependencies
@@ -413,6 +413,6 @@ install_graphalytics
 #run_benchmark
 
 ### Run the PowerGraph benchmark
-install_PowerGraph
-run_benchmark
+#install_PowerGraph
+#run_benchmark
 

### 2016-11-15 get-graphalytics.sh
Made NUM_THREADS standard across all experiments for compatibility.

@@ -20,15 +20,15 @@ HADOOP_HOME="$HOME/hadoop-2.7.3"
 ### Set up some variables and script options
 # So ! doesn't expand in the shell
 set +o histexpand
-# So Ctrl-C works
+# So Ctrl-C works and subroutines can exit the whole script.
 trap "exit 1" INT
-# So we can bail if something goes wrong
 export TOP_PID=$$ 
+MAX_THREADS=64 # This is a PowerGraph limitation.
 
-osname=$(uname)
-if [ "$JAVA_HOME" = "" -a "$osname" = "Linux" ]; then
+OSNAME=$(uname)
+if [ "$JAVA_HOME" = "" -a "$OSNAME" = "Linux" ]; then
 	export JAVA_HOME=$(update-java-alternatives -l | awk '{print $3}')
-elif [ -z "$JAVA_HOME" -a "$osname" = "Darwin" ]; then
+elif [ -z "$JAVA_HOME" -a "$OSNAME" = "Darwin" ]; then
 	export JAVA_HOME=$(/usr/libexec/java_home)
 elif [ -z $"JAVA_HOME" ]; then
 	echo "Please set the environment variable JAVA_HOME. This is the directory where jdk is installed."
@@ -38,19 +38,18 @@ if [ "$HADOOP_HOME" = "" -o "$BASE_DIR" = "" ]; then
 	echo "You need to set HADOOP_HOME and BASE_DIR before this will work"
 	exit 1
 fi
-if [ "$osname" = Darwin ]; then
-	NUM_CORES=$(sysctl -n hw.ncpu)
-	NUM_THREADS=$NUM_CORES
+if [ "$OSNAME" = Darwin ]; then
+	NUM_CORES=$(sysctl -n hw.ncpu) # The number of virtual cores (2x for hyperthreading is counted)
 	MEM_KB=$(($(vm_stat | grep "Pages free:" | awk '{print $3}' | tr -d .) * $(vm_stat | head -n 1 | grep -E -o [0-9]+) / 1024 ))
-	NUM_SOCKETS=1 # Is this always true?
+	#NUM_SOCKETS=1 # Is this always true on Mac?
 	NUM_NODES=1
 else
-	NUM_CORES=$(grep -c ^processor /proc/cpuinfo)
-	NUM_THREADS=$NUM_CORES # Intel already counts 2x for hyperthreaded cores
[…]

### 2016-11-15 report/survey.tex
Started framework for survey paper. Need to automatically generate the tables.

@@
 
+\section{Arya}
+Below are the preliminary results by running two benchmarks on the research computer Arya.
+
+Machine Specifications
+
+% TODO: Get this to be filled out automatically wherever possible.
+
+\begin{tabular}{lr}
+CPU & 72 Core Intel Xeon E5 2699 v3\\ \hline
+RAM & 256 GB DDR4 ? MHZ
+\end{tabular}
+
+Report Card
+
+\begin{tabular}{lr}
+Transport & \\ \hline
+Network Topology & \\ \hline
+Local Scheduling & \\ \hline
+Runtime Feedback & \\ \hline
+Approach & \\ \hline
+Algorithmic Considerations & \\ \hline
+\end{tabular}
+
+Performance in millions of traversed edges per second (MTEPS)
+
+\begin{tabular}{l|r|r}
+ & PowerGraph & OpenG \\ \hline
+BFS & $87.4$ & $341$ \\ \hline
+SSSP & $1.09$ & $3.08$ \\ \hline
+\end{tabular}
+
+The
+
 \end{document}

### 2016-11-17 report/survey.tex
Got a table automatically generated by LaTeX. Only put one row in it.

@@
 \usepackage{hyperref}
+\usepackage{pgfplotstable}
+\pgfplotsset{compat=1.12}
+\usepackage{booktabs}
+\usepackage{footnote}
 
@@
 \setlist{parsep=0pt, listparindent=0.5cm}
-
-\usepackage{titlesec}
-\titleformat*{\section}{\large\bfseries}
 
@@
 
-The
+\section{Graph Processing Taxonomy}
+This is in the spirit of \cite{Doekemeijer:2015:GPFSurvey}.
+\begin{savenotes}
+	\begin{table}[htb]
+		\centering
+		% \small
+		\pgfplotstabletypeset[
+			col sep=comma,
+			string type, % Makes the .style={string type} redundant
+			columns/Name/.style={string type},
+			columns/Type/.style={string type},
+			columns/HPC/.style={string type},
+			columns/Parallelism/.style={string type},
+			columns/Arch./.style={string type},
+			columns/Free/.style={string type},
+			columns/Source/.style={string type},
+			columns/Notes/.style={
+				preproc cell content/.style={@cell content=\footnote{##1}}
+			},
+		]{platforms.csv}
+	\end{table}
+\end{savenotes}
+
+\bibliographystyle{acm}
+\bibliography{drp}
 

### 2016-11-20 report/survey.tex
Tiny change to survey.

--- 2016-11-15 report/survey.tex
@@
 
-The
+\section{Graph Processing Taxonomy}
+This is in the spirit of \cite{Doekemeijer:2015:GPFSurvey}.
+\begin{savenotes}
+	\begin{table}[htb]
+		\centering
+		% \small
+		\pgfplotstabletypeset[
+			col sep=comma,
+			string type, % Makes the .style={string type} redundant
+			every head row/.style={after row=\midrule},
+			columns/Name/.style={string type, column type={l|}},
+			columns/Type/.style={string type},
+			columns/HPC/.style={string type},
+			columns/Parallelism/.style={string type},
+			columns/Arch./.style={string type},
+			columns/Free/.style={string type},
+			columns/Source/.style={string type},
+			columns/Notes/.style={
+				preproc cell content/.style={@cell content=\footnote{##1}}
+			},
+		]{platforms.csv}
+	\end{table}
+\end{savenotes}
+
+\bibliographystyle{acm}
+\bibliography{drp}
 

### 2016-11-21 get-graphalytics.sh
Expanded get-graphalytics to move output folders
into an experiments section. Fixed a few small other things.

@@ -67,7 +67,7 @@ install_graphalytics()
 	else
 		echo "I found an ldbc_graphalytics directory. I assume everything is built in there."
 	fi
-	cp -r config-template config
+	cp -r ldbc_graphalytics/config-template ldbc_graphalytics/config
 	echo "Changing graphs.root-directory in config to $DATASET_DIR"
 	echo "$OSNAME" | grep -q '.*BSD'
 	if [ $? -eq 0 -o "$OSNAME" = "Darwin" ]; then
@@ -382,17 +382,24 @@ download_datasets()
 {
 	cd "$BASE_DIR"
 	# The full dataset is at http://atlarge.ewi.tudelft.nl/graphalytics/zip/dota-league.zip for example.
-	wget -r -l 1 --no-clobber http://atlarge.ewi.tudelft.nl/graphalytics/data/ $DATASET_DIR
+	wget -r -l 1 --no-clobber --reject 'index.html*' --exclude-directories=icons http://atlarge.ewi.tudelft.nl/graphalytics/data/ $DATASET_DIR
 	# Get the reference solutions
-	wget -r -l 1 --no-clobber http://atlarge.ewi.tudelft.nl/graphalytics/ref/ $DATASET_DIR
+	wget -r -l 1 --no-clobber --reject 'index.html' --exclude-directoryes=icons http://atlarge.ewi.tudelft.nl/graphalytics/ref/ $DATASET_DIR
 }
 
-# Runs the benchmark $platform
-# Requirements: $PKGDIR be defined.
+# Runs and logs the results of running a given $platform.
+# NOTE: This must be done after calling an install_$platform.
+#       That is, ALWAYS call install before run.
 run_benchmark()
 {
 	cd "$PKGDIR" # Very important that you're in the directory you un-tar'd
-	./run-benchmark.sh # calls prepare-benchmark.sh
+	mkdir -p "$BASE_DIR/experiments"
+	LOG_FILE="$BASE_DIR/experiments/${platform}-log.txt"
+	./run-benchmark.sh | tee "$LOG_FILE" # calls prepare-benchmark.sh
+	# Parse through logs and move the html outputs to a more convenient location
+	OUTPUT=$(dirname $(awk -F '"' '/Wrote benchmark report/{print $(NF-1)}' "$LOG_FILE"))
+	echo "$PKGDIR/$OUTPUT"
+	mv "$PKGDIR/$OUTPUT" "$BASE_DIR/experiments"
 }
 # For each platform repository package up an executable
 

### 2016-11-21 report/survey.tex
Added one platform and expanded explanation.
Added the bib file. Still need to make auto-generation of results,
and the same for hardware information and other stuff.

--- 2016-11-15 report/survey.tex
@@
 
-\begin{tabular}{l|r|r}
+\begin{tabular}{l|r|r|}
  & PowerGraph & OpenG \\ \hline
@@
 
-The
+\section{Graph Processing Taxonomy}
+This is in the spirit of \cite{Doekemeijer:2015:GPFSurvey}.
+Here, ``|'' means ``or'' and ``+'' means ``and.'' FOSS means Free and Open Source Software.
+The quotes around ``yes'' for HPC mean that the product claims to be amenable to high performance computing.
+Whether these actually achieve their goal is one of the purposes of this project.
+\begin{savenotes}
+	\begin{table}[htb]
+		\centering
+		% \small
+		\pgfplotstabletypeset[
+			col sep=comma,
+			string type, % Makes the .style={string type} redundant
+			every head row/.style={after row=\midrule},
+			columns/Name/.style={string type, column type={l|}},
+			columns/Type/.style={string type},
+			columns/HPC/.style={string type},
+			columns/Parallelism/.style={string type},
+			columns/Target/.style={string type},
+			columns/FOSS/.style={string type},
+			columns/Source/.style={string type},
+			columns/Notes/.style={
+				preproc cell content/.style={@cell content=
+					\ifx&##1&% Only make a footnote if the cell is nonempty
+						##1
+					\else
+						\footnote{##1}
+					\fi}
+			},
+		]{platforms.csv}
+	\end{table}
+\end{savenotes}
+
+\bibliographystyle{acm}
+\bibliography{drp}
 

### 2016-11-23 report/survey.tex
Added get-hwinfo.sh and generated tables.
get-hwinfo only works for Linux. Updated some citations for drp.bib.
Added real \begin{tables} for survey. Now the hardware info is
automatically generated and can be sucked in via specs.csv.
Added specs.csv for SamXu2. Added conclustion to survey.

@@
 
-\section{Arya}
+\section{Machine Specifications}
 Below are the preliminary results by running two benchmarks on the research computer Arya.
 
-Machine Specifications
+% TODO: Make this a pgfplotstable with an automatically generated csv.
+\begin{table}[htb]
+	\centering
+% Arya
+%	\begin{tabular}{l|r}
+%		CPU & 72 Core Intel Xeon E5 2699 v3\\ \hline
+%		RAM & 256 GB DDR4 2133 MHz
+%	\end{tabular}
 
-% TODO: Get this to be filled out automatically wherever possible.
+% Keep in mind you can do this at the beginning: string replace={s1}{s2}
+	\pgfplotstabletypeset[
+	header=false,
+	col sep=tab,
+	string type,
+    every head row/.style={output empty row, before row=\bottomrule},
+	columns/0/.style={column type={|r|}},
+	columns/1/.style={column type={l|}},
+	every last row/.style={after row=\toprule},
+	]{specs.csv}
+	\caption{Machine specifications.}
+	\label{tab:specs}
+\end{table}
 
-\begin{tabular}{lr}
-CPU & 72 Core Intel Xeon E5 2699 v3\\ \hline
-RAM & 256 GB DDR4 ? MHZ
-\end{tabular}
-
-Report Card
-
+\begin{table}[htb]
+	\centering
 	\begin{tabular}{lr}
@@
 	\end{tabular}
+	\caption{Middleware specifications}
+	\label{tab:reportcard}
+\end{table}
 
@@
 
+% TODO: This should be unnecessary once autogeneration is used.
+\begin{table}[htb]
+	\centering
 	\begin{tabular}{l|r|r|}
@@
 	\end{tabular}
+	\caption{Performance Results}
+	\label{tab:perf}
+\end{table}
 
@@
 \begin{savenotes}
-	\begin{table}[htb]
+	\begin{table}
 		\centering
@@
 		]{platforms.csv}
+		\caption{Tools used for graph processing}
+		\label{tab:frameworks}
 	\end{table}
 \end{savenotes}
+
+\section{Conclusion} % Begin with the end in mind...
+We have presented an updated survey of parallel graph processing frameworks supplementary to \cite{Doekemeijer:2015:GPFSurvey}.
+From this, we have selected a representative subset of frameworks on which performance is analyzed and have stored these results in a database.
+To facilitate parallel graph processing, hardware information and performance results are automatically populated (as were all the tables in this paper).
+These performance results are then used to provide simple recommendations of the optimally-performing framework given a particular algorithm and problem size.
+% We have developed a simple model of hardware and its correlation with performance to predict performance on other architectures
 

### 2016-11-23 get-hwinfo.sh

@@ -1,0 +1,50 @@
+#!/bin/bash
+# Gets hardware data and prints to standard output.
+# Note: This utility will get better data if run as superuser.
+
+OSNAME=$(uname)
+dlm="\t"
+
+# Get the data
+if [ "$OSNAME" = Darwin ]; then
+	CPU_SOCKETS=1 # This appears to be true on all Mac models so far.
+	CPU_CORES=$(sysctl -n hw.ncpu) # The number of virtual cores (2x for hyperthreading is counted)
+	#CHANGE TO RAM_SIZE: MEM_KB=$(($(vm_stat | grep "Pages free:" | awk '{print $3}' | tr -d .) * $(vm_stat | head -n 1 | grep -E -o [0-9]+) / 1024 ))
+else
+	CPU_MODEL=$(lscpu | grep "Model name" | awk -F ":" '{print $2}' | sed 's/^[[:space:]]*//')
+	CPU_SOCKETS=$(grep -i "physical id" /proc/cpuinfo | sort -u | wc -l)
+	CPU_CORES=$(grep -c ^processor /proc/cpuinfo) # The number of virtual cores are counted (2x for hyperthreading)
+	CPU_CLOCK=$(printf "%.0f%s" $(lscpu | grep "CPU max MHz" | awk -F ":" '{print $2}' | sed 's/^\s*//') "MHz")
+	RAM_SIZE=$(cat /proc/meminfo | grep MemTotal | awk '{print $2 / 1024 / 1024 "Gib"}')
+fi
+
+# Get data that requires superuser.
+if [ -f "lshw.txt" ]; then
+	RAM_FREQ=$(grep -E -m 1 'clock: [0-9]+.*\(' lshw.txt | awk '{print $2}')
+	GPU_MODEL=$(grep -E -A 11 '\*-display' lshw.txt | grep product | sed 's/^\s*product:\s//')
+else
+	echo -e "To get more in-depth hardware info run the command\n"
+	echo -e "sudo lshw > lshw.txt\n"
+fi
+
+# Prints a record. Requires 2 arguments: The string and the value associated with that.
+print_record()
+{
+	if [ "$#" -lt 2 ]; then
+		exit # Do nothing since the value is empty
+	fi
+	printf "$1${dlm}$2\n"
+}
+
+# Print the data
+print_record "CPU Model" "$CPU_MODEL"
+print_record "CPU Sockets" "$CPU_SOCKETS"
[…]

### 2016-11-23 get-hwinfo.sh
Got hwinfo to work on arya by removing extra \ns

@@ -14,14 +14,14 @@ else
 	CPU_MODEL=$(lscpu | grep "Model name" | awk -F ":" '{print $2}' | sed 's/^[[:space:]]*//')
 	CPU_SOCKETS=$(grep -i "physical id" /proc/cpuinfo | sort -u | wc -l)
 	CPU_CORES=$(grep -c ^processor /proc/cpuinfo) # The number of virtual cores are counted (2x for hyperthreading)
-	CPU_CLOCK=$(printf "%.0f%s" $(lscpu | grep "CPU max MHz" | awk -F ":" '{print $2}' | sed 's/^\s*//') "MHz")
 	RAM_SIZE=$(cat /proc/meminfo | grep MemTotal | awk '{print $2 / 1024 / 1024 "Gib"}')
+	CPU_CLOCK=$(printf "%.0f%s" $(lscpu | grep "CPU max MHz" | awk -F ":" '{print $2}' | sed 's/^\s*//') "MHz")
 fi
 
 # Get data that requires superuser.
 if [ -f "lshw.txt" ]; then
 	RAM_FREQ=$(grep -E -m 1 'clock: [0-9]+.*\(' lshw.txt | awk '{print $2}')
-	GPU_MODEL=$(grep -E -A 11 '\*-display' lshw.txt | grep product | sed 's/^\s*product:\s//')
+	GPU_MODEL=$(grep -E -A 11 '\*-display' lshw.txt | grep product | sed 's/^\s*product:\s//' | tr "\n" " ")
 else
 	echo -e "To get more in-depth hardware info run the command\n"
 	echo -e "sudo lshw > lshw.txt\n"

### 2016-11-27 report/report.tex
Added skeleton of automatically generated report,
report.tex. This will be eventually run with a Makefile for each
machine to compile results to a human-readable format. Added
the "Comm." column to platforms.csv for the communication paradigms,
though this doesn't quite fit in the survey.tex table yet.

--- 2016-11-20 report/survey.tex
@@
 \usepackage{hyperref}
-\usepackage{pgfplotstable}
+\usepackage{pgfplotstable, booktabs}
 \pgfplotsset{compat=1.12}
-\usepackage{booktabs}
-\usepackage{footnote}
+\usepackage{datetime}
 
@@
 
-\usepackage{enumitem}
-\setlist{parsep=0pt, listparindent=0.5cm}
+\begin{document}
+\title{Parallel Graph Processing Report\vspace{-2em}}
+\author{}
+\date{Generated \today\ at \currenttime}
+\maketitle
 
-\begin{document}
-\begin{center}
-{ \huge
-	Simplifying Parallel Graph Processing: \\
-}
-{ \Large
-	Survey of Existing Platforms \\
-}
-Sam Pollard (\href{mailto:spollard@cs.uoregon.edu}{spollard@cs.uoregon.edu}), University of Oregon \\
-\today
-\end{center}
+This report was automatically generated from \verb|PATH_TO_MAKEFILE| on \verb|HOSTNAME|.
 
-This is a survey of existing graph analytics frameworks.
+\section{Machine Specifications}
+These data were gathered from \verb|specs.csv|. If these results are not complete enough, try running \verb|get-hwinfo.sh| again and run the command it suggests as root.
 
-\section{Arya}
-Below are the preliminary results by running two benchmarks on the research computer Arya.
+% TODO: Make this a pgfplotstable with an automatically generated csv.
+\begin{table}[htb]
+	\centering
+% Keep in mind you can do this at the beginning: string replace={s1}{s2}
+	\pgfplotstabletypeset[
+	header=false,
+	col sep=tab,
+	string type,
+    every head row/.style={output empty row, before row=\bottomrule},
+	columns/0/.style={column type={|r|}},
+	columns/1/.style={column type={l|}},
+	every last row/.style={after row=\toprule},
+	]{specs.csv}
+	\caption{Machine specifications.}
+	\label{tab:specs}
+\end{table}
 
-Machine Specifications
-
-% TODO: Get this to be filled out automatically wherever possible.
-
-\begin{tabular}{lr}
-CPU & 72 Core Intel Xeon E5 2699 v3\\ \hline
-RAM & 256 GB DDR4 ? MHZ
-\end{tabular}
-
-Report Card
-
-\begin{tabular}{lr}
-Transport & \\ \hline
-Network Topology & \\ \hline
-Local Scheduling & \\ \hline
-Runtime Feedback & \\ \hline
-Approach & \\ \hline
-Algorithmic Considerations & \\ \hline
-\end{tabular}
-
+\section{Performance Results}
 Performance in millions of traversed edges per second (MTEPS)
 
-\begin{tabular}{l|r|r}
+% TODO: Actually get this to autogenerate.
+\begin{table}[htb]
+	\centering
+	\begin{tabular}{l|r|r|}
 	 & PowerGraph & OpenG \\ \hline
@@
 	\end{tabular}
-
-\section{Graph Processing Taxonomy}
-This is in the spirit of \cite{Doekemeijer:2015:GPFSurvey}.
-\begin{savenotes}
-	\begin{table}[htb]
-		\centering
-		% \small
-		\pgfplotstabletypeset[
-			col sep=comma,
-			string type, % Makes the .style={string type} redundant
-			every head row/.style={after row=\midrule},
-			columns/Name/.style={string type, column type={l|}},
-			columns/Type/.style={string type},
-			columns/HPC/.style={string type},
-			columns/Parallelism/.style={string type},
-			columns/Arch./.style={string type},
-			columns/Free/.style={string type},
-			columns/Source/.style={string type},
-			columns/Notes/.style={
-				preproc cell content/.style={@cell content=\footnote{##1}}
-			},
-		]{platforms.csv}
+	\caption{Performance Results}
+	\label{tab:perf}
 \end{table}
-\end{savenotes}
-
-\bibliographystyle{acm}
-\bibliography{drp}
 

### 2016-11-27 report/survey.tex
Added skeleton of automatically generated report,
report.tex. This will be eventually run with a Makefile for each
machine to compile results to a human-readable format. Added
the "Comm." column to platforms.csv for the communication paradigms,
though this doesn't quite fit in the survey.tex table yet.

@@
 \usepackage{booktabs}
-\usepackage{footnote}
 
@@
 \begin{document}
-\begin{center}
-{ \huge
-	Simplifying Parallel Graph Processing: \\
+\title{Simplifying Parallel Graph Processing: \\
+	\large Survey of Existing Platforms
 }
-{ \Large
-	Survey of Existing Platforms \\
-}
-Sam Pollard (\href{mailto:spollard@cs.uoregon.edu}{spollard@cs.uoregon.edu}), University of Oregon \\
-\today
-\end{center}
+\author{Samuel Pollard (\href{mailto:spollard@cs.uoregon.edu}{spollard@cs.uoregon.edu})}
+\date{\today}
+\maketitle
 
@@
 Whether these actually achieve their goal is one of the purposes of this project.
-\begin{savenotes}
 \begin{table}
+	\begin{minipage}{\linewidth} % So the footnotes all get printed on the same page
 		\centering
@@
 			string type, % Makes the .style={string type} redundant
+			columns={Name,Type,HPC,Parallelism,Target,FOSS,Source,Notes},
 			every head row/.style={after row=\midrule},
@@
 			columns/Parallelism/.style={string type},
+%			columns/Comm./.style={%
+%				string type,
+%				postproc cell content/.style={%
+%					@cell content/.add={\footnotesize}
+%				}
+%			},
 			columns/Target/.style={string type},
@@
 		\label{tab:frameworks}
+	\end{minipage}
 \end{table}
-\end{savenotes}
 

### 2016-11-27 get-hwinfo.sh

@@ -12,6 +12,7 @@ if [ "$OSNAME" = Darwin ]; then
 	CPU_CORES=$(sysctl -n hw.ncpu) # The number of virtual cores (2x for hyperthreading is counted)
 	RAM_SIZE=$(sysctl -n hw.memsize | awk '{print $0 / 1024 / 1024 / 1024 "Gib"}')
 	CPU_CLOCK=$(printf "%.0f %s" $(echo $(sysctl -n hw.cpufrequency_max) "/ 1000000" | bc) "MHz")
+	echo "Running system_profiler and saving to /tmp/profile.txt. This will be deleted afterwards."
 	system_profiler > /tmp/profile.txt
 	RAM_FREQ=$(cat /tmp/profile.txt | grep -A 16 "Memory:$" | grep "Speed" | awk -F ":" '{print $2}' | sed 's/^[[:space:]]*//')
 	GPU_MODEL=$(cat /tmp/profile.txt | grep -A 11 "Graphics/Displays" | grep "Chipset" | awk -F ":" '{print $2}' | sed 's/^[[:space:]]*//')
@@ -22,6 +23,7 @@ else
 	CPU_CORES=$(grep -c ^processor /proc/cpuinfo) # The number of virtual cores are counted (2x for hyperthreading)
 	RAM_SIZE=$(cat /proc/meminfo | grep MemTotal | awk '{print $2 / 1024 / 1024 "Gib"}')
 	CPU_CLOCK=$(printf "%.0f %s" $(lscpu | grep "CPU max MHz" | awk -F ":" '{print $2}' | sed 's/^\s*//') "MHz")
+	CPU_SIMD=$(sysctl -n machdep.cpu.features)
 fi
 
 # Get data that requires superuser.

### 2016-11-28 report/survey.tex
Updated get graphalytics. Still doesn't work on mac.
Updated get-hwinfo and report specs towards the aim of getting
a csv output from graphalytics. Notices TEPS is not very well-
defined for graphalytics.

@@
 \section{Machine Specifications}
-Below are the preliminary results by running two benchmarks on the research computer Arya.
+Table~\ref{tab:specs} shows the specifications of the research computer (named Arya).
 
 % TODO: Make this a pgfplotstable with an automatically generated csv.
-\begin{table}[htb]
+\begin{table}[!htb]
 	\centering
@@
 
-\begin{table}[htb]
-	\centering
-	\begin{tabular}{lr}
-		Transport & \\ \hline
-		Network Topology & \\ \hline
-		Local Scheduling & \\ \hline
-		Runtime Feedback & \\ \hline
-		Approach & \\ \hline
-		Algorithmic Considerations & \\ \hline
-	\end{tabular}
-	\caption{Middleware specifications}
-	\label{tab:reportcard}
-\end{table}
+%\begin{table}[htb]
+%	\centering
+%	\begin{tabular}{lr}
+%		Transport & \\ \hline
+%		Network Topology & \\ \hline
+%		Local Scheduling & \\ \hline
+%		Runtime Feedback & \\ \hline
+%		Approach & \\ \hline
+%		Algorithmic Considerations & \\ \hline
+%	\end{tabular}
+%	\caption{Middleware specifications}
+%	\label{tab:reportcard}
+%\end{table}
 
-Performance in millions of traversed edges per second (MTEPS)
+\section{Performance}
+Graphalytics without the use of the Granula plugin produces performance measurement in two forms: runtime in seconds and traversed edges per second.
+
+Table~\ref{tab:perf} lists performance in millions of traversed edges per second (MTEPS) according to the graphalytics output.
+While this metric is simple to compute it is invariant across algorithms: for example, computing the local clustering coefficient involves traversing each edge multiple times (proportional to the sparsity of the graph), while breadth first search (BFS) traverses each edge exactly once, and on na\:ive implementations single-source shortest paths (SSSP) may have $O(|E| + |V|^2)$ traversed edges.
+% TODO: Cite Comer algorithm book for Dijkstra's algorithm runtime.
 
 % TODO: This should be unnecessary once autogeneration is used.
-\begin{table}[htb]
+\begin{table}[!htb]
 	\centering
@@
 	 & PowerGraph & OpenG \\ \hline
-	BFS & $87.4$ & $341$ \\ \hline
-	SSSP & $1.09$ & $3.08$ \\ \hline
+	BFS & 81.8 & 341 \\ \hline
+	SSSP & 1.64 & 15.0 \\ \hline
+	LCC & 54.6 & 142 \\ \hline
 	\end{tabular}
-	\caption{Performance Results}
+	\caption{Performance Results for the \texttt{dota-league} dataset with 61,670 vertices and 50,870,313 edges.
+BFS is breadth-first search, SSSP is single-source shortest paths, LCC is local clustering coefficient.}
 	\label{tab:perf}
@@
 Whether these actually achieve their goal is one of the purposes of this project.
-\begin{table}
+\begin{table}[!htb]
 	\begin{minipage}{\linewidth} % So the footnotes all get printed on the same page

### 2016-11-28 get-graphalytics.sh

@@ -41,18 +41,13 @@ fi
 if [ "$OSNAME" = Darwin ]; then
 	NUM_CORES=$(sysctl -n hw.ncpu) # The number of virtual cores (2x for hyperthreading is counted)
 	MEM_KB=$(($(vm_stat | grep "Pages free:" | awk '{print $3}' | tr -d .) * $(vm_stat | head -n 1 | grep -E -o [0-9]+) / 1024 ))
-	#NUM_SOCKETS=1 # Is this always true on Mac?
-	NUM_NODES=1
 else
 	NUM_CORES=$(grep -c ^processor /proc/cpuinfo) # The number of virtual cores (2x for hyperthreading is counted)
 	MEM_KB=$(cat /proc/meminfo | grep MemAvailable | awk '{print $2}')
-	#NUM_SOCKETS=$(grep -i "physical id" /proc/cpuinfo | sort -u | wc -l) # Irrelevant for the current setup.
-	NUM_NODES=1
 fi
 NUM_THREADS=$(if [ "$NUM_CORES" -lt $MAX_THREADS ]; then echo $NUM_CORES; else echo $MAX_THREADS; fi)
 GA_DIR="$BASE_DIR/ldbc_graphalytics"
 DATASET_DIR=$BASE_DIR/datasets
-
 
 install_graphalytics()
 {
@@ -188,7 +183,6 @@ install_GraphX()
 {
 	# Assumes HDFS and YARN are running.
 	echo "Installing GraphX..."
-	OLDWD=$(pwd)
 	cd "$BASE_DIR"
 	if [ ! $(find "$BASE_DIR" -maxdepth 1 -type d -name graphalytics-platforms-graphx) ]; then
 		git clone https://github.com/tudelft-atlarge/graphalytics-platforms-graphx.git
@@ -199,7 +193,7 @@ install_GraphX()
 	cd "$GRAPHX_DIR"
 	mvn package
 
-	PKGNAME=$(basename $(find $GRAPHX_DIR -maxdepth 1 -name *.tar.gz))
+	PKGNAME=$(basename $(find $GRAPHX_DIR -maxdepth 1 -name '*.tar.gz'))
 	VERSION=$(echo $PKGNAME | awk -F '-' '{print $4}')
 	GA_VERSION=$(echo $PKGNAME | awk -F '-' '{print $2}')
 	platform=$(echo $PKGNAME | awk -F '-' '{print $3}')
@@ -227,14 +221,7 @@ install_GraphX()
[…]

### 2016-11-28 get-hwinfo.sh

@@ -17,13 +17,13 @@ if [ "$OSNAME" = Darwin ]; then
 	RAM_FREQ=$(cat /tmp/profile.txt | grep -A 16 "Memory:$" | grep "Speed" | awk -F ":" '{print $2}' | sed 's/^[[:space:]]*//')
 	GPU_MODEL=$(cat /tmp/profile.txt | grep -A 11 "Graphics/Displays" | grep "Chipset" | awk -F ":" '{print $2}' | sed 's/^[[:space:]]*//')
 	rm /tmp/profile.txt
+	#CPU_SIMD=$(sysctl -n machdep.cpu.features)
 else
 	CPU_MODEL=$(lscpu | grep "Model name" | awk -F ":" '{print $2}' | sed 's/^[[:space:]]*//')
 	CPU_SOCKETS=$(grep -i "physical id" /proc/cpuinfo | sort -u | wc -l)
 	CPU_CORES=$(grep -c ^processor /proc/cpuinfo) # The number of virtual cores are counted (2x for hyperthreading)
 	RAM_SIZE=$(cat /proc/meminfo | grep MemTotal | awk '{print $2 / 1024 / 1024 "Gib"}')
 	CPU_CLOCK=$(printf "%.0f %s" $(lscpu | grep "CPU max MHz" | awk -F ":" '{print $2}' | sed 's/^\s*//') "MHz")
-	CPU_SIMD=$(sysctl -n machdep.cpu.features)
 fi
 
 # Get data that requires superuser.

### 2016-11-28 get-graphalytics.sh
Fixed get-graphalytics again, but not for Mac.
There are a lot of problems with graphalytics not working with BSD-
like systems. Touched up a few things, fixed powergraph directory
configuration files, simplified dependency checking. I believe
since powergraph installs everything itself you don't actually
need to install all of the packages.

@@ -55,9 +55,9 @@ install_graphalytics()
 	if [ ! $(find $BASE_DIR -type d -name ldbc_graphalytics) ]; then
 		echo "Cloning and mvn-ing graphalytics into $GA_DIR... "
 		cd "$BASE_DIR"
-		git clone https://github.com/ldbc/ldbc_graphalytics.git
+		git clone https://github.com/sampollard/ldbc_graphalytics.git
 		cd ldbc_graphalytics
-		mvn install
+		mvn install -Dlicence.skip=true
 		cd ..
 	else
 		echo "I found an ldbc_graphalytics directory. I assume everything is built in there."
@@ -224,6 +224,7 @@ install_GraphX()
 	fix_readlink
 }
 
+# Installs and prepares OpenG, a part of GraphBIG, for execution.
 install_OpenG()
 {
 	if [ "$OSNAME" != "Linux" ]; then
@@ -312,27 +313,11 @@ install_PowerGraph()
 	export LD_LIBRARY_PATH=/usr/lib/jvm/java-8-openjdk-amd64/jre/lib/amd64/server/
 	# For libhdfs, you need to change graphalytics-0.3-powergraph0.1/bin/standard/CMakeFile/main.dir/link.txt to have -lhdfs
 	# There are a few dependencies to be satisfied since you're not running inside PowerGraph/apps
-	check_for_lib libtcmalloc
-	if [ $? -ne 0 ]; then
-		echo "Please install libtcmalloc or link to it in your LD_LIBRARY_PATH"
-		exit 1 # kill -s TERM $TOP_PID
-	fi
-
-	check_for_lib libevent
-	if [ $? -ne 0 ]; then
-		echo "Please install libevent or link to it in your LD_LIBRARY_PATH"
-		exit 1 # kill -s TERM $TOP_PID
-	fi	
-	check_for_lib libjson
-	if [ $? -ne 0 ]; then
-		echo "Please install libjson or link to it in your LD_LIBRARY_PATH"
-		exit 1 # kill -s TERM $TOP_PID
-	fi
-	check_for_lib libboost
-	if [ $? -ne 0 ]; then
-		echo "Please install libboost or link to it in your LD_LIBRARY_PATH"
-		exit 1 # kill -s TERM $TOP_PID
-	fi	
+	# XXX Seems to work fine for this.
+	# check_for_lib libtcmalloc
+	# check_for_lib libevent
[…]

### 2016-12-01 get-graphalytics.sh
Fixed issue with powergraph graphs.properties location.

@@ -362,6 +362,7 @@ install_PowerGraph()
 
 	# Set configuration in the packaged file
 	cp -r "$PKGDIR/config-template/" "$PKGDIR/config"
+	cp "$GA_DIR/config/graphs.properties" "$PKGDIR/config"
 	echo "powergraph.home = $POWERGRAPH_DIR" > "$PKGDIR/config/powergraph.properties"
 	echo "powergraph.num-threads = $GRAPHLAB_THREADS_PER_WORKER" >> "$PKGDIR/config/powergraph.properties"
 	#echo "powergraph.disable_mpi = false" >> "$PKGDIR/config/powergraph.properties" # TODO: Insert if > 1 node

### 2016-12-01 get-graphalytics.sh
Got CSV working.

@@ -414,8 +414,8 @@ install_graphalytics
 #download_datasets
 
 ### Run the GraphBIG OpenG benchmark
-#install_OpenG
-#run_benchmark
+install_OpenG
+run_benchmark
 
 ### Run the GraphX benchmark
 #check_hadoop_dependencies

### 2016-12-01 report/survey.tex
Got autogeneration for survey.tex sort of working.
It now only works if you copy the correct csv(s) generated
from graphalytics. Correct means only one dataset was run, but
this may be worked around by running graphalytics multiple times,
copying various files into the "datasets" directory.

@@
 	\centering
-	\begin{tabular}{l|r|r|}
-	 & PowerGraph & OpenG \\ \hline
-	BFS & 81.8 & 341 \\ \hline
-	SSSP & 1.64 & 15.0 \\ \hline
-	LCC & 54.6 & 142 \\ \hline
-	\end{tabular}
+%	\begin{tabular}{l|r|r|}
+%	 & PowerGraph & OpenG \\ \hline
+%	BFS & 81.8 & 341 \\ \hline
+%	SSSP & 1.64 & 15.0 \\ \hline
+%	LCC & 54.6 & 142 \\ \hline
+%	\end{tabular}
+
+		\centering
+		\pgfplotstabletypeset[
+			col sep=comma,
+			string type, % Makes the .style={string type} redundant
+			columns={[index]0,openg,powergraph},
+			every head row/.style={after row=\midrule},
+			columns/0/.style={string type, column type={l|}, column name={}},
+			columns/openg/.style={string type, column type={r}},
+			columns/powergraph/.style={string type, column type={r}}
+		]{runtime.csv}
 	\caption{Performance Results for the \texttt{dota-league} dataset with 61,670 vertices and 50,870,313 edges.

### 2016-12-01 report/merge-csv.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+# Merges two csv files, given that they both only use one dataset for each run.
+# To use with latex: ./merge-csv.sh > runtimes.csv
+
+first=1
+outfile=runtime.csv
+for f in *-runtime.csv; do
+	if [ "$first" -eq 1 ]; then
+		first=0
+		experiment1=${f%-runtime.csv}
+		dataset1=$(grep '^,' $f | cut -b 2-)
+		cat "$f" | sed "s/$dataset1/$experiment1/" > $outfile
+	else
+		experimentn=${f%-runtime.csv}
+		cat "$f" | cut -d ',' -f 2- | sed "s/$dataset1/$experimentn/" | lam runtime.csv -s, -
+	fi
+done
+
+

### 2016-12-02 report/survey.tex
Expanded and corrected hardware reporting.
Now it reports physical RAM when possible.

@@
 	]{specs.csv}
-	\caption{Machine specifications.}
+	\caption{Machine specifications.
+The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
+The manufacturer's published maximum clock speeds can be found at \url{http://ark.intel.com}.}
 	\label{tab:specs}
@@
 
-Table~\ref{tab:perf} lists performance in millions of traversed edges per second (MTEPS) according to the graphalytics output.
-While this metric is simple to compute it is invariant across algorithms: for example, computing the local clustering coefficient involves traversing each edge multiple times (proportional to the sparsity of the graph), while breadth first search (BFS) traverses each edge exactly once, and on na\:ive implementations single-source shortest paths (SSSP) may have $O(|E| + |V|^2)$ traversed edges.
+Table~\ref{tab:perf} lists performance in milliseconds of runtime according to the graphalytics output.
+Graphalytics also outputs MTEPS or millions of traversed edges pers econd.
+However, the graphalytics version does not make sense in all cases: for example, computing the local clustering coefficient involves traversing each edge multiple times (proportional to the sparsity of the graph), while breadth first search (BFS) traverses each edge exactly once, and on na\:ive implementations single-source shortest paths (SSSP) may have $O(|E| + |V|^2)$ traversed edges.
 % TODO: Cite Comer algorithm book for Dijkstra's algorithm runtime.
+
+In Table~\ref{tab:perf}, BFS is breadth-first search, SSSP is single-source shortest paths, LCC is local clustering coefficient, PR is PageRank, CDLP is community detection using label propagation, and WCC is weakly connected components.
+For the algorithms used, see \cite{Iosup:2016:Graphalyticstech}.
 
@@
 		]{runtime.csv}
-	\caption{Performance Results for the \texttt{dota-league} dataset with 61,670 vertices and 50,870,313 edges.
-BFS is breadth-first search, SSSP is single-source shortest paths, LCC is local clustering coefficient.}
+	\caption{Performance Results for the \texttt{dota-league} dataset with 61,670 vertices and 50,870,313 edges.}
 	\label{tab:perf}

### 2016-12-02 get-hwinfo.sh

@@ -10,7 +10,7 @@ if [ "$OSNAME" = Darwin ]; then
 	CPU_MODEL=$(sysctl -n machdep.cpu.brand_string)
 	CPU_SOCKETS=1 # This appears to be true on all Mac models.
 	CPU_CORES=$(sysctl -n hw.ncpu) # The number of virtual cores (2x for hyperthreading is counted)
-	RAM_SIZE=$(sysctl -n hw.memsize | awk '{print $0 / 1024 / 1024 / 1024 "Gib"}')
+	RAM_SIZE=$(sysctl -n hw.memsize | awk '{print $0 / 1024 / 1024 / 1024 "GB"}') # Gives actualy physical RAM
 	CPU_CLOCK=$(printf "%.0f %s" $(echo $(sysctl -n hw.cpufrequency_max) "/ 1000000" | bc) "MHz")
 	echo "Running system_profiler and saving to /tmp/profile.txt. This will be deleted afterwards."
 	system_profiler > /tmp/profile.txt
@@ -19,20 +19,25 @@ if [ "$OSNAME" = Darwin ]; then
 	rm /tmp/profile.txt
 	#CPU_SIMD=$(sysctl -n machdep.cpu.features)
 else
-	CPU_MODEL=$(lscpu | grep "Model name" | awk -F ":" '{print $2}' | sed 's/^[[:space:]]*//')
+	CPU_MODEL=$(lscpu | grep "Model name" | cut -d' ' -f 3-)
 	CPU_SOCKETS=$(grep -i "physical id" /proc/cpuinfo | sort -u | wc -l)
 	CPU_CORES=$(grep -c ^processor /proc/cpuinfo) # The number of virtual cores are counted (2x for hyperthreading)
-	RAM_SIZE=$(cat /proc/meminfo | grep MemTotal | awk '{print $2 / 1024 / 1024 "Gib"}')
-	CPU_CLOCK=$(printf "%.0f %s" $(lscpu | grep "CPU max MHz" | awk -F ":" '{print $2}' | sed 's/^\s*//') "MHz")
+	# Gives Total usable RAM (i.e., physical RAM minus a few reserved bits and the kernel binary code).
+	RAM_SIZE=$(cat /proc/meminfo | grep MemTotal | awk '{print $2 / 1024 / 1024 "GB"}')
+	CPU_CLOCK=$(echo $(cat /sys/devices/system/cpu/cpu0/cpufreq/scaling_max_freq) "/ 1000" | bc)
[…]

### 2016-12-02 get-hwinfo.sh
Polished up formatting for get-hwinfo.sh.

@@ -11,7 +11,7 @@ if [ "$OSNAME" = Darwin ]; then
 	CPU_SOCKETS=1 # This appears to be true on all Mac models.
 	CPU_CORES=$(sysctl -n hw.ncpu) # The number of virtual cores (2x for hyperthreading is counted)
 	RAM_SIZE=$(sysctl -n hw.memsize | awk '{print $0 / 1024 / 1024 / 1024 "GB"}') # Gives actualy physical RAM
-	CPU_CLOCK=$(printf "%.0f %s" $(echo $(sysctl -n hw.cpufrequency_max) "/ 1000000" | bc) "MHz")
+	CPU_CLOCK=$(printf "%.0f%s" $(echo $(sysctl -n hw.cpufrequency_max) "/ 1000000" | bc) "MHz")
 	echo "Running system_profiler and saving to /tmp/profile.txt. This will be deleted afterwards."
 	system_profiler > /tmp/profile.txt
 	RAM_FREQ=$(cat /tmp/profile.txt | grep -A 16 "Memory:$" | grep "Speed" | awk -F ":" '{print $2}' | sed 's/^[[:space:]]*//')
@@ -19,21 +19,21 @@ if [ "$OSNAME" = Darwin ]; then
 	rm /tmp/profile.txt
 	#CPU_SIMD=$(sysctl -n machdep.cpu.features)
 else
-	CPU_MODEL=$(lscpu | grep "Model name" | cut -d' ' -f 3-)
+	CPU_MODEL=$(lscpu | grep "Model name" | cut -d' ' -f 3- | sed 's/^[[:space:]]*//')
 	CPU_SOCKETS=$(grep -i "physical id" /proc/cpuinfo | sort -u | wc -l)
 	CPU_CORES=$(grep -c ^processor /proc/cpuinfo) # The number of virtual cores are counted (2x for hyperthreading)
 	# Gives Total usable RAM (i.e., physical RAM minus a few reserved bits and the kernel binary code).
 	RAM_SIZE=$(cat /proc/meminfo | grep MemTotal | awk '{print $2 / 1024 / 1024 "GB"}')
-	CPU_CLOCK=$(echo $(cat /sys/devices/system/cpu/cpu0/cpufreq/scaling_max_freq) "/ 1000" | bc)
[…]

### 2016-12-09 get-graphalytics.sh
Improved documentation, fixed issue with wget.
Still doesn't get the file formats correct though.

@@ -382,11 +382,12 @@ install_Giraph()
 # Downloads em all
 download_datasets()
 {
-	cd "$BASE_DIR"
+	cd "$DATASET_DIR"
 	# The full dataset is at http://atlarge.ewi.tudelft.nl/graphalytics/zip/dota-league.zip for example.
-	wget -r -l 1 --no-clobber --reject 'index.html*' --exclude-directories=icons http://atlarge.ewi.tudelft.nl/graphalytics/data/ $DATASET_DIR
+	wget -r -l 1 --no-clobber --reject 'index.html*' --exclude-directories=icons http://atlarge.ewi.tudelft.nl/graphalytics/data/
 	# Get the reference solutions
-	wget -r -l 1 --no-clobber --reject 'index.html' --exclude-directoryes=icons http://atlarge.ewi.tudelft.nl/graphalytics/ref/ $DATASET_DIR
+	wget -r -l 1 --no-clobber --reject 'index.html' --exclude-directoryes=icons http://atlarge.ewi.tudelft.nl/graphalytics/ref/
+	cd "$BASE_DIR"
 }
 
 # Runs and logs the results of running a given $platform.

### 2016-12-09 report/survey.tex
Got survey into ACM sig proceedings format.
Tweaked get-hwinfo and the specs to be easier to fit in 2-column
mode. Got tools table to be two columns wide.

Added sty and cls files necessary for ACM format.

@@
-\documentclass[11pt]{article}
+\documentclass{sig-alternate}
+%\documentclass[11pt]{article}
+%\usepackage[margin=1in, headheight=15pt]{geometry} % No geometry with sig-alternate
+%\usepackage{fancyhdr, setspace}
+%\renewcommand{\headrulewidth}{0pt}
+%\pagestyle{fancy}
+%\setlength{\headheight}{15pt} % Removes warnings. Set to 30pt if 2 lines on the header.
+%\lhead{}
+%\rhead{}
+%\usepackage{totcount}
+%\regtotcounter{page}
+%\cfoot{\ifnum\totvalue{page} > 1 \thepage \else\fi}
+%\usepackage{amsmath, amsthm, amsfonts}
+
 \usepackage[T1]{fontenc}
 \usepackage{lmodern}
-\usepackage[margin=1in, headheight=15pt]{geometry}
-\usepackage{amsmath, amsthm, amsfonts}
 \usepackage{hyperref}
@@
 
-\usepackage{fancyhdr, setspace}
-\renewcommand{\headrulewidth}{0pt}
-\pagestyle{fancy}
-\setlength{\headheight}{15pt} % Removes warnings. Set to 30pt if 2 lines on the header.
-\lhead{}
-\rhead{}
-\usepackage{totcount}
-\regtotcounter{page}
-\cfoot{\ifnum\totvalue{page} > 1 \thepage \else\fi}
-
 \usepackage{enumitem}
@@
 \begin{document}
-\title{Simplifying Parallel Graph Processing: \\
-	\large Survey of Existing Platforms
+%%% ACM title stuff %%%
+\setcopyright{acmcopyright}
+\doi{?}
+\isbn{?}
+%\acmPrice{}
+\conferenceinfo{Load Testing and Benchmarking of Software Systems (LTB)}{2017, L'Aquila, Italy}
+\title{Simplifying Parallel Graph Processing}
+\subtitle{A Survey of Existing Platforms}
+\numberofauthors{2}
+\author{
+\alignauthor
+	Samuel Pollard\\
+	\affaddr{University of Oregon}\\
+	\affaddr{Eugene, OR, USA}\\
+	\email{spollard@cs.uoregon.edu}
+\alignauthor
+	Boyana Norris\\
+	\affaddr{University of Oregon}\\
+	\affaddr{Eugene, OR, USA}\\
+	\email{norris@cs.uoregon.edu}
 }
-\author{Samuel Pollard (\href{mailto:spollard@cs.uoregon.edu}{spollard@cs.uoregon.edu})}
-\date{\today}
+% Generated from http://dl.acm.org/ccs/ccs.cfm
+\begin{CCSXML}
+	<ccs2012>
+	<concept>
+	<concept_id>10011007.10010940.10011003.10011002</concept_id>
+	<concept_desc>Software and its engineering~Software performance</concept_desc>
+	<concept_significance>500</concept_significance>
+	</concept>
+	</ccs2012>
+\end{CCSXML}
+\ccsdesc[500]{Software and its engineering~Software performance}
+
+%%% Normal title stuff %%%
+%\title{Simplifying Parallel Graph Processing: \\
+%	\large Survey of Existing Platforms
+%}
+%\author{Samuel Pollard (\href{mailto:spollard@cs.uoregon.edu}{spollard@cs.uoregon.edu})}
+%\date{\today}
 \maketitle
 
+\begin{abstract}
 This is a survey of existing graph analytics frameworks.
+\end{abstract}
 
@@
 Whether these actually achieve their goal is one of the purposes of this project.
-\begin{table}[!htb]
+\begin{table*}[t]
 	\begin{minipage}{\linewidth} % So the footnotes all get printed on the same page
 		\centering
-		% \small
+		%\small
 		\pgfplotstabletypeset[
@@
 	\end{minipage}
-\end{table}
+\end{table*}
 
@@
 
-\bibliographystyle{acm}
+%\bibliographystyle{acm}
+\bibliographystyle{abbrv}
 \bibliography{drp}
-
 \end{document}

### 2016-12-09 get-hwinfo.sh

@@ -26,6 +26,7 @@ else
 	RAM_SIZE=$(cat /proc/meminfo | grep MemTotal | awk '{print $2 / 1024 / 1024 "GB"}')
 	CPU_CLOCK=$(printf "%.0f%s" $(echo $(cat /sys/devices/system/cpu/cpu0/cpufreq/scaling_max_freq) "/ 1000" | bc) "MHz")
 fi
+CPU_MODEL=$(echo $CPU_MODEL | sed 's/CPU[[:space:]]*//') # We assume it's a CPU
 
 # Get data that requires superuser.
 if [ -f "lshw.txt" -a -f "dmidecode.txt" ]; then

### 2016-12-14 taxonomy/taxonomy.tex
Added a template for graph processing taxonomy.
This will eventually describe the current state of the art.

--- 2016-11-27 report/report.tex
@@
-\documentclass[11pt]{article}
+\documentclass[11pt]{scrartcl}
 \usepackage[T1]{fontenc}
@@
 \begin{document}
-\title{Parallel Graph Processing Report\vspace{-2em}}
-\author{}
-\date{Generated \today\ at \currenttime}
+\title{The State of Graph Processing}
+\subtitle{APIs, Libraries, Platforms, and Programming Languages}
+\author{Samuel Pollard}
+\date{\today}
 \maketitle
 
-This report was automatically generated from \verb|PATH_TO_MAKEFILE| on \verb|HOSTNAME|.
-
-\section{Machine Specifications}
-These data were gathered from \verb|specs.csv|. If these results are not complete enough, try running \verb|get-hwinfo.sh| again and run the command it suggests as root.
-
-% TODO: Make this a pgfplotstable with an automatically generated csv.
-\begin{table}[htb]
-	\centering
-% Keep in mind you can do this at the beginning: string replace={s1}{s2}
-	\pgfplotstabletypeset[
-	header=false,
-	col sep=tab,
-	string type,
-    every head row/.style={output empty row, before row=\bottomrule},
-	columns/0/.style={column type={|r|}},
-	columns/1/.style={column type={l|}},
-	every last row/.style={after row=\toprule},
-	]{specs.csv}
-	\caption{Machine specifications.}
-	\label{tab:specs}
-\end{table}
-
-\section{Performance Results}
-Performance in millions of traversed edges per second (MTEPS)
-
-% TODO: Actually get this to autogenerate.
-\begin{table}[htb]
-	\centering
-	\begin{tabular}{l|r|r|}
-	 & PowerGraph & OpenG \\ \hline
-	BFS & $87.4$ & $341$ \\ \hline
-	SSSP & $1.09$ & $3.08$ \\ \hline
-	\end{tabular}
-	\caption{Performance Results}
-	\label{tab:perf}
-\end{table}
+\section{Introduction}
 

### 2016-12-14 taxonomy/taxonomy.tex
Updated taxonomy, added pdf.
Now there are sections for the types of graph processing we care
about. There are a whole bunch out there so I'm not being complete
yet, but a good format may be something semicolon-delimited, stating
the paper (or a citation eventually), when the project began, its
architecture, last activity, etc.

@@
-\documentclass[11pt]{scrartcl}
+\documentclass[11pt]{article}
 \usepackage[T1]{fontenc}
@@
 \pgfplotsset{compat=1.12}
-\usepackage{datetime}
 
-\usepackage{fancyhdr, setspace}
+\usepackage{fancyhdr}
 \renewcommand{\headrulewidth}{0pt}
@@
 
+\usepackage{enumitem}
+
 \begin{document}
-\title{The State of Graph Processing}
-\subtitle{APIs, Libraries, Platforms, and Programming Languages}
+\title{The State of Graph Processing \\ \large APIs, Libraries, Platforms, and Programming Languages \vspace{-1em}}
 \author{Samuel Pollard}
@@
 
-\section{Introduction}
+\section{APIs and Libraries}
+\begin{itemize}
+	\item \textbf{Pregel.} API; developed by Google; distributed; vertex centric, Bulk-Synchronous Parallel model; inspired many other platforms such as Giraph and GPS; \href{http://dl.acm.org/citation.cfm?id=1807184}{original paper}; 20.
+
+	\item \textbf{GPS (Graph Processing System).} API; developed by Stanford; distributed; vertex centric, Bulk-Synchronous Parallel model; open source; similar to Pregel but with dynamic graph repartitioning and other enhancements.
+\href{http://ilpubs.stanford.edu:8090/1039/7/gps_ssdbm.pdf}{original paper}; \href{http://infolab.stanford.edu/gps/}{website}; First appeared 2013; apprears inactive.
+
+	\item \textbf{GraphX} Library.
+\end{itemize}
+
+\section{Programming Languages}
+
+\section{Benchmarks}
+\begin{itemize}
+	\item \textbf{Graphalytics.} CPU and GPU; Supports GraphMat, GraphX, PowerGraph
+	\item \textbf{GAP (Graph Algorithm Platform).} CPU; shared Memory (OpenMP); Berkeley; last active October 2016 on Github.
+	\item \textbf{GraphBIG.} CPU and GPU; shared Memory and CUDA; last active February 2016.
+	\item \textbf{Lonestar.} CPU and GPU; shared memory and CUDA; part of Galois. First appeared 2011; last update appears to be in 2015.
+\end{itemize}
+\section{Dynamic Graphs}
+TODO: Determine if they are parallel or serial.
+\begin{itemize}
+	\item \textbf{STINGER.} Data structure and library; Georgia Institute of Technology and various national laboratories; streaming model; parallel or serial, distributed or shared; \href{http://cass-mt.pnnl.gov/docs/pubs/pnnlgeorgiatechsandiastinger-u.pdf}{original paper}; \href{http://www.stingergraph.com/}{website}; First appeared 2009; still active on Github.
+
+	\item \textbf{DynoGraph.} \href{http://sc16.supercomputing.org/sc-archive/tech_poster/poster_files/post214s2-file3.pdf}{original paper}.
+
+	\item \textbf{GraphJet.} \href{http://www.vldb.org/pvldb/vol9/p1281-sharma.pdf}{original paper}.
+
+	\item \textbf{GraphStream.} Java library; I'm not sure but it appears to be serial; \href{https://arxiv.org/abs/0803.2093}{original paper}; first active 2007; appears still active;
+\end{itemize}
+
+\section{Visualization}
+There are many graph visualization tools out there, most notably Graphviz (and its associated DOT file format).
+Likewise, databases such as Neo4j have their own visualization tools.
+This is concerned only with visualizations which are scalable to a large number of vertices and edges.
+
+\begin{itemize}
+	\item \textbf{Gephi.} \href{https://gephi.org/}{website}
+\end{itemize}
 

### 2016-12-15 taxonomy/taxonomy.tex
Updated taxonomy and broke up into 5 sections.
Added a bit for dynamic graphs. The APIs and libraries
are definitely not complete, nor is the programming languages.
Next order of priority is to update programming languages.

@@
 \usepackage{amsmath, amsthm, amsfonts}
+\usepackage[hyphens]{url}
 \usepackage{hyperref}
@@
 \begin{document}
-\title{The State of Graph Processing \\ \large APIs, Libraries, Platforms, and Programming Languages \vspace{-1em}}
+\title{The State of Graph Processing \\ \large APIs, Libraries, Benchmarks, and Programming Languages \vspace{-1em}}
 \author{Samuel Pollard}
@@
 \begin{itemize}
-	\item \textbf{Pregel.} API; developed by Google; distributed; vertex centric, Bulk-Synchronous Parallel model; inspired many other platforms such as Giraph and GPS; \href{http://dl.acm.org/citation.cfm?id=1807184}{original paper}; 20.
+	\item \textbf{Pregel.} API; developed by Google; distributed; vertex centric, Bulk-Synchronous Parallel model; inspired many other platforms such as Giraph and GPS; \href{http://dl.acm.org/citation.cfm?id=1807184}{original paper}; 2010.
 
@@
 	\item \textbf{GraphX} Library.
+
+	% OpenCL: http://www.lidi.info.unlp.edu.ar/WorldComp2011-Mirror/PDP3248.pdf - Appears inactive since 2011.
 \end{itemize}
@@
 \section{Programming Languages}
+\begin{itemize}
+	\item \textbf{GP (Graph Programs).} Nondeterministic; serial; original paper: \cite{Plump:2009:GPL}; website: \url{https://www.cs.york.ac.uk/plasma/wiki/index.php?title=GP_(Graph_Programs)}; appears to be more theoretical and used for program verification; still active.
+
+	\item \textbf{Gremlin.} functional, data flow; distributed; a way to interact with graph databases; \cite{Rodriguez:2015:Gremlin}; website: \url{http://tinkerpop.apache.org/gremlin.html}.
+\end{itemize}
 
@@
 \begin{itemize}
-	\item \textbf{Graphalytics.} CPU and GPU; Supports GraphMat, GraphX, PowerGraph
-	\item \textbf{GAP (Graph Algorithm Platform).} CPU; shared Memory (OpenMP); Berkeley; last active October 2016 on Github.
+	\item \textbf{Graphalytics.} CPU and GPU; supports GraphMat, PowerGraph, GraphBIG, Giraph, GraphX, Neo4j, and MapReduce; \cite{Capota:2015:Graphalytics}; still active; so far I can only get PowerGraph and GraphBIG running; website: \url{http://graphalytics.ewi.tudelft.nl}.
+
+	\item \textbf{GAP (Graph Algorithm Platform).} CPU; shared Memory (OpenMP); \url{http://gap.cs.berkeley.edu/benchmark.html}; last active October 2016 on Github. \cite{Beamer:2015:GAPBench}
+
 	\item \textbf{GraphBIG.} CPU and GPU; shared Memory and CUDA; last active February 2016.
-	\item \textbf{Lonestar.} CPU and GPU; shared memory and CUDA; part of Galois. First appeared 2011; last update appears to be in 2015.
+
+	\item \textbf{Lonestar.} CPU and GPU; shared memory and CUDA; part of Galois.
+First appeared 2011; last update appears to be in 2015 though I am in recent (December 2016) contact with someone working on the project.
 \end{itemize}
+
 \section{Dynamic Graphs}
-TODO: Determine if they are parallel or serial.
 \begin{itemize}
@@
 
-	\item \textbf{DynoGraph.} \href{http://sc16.supercomputing.org/sc-archive/tech_poster/poster_files/post214s2-file3.pdf}{original paper}.
+	\item \textbf{GraphJet.} Java library; parallel but single-machine; original purpose was for real time recommendations from Twitter; \url{https://github.com/twitter/GraphJet}; \cite{Sharma:2016:GraphJet}.
 
-	\item \textbf{GraphJet.} \href{http://www.vldb.org/pvldb/vol9/p1281-sharma.pdf}{original paper}.
+	\item \textbf{GraphStream.} Java library; I'm not sure but it appears to be serial; \url{http://graphstream-project.org/}; \cite{Dutot:2007:GraphStream} first active 2007; appears still active.
 
-	\item \textbf{GraphStream.} Java library; I'm not sure but it appears to be serial; \href{https://arxiv.org/abs/0803.2093}{original paper}; first active 2007; appears still active;
+	\item \textbf{Kineograph.} API; parallel and distributed; \cite{Cheng:2012:Kineograph}; 2012; appears to be inactive but influenced GraphX, GraphChi, and PowerGraph.
+
+	\item \textbf{DynoGraph.} Performance Analysis; \url{https://github.com/sirpoovey/DynoGraph} \cite{Hein:2016:DynoGraph}; connected components, PageRank, BFS; right now just a repo and a poster at SC'16.
 \end{itemize}
@@
 
+% Other stuff: Giraph++: http://dl.acm.org/citation.cfm?id=2732238
+% GraphBuilder: http://dl.acm.org/citation.cfm?id=2484429 and https://01.org/graphbuilder/ - Active from 2013--2014
+% Chronos: http://dl.acm.org/citation.cfm?id=2592799
+% X-Stream: http://dl.acm.org/citation.cfm?id=2522740
+
+\bibliographystyle{plain}
+\bibliography{../report/drp}
+
 \end{document}

### 2016-12-15 get-gap.sh

@@ -107,6 +107,10 @@ for d in "${DATASETS[@]}"; do
 	done
 done
 
+# TODO: Run the PowerGraph benchmark.
+# We do this instead of graphalytics?
+# Graphalytics measures the time of reading the graph.
+
 # Clean up
 cd "$BASE_DIR"
 

### 2016-12-15 get-gap.sh
Got get-gap to convert graphalytics -> graphBIG.
Still need to do this for more datasets, run it on gap,
and then parse the output.

@@ -47,9 +47,23 @@ if [ $FOUND -eq 0 ]; then
 	exit 1
 fi
 
-### Run the benchmarks
-# The general loop structure is a doubly-nested for loop
-# looping over each dataset then each algorithm.
+# Transform datasets from Graphalytics format to GraphBIG format
+# (We later transform from GraphBIG to GAP)
+GRAPHALYTICS_DATASETS=( dota-league cit-Patents )
+for d in ${GRAPHALYTICS_DATASETS[@]}; do
+	mkdir -p "$DATASET_DIR/$d"
+	if [ $(awk '{print NF; exit}' "$DATASET_DIR/$d.e") -eq 2 ]; then
+		echo "SRC,DEST" > "$DATASET_DIR/$d/edge.csv"
+	elif [ $(awk '{print NF; exit}' "$DATASET_DIR/$d.e") -eq 3 ]; then
+		echo "SRC,DEST,WEIGHT" > "$DATASET_DIR/$d/edge.csv"
+	else
+		echo "File format not recognized"
+		exit 1
+	fi
+	sed 's/[:space:]+/,/' "$DATASET_DIR/$d.e" >> "$DATASET_DIR/$d/edge.csv"
+	echo "ID" > "$DATASET_DIR/$d/vertex.csv"
+	sed 's/[:space:]+/,/' "$DATASET_DIR/$d.v" >> "$DATASET_DIR/$d/vertex.csv"
+done
 
 # TODO: Get other datasets converted to the right formats.
 # Right now: LDBC Graphalytics datasets are not compatible since the edge numbers skip
@@ -61,6 +75,10 @@ GAP_ALGORITHMS=( bfs pr sssp )
 GRAPHBIG_ALGORITHMS=( bfs pagerank sssp )
 ALGORITHM_DIRS=( bench_BFS bench_pageRank bench_shortestPath ) # Just used for GraphBIG
 
+### Run the benchmarks
+# The general loop structure is a doubly-nested for loop
+# looping over each dataset then each algorithm.
+
 # Run GAP Benchmark
 mkdir "$BASE_DIR/output"
 for d in "${DATASETS[@]}"; do
@@ -68,6 +86,7 @@ for d in "${DATASETS[@]}"; do
 	# Transform the datasets from GraphBIG (CSV) -> GAP ([w]el) format
[…]

### 2016-12-15 get-graphalytics.sh

@@ -380,6 +380,7 @@ install_Giraph()
 }
 
 # Downloads em all
+# NOTE: This is finnicky. If you need the datasets, please submit an issue on GitHub.
 download_datasets()
 {
 	cd "$DATASET_DIR"

### 2016-12-28 taxonomy/taxonomy.tex
Added some more sources to dynamic graph taxonomy.

@@
 \section{Dynamic Graphs}
+The primary focuses for dynamic graphs encountered thus far are: dynamic updating of the graph objects themselves and the dynamic partitioning and re-partitioning of the graphs across a distributed architecture.
+The distinction is made clear below:\footnote{These resources are in part retrieved from \url{http://scicomp.stackexchange.com/questions/4722/i-am-looking-for-a-parallel-dynamic-graph-library-in-c}.}
+
+Dynamic Partitioning
 \begin{itemize}
-	\item \textbf{STINGER.} Data structure and library; Georgia Institute of Technology and various national laboratories; streaming model; parallel or serial, distributed or shared; \href{http://cass-mt.pnnl.gov/docs/pubs/pnnlgeorgiatechsandiastinger-u.pdf}{original paper}; \href{http://www.stingergraph.com/}{website}; First appeared 2009; still active on Github.
+	\item \textbf{Zoltan.} \url{http://www.cs.sandia.gov/zoltan/} \cite{Devine:2000:Zoltan}.
+	\item \textbf{PTScotch.} \url{https://www.labri.fr/perso/pelegrin/scotch/} \cite{Chevalier:2008:PTScotch}.
+	\item \textbf{ParMETIS.} \url{http://glaros.dtc.umn.edu/gkhome/metis/parmetis/overview}; \cite{Karypis:1998:ParMETIS}.
+\end{itemize}
+
+Streaming graphs/support for dynamic updating of graphs.
+Many of these are advertised to social network websites or consumers of large amounts of streaming data (for example, see \url{http://storm-project.net}).
+The key word here is \emph{analytics}.
+\begin{itemize}
+	\item \textbf{STINGER.} Data structure and library; Georgia Institute of Technology and various national laboratories; streaming model; parallel or serial, distributed or shared; \url{http://cass-mt.pnnl.gov/docs/pubs/pnnlgeorgiatechsandiastinger-u.pdf}; \url{http://www.stingergraph.com/}; First appeared 2009; still active on Github.
 
@@
 
-	\item \textbf{DynoGraph.} Performance Analysis; \url{https://github.com/sirpoovey/DynoGraph} \cite{Hein:2016:DynoGraph}; connected components, PageRank, BFS; right now just a repo and a poster at SC'16.
+	\item \textbf{PHISH.} Streaming graph processing; \url{http://www.sandia.gov/~sjplimp/phish.html}
+
+	\item \textbf{PGX.} Oracle Project. Supports incremental updates though the dynamic aspect is not forefront.
+	% \item \textbf{DynoGraph.} Performance Analysis; \url{https://github.com/sirpoovey/DynoGraph} \cite{Hein:2016:DynoGraph}; connected components, PageRank, BFS; right now just a repo and a poster at SC'16.
 \end{itemize}

### 2017-01-13 papers/gabb17/gabb17-pollard.tex
Started GABB17 paper. Moved writeups to papers/
Next order of business is to get some reference impelementations
all going on ACISS or some standard platform.

--- 2016-12-02 report/survey.tex
@@
-\documentclass[11pt]{article}
+\documentclass[conference]{IEEEtran}
 \usepackage[T1]{fontenc}
 \usepackage{lmodern}
-\usepackage[margin=1in, headheight=15pt]{geometry}
-\usepackage{amsmath, amsthm, amsfonts}
 \usepackage{hyperref}
@@
 
+\title{A Comparison of Parallel Graph Processing Platforms}
+\author{
+	\IEEEauthorblockN{Samuel Pollard}
+	\IEEEauthorblockA{Department of Computer and Information Science \\
+		University of Oregon \\
+		Eugene, OR, USA \\
+		Email: spollard@cs.uoregon.edu
+	}
+	\and
+	\IEEEauthorblockN{Boyana Norris}
+	\IEEEauthorblockA{Department of Computer and Information Science \\
+	University of Oregon \\
+	Eugene, OR, USA \\
+	Email: norris@cs.uoregon.edu
+	}
+}
+
 \begin{document}
-\title{Simplifying Parallel Graph Processing: \\
-	\large Survey of Existing Platforms
-}
-\author{Samuel Pollard (\href{mailto:spollard@cs.uoregon.edu}{spollard@cs.uoregon.edu})}
-\date{\today}
 \maketitle
-
+\begin{abstract}
 This is a survey of existing graph analytics frameworks.
+\end{abstract}
 
@@
 	\centering
-% Arya
-%	\begin{tabular}{l|r}
-%		CPU & 72 Core Intel Xeon E5 2699 v3\\ \hline
-%		RAM & 256 GB DDR4 2133 MHz
-%	\end{tabular}
-
 	% Keep in mind you can do this at the beginning: string replace={s1}{s2}
@@
 	every last row/.style={after row=\toprule},
-	]{specs.csv}
+	]{../report/specs.csv}
 	\caption{Machine specifications.
@@
 
-%\begin{table}[htb]
-%	\centering
-%	\begin{tabular}{lr}
-%		Transport & \\ \hline
-%		Network Topology & \\ \hline
-%		Local Scheduling & \\ \hline
-%		Runtime Feedback & \\ \hline
-%		Approach & \\ \hline
-%		Algorithmic Considerations & \\ \hline
-%	\end{tabular}
-%	\caption{Middleware specifications}
-%	\label{tab:reportcard}
-%\end{table}
+\begin{table}[htb]
+	\centering
+	\begin{tabular}{l|c|c|c}
+			& BFS & SSSP & PR \\ \hline
+		Graph500 & & & \\ \hline
+		PBGL  & & & \\ \hline
+		GAP  & & & \\ \hline
+		GraphBIG  & & & \\ \hline
+		Galois  & & & \\ \hline
+		PowerGraph  & & &
+	\end{tabular}
+	\caption{Performance. Note that implementations of every algorithm are not availabe for every platform.}
+	\label{tab:reportcard}
+\end{table}
 
@@
 
-Table~\ref{tab:perf} lists performance in milliseconds of runtime according to the graphalytics output.
+%Table~\ref{tab:perf} lists performance in milliseconds of runtime according to the graphalytics output.
 Graphalytics also outputs MTEPS or millions of traversed edges pers econd.
@@
 			columns/powergraph/.style={string type, column type={r}}
-		]{runtime.csv}
+		]{../report/runtime.csv}
 	\caption{Performance Results for the \texttt{dota-league} dataset with 61,670 vertices and 50,870,313 edges.}
@@
 			},
-		]{platforms.csv}
+		]{../report/platforms.csv}
 		\caption{Tools used for graph processing}
@@
 
-\bibliographystyle{acm}
-\bibliography{drp}
-
+%\bibliographystyle{acm}
+\bibliographystyle{IEEEtran}
+\bibliography{../drp}
 \end{document}

### 2017-01-13 papers/taxonomy/taxonomy.tex
Started GABB17 paper. Moved writeups to papers/
Next order of business is to get some reference impelementations
all going on ACISS or some standard platform.

--- 2016-12-28 taxonomy/taxonomy.tex
@@
 
-	\item \textbf{GraphStream.} Java library; I'm not sure but it appears to be serial; \url{http://graphstream-project.org/}; \cite{Dutot:2007:GraphStream} first active 2007; appears still active.
+	\item \textbf{GraphStream.} Java library; appears to be serial and focuses mainly on visualization; \url{http://graphstream-project.org/}; \cite{Dutot:2007:GraphStream} first active 2007 but appears active as of December 2016.
 
@@
 
-	\item \textbf{PGX.} Oracle Project. Supports incremental updates though the dynamic aspect is not forefront.
+	\item \textbf{PGX.} Oracle Project; integration with Groovy, Green Marl, Spark, Hadoop; Supports incremental updates though the dynamic aspect is listed as future work. \cite{Hong:2015:PGX}.
 	% \item \textbf{DynoGraph.} Performance Analysis; \url{https://github.com/sirpoovey/DynoGraph} \cite{Hein:2016:DynoGraph}; connected components, PageRank, BFS; right now just a repo and a poster at SC'16.
@@
 \bibliographystyle{plain}
-\bibliography{../report/drp}
+\bibliography{../drp}
 

### 2017-01-13 papers/report/merge-csv.sh

@@ -1,0 +1,36 @@
+#!/bin/bash
+# Merges two csv files, given that they both only use one dataset for each run.
+# To use with latex: ./merge-csv.sh > runtimes.csv
+# Grabs all of the runs for a given dataset and averages them.
+# To specify: The EXPERIMENT_DIR directory specicies where the output is saved.  
+EXPERIMENT_DIR=/home/users/spollard/graphalytics/experiments
+first=1
+outfile=runtime.csv
+N_EXPERIMENTS=0
+if [ -z "$EXPERIMENT_DIR" ]; then
+	echo "Please specify where the output csv files are located."
+	exit 1
+fi
+PLATFORMS="openg powergraph"
+# Compute the mean runtime.
+for p in $PLATFORMS; do
+	P_RUNS=$(find "$EXPERIMENT_DIR" -path */"${p}"-report-*/runtime.csv)
+	awk -F ',' '/^,/{if (ARGIND==1) print $0} !/^,/{A[$1] += $2} END{for (alg in A) print alg "," (A[alg] / ARGIND)}' $P_RUNS > "${p}-runtime.csv"
+	N_EXPERIMENTS=$(awk 'BEGIN{print ARGC}' $P_RUNS)
+done
+# Assumes everything is in the same directory named "openg-runtime.csv" etc.
+# Assumes there is only one dataset---named $dataset1.
+echo "Writing"
+for f in *-runtime.csv; do
+	if [ "$first" -eq 1 ]; then
+ 		first=0
+ 		experiment1=${f%-runtime.csv}
+ 		dataset1=$(grep '^,' $f | cut -b 2-)
+ 		cat "$f" | sed "s/$dataset1/$experiment1/" > $outfile
+ 	else
+ 		experimentn=${f%-runtime.csv}
+		cat "$f" | cut -d ',' -f 2- | sed "s/$dataset1/$experimentn/" | paste -d, runtime.csv - | tee "$outfile"
+ 	fi
+done
+echo "to $outfile using dataset $dataset1 over the course of $N_EXPERIMENTS runs"
+

### 2017-01-13 papers/drp-contract/drp-contract-pollard.tex
Added drp contract folder.

\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{hyperref}

\usepackage{fancyhdr, setspace}
\renewcommand{\headrulewidth}{0pt}
\pagestyle{fancy}
\setlength{\headheight}{15pt} % Removes warnings. Set to 30pt if 2 lines on the header.
\lhead{}
\rhead{}
\usepackage{totcount}
\regtotcounter{page}
\cfoot{\ifnum\totvalue{page} > 1 \thepage \else\fi}

\usepackage{enumitem}
\setlist{parsep=0pt, listparindent=0.5cm}

\usepackage{titlesec}
\titleformat*{\section}{\large\bfseries}

\begin{document}
\begin{center}
{ \huge
	Simplifying Parallel Graph Processing: \\
}
{ \Large
	Directed Research Project Contract \\
}
Sam Pollard (\href{mailto:spollard@cs.uoregon.edu}{spollard@cs.uoregon.edu}), University of Oregon \\
\today
\end{center}

\section{Introduction}
Graph processing has important differences from other High Performance Computing (HPC) applications. For example, floating point operations per second mean little when performing a breadth first search. Furthermore, properties of the input graph such as its sparsity can play a profound role in an algorithm's performance.
% bfs turns out to be memory latency-bound instead of compute or memory bandwidth-bound \cite{Beamer:2016:GAP, Eisenman:2016:Prejudice}.
Because of these differences, there has been movement toward standardizing performance measurements specifically directed at graph processing. The largest example is the Graph500 \cite{Murphy:2010:Graph500}, a ranking of supercomputers in the spirit of the Top500. Other projects such as Graphalytics \cite{Guo:2014:Graphalytics} attempt to compare performance across various platforms while projects such as the Graph Algorithm Platform Benchmark Suite \cite{Beamer:2016:GAP} and GraphBIG \cite{Nai:2015:Graphbig} attempt to provide high quality, reference implementations next to which performance can be compared. Here, we use the definition of \emph{platform} defined in \cite{Guo:2014:Graphalytics}: ``the combined hardware, software, and programming system that is being used to complete a graph processing task.''

There are a plethora of programming paradigms, domain specific languages, and libraries for parallel graph processing. Together these cover a diverse range of applications, programming languages, and target architectures. Whereas the Graph500 measures scalability, there is impressive research on achieving the highest performance per core on shared memory systems. For example, the GraphChi platform in particular showed performance on a personal computer can have performance comparable to a medium-sized cluster \cite{kyrola:2012:Graphchi}.

Publications introducing new platforms generally compare their performance to popular existing platforms on a few standard datasets. However, these performance measurements are naturally biased towards the authors' works. Thus, when attempting to solve a graph-based problem, the question of \emph{which} platform to use is daunting. Beyond this, platforms have unique input file formats, programming paradigms, configuration demands, and dependencies, all of which combine to yield a steep learning curve and a lack of portability, which are concerns applicable to HPC in general.

\section{Project Description}
This project aims to analyze the performance of graph algorithms and recommends an optimal platform to the user based on this analysis. The project aims to look beyond the number of traversed edges per second (TEPS) on just breadth-first search as in the Graph500. Potential measurements include strong and weak scalability, load balancing, memory utilization, and disk usage. The Tuning and Analysis Utilities (TAU) will be investigated as a potential source for these measurements \cite{Shende:2006:Tau}. Beyond performance analysis, hardware and middleware specifications will be reported such as data transport mechanisms and scheduling paradigms as inspired by \cite{Firoz:2016:Reportcard}.

% Idea: Measure scalability for different problem sizes differently: measure 2 classes: one in which distributed systems make the most sense and one for which shared memory systems make the most sense (and find that magic number!). Define \emph{scale} as $\log_10(|E| + |V|)$ \cite{Iosup:2016:Graphalyticstech}.

% Also: May want to list all the existing graph processing benchmarks: GAP, Graphalytics, GraphBIG, BigBench 2.0 (in development).

%The goal of such detailed measurement is to make the selection of a particular graph processing platform easier; these data can be used to moref intelligently direct programmer effort. This project aims to create a utility to make these measurements, but also takes in user-specified architecture and graph properties to intelligently select the optimal platform from a performance standpoint.

This project will begin by building on top of the graph benchmarking tool Graphalytics (described in \cite{Guo:2014:Graphalytics}) with increased focus on describing the hardware, communication paradigms, scheduling, and other lower-level details as recommended by \cite{Firoz:2016:Reportcard}. Preliminary research has shown Graphalytics to be unportable and challenging to extend, so this project will provide a much-needed simplification over existing frameworks. Furthermore, this project will standardize performance analysis for each platform. This will be accomplished through comprehensive analysis of all prominent platforms. Additionally, modeling of the target architectures will be used to explain the observations performance and to predict performance on new architectures.

Lastly, this project aims to use these data to provide recommendations so users can make informed decisions about which platform is optimal for a given architecture, problem size, and algorithm. At their simplest, these recommendations will be based on existing measurements and of the form: ``given machine $X$ running algorithm $Y$ on a problem of size $Z$, the optimal platform is $P$.'' This format is restrictive (the user may not know $X$, $Y$, or $Z$) and does not address the cost of programming on a new platform for each problem. The first step to address this would be to use a model of graph processing performance to predict the optimal platform with less information. Time permitting, the use of a domain specific language to translate high level algorithm specifications to the optimal platform will be investigated.
% To foster increased developer interaction, an online leaderboard could be created which ranks platforms by various metrics such as scalability (for small problem sizes), scalability (for large problem sizes), power efficiency, simplicity of implementation (number of steps required to get the problem solved), and performance for various algorithms. This would be done to spark competition and to motivate developers to provide high quality implementations of each algorithm.

\section{Timeline}
	\begin{enumerate}
		\item[Fall '16.] Continue literature search about benchmarking for graph processing algorithms. Decide on two or three platforms as starting points to analyze performance and tabulate basic performance measurements. Look into methods to increase portability such as Spack \cite{Gamblin:2015:Spack}.
		% \item Apply for the NDSEG fellowship by \textbf{December 9}.
		\item[Winter.]  Get preliminary measurements and begin to analyze which platforms are optimal for a given input, algorithm, and hardware. Create a model to explain observed performance differences. Use these data to make simple platform recommendations. Coalesce all noteworthy graph processing platforms, prepare an exhaustive list of these platforms. This may be the beginnings of an area exam survey paper. Time and resource-permitting: install a power monitor on the research hardware to increase quality and depth of performance measurement.
		% 		\item Select the most prominent and promising of these platforms for detailed analysis.
		\item[Spring.] Keep up with the state of the art by adding more platforms and architectures to the measurements. Enhance portability by testing on various architectures. Compile performance results and begin writing the paper.
		\item[Summer.] Enhance model of platform, algorithm, and dataset performance. Improve recommender to work in cases where the user does not have complete knowledge of the problem to be solved. Write paper and investigate which conference or journal is appropriate.
		\item[Fall '17.] Submit the paper to a conference or journal. Present to DRP committee. Time permitting: investigate viability of a domain specific language (DSL). Develop the DSL so it can express the most popular algorithms on the most popular platforms.
		% \item (Time Permitting) Make an online leaderboard which measures a platform, not a supercomputer: this would rank graph processing frameworks by various aspects such as scalability and computational efficiency.
		% \item (Time Permitting) Make an interative tool where you select some number of choices (always including "I don't know" as a possible choice), and the tool spits out which platform to use.
	\end{enumerate}

\section{Deliverables}
	\begin{enumerate}
		\item A comprehensive list of graph processing platforms and descriptions of them. Categorize these by prominence so analysis and porting effort is well-spent.
		\item Machine and human readable, complete, precise performance analysis for several graph processing platforms on the existing research hardware\footnote{This machine is named Arya and is an Intel Xeon-based machine with 72 cores, an NVIDIA GTX 980 GPU and 256GB RAM} and the UO cluster. These measurements will be performed on several standardized datasets and the most common algorithms such as breadth first search, local clustering coefficients, single-source shortest paths, PageRank, and community detection.
		\item A recommender tool which can be easily run on various architectures.
		\item A paper to submit to a conference or journal.
		\item A workshop-style tutorial of this recommender tool.
		\item Time permitting: a proof of concept DSL.
	\end{enumerate}

\section{Committee Members}
\begin{enumerate}
	\item Boyana Norris
	\item Allen Malony
	\item Zena Ariola
\end{enumerate}
\bibliographystyle{acm}
\bibliography{drp}
\end{document}

### 2017-01-14 papers/gabb17/gabb17-pollard.tex
GABB17: Added more sources and expanded introduction.
Added a small description of why and what algorithms are being
used.

@@
 
+\section{Introduction}
+
+[Talk about WHY we care about graph processing, and parallel in particular. Talk about what parallel graph processing is.]
+
+Our research is motivated by the current state of parallel graph processing.
+The most comprehensive survey, released in 2014, identified and taxonomized over 80 different parallel graph processing systems \cite{Doekemeijer:2015:GPFSurvey}.
+These systems operate with a wide range of parallelism paradigms such as GPU [Medusa], shared memory [pretty much everything], a combination of CPU and GPU \cite{Gharaibeh:2012:Totem}, distributed database querying, \cite{Rodriguez:2015:Gremlin}, distributed filesystem based approaches \cite{Xin:2013:GraphX}, distributed memory with MPI [cite], domain-specific languages [Green Marl], as well as novel communication schemes [activemessages, sockets].
+
+Since 2014, the problem has compounded with the addition of even more proprietary and open source projects such as \cite{Cheramangalath:2015:Falcon}, \cite{Perez:2015:Ringo} [name some more].
+At the outset, this plethora of choices makes the question, ``which system is the best for my problem?'' incredibly difficult.
+There has even been a propagation of so-called ``reference implementations'' which provide implementations of the most common graph algorithms [GAP, GraphBIG].
+Thus, even selecting a standard and a benchmark over which to compare various implementations is nontrivial.
+To quote Andrew Tanenbaum, ``The nice thing about standards is that you have so many to choose from.''
+
+Another issue with the parallel graph processing library designers is the lack of comprehensive comparison.
+One possible reason for this is the considerable effort of getting each library and package to work, satisfying dependencies, and ensuring the data is in the correct format for each system.
+Beyond this, each system may have a different method of measuring performance.
+Thus, one of the contributions of this paper is to provide a ``level playing field'' for each graph processing system.
+Graphalytics \cite{Capota:2015:Graphalytics} attempts to remedy this but has not seen widespread adoption.
+
+%This paper aims to make that selection process just a little easier.
+
 \section{Machine Specifications}
@@
 
-% TODO: Make this a pgfplotstable with an automatically generated csv.
 \begin{table}[!htb]
@@
 \end{table}
+
+\section{Algorithms and Systems}
+The canonical performance leaderboard for paralell graph processing is the Graph500 \cite{Murphy:2010:Graph500}.
+The advantage of the Graph500 is it provides standardized measurement requirements and dataset generation.
+The primary drawback with using reference implementations for the Graph500 is the standard only supports a single algorithm: breadth first search.
+
+This report explores a small sample of the existing graph processing platforms with a focus on the so-called ``reference implementations:'' the Graph500 (from \url{http://www.graph500.org}), GAP \cite{Beamer:2015:GAPBench}, and GraphBIG \cite{Nai:2015:Graphbig}.
+Parallel Boost Graph Library \cite{Gregor:2005:PBGL} and PowerGraph \cite{Gonzalez:2012:Powergraph} implementations are also provided because of their popularity and the availability of high quality reference implementationsF.
 

### 2017-01-16 papers/gabb17/gabb17-pollard.tex
Explained graph500 in paper, added graph gen.
The graph generator octave script has not been tested.
The next order of business is to test it, then convert
the file to formats accepted by Galois, GraphBIG, & GAP.

@@
 \begin{abstract}
-This is a survey of existing graph analytics frameworks.
+This is a survey of existing graph analytics frameworks and an analysis of their performance.
 \end{abstract}
@@
 At the outset, this plethora of choices makes the question, ``which system is the best for my problem?'' incredibly difficult.
-There has even been a propagation of so-called ``reference implementations'' which provide implementations of the most common graph algorithms [GAP, GraphBIG].
+There has even been a propagation of so-called ``reference implementations'' which provide implementations of the most common graph algorithms [GAP, GraphBIG, Galois].
 Thus, even selecting a standard and a benchmark over which to compare various implementations is nontrivial.
 To quote Andrew Tanenbaum, ``The nice thing about standards is that you have so many to choose from.''
+
+[This paper aims to simplify the selection process for graph processing.]
 
@@
 Thus, one of the contributions of this paper is to provide a ``level playing field'' for each graph processing system.
-Graphalytics \cite{Capota:2015:Graphalytics} attempts to remedy this but has not seen widespread adoption.
+Graphalytics \cite{Capota:2015:Graphalytics} attempts to remedy this but has not seen widespread adoption---only eight platforms are supported\footnote{According to \url{https://github.com/tudelft-atlarge}}.
+[TODO: Say more about why this is justified instead of just using Graphalytics]
 
-%This paper aims to make that selection process just a little easier.
+\section{Algorithms}
+The canonical performance leaderboard for paralell graph processing is the Graph500 \cite{Murphy:2010:Graph500}.
+The advantage of the Graph500 is it provides standardized measurement requirements and dataset generation.
+The primary drawback with using reference implementations for the Graph500 is the standard only supports a single algorithm: breadth first search (BFS).
+
+This report attempts to add similar rigor to other graph algorithms.
+This is done by borrowing heavily from the Graph500 specification; we use a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19,
+C = 0.19,$ and $D = 1-(A+B+C) = 0.05$.
+The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}.
+Some clarifications are in order: for the first kernel, the time to read in the graph from disk is not considered.
+Furthermore, these data structures need only be created once and BFS is performed on 64 roots.
+The second kernel creates a tree with distances from the root, but need only label the existing data structure (and not output the entire graph, for example).
+
+One straightforward extension is computing the Single-Source Shortest Paths algorithm (SSSP). We use the same graph and the same source vertices as in BFS.
+
+\section{The Existing Systems}
+This report explores a small sample of the existing graph processing platforms with a focus on the so-called ``reference implementations:'' the Graph500\footnote{We used version 2.1.4 from \url{http://graph500.org/referencecode}.}, GAP \cite{Beamer:2015:GAPBench}, and GraphBIG \cite{Nai:2015:Graphbig}.
+
+Parallel Boost Graph Library \cite{Gregor:2005:PBGL} and PowerGraph \cite{Gonzalez:2012:Powergraph} implementations are also provided because of their popularity and the availability of high quality reference implementations.
 
@@
 The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
-The manufacturer's published maximum clock speeds can be found at \url{http://ark.intel.com}.}
+We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.}
 	\label{tab:specs}
 \end{table}
+
+\section{Performance}
 
@@
 
-\section{Algorithms and Systems}
-The canonical performance leaderboard for paralell graph processing is the Graph500 \cite{Murphy:2010:Graph500}.
-The advantage of the Graph500 is it provides standardized measurement requirements and dataset generation.
-The primary drawback with using reference implementations for the Graph500 is the standard only supports a single algorithm: breadth first search.
-
-This report explores a small sample of the existing graph processing platforms with a focus on the so-called ``reference implementations:'' the Graph500 (from \url{http://www.graph500.org}), GAP \cite{Beamer:2015:GAPBench}, and GraphBIG \cite{Nai:2015:Graphbig}.
-Parallel Boost Graph Library \cite{Gregor:2005:PBGL} and PowerGraph \cite{Gonzalez:2012:Powergraph} implementations are also provided because of their popularity and the availability of high quality reference implementationsF.
-
-\section{Performance}
 Graphalytics without the use of the Granula plugin produces performance measurement in two forms: runtime in seconds and traversed edges per second.

### 2017-01-16 prepare-compare.m

@@ -1,0 +1,23 @@
+% GNU Octave Script
+% Prepare to compare and run performance analysis.
+% Must be run in the graph500-2.1.4 reference implementation directory
+% Code taken from graph500-2.1.4/octave
+SCALE = 10;
+edgefactor = 16;
+NBFS = 64;
+rand("seed", 103);
+ij = kronecker_generator (SCALE, edgefactor);
+filename = sprintf ('kron-%d-%d.el',SCALE,edgefactor);
+dlmwrite (filename, ij', ' ')
+G = kernel_1 (ij);
+N = size (G, 1);
+coldeg = full (gpstats (G));
+search_key = randperm (N);
+search_key(coldeg(search_key) == 0) = [];
+if length (search_key) > NBFS,
+  search_key = search_key(1:NBFS);
+else
+  NBFS = length (search_key);
+end
+search_key = search_key - 1;  
+dlmwrite (sprintf ('kron-%d-%d.roots',SCALE,edgefactor), search_key, ' ')

### 2017-01-26 misc/parse-output.sh
Added an output parser.
It is about as non-robust as you can get.

@@ -1,0 +1,85 @@
+#!/bin/bash
+# Parse the output from run-experiment.sh
+
+FN=$1
+NROOTS=64
+if [ -z "$FN" ]; then
+	echo "usage: parse-output.sh <output_file>"
+	exit 2
+fi
+
+# Graph500
+echo -n 'Graph500,BFS,graph construction,'
+echo $(grep "construction_time" "$FN" | awk '{print $2}') 
+echo -n 'Graph500,BFS,graph generation,'
+echo $(grep "generation_time" "$FN" | awk '{print $2}')
+echo -n 'Graph500,BFS,Time,'
+grep "mean_time" "$FN" | awk '{print $2}'
+
+# GraphBIG
+echo -n 'GraphBIG,BFS,File reading,'
+grep -A 7 "Benchmark: BFS" "$FN" | awk -v T=0 '/time/{if(T%2==0){print $3} T++}' | awk '{s+=$1}END{print s/NR}'
+echo -n 'GraphBIG,BFS,Time,'
+grep -A 7 "Benchmark: BFS" "$FN" | awk -v T=1 '/time/{if(T%2==0){print $3} T++}' | awk '{s+=$1}END{print s/NR}'
+echo -n 'GraphBIG,SSSP,File reading,'
+grep -A 7 'Benchmark: sssp' "$FN" | awk -v T=0 '/time/{if(T%2==0){print $3} T++}' | awk '{s+=$1}END{print s/NR}'
+echo -n 'GraphBIG,SSSP,Time,'
+grep -A 7 'Benchmark: sssp' "$FN" | awk -v T=1 '/time/{if(T%2==0){print $3} T++}' | awk '{s+=$1}END{print s/NR}'
+# XXX: PageRank only does one trial
+echo -n 'GraphBIG,PageRank,File reading,'
+grep -A 7 "Degree Centrality" "$FN" | awk -v T=0 '/time/{if(T%2==0){print $3} T++}' | awk '{s+=$1}END{print s/NR}'
+echo -n 'GraphBIG,PageRank,Iterations,'
+grep 'iteration #' "$FN" | awk '{print $4}'
+echo -n 'GraphBIG,PageRank,Time,'
+grep -A 11 "Degree Centrality" "$FN" | awk -v T=1 '/time/{if(T%2==0){print $3} T++}' | awk '{s+=$1}END{print s/NR}'
+
+
+# GraphMat
+echo -n 'GraphMat,BFS,File reading,'
[…]

### 2017-01-26 papers/gabb17/gabb17-pollard.tex
Updated gabb17 paper.

@@
 \begin{abstract}
-This is a survey of existing graph analytics frameworks and an analysis of their performance.
+In this paper we analyze the performance of four parallel graph processing systems to determine their performance and scalability using synthetic and real-world datasets.
+We perform analysis for three algorithms: breadth first serach, single source shortest paths, and PageRank.
+This paper examines previously overlooked aspects of parallel programming performance such as file I/O in addition to a more detailed performance analysis.
 \end{abstract}
@@
 
-[Talk about WHY we care about graph processing, and parallel in particular. Talk about what parallel graph processing is.]
-
 Our research is motivated by the current state of parallel graph processing.
 The most comprehensive survey, released in 2014, identified and taxonomized over 80 different parallel graph processing systems \cite{Doekemeijer:2015:GPFSurvey}.
-These systems operate with a wide range of parallelism paradigms such as GPU [Medusa], shared memory [pretty much everything], a combination of CPU and GPU \cite{Gharaibeh:2012:Totem}, distributed database querying, \cite{Rodriguez:2015:Gremlin}, distributed filesystem based approaches \cite{Xin:2013:GraphX}, distributed memory with MPI [cite], domain-specific languages [Green Marl], as well as novel communication schemes [activemessages, sockets].
+These systems operate with a wide range of parallelism paradigms and target architectures such as GPU \cite{Zhong:2014:Medusa}, \cite{Kang:2009:Pegasus}, shared memory [pretty much everything], a combination of CPU and GPU \cite{Gharaibeh:2012:Totem}, distributed database querying, \cite{Rodriguez:2015:Gremlin}, distributed filesystem based approaches \cite{Xin:2013:GraphX}, distributed memory with MPI \cite{Hong:2015:PGX}, domain-specific languages \cite{Hong:2012:GreenMarl}, as well as novel communication schemes \cite{Edmonds:2013:ActiveMessages}.
 
 Since 2014, the problem has compounded with the addition of even more proprietary and open source projects such as \cite{Cheramangalath:2015:Falcon}, \cite{Perez:2015:Ringo} [name some more].
-At the outset, this plethora of choices makes the question, ``which system is the best for my problem?'' incredibly difficult.
-There has even been a propagation of so-called ``reference implementations'' which provide implementations of the most common graph algorithms [GAP, GraphBIG, Galois].
+At the outset, this plethora of choices makes the question, ``which system is the best for my problem?'' daunting.
+There has even been a propagation of so-called ``reference implementations'' which  implement the most common graph algorithms [cite GAP, GraphBIG, Galois].
 Thus, even selecting a standard and a benchmark over which to compare various implementations is nontrivial.
@@
 
-[This paper aims to simplify the selection process for graph processing.]
-
-Another issue with the parallel graph processing library designers is the lack of comprehensive comparison.
-One possible reason for this is the considerable effort of getting each library and package to work, satisfying dependencies, and ensuring the data is in the correct format for each system.
+Another issue among parallel graph processing systems is the lack of comprehensive comparisons.
+One possible reason for this is the considerable effort involved in getting each system to run: satisfying dependencies and ensuring correctly formatted data the data are nontrivial for most tasks.
 Beyond this, each system may have a different method of measuring performance.
 Thus, one of the contributions of this paper is to provide a ``level playing field'' for each graph processing system.
-Graphalytics \cite{Capota:2015:Graphalytics} attempts to remedy this but has not seen widespread adoption---only eight platforms are supported\footnote{According to \url{https://github.com/tudelft-atlarge}}.
+Graphalytics \cite{Capota:2015:Graphalytics} attempts to remedy this but
+[state the limitations of graphalytics]
+%has not seen widespread adoption---only eight platforms are supported\footnote{According to \url{https://github.com/tudelft-atlarge}}.
 [TODO: Say more about why this is justified instead of just using Graphalytics]
@@
 \section{Algorithms}
-The canonical performance leaderboard for paralell graph processing is the Graph500 \cite{Murphy:2010:Graph500}.
-The advantage of the Graph500 is it provides standardized measurement requirements and dataset generation.
-The primary drawback with using reference implementations for the Graph500 is the standard only supports a single algorithm: breadth first search (BFS).
+The canonical performance leaderboard for parallel graph processing is the Graph500 \cite{Murphy:2010:Graph500}.
+The advantage of the Graph500 is it provides standardized measurement specifications and dataset generation.
+The primary drawback with the Graph500 is it measures a single algorithm: breadth first search (BFS).
 
@@
 The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}.
-Some clarifications are in order: for the first kernel, the time to read in the graph from disk is not considered.
+Some clarifications are in order: for the first kernel, the time to read the graph from disk is not considered.
 Furthermore, these data structures need only be created once and BFS is performed on 64 roots.
@@
 
-One straightforward extension is computing the Single-Source Shortest Paths algorithm (SSSP). We use the same graph and the same source vertices as in BFS.
+One straightforward extension is computing the Single-Source Shortest Paths algorithm (SSSP).
+We use the same graph and the same source vertices as in BFS.
+We use SSSP and PageRank in this paper because of their simplicity and ubiquity.
 
@@
 This report explores a small sample of the existing graph processing platforms with a focus on the so-called ``reference implementations:'' the Graph500\footnote{We used version 2.1.4 from \url{http://graph500.org/referencecode}.}, GAP \cite{Beamer:2015:GAPBench}, and GraphBIG \cite{Nai:2015:Graphbig}.
+GraphMat is also used because it has been cited as the most performant [cite].
+Our target is shared memory CPU processing.
 
-Parallel Boost Graph Library \cite{Gregor:2005:PBGL} and PowerGraph \cite{Gonzalez:2012:Powergraph} implementations are also provided because of their popularity and the availability of high quality reference implementations.
+Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} %and PowerGraph \cite{Gonzalez:2012:Powergraph} 
+are not considered here because they are a PITA to get working and the authors never responded to my emails.
 
@@
 
-\begin{table}[htb]
-	\centering
-	\begin{tabular}{l|c|c|c}
-		& BFS & SSSP & PR \\ \hline
-		Graph500 & & & \\ \hline
-		PBGL  & & & \\ \hline
-		GAP  & & & \\ \hline
-		GraphBIG  & & & \\ \hline
-		Galois  & & & \\ \hline
-		PowerGraph  & & &
-	\end{tabular}
-	\caption{Performance. Note that implementations of every algorithm are not availabe for every platform.}
-	\label{tab:reportcard}
-\end{table}
-
-Graphalytics without the use of the Granula plugin produces performance measurement in two forms: runtime in seconds and traversed edges per second.
+%\begin{table}[htb]
+%	\centering
+%	\begin{tabular}{l|c|c|c}
+%		& BFS & SSSP & PR \\ \hline
+%		Graph500 & & & \\ \hline
+%		PBGL  & & & \\ \hline
+%		GAP  & & & \\ \hline
+%		GraphBIG  & & & \\ \hline
+%		Galois  & & & \\ \hline
+%		PowerGraph  & & &
+%	\end{tabular}
+%	\caption{Performance. Note that implementations of every algorithm are not availabe for every platform.}
+%	\label{tab:reportcard}
+%\end{table}
 
@@
 \end{table}
+
+\begin{table}[!htb]
+	\centering
+	\begin{tabular}{l|c|c|c}
+		System & Load Graph & Make Data Structure & Run BFS \\ \hline
+		Graph500 & 0.3474 & 0.3971 & 0.003380 \\
+		GraphBIG & 37.22  & N/A    & 0.1528 \\
+		GraphMat & 0.1511 & 1.101  & 0.1031 \\
+		GAP      & 2.351  & 0.1935 & 0.001393 \\
+	\end{tabular}
+	\caption{Times are in seconds.
+The Graph500 generates the graph instead of loading it into a file.
+GraphBIG builds the graph and reads in the file simultaneously.
+These results were averaged across 64 roots.
+[TODO: Why does the first root often have degree 0?]}
+	\label{tab:fileload}
+\end{table}
+
+[Compare lines of code for implementation]
 

### 2017-01-26 misc/gen-datasets.sh
Updated comments and documentation.
Now run-experiment will point you in the right direction for
all the repositories. README now talks about the scripts in misc.

@@ -1,6 +1,6 @@
 #!/bin/bash
 # Generate an unweighted, undirected Kronecker graph in the file formats for
-# graph500, GraphBIG, and GAP [add other systems when ready]
+# graph500, GraphMat, GraphBIG, and GAP
 # for BFS [add other algorithms when ready]
 # This only needs to be done once per scale.
 # For example, with S=20 the structure will look like
@@ -16,8 +16,8 @@
 # NOTE: These must be set before you run this script
 # It is also assumed these are all built.
 # The building instructions are available in the respective repositories.
-GRAPH500DIR=
-GRAPHMATDIR=
+GRAPH500DIR=$HOME/graph500
+GRAPHMATDIR=$HOME/graphalytics/GraphMat
 DDIR= # Where to save the dataset
 
 USAGE="gen-datasets.sh <scale>"
@@ -39,10 +39,6 @@ awk 'BEGIN{print "SRC,DEST"} {printf "%d,%d\n", $1, $2}' "$DDIR/$BASE_FN.el" > "
 echo ID > "$DDIR/$BASE_FN/vertex.csv"
 cat "$DDIR/$BASE_FN.el" | tr ' ' '\n' | sort -n | uniq >> "$DDIR/$BASE_FN/vertex.csv"
 
-# GraphMat can be installed using the following commands on arya
-#module load intel/17
-#cd ~/graphalytics/GraphMat
-#make
 # Convert to GraphMat format
 # GraphMat requires edge weights---Just make them all 1 for the .1wel format
 # TODO: See what happens when you remove selfloops and duplicated edges.

### 2017-01-26 papers/gabb17/gabb17-pollard.tex

@@
 			string type, % Makes the .style={string type} redundant
-			columns={Name,Type,HPC,Parallelism,Target,FOSS,Source,Notes},
+			columns={Name,Type,Parallelism,Target,FOSS,Source,Notes},
 			every head row/.style={after row=\midrule},
@@
 			columns/Type/.style={string type},
-			columns/HPC/.style={string type},
+			% columns/HPC/.style={string type},
 			columns/Parallelism/.style={string type},
@@
 %\bibliographystyle{acm}
-\bibliographystyle{IEEEtran}
+\bibliographystyle{IEEEtranS}
 \bibliography{../drp}

### 2017-01-27 misc/parse-output.sh
Parses all trials not just an average.
Added the same number of trials for PageRank too.

@@ -19,67 +19,48 @@ grep "mean_time" "$FN" | awk '{print $2}'
 # GraphBIG
 echo -n 'GraphBIG,BFS,File reading,'
 grep -A 7 "Benchmark: BFS" "$FN" | awk -v T=0 '/time/{if(T%2==0){print $3} T++}' | awk '{s+=$1}END{print s/NR}'
-echo -n 'GraphBIG,BFS,Time,'
-grep -A 7 "Benchmark: BFS" "$FN" | awk -v T=1 '/time/{if(T%2==0){print $3} T++}' | awk '{s+=$1}END{print s/NR}'
+grep -A 7 "Benchmark: BFS" "$FN" | awk -v T=1 '/time/{if(T%2==0){print "GraphBIG,BFS,Time," $3} T++}'
 echo -n 'GraphBIG,SSSP,File reading,'
 grep -A 7 'Benchmark: sssp' "$FN" | awk -v T=0 '/time/{if(T%2==0){print $3} T++}' | awk '{s+=$1}END{print s/NR}'
-echo -n 'GraphBIG,SSSP,Time,'
-grep -A 7 'Benchmark: sssp' "$FN" | awk -v T=1 '/time/{if(T%2==0){print $3} T++}' | awk '{s+=$1}END{print s/NR}'
+grep -A 7 'Benchmark: sssp' "$FN" | awk -v T=1 '/time/{if(T%2==0){print "GraphBIG,SSSP,Time," $3} T++}'
 # XXX: PageRank only does one trial
 echo -n 'GraphBIG,PageRank,File reading,'
 grep -A 7 "Degree Centrality" "$FN" | awk -v T=0 '/time/{if(T%2==0){print $3} T++}' | awk '{s+=$1}END{print s/NR}'
-echo -n 'GraphBIG,PageRank,Iterations,'
-grep 'iteration #' "$FN" | awk '{print $4}'
-echo -n 'GraphBIG,PageRank,Time,'
-grep -A 11 "Degree Centrality" "$FN" | awk -v T=1 '/time/{if(T%2==0){print $3} T++}' | awk '{s+=$1}END{print s/NR}'
-
+grep 'iteration #' "$FN" | awk '{print "GraphBIG,PageRank,Iterations," $4}'
+grep -A 11 "Degree Centrality" "$FN" | awk -v T=1 '/time/{if(T%2==0){print "GraphBIG,PageRank,Time," $3} T++}'
 
 # GraphMat
 echo -n 'GraphMat,BFS,File reading,'
[…]

### 2017-01-27 misc/run-experiment.sh

@@ -11,18 +11,18 @@
 # It is assumed these are all built and the gen-dataset.sh script has been run.
 # The building instructions are available in the respective repositories.
 # Here, edge factor (number of edges per vertex) is the default of 16.
-S= # Scale. 2^S = Number of vertices.
+S=20 # Scale. 2^S = Number of vertices.
 DDIR= # Dataset directory
-GAPDIR=/home/users/spollard/gap/gapbs
-GRAPHBIGDIR=/home/users/spollard/gap/graphBIG
-GRAPH500DIR=/home/users/spollard/graph500
-GRAPHMATDIR=/home/users/spollard/graphalytics/GraphMat
+GAPDIR=
+GRAPHBIGDIR=
+GRAPH500DIR=
+GRAPHMATDIR=
 
 # PageRank is usually represented as a 32-bit float,
 # so ~6e-8*nvertices is the minimum absolute error detectable
 # We set alpha = 0.15 in the respective source codes.
 # NOTE: GraphMat doesn't seem to compute iterations in the same way.
-MAXITER=20 # Maximum iterations for PageRank
+MAXITER=50 # Maximum iterations for PageRank
 TOL=0.00000006
 if [ -z $S -o -z "$DDIR" ]; then
 	echo 'Please set S and the parameters of the form *DIR in run-experiment.sh'
@@ -108,17 +108,24 @@ done <"$DDIR/kron-${S}-roots.1v"
 
 # Run GAP PageRank
 # error = sum(|newPR - oldPR|)
-"$GAPDIR"/pr -f "$DDIR/kron-${S}.el" -i $MAXITER -t $TOL
+# Note: ROOT is a dummy variable for pagerank; we just ensure the same trials are computed.
+while read ROOT; do
+	"$GAPDIR"/pr -f "$DDIR/kron-${S}.el" -i $MAXITER -t $TOL
+done <"$DDIR/kron-${S}-roots.1v"
 
 # Run GraphBIG PageRank
 # GraphBIG has --quad = sqrt(sum((newPR - oldPR)^2))
 # GraphBIG error has been modified to now be sum(|newPR - oldPR|)
-"$GRAPHBIGDIR/benchmark/bench_pageRank/pagerank" --dataset "$DDIR/kron-${S}" --maxiter $MAXITER --quad $TOL
+while read ROOT; do
[…]

### 2017-01-27 papers/gabb17/plot_data.R
Added a rudimentary draft of plot_data.R.

@@ -1,0 +1,26 @@
+# This is some bad bad bad R code.
+# Might want to double check it.
+scale <- 20
+filename <- paste0("parsed",scale,".csv")
+x <- read.csv(filename, header = FALSE)
+bfs_time <- subset(x, x[[2]] == "BFS" & x[[3]] == "Time", c(V1,V4))
+colnames(bfs_time) <- c("System","Time")
+# Boxplot log = "y" may be more useful.
+# TODO: Figure out why the times are so low sometimes!
+boxplot(Time~System, bfs_time, ylab = "Time (seconds)")
+
+# File reading not as useful
+file_reading <- subset(x, x[[3]] == "File reading", select = c(V1,V4))
+file_reading_times <- c(0,0,0)
+file_reading_times[1] <- mean(file_reading[[2]][1:3])
+file_reading_times[2] <- mean(file_reading[[2]][4:6])
+file_reading_times[3] <- mean(file_reading[[2]][7:9])
+
+ds_const <- subset(x, x[[3]] == "BFS" & x[[3]] == "Data structure build",
+		select = c(V1,V4))
+
+# This doesn't work well
+#barplot(file_reading_times,
+#		names.arg = as.character(unique(file_reading[[1]])),
+#		ylab="Time (seconds)",
+#		log="y")

### 2017-01-28 papers/gabb17/gabb17-pollard.tex
Added more plots and plot-generation R code.
Also updated some littel bits fo the writeup and some more TODOS.
Makefile now points to the graphics directory but I didn't add any
graphics to the repo yet since they most certainly will change.
Removed execute permission on IEEtran.cls.

@@
 \section{Algorithms}
+
 The canonical performance leaderboard for parallel graph processing is the Graph500 \cite{Murphy:2010:Graph500}.
@@
 
-This report attempts to add similar rigor to other graph algorithms.
-This is done by borrowing heavily from the Graph500 specification; we use a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19,
-C = 0.19,$ and $D = 1-(A+B+C) = 0.05$.
-The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}.
+This report attempts to add similar rigor to other graph algorithms by borrowing heavily from the Graph500 specification.
+The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the actual BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}.
 Some clarifications are in order: for the first kernel, the time to read the graph from disk is not considered.
 Furthermore, these data structures need only be created once and BFS is performed on 64 roots.
-The second kernel creates a tree with distances from the root, but need only label the existing data structure (and not output the entire graph, for example).
 
@@
 This report explores a small sample of the existing graph processing platforms with a focus on the so-called ``reference implementations:'' the Graph500\footnote{We used version 2.1.4 from \url{http://graph500.org/referencecode}.}, GAP \cite{Beamer:2015:GAPBench}, and GraphBIG \cite{Nai:2015:Graphbig}.
-GraphMat is also used because it has been cited as the most performant [cite].
+GraphMat is also included because it has been cited as the most performant \cite{Sundaram:2015:GraphMat}.
 Our target is shared memory CPU processing.
 
-Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} %and PowerGraph \cite{Gonzalez:2012:Powergraph} 
-are not considered here because they are a PITA to get working and the authors never responded to my emails.
+% Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} %and PowerGraph \cite{Gonzalez:2012:Powergraph} 
+% are not considered here because they are a PITA to get working and the authors never responded to my emails.
 
-\section{Machine Specifications}
-Table~\ref{tab:specs} shows the specifications of the research computer (named Arya).
+\section{Experimental Setup}
+\subsection{Machine Specifications}
+Table~\ref{tab:specs} shows the specifications of the research machine.
 
@@
 
-\section{Performance}
+\subsection{Datasets}
+We use a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19, C = 0.19,$ and $D = 1-(A+B+C) = 0.05$.
+
+\section{Perf. Anal.}
+
+\begin{figure}[!htb]
+	\includegraphics[width=\columnwidth]{/Users/spollard/Desktop/modern-aRt.pdf}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\columnwidth]{graphics/bfs_time.pdf}
+	\vspace{-18pt}
+	\caption{The $y$-axis is logarithmic. [TODO: Replace this with the data with non-zero degree vertices]}
+	\label{fig:bfs-time}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\columnwidth]{graphics/bfs_dsc.pdf}
+	\vspace{-18pt}
+	\caption{The $y$-axis is logarithmic. GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
+	\label{fig:bfs-dsc}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\columnwidth]{graphics/sssp_time.pdf}
+	\vspace{-18pt}
+	\caption{The $y$-axis is logarithmic.}
+	\label{fig:sssp-time}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\columnwidth]{graphics/pr_time.pdf}
+	\vspace{-18pt}
+	\caption{The $y$-axis is logarithmic. GraphMat continues to run until none of the vertices' ranks change. [There is only one data point]}
+	\label{fig:pr-time}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\columnwidth]{graphics/pr_iters.pdf}
+	\vspace{-18pt}
+	\caption{The $y$-axis \emph{not} logarithmic.
+GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs unitl none of the vertices' ranks change.
+[There is only one data point]}
+	\label{fig:pr-iters}
+\end{figure}
 
@@
 	\begin{tabular}{l|c|c|c}
-		System & Load Graph & Make Data Structure & Run BFS \\ \hline
+		System & Load Graph & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular} & Run BFS \\ \hline
 		Graph500 & 0.3474 & 0.3971 & 0.003380 \\
@@
 	\end{tabular}
-	\caption{Times are in seconds.
+	\caption{The above table shows times for $2^{20}$ verties and the times are in seconds.
 The Graph500 generates the graph instead of loading it into a file.
 GraphBIG builds the graph and reads in the file simultaneously.
-These results were averaged across 64 roots.
-[TODO: Why does the first root often have degree 0?]}
+These results were averaged across 64 roots.}
 	\label{tab:fileload}

### 2017-01-28 papers/gabb17/plot_data.R

@@ -1,13 +1,76 @@
 # This is some bad bad bad R code.
 # Might want to double check it.
 scale <- 20
+bpc <- "chartreuse"
 filename <- paste0("parsed",scale,".csv")
 x <- read.csv(filename, header = FALSE)
-bfs_time <- subset(x, x[[2]] == "BFS" & x[[3]] == "Time", c(V1,V4))
-colnames(bfs_time) <- c("System","Time")
-# Boxplot log = "y" may be more useful.
-# TODO: Figure out why the times are so low sometimes!
-boxplot(Time~System, bfs_time, ylab = "Time (seconds)")
+colnames(x) <- c("Sys","Algo","Metric","Time")
+
+bfs_time <- subset(x, x$Algo == "BFS" & x$Metric == "Time",
+		c("Sys","Time"))
+bfs_dsc <- subset(x, x$Algo == "BFS" & x$Metric == "Data structure build",
+		c("Sys","Time"))
+bfs_dsc$Sys <- factor(bfs_dsc$Sys)
+
+sssp_time <- subset(x, x$Algo == "SSSP" & x$Metric == "Time",
+		c("Sys","Time"))
+sssp_dsc <- subset(x, x$Algo == "SSSP" & x$Metric == "Data structure build",
+		c("Sys","Time"))
+sssp_time$Sys <- factor(sssp_time$Sys)
+sssp_dsc$Sys <- factor(sssp_dsc$Sys)
+
+pr_time <- subset(x, x$Algo == "PageRank" & x$Metric == "Time",
+		c("Sys","Time"))
+pr_dsc <- subset(x, x$Algo == "PageRank" & x$Metric == "Data structure build",
+		c("Sys","Time"))
+pr_iters <- subset(x, x$Algo == "PageRank" & x$Metric == "Iterations",
+		c("Sys","Time"))
+pr_time$Sys <- factor(pr_time$Sys)
+pr_dsc$Sys <- factor(pr_dsc$Sys)
+
+pdf("graphics/bfs_time.pdf")
+boxplot(Time~Sys, bfs_time, ylab = "Time (seconds)",
+		main = "BFS Time", log = "y", col=bpc)
+dev.off()
+
+pdf("graphics/bfs_dsc.pdf")
+boxplot(Time~Sys, bfs_dsc, ylab = "Time (seconds)",
+		main = "BFS Data Structure Construction", col=bpc, log = "y")
+dev.off()
+
+pdf("graphics/sssp_time.pdf")
[…]

### 2017-01-28 misc/run-experiment.sh
Added thread numbers and standardized # of trials.
Now, GraphBIG runs on the same number of threads,
the script sets OMP_NUM_THREADS, and GAP will only do
1 trial, because previously the was lower because it did 16.

@@ -11,7 +11,7 @@
 # It is assumed these are all built and the gen-dataset.sh script has been run.
 # The building instructions are available in the respective repositories.
 # Here, edge factor (number of edges per vertex) is the default of 16.
-S=20 # Scale. 2^S = Number of vertices.
+S= # Scale. 2^S = Number of vertices.
 DDIR= # Dataset directory
 GAPDIR=
 GRAPHBIGDIR=
@@ -28,6 +28,11 @@ if [ -z $S -o -z "$DDIR" ]; then
 	echo 'Please set S and the parameters of the form *DIR in run-experiment.sh'
 	exit 2
 fi
+if [ "$1" -lt 1 ]; then
+	echo "usage: run-experiment <num_threads>"
+	exit 2
+fi
+export OMP_NUM_THREADS=$1
 
 # Load all the modules here
 module load intel/17
@@ -74,14 +79,15 @@ module load intel/17
 
 # Run for GAP BFS
 # It would be nice if you could read in a file for the roots
+# Just do one trial to be the same as the rest of the experiments
 while read ROOT; do
-	"$GAPDIR"/bfs -r $ROOT -f "$DDIR/kron-${S}.el"
+	"$GAPDIR"/bfs -r $ROOT -f "$DDIR/kron-${S}.el" -n 1
 done <"$DDIR/kron-${S}-roots.v"
 
 # Run the GraphBIG BFS
 # For this, one needs a vertex.csv file and and an edge.csv.
 while read ROOT; do
-	"$GRAPHBIGDIR/benchmark/bench_BFS/bfs" --dataset "$DDIR/kron-${S}" --root $ROOT
+	"$GRAPHBIGDIR/benchmark/bench_BFS/bfs" --dataset "$DDIR/kron-${S}" --root $ROOT --threadnum $OMP_NUM_THREADS
 done <"$DDIR/kron-${S}-roots.v"
 
 # Run the GraphMat BFS
@@ -92,12 +98,12 @@ done <"$DDIR/kron-${S}-roots.1v"
 
 # Run the GAP SSSP for each root
 while read ROOT; do
-	"$GAPDIR"/sssp -r $ROOT -f "$DDIR/kron-${S}.el"
+	"$GAPDIR"/sssp -r $ROOT -f "$DDIR/kron-${S}.el" -n 1
[…]

### 2017-01-30 misc/parse-output.sh
Fixed consistency with graph500 naming and serial.
Now instead of omp-csr we use seq-csr when OMP_NUM_THREADS=1.

@@ -9,10 +9,10 @@ if [ -z "$FN" ]; then
 fi
 
 # Graph500
-echo -n 'Graph500,BFS,graph construction,'
-echo $(grep "construction_time" "$FN" | awk '{print $2}') 
 echo -n 'Graph500,BFS,graph generation,'
 echo $(grep "generation_time" "$FN" | awk '{print $2}')
+echo -n 'Graph500,BFS,Data structure build,'
+echo $(grep "construction_time" "$FN" | awk '{print $2}') 
 awk '/bfs_time\[/{print "Graph500,BFS,Time," $2}' "$FN"
 
 # GraphBIG
@@ -58,7 +58,7 @@ awk -v NRT=$NRT '/Build Time:/{i++; if(i>NRT && i<=2*NRT)print "GAP,SSSP,Data st
 awk -v NRT=$NRT '/Average Time:/{i++; if(i>NRT && i<=2*NRT)print "GAP,SSSP,Time," $3}' "$FN"
 echo -n 'GAP,PageRank,File reading,'
 awk -v NRT=$NRT '/Read Time:/{i++; if(i>2*NRT)print $3}' "$FN" | awk '{s+=$1}END{print s/NR}'
-awk '/Build Time:/{i++; if(i>128)print "GAP,PageRank,Data structure build," $3}' "$FN"
+awk -v NRT=$NRT '/Build Time:/{i++; if(i>2*NRT)print "GAP,PageRank,Data structure build," $3}' "$FN"
 grep -B 2 'Average Time:' "$FN" | awk '/^\s*[0-9]+/{print "GAP,PageRank,Iterations," $1}'
-awk '/Average Time:/{i++; if(i>128)print "GAP,PageRank,Time," $3}' "$FN"
+awk -v NRT=$NRT '/Average Time:/{i++; if(i>2*NRT)print "GAP,PageRank,Time," $3}' "$FN"
 

### 2017-01-30 misc/run-experiment.sh

@@ -76,7 +76,11 @@ module load intel/17
 # Run the Graph500 BFS: OpenMP
 # This isn't working, so just regenerate the data
 #omp-csr/omp-csr -s $S -o "$DDIR/kron-${S}.graph500" -r "$DDIR/kron-${S}.roots"
-"$GRAPH500DIR/omp-csr/omp-csr" -s $S 
+if [ "$OMP_NUM_THREADS" -gt 1 ]; then
+	"$GRAPH500DIR/omp-csr/omp-csr" -s $S 
+else
+	"$GRAPH500DIR/seq-csr/seq-csr" -s $S 
+fi
 
 # Run for GAP BFS
 # It would be nice if you could read in a file for the roots

### 2017-01-30 papers/gabb17/gabb17-pollard.tex
Added graphics to paper and cleaned up tables.
It won't be ready for GABB17 unless I decide I don't need to sleep
for the next 40 hours. Oh well. I tried my best.

Really, what needs to happen next is I need to dust off those old
shell scripts like get-graphalytics, get graphalytics running
again on a wider range of datasets, and then get PowerGraph
working in my own framework and get GraphMat running in
the Graphalytics framework. That way it can be a more interesting
comparison and it can actually show some of the benefits and
limitations of both approaches. Another nice-to-have would be
an

@@
 
-\title{A Comparison of Parallel Graph Processing Platforms}
+\title{A Comparison of Parallel Graph Processing Benchmarks}
 \author{
-	\IEEEauthorblockN{Samuel Pollard}
+	\IEEEauthorblockN{Sam D. Pollard}
 	\IEEEauthorblockA{Department of Computer and Information Science \\
@@
 Our research is motivated by the current state of parallel graph processing.
-The most comprehensive survey, released in 2014, identified and taxonomized over 80 different parallel graph processing systems \cite{Doekemeijer:2015:GPFSurvey}.
-These systems operate with a wide range of parallelism paradigms and target architectures such as GPU \cite{Zhong:2014:Medusa}, \cite{Kang:2009:Pegasus}, shared memory [pretty much everything], a combination of CPU and GPU \cite{Gharaibeh:2012:Totem}, distributed database querying, \cite{Rodriguez:2015:Gremlin}, distributed filesystem based approaches \cite{Xin:2013:GraphX}, distributed memory with MPI \cite{Hong:2015:PGX}, domain-specific languages \cite{Hong:2012:GreenMarl}, as well as novel communication schemes \cite{Edmonds:2013:ActiveMessages}.
+The most comprehensive survey, released in 2014, identified and taxonomized over 80 different parallel graph processing systems without including domain specific languages \cite{Doekemeijer:2015:GPFSurvey}.
+The systems described operate with a wide range of parallelism paradigms and target architectures such as GPU \cite{Zhong:2014:Medusa, Kang:2009:Pegasus}, shared memory CPU \cite{Shun:2013:Ligra, kyrola:2012:Graphchi, Nguyen:2013:Galois}, a combination of CPU and GPU \cite{Gharaibeh:2012:Totem}, distributed filesystem based approaches \cite{Xin:2013:GraphX}, and distributed memory with MPI \cite{Gregor:2005:PBGL}.
 
-Since 2014, the problem has compounded with the addition of even more proprietary and open source projects such as \cite{Cheramangalath:2015:Falcon}, \cite{Perez:2015:Ringo} [name some more].
+Beyond the systems described by Doekemeijer and Varbanescu, the problem has compounded with the addition of even more proprietary and open source projects such as \cite{Cheramangalath:2015:Falcon}, \cite{Perez:2015:Ringo}, distributed memory approaches such as \cite{Hong:2015:PGX}.
+domain-specific languages \cite{Hong:2012:GreenMarl}, distributed database querying, \cite{Rodriguez:2015:Gremlin}, as well as novel communication schemes \cite{Edmonds:2013:ActiveMessages}.
 At the outset, this plethora of choices makes the question, ``which system is the best for my problem?'' daunting.
-There has even been a propagation of so-called ``reference implementations'' which  implement the most common graph algorithms [cite GAP, GraphBIG, Galois].
+
+In addition to libraries with associated APIs there has also been a propagation of ``reference implementations'' which implement the most common graph algorithms such as \cite{Beamer:2015:GAPBench, Nai:2015:Graphbig}.
 Thus, even selecting a standard and a benchmark over which to compare various implementations is nontrivial.
@@
 Another issue among parallel graph processing systems is the lack of comprehensive comparisons.
-One possible reason for this is the considerable effort involved in getting each system to run: satisfying dependencies and ensuring correctly formatted data the data are nontrivial for most tasks.
-Beyond this, each system may have a different method of measuring performance.
-Thus, one of the contributions of this paper is to provide a ``level playing field'' for each graph processing system.
-Graphalytics \cite{Capota:2015:Graphalytics} attempts to remedy this but
-[state the limitations of graphalytics]
-%has not seen widespread adoption---only eight platforms are supported\footnote{According to \url{https://github.com/tudelft-atlarge}}.
-[TODO: Say more about why this is justified instead of just using Graphalytics]
+One possible reason is the considerable effort involved in getting each system to run: satisfying dependencies and ensuring data are correctly formatted are generally nontrivial tasks.
+Beyond this, there are optimizations for each system which are not exploited in a na\"ive implementation.
+Addressing these issues helps provide an environment in which no graph processing system has an advantage.
 
-\section{Algorithms}
+Graphalytics \cite{Capota:2015:Graphalytics} also attempts to remedy these problems.
+However, in order for Graphalytics to analyze a graph processing system---Graphalytics calls these platforms---a programmer must provide Java methods to call the particular implementations and initialization shell scripts to satisfy dependencies and set environment variables.
+These all require some knowledge of the inner workings of each system in addition to familiarity with the Graphalytics API.
+
+With a plugin to Graphalytics called Granula \cite{Ngai:2015:Granula} one can explicitly specify a performance model to analyze specific execution behavior such as the amount of communication or runtime of particular kernels of execution.
+As with Graphalytics, the initial development effort is high but Granula allows for multiple experiments to be run easily.
+
+One drawback of Granula is it requires knowledge of the source code and execution model which is not always available.
+Furthermore, this requires a high level of expertise with the given system and with Granula\footnote{An example of Granula can be seen at \url{https://github.com/tudelft-atlarge/graphalytics-platforms-graphx/tree/master/granula-model-graphx}}.
+
+To address these concerns we analyze the performance of several graph processing benchmarks to facilitate the selection among such systems.
+Our analysis is done without modification of the existing source code.
+
+\section{Experimental Setup}
+
+\subsection{Graph Processing Systems}
+
+This report explores four shared memory parallel graph processing platforms.
+The first three are so-called ``reference implementations'' while the fourth is included because it has been cited as highly performant \cite{Sundaram:2015:GraphMat}.
+Our target is shared memory CPU processing.
+Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} are not considered here because the authors do not provide reference implementations for all algorithms.
+The systems are:
+
+\begin{enumerate}
+	\item The Graph500\footnote{We used the most recent version from \url{https://github.com/graph500/graph500}, most similar to release 2.1.4.} \cite{Murphy:2010:Graph500}
+	\item The Graph Algorithm Platform (GAP) Benchmark Suite \cite{Beamer:2015:GAPBench}
+	\item GraphBIG \cite{Nai:2015:Graphbig}
+	\item GraphMat \cite{Sundaram:2015:GraphMat}
+\end{enumerate}
+
+\subsection{Algorithms}
+
+We consider three algorithms:
+
+\begin{enumerate}
+	\item Breadth First Search (BFS)
+	\item Single Source Shortest Paths (SSSP)
+	\item PageRank
+\end{enumerate}
 
@@
 The advantage of the Graph500 is it provides standardized measurement specifications and dataset generation.
-The primary drawback with the Graph500 is it measures a single algorithm: breadth first search (BFS).
+The primary drawback with the Graph500 is it measures a single algorithm: breadth first search.
 
-This report attempts to add similar rigor to other graph algorithms by borrowing heavily from the Graph500 specification.
+This report aims to add similar rigor to other graph algorithms by borrowing heavily from the Graph500 specification.
 The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the actual BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}.
-Some clarifications are in order: for the first kernel, the time to read the graph from disk is not considered.
-Furthermore, these data structures need only be created once and BFS is performed on 64 roots.
+We run the BFS using 32 random roots.
 
-One straightforward extension is computing the Single-Source Shortest Paths algorithm (SSSP).
-We use the same graph and the same source vertices as in BFS.
-We use SSSP and PageRank in this paper because of their simplicity and ubiquity.
+One straightforward extension to BFS and our second algorithm is the Single-Source Shortest Paths algorithm (SSSP). We use the same graph and the same source vertices as in BFS.
 
-\section{The Existing Systems}
-This report explores a small sample of the existing graph processing platforms with a focus on the so-called ``reference implementations:'' the Graph500\footnote{We used version 2.1.4 from \url{http://graph500.org/referencecode}.}, GAP \cite{Beamer:2015:GAPBench}, and GraphBIG \cite{Nai:2015:Graphbig}.
-GraphMat is also included because it has been cited as the most performant \cite{Sundaram:2015:GraphMat}.
-Our target is shared memory CPU processing.
+The third algorithm is PageRank.
+These three algorithms are used because of their popularity; most libraries provide reference implementations.
+All the experiments performed use the developer-provided implementations because we assume the vendor will provide a more performant implementation than us.
+While this limits the scope of the experiments it mitigates the bias inherent in any developer's programming skills.
 
-% Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} %and PowerGraph \cite{Gonzalez:2012:Powergraph} 
-% are not considered here because they are a PITA to get working and the authors never responded to my emails.
-
-\section{Experimental Setup}
 \subsection{Machine Specifications}
 Table~\ref{tab:specs} shows the specifications of the research machine.
+The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
+We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.
 
-\begin{table}[!htb]
+\begin{table}
 	\centering
 	% Keep in mind you can do this at the beginning: string replace={s1}{s2}
+	% For arya I deleted Max RAM Freq	2133MHz
 	\pgfplotstabletypeset[
@@
 	]{../report/specs.csv}
-	\caption{Machine specifications.
-The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
-We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.}
+	\caption{The operating system is GNU/Linux version 4.4.0-22.}
 	\label{tab:specs}
@@
 \subsection{Datasets}
-We use a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19, C = 0.19,$ and $D = 1-(A+B+C) = 0.05$.
+We use the Graph500 synthetic graph generator which creates a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19, C = 0.19,$ and $D = 1-(A+B+C) = 0.05$.
 
-\section{Perf. Anal.}
+The Graphalytics results in Table~\ref{tab:graphalytics} were performed on the Dota-League dataset.
+This dataset contains 61,670 vertices and 50,870,313 edges.
+This dataset is sourced from the Game Trace Archive\cite{Guo:2012:GTA} and modified for Graphalytics\footnote{This dataset is available at \url{https://atlarge.ewi.tudelft.nl/graphalytics/}.}.
 
-\begin{figure}[!htb]
-	\includegraphics[width=\columnwidth]{/Users/spollard/Desktop/modern-aRt.pdf}
-\end{figure}
+\section{Performance Analysis}
+
+In Table~\ref{tab:graphalytics} we show the results from running Graphalytics. For an explanation of each algorithm used, see \cite{Iosup:2016:Graphalyticstech}.
+
+From this we get only a broad overview of the two systems: GraphBIG and PowerGraph.
+In general, we see GraphBIG is more performant.
+However, results such as these are preliminary and do not show the complete picture.
+
+[How to make this a paper worth submitting: add in PowerGraph as a platform and see how it fares using our method compared to Graphalytics. Also run this on different datasets and compare results.]
+
+%\begin{table}[htb]
+%	\centering
+%	\begin{tabular}{l|c|c|c}
+%		& BFS & SSSP & PR \\ \hline
+%		Graph500 & & & \\ \hline
+%		PBGL  & & & \\ \hline
+%		GAP  & & & \\ \hline
+%		GraphBIG  & & & \\ \hline
+%		Galois  & & & \\ \hline
+%		PowerGraph  & & &
+%	\end{tabular}
+%	\caption{Performance. Note that implementations of every algorithm are not availabe for every platform.}
+%	\label{tab:reportcard}
+%\end{table}
+
+%Table~\ref{tab:perf} lists performance in milliseconds of runtime according to the graphalytics output.
+Graphalytics also outputs MTEPS or millions of traversed edges pers econd.
+However, the graphalytics version does not make sense in all cases: for example, computing the local clustering coefficient involves traversing each edge multiple times (proportional to the sparsity of the graph), while breadth first search (BFS) traverses each edge exactly once, and on na\:ive implementations single-source shortest paths (SSSP) may have $O(|E| + |V|^2)$ traversed edges.
+
+\begin{table}
+%	\begin{tabular}{l|r|r|}
+%	 & PowerGraph & OpenG \\ \hline
+%	BFS & 81.8 & 341 \\ \hline
+%	SSSP & 1.64 & 15.0 \\ \hline
+%	LCC & 54.6 & 142 \\ \hline
+%	\end{tabular}
+		\centering
+		\pgfplotstabletypeset[
+			col sep=comma,
+			string type, % Makes the .style={string type} redundant
+			columns={[index]0,GraphBIG,PowerGraph},
+			every head row/.style={after row=\midrule},
+			columns/0/.style={string type, column type={l|}, column name={}},
+			columns/openg/.style={string type, column type={r}},
+			columns/powergraph/.style={string type, column type={r}}
+		]{../report/runtime.csv}
+	\caption{Performance results are in milliseconds. LCC is local clustering coefficient, PR is PageRank, CDLP is community detection using label propagation, and WCC is weakly connected components}
+	\label{tab:graphalytics}
+\end{table}
+
+The following figures are not useful.
+They were are merely placeholders while further experiments are being run using the correct number of threads and roots.
+These issues are especially noticeable in GraphBIG.
+\begin{table}[!htb]
+	\centering
+	\begin{tabular}{l|c|c|c}
+		System & Load Graph & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular} & Run BFS \\ \hline
+		Graph500 & 0.3474 & 0.3971 & 0.003380 \\
+		GraphBIG & 37.22  & N/A    & 0.1528 \\
+		GraphMat & 0.1511 & 1.101  & 0.1031 \\
+		GAP      & 2.351  & 0.1935 & 0.001393 \\
+	\end{tabular}
+	\caption{The above table shows times for $2^{20}$ verties and the times are in seconds.
+The Graph500 generates the graph instead of loading it into a file.
+GraphBIG builds the graph and reads in the file simultaneously.
+These results were averaged across 32 roots.}
+	\label{tab:fileload}
+\end{table}
 
@@
 	\vspace{-18pt}
+	\caption{The $y$-axis \emph{not} logarithmic. GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs unitl none of the vertices' ranks change.}
+	\label{fig:pr-iters}
+\end{figure}
+[Note that for figure Fig.\ref{fig:pr-iters}, mention time per iteration]
+
+[Compare lines of code for implementation]
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\columnwidth]{graphics/bfs_ss.pdf}
+	\vspace{-18pt}
 	\caption{The $y$-axis \emph{not} logarithmic.
@@
 [There is only one data point]}
-	\label{fig:pr-iters}
+	\label{fig:bfs-ss}
 \end{figure}
 
-%\begin{table}[htb]
-%	\centering
-%	\begin{tabular}{l|c|c|c}
-%		& BFS & SSSP & PR \\ \hline
-%		Graph500 & & & \\ \hline
-%		PBGL  & & & \\ \hline
-%		GAP  & & & \\ \hline
-%		GraphBIG  & & & \\ \hline
-%		Galois  & & & \\ \hline
-%		PowerGraph  & & &
-%	\end{tabular}
-%	\caption{Performance. Note that implementations of every algorithm are not availabe for every platform.}
-%	\label{tab:reportcard}
-%\end{table}
-
-%Table~\ref{tab:perf} lists performance in milliseconds of runtime according to the graphalytics output.
-Graphalytics also outputs MTEPS or millions of traversed edges pers econd.
-However, the graphalytics version does not make sense in all cases: for example, computing the local clustering coefficient involves traversing each edge multiple times (proportional to the sparsity of the graph), while breadth first search (BFS) traverses each edge exactly once, and on na\:ive implementations single-source shortest paths (SSSP) may have $O(|E| + |V|^2)$ traversed edges.
-% TODO: Cite Comer algorithm book for Dijkstra's algorithm runtime.
-
-In Table~\ref{tab:perf}, BFS is breadth-first search, SSSP is single-source shortest paths, LCC is local clustering coefficient, PR is PageRank, CDLP is community detection using label propagation, and WCC is weakly connected components.
-For the algorithms used, see \cite{Iosup:2016:Graphalyticstech}.
-
-% TODO: This should be unnecessary once autogeneration is used.
-\begin{table}[!htb]
-	\centering
-%	\begin{tabular}{l|r|r|}
-%	 & PowerGraph & OpenG \\ \hline
-%	BFS & 81.8 & 341 \\ \hline
-%	SSSP & 1.64 & 15.0 \\ \hline
-%	LCC & 54.6 & 142 \\ \hline
-%	\end{tabular}
-
-		\centering
-		\pgfplotstabletypeset[
-			col sep=comma,
-			string type, % Makes the .style={string type} redundant
-			columns={[index]0,openg,powergraph},
-			every head row/.style={after row=\midrule},
-			columns/0/.style={string type, column type={l|}, column name={}},
-			columns/openg/.style={string type, column type={r}},
-			columns/powergraph/.style={string type, column type={r}}
-		]{../report/runtime.csv}
-	\caption{Performance Results for the \texttt{dota-league} dataset with 61,670 vertices and 50,870,313 edges.}
-	\label{tab:perf}
-\end{table}
-
-\begin{table}[!htb]
-	\centering
-	\begin{tabular}{l|c|c|c}
-		System & Load Graph & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular} & Run BFS \\ \hline
-		Graph500 & 0.3474 & 0.3971 & 0.003380 \\
-		GraphBIG & 37.22  & N/A    & 0.1528 \\
-		GraphMat & 0.1511 & 1.101  & 0.1031 \\
-		GAP      & 2.351  & 0.1935 & 0.001393 \\
-	\end{tabular}
-	\caption{The above table shows times for $2^{20}$ verties and the times are in seconds.
-The Graph500 generates the graph instead of loading it into a file.
-GraphBIG builds the graph and reads in the file simultaneously.
-These results were averaged across 64 roots.}
-	\label{tab:fileload}
-\end{table}
-
-[Compare lines of code for implementation]
-
-\section{Graph Processing Taxonomy}
-This is in the spirit of \cite{Doekemeijer:2015:GPFSurvey}.
+%\section{Graph Processing Taxonomy}
+%This is in the spirit of \cite{Doekemeijer:2015:GPFSurvey}.
 Here, ``|'' means ``or'' and ``+'' means ``and.'' FOSS means Free and Open Source Software.
@@
 Whether these actually achieve their goal is one of the purposes of this project.
-\begin{table*}[t]
-	\begin{minipage}{\linewidth} % So the footnotes all get printed on the same page
-		\centering
-		%\small
-		\pgfplotstabletypeset[
-			col sep=comma,
-			string type, % Makes the .style={string type} redundant
-			columns={Name,Type,Parallelism,Target,FOSS,Source,Notes},
-			every head row/.style={after row=\midrule},
-			columns/Name/.style={string type, column type={l|}},
-			columns/Type/.style={string type},
-			% columns/HPC/.style={string type},
-			columns/Parallelism/.style={string type},
-%			columns/Comm./.style={%
-%				string type,
-%				postproc cell content/.style={%
-%					@cell content/.add={\footnotesize}
-%				}
+%\begin{table*}[t]
+%	\begin{minipage}{\linewidth} % So the footnotes all get printed on the same page
+%		\centering
+%		%\small
+%		\pgfplotstabletypeset[
+%			col sep=comma,
+%			string type, % Makes the .style={string type} redundant
+%			columns={Name,Type,Parallelism,Target,FOSS,Source,Notes},
+%			every head row/.style={after row=\midrule},
+%			columns/Name/.style={string type, column type={l|}},
+%			columns/Type/.style={string type},
+%			% columns/HPC/.style={string type},
+%			columns/Parallelism/.style={string type},
+%%			columns/Comm./.style={%
+%%				string type,
+%%				postproc cell content/.style={%
+%%					@cell content/.add={\footnotesize}
+%%				}
+%%			},
+%			columns/Target/.style={string type},
+%			columns/FOSS/.style={string type},
+%			columns/Source/.style={string type},
+%			columns/Notes/.style={
+%				preproc cell content/.style={@cell content=
+%					\ifx&##1&% Only make a footnote if the cell is nonempty
+%						##1
+%					\else
+%						\footnote{##1}
+%					\fi}
 %			},
-			columns/Target/.style={string type},
-			columns/FOSS/.style={string type},
-			columns/Source/.style={string type},
-			columns/Notes/.style={
-				preproc cell content/.style={@cell content=
-					\ifx&##1&% Only make a footnote if the cell is nonempty
-						##1
-					\else
-						\footnote{##1}
-					\fi}
-			},
-		]{../report/platforms.csv}
-		\caption{Tools used for graph processing}
-		\label{tab:frameworks}
-	\end{minipage}
-\end{table*}
+%		]{../report/platforms.csv}
+%		\caption{Tools used for graph processing}
+%		\label{tab:frameworks}
+%	\end{minipage}
+%\end{table*}
 

### 2017-01-30 papers/gabb17/plot_data.R

@@ -1,7 +1,8 @@
-# This is some bad bad bad R code.
-# Might want to double check it.
+# Generate some plots using the data parsed by ../../misc/parse-output.sh
+# Make this section be the results from the optimal number of threads
+# TODO: Make the first filename compared the highest-performing one.
 scale <- 20
-bpc <- "chartreuse"
+bpc <- "cyan"
 filename <- paste0("parsed",scale,".csv")
 x <- read.csv(filename, header = FALSE)
 colnames(x) <- c("Sys","Algo","Metric","Time")
@@ -28,59 +29,90 @@ pr_iters <- subset(x, x$Algo == "PageRank" & x$Metric == "Iterations",
 pr_time$Sys <- factor(pr_time$Sys)
 pr_dsc$Sys <- factor(pr_dsc$Sys)
 
-pdf("graphics/bfs_time.pdf")
+pdf("graphics/bfs_time.pdf", width = 5.2, height = 5.2)
 boxplot(Time~Sys, bfs_time, ylab = "Time (seconds)",
 		main = "BFS Time", log = "y", col=bpc)
 dev.off()
 
-pdf("graphics/bfs_dsc.pdf")
+pdf("graphics/bfs_dsc.pdf", width = 4.5, height = 4.5)
 boxplot(Time~Sys, bfs_dsc, ylab = "Time (seconds)",
 		main = "BFS Data Structure Construction", col=bpc, log = "y")
 dev.off()
 
-pdf("graphics/sssp_time.pdf")
+pdf("graphics/sssp_time.pdf", width = 4.5, height = 4.5)
 boxplot(Time~Sys, sssp_time, ylab = "Time (seconds)",
 		main = "SSSP Time", log = "y", col=bpc)
 dev.off()
 
-pdf("graphics/sssp_dsc.pdf")
+pdf("graphics/sssp_dsc.pdf", width = 4.5, height = 4.5)
 boxplot(Time~Sys, sssp_time, ylab = "Time (seconds)",
 		main = "SSSP Data Structure Construction", log = "y", col=bpc)
 dev.off()
 
-pdf("graphics/pr_time.pdf")
-barplot(pr_time$Time, ylab = "Time (seconds)",
-		main = "PageRank Time", log = "y", col=bpc,
-		names.arg=factor(pr_time$Sys))
+pdf("graphics/pr_time.pdf", width = 4.5, height = 4.5)
[…]

### 2017-02-01 get-graphalytics.sh
Started adding graphmat to graphalytics.
It builds and configures, just doesn't run. I suspect it has to
do with the file format conversion.

@@ -181,13 +181,13 @@ start_yarn()
 
 install_GraphX()
 {
+	echo "GraphX is currently not working."
+	exit 1
 	# Assumes HDFS and YARN are running.
 	echo "Installing GraphX..."
 	cd "$BASE_DIR"
 	if [ ! $(find "$BASE_DIR" -maxdepth 1 -type d -name graphalytics-platforms-graphx) ]; then
 		git clone https://github.com/tudelft-atlarge/graphalytics-platforms-graphx.git
-		# XXX: Temporary fix until it's merged into main branch
-		# git clone https://github.com/sampollard/graphalytics-platforms-graphx
 	fi
 	GRAPHX_DIR="$BASE_DIR/graphalytics-platforms-graphx"
 	cd "$GRAPHX_DIR"
@@ -229,16 +229,11 @@ install_OpenG()
 {
 	if [ "$OSNAME" != "Linux" ]; then
 		echo "GraphBIG only works on Linux."
-		exit 1
+		kill -s TERM $TOP_PID
 	fi
 	cd "$BASE_DIR"
 	GRAPHBIG_DIR="$BASE_DIR/graphBIG_GA"
 	export OPENG_HOME="$GRAPHBIG_DIR"
-	if [ "$OSNAME" != "Linux" ]; then
-		GRAPHBIG_OPTS="PFM=0"
-	else
-		GRAPHBIG_OPTS=""
-	fi
 	if [ ! -d graphBIG_GA ]; then
 		echo "Downloading and building the GraphBIG repository"
 		git clone 'https://github.com/graphbig/graphBIG.git' $GRAPHBIG_DIR
@@ -374,13 +369,49 @@ install_PowerGraph()
 
 install_Giraph()
 {
+	echo "Giraph is currently unimplemented"
+	exit 1
 	cd "$BASE_DIR"
 	wget http://www-us.apache.org/dist/zookeeper/zookeeper-3.4.9/zookeeper-3.4.9.tar.gz
 	tar -xf zookeeper-3.4.9.tar.gz
 }
 
+install_GraphMat()
+{
+	echo "Installing GraphMat..."
+	cd "$BASE_DIR"
+	# Check for icpc
+	icpc --version
+	if [ $? -ne 0 ]
+		echo "You must have a working intel compiler to continue"
+		exit 1
+	fi
+	# If you are not a member of HPCL please use
+	# https://github.com/narayanan2004/GraphMat instead
+	git clone https://github.com/HPCL/GraphMat.git
+	cd GraphMat
+	make
[…]

### 2017-02-01 papers/gabb17/gabb17-pollard.tex
Updated scalability figure generation and pdfs.
Added .gitignore to ignore extra LaTeX files.

Slightly updated the .tex, still having issues with pgfplotstable.

@@
 \usepackage{pgfplotstable}
+\usepackage{array} % Required for pgfplotstable dec sep align
 \pgfplotsset{compat=1.12}
@@
 			col sep=comma,
-			string type, % Makes the .style={string type} redundant
 			columns={[index]0,GraphBIG,PowerGraph},
@@
 			columns/0/.style={string type, column type={l|}, column name={}},
-			columns/openg/.style={string type, column type={r}},
-			columns/powergraph/.style={string type, column type={r}}
+			columns/openg/.style={dec sep align},
+			columns/powergraph/.style={dec sep align}
 		]{../report/runtime.csv}
@@
 		Graph500 & 0.3474 & 0.3971 & 0.003380 \\
+		GAP      & 2.351  & 0.1935 & 0.001393 \\
+		GraphMat & 0.1511 & 1.101  & 0.1031 \\
 		GraphBIG & 37.22  & N/A    & 0.1528 \\
-		GraphMat & 0.1511 & 1.101  & 0.1031 \\
-		GAP      & 2.351  & 0.1935 & 0.001393 \\
 	\end{tabular}
@@
 	\vspace{-18pt}
-	\caption{The $y$-axis is logarithmic. [TODO: Replace this with the data with non-zero degree vertices]}
+	\caption{The $y$-axis is logarithmic.}
 	\label{fig:bfs-time}
@@
 	\vspace{-18pt}
-	\caption{The $y$-axis \emph{not} logarithmic. GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs unitl none of the vertices' ranks change.}
+	\caption{The $y$-axis \emph{not} logarithmic. GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.}
 	\label{fig:pr-iters}
@@
 	\vspace{-18pt}
-	\caption{The $y$-axis \emph{not} logarithmic.
-GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs unitl none of the vertices' ranks change.
-[There is only one data point]}
+	\caption{Strong Scaling for $2^{20}$ vertices and an average of 16 edges per vertex.}
 	\label{fig:bfs-ss}
+\end{figure}
+
+\begin{figure}
+	\centering
+	% trim left bottom right top
+	\includegraphics[width=\columnwidth, trim=0 18pt 18pt 0, clip]{graphics/bfs_speedup.pdf}
+	\caption{Speedup.}
+	\label{fig:bfs-speedup}
 \end{figure}

### 2017-02-02 papers/gabb17/gabb17-pollard.tex
Changed paper's introduction, tweaked plot_data
Updated some of the figures.

@@
 With a plugin to Graphalytics called Granula \cite{Ngai:2015:Granula} one can explicitly specify a performance model to analyze specific execution behavior such as the amount of communication or runtime of particular kernels of execution.
-As with Graphalytics, the initial development effort is high but Granula allows for multiple experiments to be run easily.
+The requires in-depth knowledge of the source code and execution model which is not always available.
+Furthermore, creating such a model requires a high level of expertise with the given system and with Granula.
 
-One drawback of Granula is it requires knowledge of the source code and execution model which is not always available.
-Furthermore, this requires a high level of expertise with the given system and with Granula\footnote{An example of Granula can be seen at \url{https://github.com/tudelft-atlarge/graphalytics-platforms-graphx/tree/master/granula-model-graphx}}.
+As with Graphalytics, the initial development effort is high but Granula paired with graphalytics allows automatic execution and compilation of performance results\footnote{An example of Granula can be seen at \url{https://github.com/tudelft-atlarge/graphalytics-platforms-graphx/tree/master/granula-model-graphx}}.
+Likewise, our approach provides automatic execution and performance analysis without requiring a performance and execution model for each system.
 
-To address these concerns we analyze the performance of several graph processing benchmarks to facilitate the selection among such systems.
-Our analysis is done without modification of the existing source code.
+To showcase our methods we analyze the performance of several graph processing benchmarks to facilitate the selection among such systems. Our analysis is done with minimal source code modification.
 
@@
 
-The following figures are not useful.
-They were are merely placeholders while further experiments are being run using the correct number of threads and roots.
-These issues are especially noticeable in GraphBIG.
 \begin{table}[!htb]
@@
 \section{Conclusion} % Begin with the end in mind...
-We have presented an updated survey of parallel graph processing frameworks supplementary to \cite{Doekemeijer:2015:GPFSurvey}.
-From this, we have selected a representative subset of frameworks on which performance is analyzed and have stored these results in a database.
-To facilitate parallel graph processing, hardware information and performance results are automatically populated (as were all the tables in this paper).
-These performance results are then used to provide simple recommendations of the optimally-performing framework given a particular algorithm and problem size.
-% We have developed a simple model of hardware and its correlation with performance to predict performance on other architectures
 
-%\bibliographystyle{acm}
 \bibliographystyle{IEEEtranS}

### 2017-02-02 papers/gabb17/plot_data.R

@@ -3,7 +3,7 @@
 # TODO: Make the first filename compared the highest-performing one.
 scale <- 20
 bpc <- "cyan"
-filename <- paste0("parsed",scale,".csv")
+filename <- paste0("parsed",scale,"-32.csv")
 x <- read.csv(filename, header = FALSE)
 colnames(x) <- c("Sys","Algo","Metric","Time")
 
@@ -64,8 +64,9 @@ barplot(pr_mean_iters, ylab = "Time (second)",
 		names.arg = unique(pr_iters$Sys))
 dev.off()
 
-# This isn't kind of wasteful to reread the *-1.csv but it simplifies the code
-threadcnts <- c(1,2,4)
+# Read in and average the data for BFS for each thread
+# It is wasteful to reread the parsed*-1.csv but it simplifies the code
+threadcnts <- c(1,2,4,8,16,32,64,72)
 x <- read.csv(paste0("parsed",scale,"-1.csv"), header = FALSE)
 colnames(x) <- c("Sys","Algo","Metric","Time")
 systems <- unique(x$Sys)
@@ -82,28 +83,47 @@ for (ti in seq(length(threadcnts))) {
 	bfs_scale[ti] <- aggregate(bfs_time$V4, list(bfs_time$V1), mean)[[2]]
 }
 
+colors <- rainbow(nrow(bfs_scale))
+colors <- gsub("F", "C", colors) # You want it darker
+colors <- gsub("CC$", "FF", colors) # But keep it opaque
+
 # Plot the strong scalability for BFS
 pdf("graphics/bfs_ss.pdf", width = 4.5, height = 4.5)
 bfs_ss <- bfs_scale
-colors <- rainbow(nrow(bfs_ss))
-colors <- gsub("F", "C", colors) # You want it darker
-par(lwd = 3)
 # Strong scaling for sequential is 1---we compute that last
 for (ti in rev(seq(length(threadcnts)))) {
 	bfs_ss[ti] <- bfs_ss[1] / (threadcnts[ti] * bfs_ss[ti])
 }
 plot(as.numeric(bfs_ss[1,]), xaxt = "n", type = "b", ylim = c(0,1),
 		ylab = "Scalability", xlab = "Threads", col = colors[1],
-		main = "BFS Strong Scaling", lty = 1)
+		main = "BFS Strong Scaling", lty = 1, pch = 1, lwd = 3)
 for (pli in seq(2,nrow(bfs_ss))) {
[…]

### 2017-02-07 misc/run-experiment.sh
Added PowerGraph SSSP.

@@ -6,18 +6,21 @@
 # ./run-experiment
 # or
 # qsub run-experiment.sh
+USAGE="usage: run-experiment <scale> <num-threads>" # 2^{<scale>} = Number of vertices.
 
+# Say where the packages are located
 # NOTE: YOU MUST SET THE REPOSITORY LOCATIONS AND MODULES TO LOAD
 # It is assumed these are all built and the gen-dataset.sh script has been run.
 # The building instructions are available in the respective repositories.
 # Here, edge factor (number of edges per vertex) is the default of 16.
-S= # Scale. 2^S = Number of vertices.
 DDIR= # Dataset directory
 GAPDIR=
 GRAPHBIGDIR=
 GRAPH500DIR=
 GRAPHMATDIR=
+POWERGRAPHDIR=
 
+# Set some other variables used throughout the experiment
 # PageRank is usually represented as a 32-bit float,
 # so ~6e-8*nvertices is the minimum absolute error detectable
 # We set alpha = 0.15 in the respective source codes.
@@ -25,15 +28,18 @@ GRAPHMATDIR=
 MAXITER=50 # Maximum iterations for PageRank
 TOL=0.00000006
 NRT=32 # Number of roots
+export SKIP_VALIDATION=1
+S=$1
 if [ -z $S -o -z "$DDIR" ]; then
-	echo 'Please set S and the parameters of the form *DIR in run-experiment.sh'
+	echo 'Please provide S and T and set parameters of the form *DIR in run-experiment.sh'
+	echo "$USAGE"
 	exit 2
 fi
-if [ -z "$1" -o "$1" = "-h" -o "$1" = "--help" ]; then
-	echo "usage: run-experiment <num_threads>"
+if [ "$#" -ne 2 -o "$1" = "-h" -o "$1" = "--help" ]; then
+	echo "$USAGE"
 	exit 2
 fi
-export OMP_NUM_THREADS=$1
+export OMP_NUM_THREADS=$2
 
 # Load all the modules here
 module load intel/17
@@ -54,10 +60,17 @@ module load intel/17
 # Graph500:
 # 	git clone https://github.com/sampollard/graph500.git
 #   cd graph500; make
[…]

### 2017-02-07 get-graphalytics.sh
Got get-graphalytics to work with GraphMat on SSSP
It only does SSSP and I can't figure out why.

Cleaned up a few other bugs, moved logs into the experiment
output directory so they are persistent-they have good data in em!

@@ -57,7 +57,7 @@ install_graphalytics()
 		cd "$BASE_DIR"
 		git clone https://github.com/sampollard/ldbc_graphalytics.git
 		cd ldbc_graphalytics
-		mvn install -Dlicence.skip=true
+		mvn clean install -DskipTests=true -Dlicense.skip=true
 		cd ..
 	else
 		echo "I found an ldbc_graphalytics directory. I assume everything is built in there."
@@ -191,7 +191,7 @@ install_GraphX()
 	fi
 	GRAPHX_DIR="$BASE_DIR/graphalytics-platforms-graphx"
 	cd "$GRAPHX_DIR"
-	mvn package
+	mvn clean package
 
 	PKGNAME=$(basename $(find $GRAPHX_DIR -maxdepth 1 -name '*.tar.gz'))
 	VERSION=$(echo $PKGNAME | awk -F '-' '{print $4}')
@@ -250,7 +250,7 @@ install_OpenG()
 		git clone 'https://github.com/tudelft-atlarge/graphalytics-platforms-openg.git'
 	fi
 	cd "$OPENG_DIR" # You have to be in this dir for it to work
-	mvn package
+	mvn clean package
 	PKGNAME=$(basename $(find $OPENG_DIR -maxdepth 1 -name '*.tar.gz'))
 	tar -xf "$PKGNAME"
 	VERSION=$(echo $PKGNAME | awk -F '-' '{print $4}')
@@ -348,7 +348,7 @@ install_PowerGraph()
  		export GRAPHLAB_THREADS_PER_WORKER=$NUM_THREADS
  	fi
 
-	mvn package -DskipTests
+	mvn clean package -DskipTests
     PKGNAME=$(basename $(find $PLATFORM_DIR -maxdepth 1 -name '*.tar.gz'))
     VERSION=$(echo $PKGNAME | awk -F '-' '{print $4}')
     GA_VERSION=$(echo $PKGNAME | awk -F '-' '{print $2}')
@@ -382,7 +382,7 @@ install_GraphMat()
 	cd "$BASE_DIR"
 	# Check for icpc
 	icpc --version
-	if [ $? -ne 0 ]
+	if [ $? -ne 0 ]; then
 		echo "You must have a working intel compiler to continue"
 		exit 1
 	fi
@@ -395,7 +395,7 @@ install_GraphMat()
 	GRAPHMAT_DIR="$BASE_DIR/graphalytics-platforms-graphmat"
[…]

### 2017-02-07 papers/gabb17/gabb17-pollard.tex
Added ideal curve to bfs_ss, expanded abstract.
Added some more citations and cleaned up R code.

@@
 \begin{abstract}
-In this paper we analyze the performance of four parallel graph processing systems to determine their performance and scalability using synthetic and real-world datasets.
-We perform analysis for three algorithms: breadth first serach, single source shortest paths, and PageRank.
-This paper examines previously overlooked aspects of parallel programming performance such as file I/O in addition to a more detailed performance analysis.
+There are an intimidating number of parallel graph processing systems.
+One survey in 2014 identified over 80.
+Since then, the landscape has changed; some packages have become inactive, even more projects have been introducted, and the previous top performers are being usurped.
+To address the high churn of projects and the complexity of selecting a system for a given problem, we analyze the performance and scalability of four parallel open source graph processing systems (GraphMat, GAP Benchmark Suite, GraphBIG, and PowerGraph) using synthetic and real-world datasets.
+We perform analysis for three algorithms: breadth first search, single source shortest paths, and PageRank.
+We compare our results to largest parallel graph processing performance and benchmarking project: Graphalytics.
+This paper examines previously overlooked aspects of parallel graph processing performance such as energy usage in addition to a detailed performance analysis modeled after the Graph500 that takes into consideration multiple runs and the phases of computation that graph programs typically undergo.
 \end{abstract}
@@
 These all require some knowledge of the inner workings of each system in addition to familiarity with the Graphalytics API.
+
+The complexity of Graphalytics obfuscates the true behavior of the algorithm.
+We use GraphMat, the single source shortest paths algorithm, the dota-league dataset, and the default settings as an example here.
+By default, Graphalytics generates an HTML report giving the runtimes for each dataset in seconds.
+Two separate runs for a single dataset, all else being equal, gave 5.6 seconds.
+However, perusing the log files reveals a more complete picture: of these 5.6 seconds, 0.19 was spent computing the shortest paths while the remaining time consisted of building the necessary data structures.
 
@@
 \section{Performance Analysis}
+
+We use the Performance Application Programming Interface (PAPI) \cite{Browne:2000:PAPI} to gain access to Intel's Running Average Power Limit (RAPL), which provides access to a set of hardware counters measuring energy usage.
 

### 2017-02-07 papers/gabb17/plot_data.R

@@ -1,7 +1,10 @@
 # Generate some plots using the data parsed by ../../misc/parse-output.sh
 # Make this section be the results from the optimal number of threads
-# TODO: Make the first filename compared the highest-performing one.
-scale <- 20
+
+###
+# Part 1: Generate the plots for a single problem size and multiple algorithms
+###
+scale <- 21
 bpc <- "cyan"
 filename <- paste0("parsed",scale,"-32.csv")
 x <- read.csv(filename, header = FALSE)
@@ -64,6 +67,10 @@ barplot(pr_mean_iters, ylab = "Time (second)",
 		names.arg = unique(pr_iters$Sys))
 dev.off()
 
+###
+# Part 2: Generate the plots for a single algorithm and multiple problem sizes
+###
+scale <- 20
 # Read in and average the data for BFS for each thread
 # It is wasteful to reread the parsed*-1.csv but it simplifies the code
 threadcnts <- c(1,2,4,8,16,32,64,72)
@@ -101,10 +108,12 @@ for (pli in seq(2,nrow(bfs_ss))) {
 	lines(as.numeric(bfs_ss[pli,]), col = colors[pli], type = "b",
 			lwd = 3, pch = pli, lty = pli) # XXX: lty may repeat after 8
 }
+lines(1:length(threadcnts), 1/threadcnts, lwd = 1, col = "#000000FF")
 axis(1, at = seq(length(threadcnts)), labels = threadcnts)
-legend(legend = rownames(bfs_ss), x = "topright", col = colors,
-		lty = 1:length(systems), pch = 1:length(systems),
-		box.lwd = 1, lwd = 3)
+legend(legend = c(rownames(bfs_ss), "Ideal"), x = "topright",
+		lty = c(1:length(systems), 1), pch = c(1:length(systems), NA_integer_),
+		box.lwd = 1, lwd = c(rep(3,length(systems)), 1),
+		col = c(colors, "#000000FF"))
 dev.off()
 
 # Plot the speedup for BFS

### 2017-02-07 get-graphalytics.sh
Changed what must be compiled for PowerGraph.
GraphMat still only can run sssp.

@@ -327,8 +327,10 @@ install_PowerGraph()
 		cd PowerGraph
 		./configure --no_jvm
 		# Make everything (May not need everything) using at most 16 processes.
-		cd release/toolkits # TODO: Determine what exactly needs to be made
+		cd release/toolkits/graph_analytics
 		local NP=$(if [ "$NUM_CORES" -lt 16 ]; then echo 8; else echo $NUM_CORES; fi)
+		make -j $NP
+		cd ../graph_algorithms
 		make -j $NP
 	else
 		echo "Found PowerGraph directory. I'm assuming you have everything built in there."

### 2017-02-09 rapl/power_rapl.c
Added rapl directory, modified parsing for RAPL.
Addded a baseline which sleeps for 10 seconds so you can measure
how much power the system would consume without any graph algorithm.
Updated parsing (still haven't tested completely)

Moved from misc to experiment.

@@ -1,0 +1,133 @@
+#include "power_rapl.h"
+
+void power_rapl_init(power_rapl_t* ps) {
+    int cid=-1, i, code, retval, enum_retval;
+    ps->EventSet = PAPI_NULL;
+    const PAPI_component_info_t *cmpinfo = NULL;
+    char event_name[BUFSIZ];
+    int rapl_cid;
+
+    ps->num_events = 0;
+
+
+    /* PAPI Initialization */
+    retval = PAPI_library_init( PAPI_VER_CURRENT );
+    if ( retval != PAPI_VER_CURRENT ) {
+        fprintf(stderr,"PAPI_library_init failed\n");
+        exit(1);
+    }
+
+    ps->numcmp = PAPI_num_components();
+
+    for(cid = 0; cid < ps->numcmp; cid++) {
+
+        if ( (cmpinfo = PAPI_get_component_info(cid)) == NULL) {
+            fprintf(stderr,"PAPI_get_component_info failed\n");
+            exit(1);
+        }
+
+        if (strstr(cmpinfo->name,"rapl")) {
+            rapl_cid=cid;
+            printf("Found rapl component at cid %d\n", rapl_cid);
+
+            if (cmpinfo->disabled) {
+                fprintf(stderr,"No rapl events found: %s\n",
+                        cmpinfo->disabled_reason);
+                exit(1);
+            }
+            break;
+        }
+    }
+
+    /* Component not found */
+    if (cid==ps->numcmp) {
+        fprintf(stderr,"No rapl component found\n");
+        exit(1);
+    }
+
+    /* Find Events */
+    code = PAPI_NATIVE_MASK;
+
+    enum_retval = PAPI_enum_cmp_event( &code, PAPI_ENUM_FIRST, cid );
+
+    while ( enum_retval == PAPI_OK ) {
+
+        retval = PAPI_event_code_to_name( code, event_name );
+        if ( retval != PAPI_OK ) {
+            printf("Error translating %#x\n",code);
+            exit(1);
+        }
+
+        printf("Found: %s\n", event_name);
+        strncpy(ps->events[ps->num_events], event_name, BUFSIZ);
+        sprintf(ps->filenames[ps->num_events], "results.%s", event_name);
+        ps->num_events++;
+
+        if (ps->num_events == MAX_RAPL_EVENTS) {
+            printf("Too many events! %d\n",ps->num_events);
+            exit(1);
[…]

### 2017-02-09 rapl/power_rapl.h

@@ -1,0 +1,43 @@
+/** 
+ * @author Boyana Norris
+ */
+#ifndef POWER_RAPL_H
+#define POWER_RAPL_H
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <unistd.h>
+
+#include "papi.h"
+
+#define MAX_RAPL_EVENTS 128
+
+typedef struct {
+  char events[MAX_RAPL_EVENTS][BUFSIZ];
+  char filenames[MAX_RAPL_EVENTS][BUFSIZ];
+  FILE *fff[MAX_RAPL_EVENTS];
+  int numcmp;
+  int EventSet;
+  long long values[MAX_RAPL_EVENTS];
+  long long start_time;
+  long long before_time;
+  double total_time;
+  double elapsed_time;
+  int num_events;
+} power_rapl_t; 
+
+void power_rapl_init(power_rapl_t *);
+void power_rapl_start(power_rapl_t *);
+void power_rapl_end(power_rapl_t *);
+void power_rapl_print(power_rapl_t *);
+
+#ifdef __cplusplus
+} /* extern C */
+#endif
+
+#endif

### 2017-02-09 rapl/sleep_baseline.c

@@ -1,0 +1,15 @@
+/* Sleeps for 10 seconds, measuring the power consumption */
+#include <unistd.h>
+#include "power_rapl.h"
+
+int main(int arc, char *argv[])
+{
+	power_rapl_t ps;
+	power_rapl_init(&ps);
+	printf("Monitoring power with RAPL sleeping\n");
+	power_rapl_start(&ps);
+	sleep(10);
+	power_rapl_end(&ps);
+	power_rapl_print(&ps);
+	return 0;
+}

### 2017-02-09 papers/publication/paper-pollard.tex
Renamed and refactored paper to be in LLNCS format
The abstract is almost ready for submission but the rest of the
paper is in a rough spot.

Updated the makefile and added the .cls files.

Still need to get power/energy usage and get more platforms and
algorithms and get more architectures and write the paper.

--- 2017-02-07 papers/gabb17/gabb17-pollard.tex
@@
-\documentclass[conference]{IEEEtran}
+\documentclass{llncs}
 \usepackage[T1]{fontenc}
@@
 \usepackage{hyperref}
+\usepackage{cite} % Multiple citations in one \cite command
 \usepackage{pgfplotstable}
@@
 
+\begin{document}
 \title{A Comparison of Parallel Graph Processing Benchmarks}
-\author{
-	\IEEEauthorblockN{Sam D. Pollard}
-	\IEEEauthorblockA{Department of Computer and Information Science \\
-		University of Oregon \\
-		Eugene, OR, USA \\
-		Email: spollard@cs.uoregon.edu
+\author{Samuel D. Pollard \and Boyana Norris}
+\institute{University of Oregon \\
+	Eugene OR 97403, USA \\
+	\email{\{spollard,norris\}@cs.uoregon.edu}
 }
-	\and
-	\IEEEauthorblockN{Boyana Norris}
-	\IEEEauthorblockA{Department of Computer and Information Science \\
-	University of Oregon \\
-	Eugene, OR, USA \\
-	Email: norris@cs.uoregon.edu
-	}
-}
-
-\begin{document}
 \maketitle
@@
 One survey in 2014 identified over 80.
-Since then, the landscape has changed; some packages have become inactive, even more projects have been introducted, and the previous top performers are being usurped.
+Since then, the landscape has changed; some packages have become inactive, even more have been introduced, and previous top performers have been usurped.
 To address the high churn of projects and the complexity of selecting a system for a given problem, we analyze the performance and scalability of four parallel open source graph processing systems (GraphMat, GAP Benchmark Suite, GraphBIG, and PowerGraph) using synthetic and real-world datasets.
 We perform analysis for three algorithms: breadth first search, single source shortest paths, and PageRank.
-We compare our results to largest parallel graph processing performance and benchmarking project: Graphalytics.
-This paper examines previously overlooked aspects of parallel graph processing performance such as energy usage in addition to a detailed performance analysis modeled after the Graph500 that takes into consideration multiple runs and the phases of computation that graph programs typically undergo.
+We then compare our results to the most comprehensive parallel graph processing performance and benchmarking project: Graphalytics.
+We examine previously overlooked aspects of parallel graph processing performance such as energy usage in addition to a performance analysis model based on the Graph500.
 \end{abstract}
 
-\section{Introduction}
+\section{Introduction and Related Work}
 
@@
 Graphalytics \cite{Capota:2015:Graphalytics} also attempts to remedy these problems.
-However, in order for Graphalytics to analyze a graph processing system---Graphalytics calls these platforms---a programmer must provide Java methods to call the particular implementations and initialization shell scripts to satisfy dependencies and set environment variables.
+However, in order for Graphalytics to analyze a graph processing system---Graphalytics calls these platforms---a programmer must implement Java classes as wrappers to call the particular implementations, satisfy dependencies, and set environment variables.
 These all require some knowledge of the inner workings of each system in addition to familiarity with the Graphalytics API.
@@
 Our target is shared memory CPU processing.
-Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} are not considered here because the authors do not provide reference implementations for all algorithms.
+Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} are not considered here because the authors do not provide reference implementations.
 The systems are:
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{graphics/bfs_time.pdf}
-	\vspace{-18pt}
+	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_time.pdf}
 	\caption{The $y$-axis is logarithmic.}
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{graphics/bfs_dsc.pdf}
-	\vspace{-18pt}
+	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_dsc.pdf}
 	\caption{The $y$-axis is logarithmic. GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{graphics/sssp_time.pdf}
-	\vspace{-18pt}
+	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/sssp_time.pdf}
 	\caption{The $y$-axis is logarithmic.}
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{graphics/pr_time.pdf}
-	\vspace{-18pt}
+	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/pr_time.pdf}
 	\caption{The $y$-axis is logarithmic. GraphMat continues to run until none of the vertices' ranks change. [There is only one data point]}
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{graphics/pr_iters.pdf}
-	\vspace{-18pt}
+	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/pr_iters.pdf}
 	\caption{The $y$-axis \emph{not} logarithmic. GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.}
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{graphics/bfs_ss.pdf}
-	\vspace{-18pt}
+	\includegraphics[width=0.8\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ss.pdf}
 	\caption{Strong Scaling for $2^{20}$ vertices and an average of 16 edges per vertex.}
@@
 	% trim left bottom right top
-	\includegraphics[width=\columnwidth, trim=0 18pt 18pt 0, clip]{graphics/bfs_speedup.pdf}
+	\includegraphics[width=\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_speedup.pdf}
 	\caption{Speedup.}
@@
 
-\bibliographystyle{IEEEtranS}
+\bibliographystyle{splncs03}
 \bibliography{../drp}

### 2017-02-09 experiment/run-multi-thread.sh
Got rapl printing and parsing consistent.
Added files for running experiments, though they will only work
on arya.

@@ -1,0 +1,7 @@
+#!/bin/bash
+# Runs run-experiment.sh for each thread number.
+T=32
+for S in $(echo 18 19 21); do
+	./run-experiment.sh $S $T > out${S}-${T}.log 2> out${S}-${T}.err
+done
+

### 2017-02-09 experiment/run-power.sh

@@ -1,0 +1,40 @@
+#!/bin/bash
+# Measure Power Consumption
+# Just works for BFS on Graph500 and GAP
+# Requires root
+
+# Set some parameters
+DDIR=/home/users/spollard/graphalytics/all-datasets/gabb17 # Dataset directory
+GAPDIR=/home/users/spollard/gap/gapbs
+GRAPH500DIR=/home/users/spollard/graph500
+S=23
+NRT=32 # Number of roots that we did BFS on. GraphBIG has issues with >32.
+PKG=2 # The number of physical chips
+export OMP_NUM_THREADS=32
+T=$OMP_NUM_THREADS
+FN="out${S}-${T}-power.log"
+PFN="parsed${S}-${T}-power.csv"
+
+# Run experiments
+# GAP BFS
+for ROOT in $(head -n $NRT "$DDIR/kron-${S}-roots.v"); do
+    sudo "$GAPDIR"/bfs -r $ROOT -f "$DDIR/kron-${S}.el" -n 1 -s
+done > "$FN" 2> out${S}-${T}-power.err
+# Graph500 BFS
+sudo "$GRAPH500DIR/omp-csr/omp-csr" -s $S >> "$FN" 2>> "out${S}-${T}-power.err"
+# Baseline (do nothing, just sleep)
+sudo "$GAPDIR"/sleep_baseline >> "$FN" 2>> "out${S}-${T}-power.err"
+chown spollard "$FN"
+chown spollard "out${S}-${T}-power.err"
+
+# Parse
+# CPU Average Power
+grep -A 29 'RAPL on GAP BFS' "$FN" | awk -v PKG=$PKG '/Average.*PACKAGE_ENERGY:PACKAGE[0-9]+ \*/{t+=$3;c++}END{print "GAP,BFS,Average CPU Power (W)," t/(c*PKG)}' > "$PFN"
+grep -A 29 'RAPL on Graph500 BFS' "$FN" | awk -v PKG=$PKG '/Average.*PACKAGE_ENERGY:PACKAGE[0-9]+ \*/{t+=$3;c++}END{print "Graph500,BFS,Average CPU Power (W)," t/c*PKG}' >> "$PFN"
+# CPU Total Energy
[…]

### 2017-02-09 rapl/sleep_baseline.c

@@ -6,10 +6,11 @@ int main(int arc, char *argv[])
 {
 	power_rapl_t ps;
 	power_rapl_init(&ps);
-	printf("Monitoring power with RAPL sleeping\n");
+	printf("Monitoring baseline sleeping power for 10 seconds\n");
 	power_rapl_start(&ps);
 	sleep(10);
 	power_rapl_end(&ps);
 	power_rapl_print(&ps);
 	return 0;
 }
+

### 2017-02-09 papers/publication/paper-pollard.tex
msc

@@
 \begin{abstract}
-There are an intimidating number of parallel graph processing systems.
-One survey in 2014 identified over 80.
-Since then, the landscape has changed; some packages have become inactive, even more have been introduced, and previous top performers have been usurped.
-To address the high churn of projects and the complexity of selecting a system for a given problem, we analyze the performance and scalability of four parallel open source graph processing systems (GraphMat, GAP Benchmark Suite, GraphBIG, and PowerGraph) using synthetic and real-world datasets.
-We perform analysis for three algorithms: breadth first search, single source shortest paths, and PageRank.
-We then compare our results to the most comprehensive parallel graph processing performance and benchmarking project: Graphalytics.
-We examine previously overlooked aspects of parallel graph processing performance such as energy usage in addition to a performance analysis model based on the Graph500.
+The increasing popularity of large network analysis problems has led to the emergence of many parallel and distributed graph processing systems (one survey in 2014 identified over 80).
+Since then, the landscape has evolved---some packages have become inactive while others are being added.
+The challenge of trying to determine the best approach for a given problem is infeasible for most developers.
+To enable \emph{easy}, \emph{rigorous} and \emph{repeatable} comparison of parallel graph software capabilities, we have developed an approach and associated software for analyzing the performance and scalability of parallel open source graph processing systems.
+We demonstrate our approach on four graph frameworks: GraphMat, GAP Benchmark Suite, GraphBIG, and PowerGraph using synthetic and real-world datasets.
+We analyze three algorithms: breadth first search, single source shortest paths, and PageRank and compare our results to Graphalytics.
+We examine previously overlooked aspects of parallel graph processing performance such as energy usage.
 \end{abstract}

### 2017-02-09 papers/publication/paper-pollard.tex
Updated abstract.

--- 2017-02-09 papers/publication/paper-pollard.tex
@@
 \begin{abstract}
-There are an intimidating number of parallel graph processing systems.
-One survey in 2014 identified over 80.
-Since then, the landscape has changed; some packages have become inactive, even more have been introduced, and previous top performers have been usurped.
-To address the high churn of projects and the complexity of selecting a system for a given problem, we analyze the performance and scalability of four parallel open source graph processing systems (GraphMat, GAP Benchmark Suite, GraphBIG, and PowerGraph) using synthetic and real-world datasets.
-We perform analysis for three algorithms: breadth first search, single source shortest paths, and PageRank.
-We then compare our results to the most comprehensive parallel graph processing performance and benchmarking project: Graphalytics.
-We examine previously overlooked aspects of parallel graph processing performance such as energy usage in addition to a performance analysis model based on the Graph500.
+The increasing popularity of large network analysis problems has led to the emergence of many parallel and distributed graph processing systems---one survey in 2014 identified over 80.
+Since then, the landscape has evolved; some packages have become inactive while more are being developed.
+Determining the best approach for a given problem is infeasible for most developers.
+To enable easy, rigorous, and repeatable comparison of the capabilities of such systems, we present an approach and associated software for analyzing the performance and scalability of parallel, open-source graph processing systems.
+We demonstrate our approach on five graph frameworks: GraphMat, the Graph500, the Graph Algorithm Platform, GraphBIG, and PowerGraph using synthetic and real-world datasets.
+We examine previously overlooked aspects of parallel graph processing performance such as phases of execution and energy usage for three algorithms: breadth first search, single source shortest paths, and PageRank and compare our results to Graphalytics.
 \end{abstract}

### 2017-02-10 papers/publication/paper-pollard.tex
Finalized abstract

@@
 Determining the best approach for a given problem is infeasible for most developers.
-To enable easy, rigorous, and repeatable comparison of the capabilities of such systems, we present an approach and associated software for analyzing the performance and scalability of parallel, open-source graph processing systems.
+To enable easy, rigorous, and repeatable comparison of the capabilities of such systems, we present an approach and associated software for analyzing the performance and scalability of parallel, open-source graph libraries.
 We demonstrate our approach on five graph frameworks: GraphMat, the Graph500, the Graph Algorithm Platform, GraphBIG, and PowerGraph using synthetic and real-world datasets.

### 2017-02-10 papers/publication/paper-pollard.tex
initial reorg of intro

--- 2017-02-09 papers/publication/paper-pollard.tex
@@
 
-\section{Introduction and Related Work}
+\section{Introduction}
+
+%% What's the problem (motivation)
+
+Another issue among parallel graph processing systems is the lack of comprehensive comparisons.
+One possible reason is the considerable effort involved in getting each system to run: satisfying dependencies and ensuring data are correctly formatted are generally nontrivial tasks.
+Beyond this, there are optimizations for each system which are not exploited in a na\"ive implementation.
+Addressing these issues helps provide an environment in which no graph processing system has an advantage.
+
+%% How do we address it?
+
+Our contributions can be summarized as follows.
+\begin{itemize}
+\item
+\item
+\item
+\end{itemize}
+
+\section{Related Work}
 
@@
 
-Beyond the systems described by Doekemeijer and Varbanescu, the problem has compounded with the addition of even more proprietary and open source projects such as \cite{Cheramangalath:2015:Falcon}, \cite{Perez:2015:Ringo}, distributed memory approaches such as \cite{Hong:2015:PGX}.
+Beyond the systems described by Doekemeijer and Varbanescu, the problem has compounded with the addition of even more proprietary and open source projects such as \cite{Cheramangalath:2015:Falcon,Perez:2015:Ringo}, distributed memory approaches such as \cite{Hong:2015:PGX}.
 domain-specific languages \cite{Hong:2012:GreenMarl}, distributed database querying, \cite{Rodriguez:2015:Gremlin}, as well as novel communication schemes \cite{Edmonds:2013:ActiveMessages}.
@@
 
-Another issue among parallel graph processing systems is the lack of comprehensive comparisons.
-One possible reason is the considerable effort involved in getting each system to run: satisfying dependencies and ensuring data are correctly formatted are generally nontrivial tasks.
-Beyond this, there are optimizations for each system which are not exploited in a na\"ive implementation.
-Addressing these issues helps provide an environment in which no graph processing system has an advantage.
-
-Graphalytics \cite{Capota:2015:Graphalytics} also attempts to remedy these problems.
+Graphalytics \cite{Capota:2015:Graphalytics} attempts to automate the setup and execution of packages in order to compare performance.
 However, in order for Graphalytics to analyze a graph processing system---Graphalytics calls these platforms---a programmer must implement Java classes as wrappers to call the particular implementations, satisfy dependencies, and set environment variables.

### 2017-02-10 experiment/run-power.sh
Added support for GraphMat RAPL on BFS.
Fixed a few small little documentation things, but the top level
README is still severly lacking on the documentation side.

@@ -3,7 +3,9 @@
 # Just works for BFS on Graph500 and GAP
 # Requires root
 
-# Set some parameters
+# Set some parameters and environment variables
+module load intel/17
+module load tau
 DDIR=/home/users/spollard/graphalytics/all-datasets/gabb17 # Dataset directory
 GAPDIR=/home/users/spollard/gap/gapbs
 GRAPH500DIR=/home/users/spollard/graph500
@@ -11,6 +13,7 @@ S=23
 NRT=32 # Number of roots that we did BFS on. GraphBIG has issues with >32.
 PKG=2 # The number of physical chips
 export OMP_NUM_THREADS=32
+export SKIP_VALIDATION=1 # Graph500 by default verifies the BFS
 T=$OMP_NUM_THREADS
 FN="out${S}-${T}-power.log"
 PFN="parsed${S}-${T}-power.csv"
@@ -22,6 +25,13 @@ for ROOT in $(head -n $NRT "$DDIR/kron-${S}-roots.v"); do
 done > "$FN" 2> out${S}-${T}-power.err
 # Graph500 BFS
 sudo "$GRAPH500DIR/omp-csr/omp-csr" -s $S >> "$FN" 2>> "out${S}-${T}-power.err"
+# GraphMat BFS
+for ROOT in $(head -n $NRT "$DDIR/kron-${S}-roots.1v"); do
+	echo "BFS root: $ROOT"
+    # WARNING: May not behave nicely if there are spaces in your DDIR or LD path
+    # TODO: Check if this is actually running on 32 threads
+    sudo bash -c "export LD_LIBRARY_PATH=$LD_LIBRARY_PATH; bin/BFS $DDIR/kron-${S}.graphmat $ROOT"
+done > "$FN" 2> out${S}-${T}-power.err
 # Baseline (do nothing, just sleep)
 sudo "$GAPDIR"/sleep_baseline >> "$FN" 2>> "out${S}-${T}-power.err"
 chown spollard "$FN"
@@ -29,12 +39,14 @@ chown spollard "out${S}-${T}-power.err"
 
 # Parse
 # CPU Average Power
[…]

### 2017-02-13 papers/publication/paper-pollard.tex
Updated R script to generate power plots.
Made a few small changes to the apper. Added the paper pdf.

@@
 	Eugene OR 97403, USA \\
-	\email{\{spollard,norris\}@cs.uoregon.edu}
+	\email{spollard@cs.uoregon.edu} \\
+	\email{norris@cs.uoregon.edu}
 }
@@
 \begin{table}
-%	\begin{tabular}{l|r|r|}
-%	 & PowerGraph & OpenG \\ \hline
-%	BFS & 81.8 & 341 \\ \hline
-%	SSSP & 1.64 & 15.0 \\ \hline
-%	LCC & 54.6 & 142 \\ \hline
-%	\end{tabular}
 	\centering
@@
 		col sep=comma,
-			columns={[index]0,GraphBIG,PowerGraph},
+		columns={[index]0,graphmat,openg,powergraph},
 		every head row/.style={after row=\midrule},
 		columns/0/.style={string type, column type={l|}, column name={}},
-			columns/openg/.style={dec sep align},
-			columns/powergraph/.style={dec sep align}
+		columns/graphmat/.style={
+			column name={GraphMat},
+			string replace={0}{},
+			dec sep align,
+			empty cells with={N/A}
+		},
+		columns/openg/.style={
+		column name={GraphBIG},
+			dec sep align},
+			columns/powergraph/.style={
+			column name={PowerGraph},
+			dec sep align}
 	]{../report/runtime.csv}
-	\caption{Performance results are in milliseconds. LCC is local clustering coefficient, PR is PageRank, CDLP is community detection using label propagation, and WCC is weakly connected components}
+	\caption{Performance results are in milliseconds.
+LCC is local clustering coefficient, PR is PageRank, CDLP is community detection using label propagation, and WCC is weakly connected components.
+At the time of this writing, Graphalytics only supports SSSP for GraphMat.}
 	\label{tab:graphalytics}

### 2017-02-13 papers/publication/plot_data.R

@@ -58,6 +58,7 @@ boxplot(Time~Sys, pr_time, ylab = "Time (seconds)",
 dev.off()
 
 pdf("graphics/pr_iters.pdf", width = 4.5, height = 4.5)
+# Could be done with aggregate instead
 pr_mean_iters <- c(
 		mean(pr_iters$Time[pr_iters$Sys == "GraphBIG"]),
 		mean(pr_iters$Time[pr_iters$Sys == "GraphMat"]),
@@ -132,6 +133,81 @@ axis(1, at = seq(length(threadcnts)), labels = threadcnts)
 legend(legend = c(rownames(bfs_spd), "Ideal"), x = "topleft",
 		col = c(colors,"#000000FF"), lwd = c(rep(3,length(systems)),1),
 		lty = c(1:length(systems), 1), pch = c(1:length(systems), NA_integer_))
+dev.off()
+
+###
+# Part 3: Power
+###
+# Just BFS for now... :(
+# Read in the data
+GRAPH500NRT <- 64 # Even though everyone else does 32.
+scale <- 16
+filename <- paste0("parsed",scale,"-32-power.csv")
+x <- read.csv(filename, header = FALSE)
+colnames(x) <- c("Sys","Algo","Metric","Value")
+bfs_cpu_pwr <- subset(x, x$Algo == "BFS" & x$Metric == "Average CPU Power (W)",
+		c("Sys","Value"))
+bfs_cpu_nrg <- subset(x, x$Algo == "BFS" & x$Metric == "Total CPU Energy (J)",
+		c("Sys","Value"))
+cpu_pwr_sleep <- subset(x,
+		x$Sys == "Baseline" & x$Metric == "Average CPU Power (W)",
+		c("Sys","Value"))
+ram_pwr_sleep <- subset(x,
+		x$Sys == "Baseline" & x$Metric == "Average DRAM Power (W)",
+		c("Sys","Value"))
+bfs_cpu_pwr$Sys <- factor(bfs_cpu_pwr$Sys)
+bfs_cpu_nrg$Sys <- factor(bfs_cpu_nrg$Sys)
+
+bfs_systems <- unique(bfs_cpu_nrg$Sys)
+bfs_cpu_nrg_per_root <- numeric(length(bfs_systems))
+for (si in seq(length(bfs_systems))) {
+	sys <- as.character(unique(bfs_cpu_nrg$Sys)[si])
+	one_sys <- subset(x,
[…]

### 2017-02-13 papers/publication/paper-pollard.tex
Updated the paper. Mostly sections 1--3.

@@
 %% What's the problem (motivation)
+Our research is motivated by the current state of parallel graph processing.
+The most comprehensive survey, released in 2014, identified and taxonomized over 80 different parallel graph processing systems without even including domain specific languages \cite{Doekemeijer:2015:GPFSurvey}.
 
-Another issue among parallel graph processing systems is the lack of comprehensive comparisons.
+An issue among these systems is the lack of comprehensive comparisons.
 One possible reason is the considerable effort involved in getting each system to run: satisfying dependencies and ensuring data are correctly formatted are generally nontrivial tasks.
-Beyond this, there are optimizations for each system which are not exploited in a na\"ive implementation.
-Addressing these issues helps provide an environment in which no graph processing system has an advantage.
+Additionally, some systems are developed with large, multi-node clusters in mind while others only work with shared memory, single node computers.
+An example of the former is GraphX \cite{Xin:2013:GraphX}, which incurs some overhead while gaining fault tolerance and an example of the latter is Ligra \cite{Shun:2013:Ligra}, a framework requiring a shared-memory architecture.
+% ``The project was motivated by the fact that the largest publicly available real-world graphs all fit in shared memory''
+
+At the human level, there are optimizations for each system which may not be apparent.
+For example, the Parallel Boost Graph Library (PBGL) \cite{Gregor:2005:PBGL} provides generic implementations of their algorithms and the programmer must provide the template specializations.
+The optimal data structures may differ across algorithms and graphs and must be determined by the programmer.
 
 %% How do we address it?
+To mitigate the inherent bias in hand-generated implementations, all the experiments performed use the author-provided implementations; we assume the developers of each system will provide the most performant implementation.
+While this limits the scope of the experiments it mitigates the bias inherent in our programming skills.
+
+%To address the challenges of getting various systems installed, we automate the process as much as possible n
+%To showcase our methods we analyze the performance of several graph processing benchmarks to facilitate the selection among such systems. Our analysis is done with minimal source code modification.
 
@@
 \begin{itemize}
-\item
-\item
-\item
+	\item Automation of the installation, configuration, and dataset conversion to ensure the experiments execute in the same manner.
+Essentially, we provide a ``level playing field'' for graph algorithms.
+	\item Comparison of algorithm runtime and scalability using a variety of data sets
+	\item Comparison of our framework with the Graphalytics \cite{Capota:2015:Graphalytics} project
+	\item Inspection of the source code of the surveyed parallel graph processing systems to ensure the same phases of execution are measured across differing execution and programming paradigms
+	\item Provision of a experimental framework distinguishing between the data structure generation and algorithm execution phases of each system
+	\item Analysis of energy consumption during the algorithm execution phase
 \end{itemize}
@@
 \section{Related Work}
+A related performance analysis project is Graphalytics \cite{Capota:2015:Graphalytics} which also attempts to automate the setup and execution of packages for performance analysis.
+However, in order for Graphalytics to analyze a graph processing system---Graphalytics calls these platforms---a programmer must implement Java classes as wrappers to call the particular implementations, satisfy dependencies, and set the correct shell commands to execute.
+These all require some knowledge of the inner workings of each system in addition to familiarity with the Graphalytics API.
 
-Our research is motivated by the current state of parallel graph processing.
-The most comprehensive survey, released in 2014, identified and taxonomized over 80 different parallel graph processing systems without including domain specific languages \cite{Doekemeijer:2015:GPFSurvey}.
-The systems described operate with a wide range of parallelism paradigms and target architectures such as GPU \cite{Zhong:2014:Medusa, Kang:2009:Pegasus}, shared memory CPU \cite{Shun:2013:Ligra, kyrola:2012:Graphchi, Nguyen:2013:Galois}, a combination of CPU and GPU \cite{Gharaibeh:2012:Totem}, distributed filesystem based approaches \cite{Xin:2013:GraphX}, and distributed memory with MPI \cite{Gregor:2005:PBGL}.
+The complexity of Graphalytics obfuscates the true behavior of the program.
+By default, Graphalytics generates an HTML report listing the runtimes for each dataset and each algorithm in seconds.
+For example, one run of a single source shortest paths algorithm took 5.6 seconds.
+However, perusing the log files reveals a more complete picture: of these 5.96 seconds, 0.21 was spent computing the shortest paths while the remaining time consisted of building the necessary data structures.
+Furthermore, data such as the number of iterations run for PageRank is not easily available using Graphalytics.
+% We use GraphMat, the single source shortest paths algorithm, the dota-league dataset from the Game Trace Archive\cite{Guo:2012:GTA}, and the default settings as an example here.
+% The full logs:
+%Finished setting ids, time: 0.157009
+%Starting sort
+%Finished sort, time: 1.39571
+%Finished setting edge pointers, time: 0.000133
+%Starting build_dcsc
+%Finished build_dcsc, time: 0.235948
+%Finished setting ids, time: 0.154693
+%Starting sort
+%Finished sort, time: 1.75676
+%Finished setting edge pointers, time: 0.000142
+%Starting build_dcsc
+%Finished build_dcsc, time: 0.238386
+%Completed reading A from memory in 4.089144 seconds.
+%Completed reading A from file in 5.114708 seconds
+%Completed 19 iterations
+%Timing results:
+%- load graph: 5.11472 sec
+%- initialize engine: 2.88486e-05 sec
+%- run algorithm: 0.212451 sec
+%- print output: 0.0799532 sec
+%- deinitialize engine: 2.88486e-05 sec
+%17:29:05.575 [INFO ] Benchmarking algorithm "Single source shortest paths" on graph "dota-league" took 5955 ms.
 
-Beyond the systems described by Doekemeijer and Varbanescu, the problem has compounded with the addition of even more proprietary and open source projects such as \cite{Cheramangalath:2015:Falcon,Perez:2015:Ringo}, distributed memory approaches such as \cite{Hong:2015:PGX}.
+With a plugin to Graphalytics called Granula \cite{Ngai:2015:Granula} one can explicitly specify a performance model to analyze specific execution behavior such as the amount of communication or runtime of particular kernels of execution.
+The requires in-depth knowledge of the source code and execution model which is not always available.
+Furthermore, creating such a model requires a high level of expertise with the given system and with Granula\footnote{An example of Granula can be seen at \url{https://github.com/tudelft-atlarge/graphalytics-platforms-graphx/tree/master/granula-model-graphx}}.
+
+As with Graphalytics, the initial development effort is high but Granula paired with graphalytics allows automatic execution and compilation of performance results.
+Likewise, our approach provides automatic execution and performance analysis without requiring a performance and execution model for each system.
+
+% The systems described in \cite{Doekemeijer:2015:GPFSurvey} operate with a wide range of parallelism paradigms and target architectures such as GPU \cite{Zhong:2014:Medusa, Kang:2009:Pegasus}, shared memory CPU \cite{Shun:2013:Ligra, kyrola:2012:Graphchi, Nguyen:2013:Galois}, a combination of CPU and GPU \cite{Gharaibeh:2012:Totem}, distributed filesystem based approaches \cite{Xin:2013:GraphX}, and distributed memory with MPI \cite{Gregor:2005:PBGL}.
+
+% Beyond the systems described by Doekemeijer and Varbanescu, the problem has compounded with the addition of even more proprietary and open source projects such as \cite{Cheramangalath:2015:Falcon, Perez:2015:Ringo}, distributed memory approaches such as \cite{Hong:2015:PGX}.
 domain-specific languages \cite{Hong:2012:GreenMarl}, distributed database querying, \cite{Rodriguez:2015:Gremlin}, as well as novel communication schemes \cite{Edmonds:2013:ActiveMessages}.
@@
 
-In addition to libraries with associated APIs there has also been a propagation of ``reference implementations'' which implement the most common graph algorithms such as \cite{Beamer:2015:GAPBench, Nai:2015:Graphbig}.
+% In addition to libraries with associated APIs there has also been a propagation of ``reference implementations'' which implement the most common graph algorithms such as \cite{Beamer:2015:GAPBench, Nai:2015:Graphbig}.
 Thus, even selecting a standard and a benchmark over which to compare various implementations is nontrivial.
 To quote Andrew Tanenbaum, ``The nice thing about standards is that you have so many to choose from.''
-
-Graphalytics \cite{Capota:2015:Graphalytics} attempts to automate the setup and execution of packages in order to compare performance.
-However, in order for Graphalytics to analyze a graph processing system---Graphalytics calls these platforms---a programmer must implement Java classes as wrappers to call the particular implementations, satisfy dependencies, and set environment variables.
-These all require some knowledge of the inner workings of each system in addition to familiarity with the Graphalytics API.
-
-The complexity of Graphalytics obfuscates the true behavior of the algorithm.
-We use GraphMat, the single source shortest paths algorithm, the dota-league dataset, and the default settings as an example here.
-By default, Graphalytics generates an HTML report giving the runtimes for each dataset in seconds.
-Two separate runs for a single dataset, all else being equal, gave 5.6 seconds.
-However, perusing the log files reveals a more complete picture: of these 5.6 seconds, 0.19 was spent computing the shortest paths while the remaining time consisted of building the necessary data structures.
-
-With a plugin to Graphalytics called Granula \cite{Ngai:2015:Granula} one can explicitly specify a performance model to analyze specific execution behavior such as the amount of communication or runtime of particular kernels of execution.
-The requires in-depth knowledge of the source code and execution model which is not always available.
-Furthermore, creating such a model requires a high level of expertise with the given system and with Granula.
-
-As with Graphalytics, the initial development effort is high but Granula paired with graphalytics allows automatic execution and compilation of performance results\footnote{An example of Granula can be seen at \url{https://github.com/tudelft-atlarge/graphalytics-platforms-graphx/tree/master/granula-model-graphx}}.
-Likewise, our approach provides automatic execution and performance analysis without requiring a performance and execution model for each system.
-
-To showcase our methods we analyze the performance of several graph processing benchmarks to facilitate the selection among such systems. Our analysis is done with minimal source code modification.
 
@@
 This report explores four shared memory parallel graph processing platforms.
-The first three are so-called ``reference implementations'' while the fourth is included because it has been cited as highly performant \cite{Sundaram:2015:GraphMat}.
+The first three are so-called ``reference implementations'' while the remaining two are included because of their performance and popularity.
 Our target is shared memory CPU processing.
@@
 	\item GraphMat \cite{Sundaram:2015:GraphMat}
+	\item PowerGraph \cite{Gonzalez:2012:Powergraph}
 \end{enumerate}
@@
 
-We consider three algorithms:
+We consider three algorithms, but not all algorithms are available on all systems.
 
@@
 The advantage of the Graph500 is it provides standardized measurement specifications and dataset generation.
-The primary drawback with the Graph500 is it measures a single algorithm: breadth first search.
+The primary drawback with the Graph500 is it measures a single algorithm: BFS.
 
-This report aims to add similar rigor to other graph algorithms by borrowing heavily from the Graph500 specification.
+Our work aims to add similar rigor to other graph algorithms by borrowing heavily from the Graph500 specification.
 The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the actual BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}.
@@
 
-The third algorithm is PageRank.
-These three algorithms are used because of their popularity; most libraries provide reference implementations.
-All the experiments performed use the developer-provided implementations because we assume the vendor will provide a more performant implementation than us.
-While this limits the scope of the experiments it mitigates the bias inherent in any developer's programming skills.
+The third algorithm is PageRank. These three algorithms are used because of their popularity; most libraries provide reference implementations.
 
 \subsection{Machine Specifications}
-Table~\ref{tab:specs} shows the specifications of the research machine.
-The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
-We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.
+Table~\ref{tab:specs} shows the specifications of the research computer.
 
@@
 	]{../report/specs.csv}
-	\caption{The operating system is GNU/Linux version 4.4.0-22.}
+	\caption{The operating system is GNU/Linux version 4.4.0-22.
+The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
+We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.}
 	\label{tab:specs}
@@
 However, results such as these are preliminary and do not show the complete picture.
-
-[How to make this a paper worth submitting: add in PowerGraph as a platform and see how it fares using our method compared to Graphalytics. Also run this on different datasets and compare results.]
-
-%\begin{table}[htb]
-%	\centering
-%	\begin{tabular}{l|c|c|c}
-%		& BFS & SSSP & PR \\ \hline
-%		Graph500 & & & \\ \hline
-%		PBGL  & & & \\ \hline
-%		GAP  & & & \\ \hline
-%		GraphBIG  & & & \\ \hline
-%		Galois  & & & \\ \hline
-%		PowerGraph  & & &
-%	\end{tabular}
-%	\caption{Performance. Note that implementations of every algorithm are not availabe for every platform.}
-%	\label{tab:reportcard}
-%\end{table}
+It is difficult to make any claims about GraphMat from Graphalytics because of its lack of algorithm diversity.
 
@@
 Graphalytics also outputs MTEPS or millions of traversed edges pers econd.
-However, the graphalytics version does not make sense in all cases: for example, computing the local clustering coefficient involves traversing each edge multiple times (proportional to the sparsity of the graph), while breadth first search (BFS) traverses each edge exactly once, and on na\:ive implementations single-source shortest paths (SSSP) may have $O(|E| + |V|^2)$ traversed edges.
+However, the graphalytics version does not make sense in all cases: for example, computing the local clustering coefficient involves traversing each edge multiple times (proportional to the sparsity of the graph), while BFS traverses each edge exactly once, and the number of edges traversed with PageRank depends on the connectivity of the graph and the number of iterations.
 

### 2017-02-14 papers/publication/paper-pollard.tex
Cleaned up some graphics and merged to save space.
Still waiting on some power results that I messed up.

@@
 
-An issue among these systems is the lack of comprehensive comparisons.
+An overarching issue among these systems is the lack of comprehensive comparisons.
 One possible reason is the considerable effort involved in getting each system to run: satisfying dependencies and ensuring data are correctly formatted are generally nontrivial tasks.
@@
 Furthermore, data such as the number of iterations run for PageRank is not easily available using Graphalytics.
-% We use GraphMat, the single source shortest paths algorithm, the dota-league dataset from the Game Trace Archive\cite{Guo:2012:GTA}, and the default settings as an example here.
+%We use GraphMat, the single source shortest paths algorithm, the dota-league dataset from the Game Trace Archive\cite{Guo:2012:GTA}, and the default settings as an example here.
 % The full logs:
@@
 
-We use the Performance Application Programming Interface (PAPI) \cite{Browne:2000:PAPI} to gain access to Intel's Running Average Power Limit (RAPL), which provides access to a set of hardware counters measuring energy usage.
-
 In Table~\ref{tab:graphalytics} we show the results from running Graphalytics. For an explanation of each algorithm used, see \cite{Iosup:2016:Graphalyticstech}.
@@
 
-\begin{table}[!htb]
+Table~\ref{tab:epg-perf}, we show the performance results of running graphalytics and our approach using the same, synthetic dataset.
+Now, our approach computes multiple runs whereas graphalytics only uses one root for BFS.
+This gives a better picture of algorithm reliability and is explained further.
+
+\begin{table}
 	\centering
@@
 These results were averaged across 32 roots.}
-	\label{tab:fileload}
+	\label{tab:epg-perf}
 \end{table}
+
+We use the Performance Application Programming Interface (PAPI) \cite{Browne:2000:PAPI} to gain access to Intel's Running Average Power Limit (RAPL), which provides access to a set of hardware counters measuring energy usage.
 
@@
 	\centering
-	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_time.pdf}
-	\caption{The $y$-axis is logarithmic.}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_time.pdf}
+	\end{minipage}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_dsc.pdf}
+	\end{minipage}
+	\caption{The $y$-axis is logarithmic. GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
 	\label{fig:bfs-time}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_dsc.pdf}
-	\caption{The $y$-axis is logarithmic. GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
-	\label{fig:bfs-dsc}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/pr_time.pdf}
-	\caption{The $y$-axis is logarithmic. GraphMat continues to run until none of the vertices' ranks change. [There is only one data point]}
-	\label{fig:pr-time}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/pr_time.pdf}
+	\end{minipage}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/pr_iters.pdf}
+	\end{minipage}
+	\caption{The $y$-axis is logarithmic.
+GraphMat continues to run until none of the vertices' ranks change.
+[There is only one data point] The $y$-axis \emph{not} logarithmic.
+GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.
+[Note that for figure Fig.\ref{fig:pr-iters}, mention time per iteration]}
+	\label{fig:pr}
 \end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/pr_iters.pdf}
-	\caption{The $y$-axis \emph{not} logarithmic. GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.}
-	\label{fig:pr-iters}
-\end{figure}
-[Note that for figure Fig.\ref{fig:pr-iters}, mention time per iteration]
 

### 2017-02-14 papers/publication/plot_data.R

@@ -35,11 +35,13 @@ pr_dsc$Sys <- factor(pr_dsc$Sys)
 pdf("graphics/bfs_time.pdf", width = 5.2, height = 5.2)
 boxplot(Time~Sys, bfs_time, ylab = "Time (seconds)",
 		main = "BFS Time", log = "y", col=bpc)
+mtext(paste0("Scale = ",scale), side = 3) # May want to remove subtitle later
 dev.off()
 
-pdf("graphics/bfs_dsc.pdf", width = 4.5, height = 4.5)
+pdf("graphics/bfs_dsc.pdf", width = 5.2, height = 5.2)
 boxplot(Time~Sys, bfs_dsc, ylab = "Time (seconds)",
 		main = "BFS Data Structure Construction", col=bpc, log = "y")
+mtext(paste0("Scale = ",scale), side = 3) # May want to remove subtitle later
 dev.off()
 
 pdf("graphics/sssp_time.pdf", width = 4.5, height = 4.5)
@@ -55,17 +57,18 @@ dev.off()
 pdf("graphics/pr_time.pdf", width = 4.5, height = 4.5)
 boxplot(Time~Sys, pr_time, ylab = "Time (seconds)",
 		main = "PageRank Time", log = "y", col=bpc)
+mtext(paste0("Scale = ",scale), side = 3) # May want to remove subtitle later
 dev.off()
 
 pdf("graphics/pr_iters.pdf", width = 4.5, height = 4.5)
 # Could be done with aggregate instead
 pr_mean_iters <- c(
+		mean(pr_iters$Time[pr_iters$Sys == "GAP"]),
 		mean(pr_iters$Time[pr_iters$Sys == "GraphBIG"]),
-		mean(pr_iters$Time[pr_iters$Sys == "GraphMat"]),
-		mean(pr_iters$Time[pr_iters$Sys == "GAP"]))
+		mean(pr_iters$Time[pr_iters$Sys == "GraphMat"]))
 barplot(pr_mean_iters, ylab = "Time (second)",
 		main = "PageRank Iterations", col=c("green","orange","cyan"),
-		names.arg = unique(pr_iters$Sys))
+		names.arg = c("GAP","GraphBIG","GraphMat"))
 dev.off()
 
 ###
@@ -141,7 +144,7 @@ dev.off()
 # Just BFS for now... :(
 # Read in the data
 GRAPH500NRT <- 64 # Even though everyone else does 32.
-scale <- 16
+scale <- 22
[…]

### 2017-02-15 papers/publication/paper-pollard.tex
Added more tables to the plot_data and paper.
Still need to get the paper with EPG scale 22 and real world
datasets.

@@
 
-We consider three algorithms, but not all algorithms are available on all systems.
+We consider three algorithms though not all algorithms are available on all systems.
 
@@
 
+\subsection{Runtime}
+
 In Table~\ref{tab:graphalytics} we show the results from running Graphalytics. For an explanation of each algorithm used, see \cite{Iosup:2016:Graphalyticstech}.
 
-From this we get only a broad overview of the two systems: GraphBIG and PowerGraph.
+Table~\ref{tab:graphalytics} shows only a broad overview of the two systems: GraphBIG and PowerGraph.
 In general, we see GraphBIG is more performant.
@@
 \begin{table}
+% ,graphmat,openg,powergraph
+%	Community Detection,0,212.5,1167
+%	PageRank,0,293.5,983
+%	Local Clustering Coefficient,0,316.5,1011
+%	Weakly Connected Components,0,89,767.5
+%	Single-Source Shortest Paths,11298.5,6378.5,34654.5
+	\caption{Performance results are in milliseconds. Community detection is performed using label propagation. At the time of this writing, Graphalytics only supports SSSP for GraphMat.}
 	\centering
@@
 	]{../report/runtime.csv}
-	\caption{Performance results are in milliseconds.
-LCC is local clustering coefficient, PR is PageRank, CDLP is community detection using label propagation, and WCC is weakly connected components.
-At the time of this writing, Graphalytics only supports SSSP for GraphMat.}
 	\label{tab:graphalytics}
@@
 
-Table~\ref{tab:epg-perf}, we show the performance results of running graphalytics and our approach using the same, synthetic dataset.
-Now, our approach computes multiple runs whereas graphalytics only uses one root for BFS.
-This gives a better picture of algorithm reliability and is explained further.
+Table~\ref{tab:epg-perf}, in contrast, gives an overview of the performance results of running graphalytics and our approach using the same dataset [TODO: GET DOTA LEAGUE WORKING FOR EPG!].
+For compactness, Table~\ref{tab:epg-perf} shows the average time across 32 roots, while Table~\ref{tab:graphalytics} shows results for a single root.
+Figures~\ref{fig:bfs-time} and \ref{fig:sssp-time} show the distribution of runtimes.
+However, even these do not give a complete picture of performance.
+Both of these show GraphMat to not be highly performant.
+However, a likely explanation is that GraphMat's underlying computation model (sparse matrix operations) paired the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation is more conducive to larger-scale graphs.
+
+\begin{figure}
+	\centering
+	\caption{The $y$-axis is logarithmic. GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_time.pdf}
+	\end{minipage}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_dsc.pdf}
+	\end{minipage}
+	\label{fig:bfs-time}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\caption{The $y$-axis is logarithmic. [TODO: FIX IT SO GRAPHBIG HAS NO DSC]}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/sssp_time.pdf}
+	\end{minipage}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/sssp_dsc.pdf}
+	\end{minipage}
+	\label{fig:sssp-time}
+\end{figure}
 
@@
 	\centering
+	\caption{The above table shows times for $2^{20}$ verties and the times are in seconds.
+The Graph500 generates the graph instead of loading it into a file.
+GraphBIG builds the graph and reads in the file simultaneously.
+These results were averaged across 32 roots.}
 	\begin{tabular}{l|c|c|c}
@@
 	\end{tabular}
-	\caption{The above table shows times for $2^{20}$ verties and the times are in seconds.
-The Graph500 generates the graph instead of loading it into a file.
-GraphBIG builds the graph and reads in the file simultaneously.
-These results were averaged across 32 roots.}
 	\label{tab:epg-perf}
@@
 
+\begin{figure}
+	\centering
+	\caption{Both figures use the kronecker graph with $2^{20}$ vertices with an average of 16 edges per vertex.}
+	\begin{minipage}{0.55\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_speedup.pdf}
+	\end{minipage}
+	\begin{minipage}{0.425\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ss.pdf}
+	\end{minipage}
+	\label{fig:bfs-scaling}
+\end{figure}
+
+\subsection{Power and Energy Consumption}
 We use the Performance Application Programming Interface (PAPI) \cite{Browne:2000:PAPI} to gain access to Intel's Running Average Power Limit (RAPL), which provides access to a set of hardware counters measuring energy usage.
+PAPI provides convenient access to these values and provides average energy in nanojoules for a given time interval.
+
+\begin{table}
+	\caption{The timeframe measured here is only when the BFS is performed.
+Sleeping Energy refers to the power (in Watts) consumed during the Unix \texttt{sleep} function, multiplied by time.
+Effectively, this measures the amount of energy that would have been consumed even if nothing was running.
+The increase over sleep is the ratio of the first and third columns.
+These are all averaged over the 32 roots.}
+	\centering
+	\begin{tabular}{l|c|c|c|c}
+			&	GAP  &    Graph500 & GraphMat & GraphBIG \\ \hline
+		Energy Per Root (J) &	1.184 & 1.830 & 112.213 & 111.104 \\
+		Time (s) &  0.01636 & 0.01884 & 1.600 & 1.424 \\
+		Sleeping Energy (J) & 0.4046  & 0.4660 & 39.591 &  35.234 \\
+		Increase over Sleep & 2.926 & 3.928 & 2.834 & 3.153
+	\end{tabular}
+\end{table}
 
@@
 	\centering
-	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_time.pdf}
-	\end{minipage}
-	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_dsc.pdf}
-	\end{minipage}
-	\caption{The $y$-axis is logarithmic. GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
-	\label{fig:bfs-time}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/sssp_time.pdf}
-	\caption{The $y$-axis is logarithmic.}
-	\label{fig:sssp-time}
-\end{figure}
-
-\begin{figure}
-	\centering
+	\caption{The $y$-axis is logarithmic.
+GraphMat continues to run until none of the vertices' ranks change.
+[There is only one data point] The $y$-axis \emph{not} logarithmic.
+GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.
+[Note that for figure Fig.\ref{fig:pr}, mention time per iteration]}
 	\begin{minipage}{0.48\linewidth}
@@
 	\end{minipage}
-	\caption{The $y$-axis is logarithmic.
-GraphMat continues to run until none of the vertices' ranks change.
-[There is only one data point] The $y$-axis \emph{not} logarithmic.
-GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.
-[Note that for figure Fig.\ref{fig:pr-iters}, mention time per iteration]}
 	\label{fig:pr}
@@
 
-[Compare lines of code for implementation]
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ss.pdf}
-	\caption{Strong Scaling for $2^{20}$ vertices and an average of 16 edges per vertex.}
-	\label{fig:bfs-ss}
-\end{figure}
-
-\begin{figure}
-	\centering
-	% trim left bottom right top
-	\includegraphics[width=\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_speedup.pdf}
-	\caption{Speedup.}
-	\label{fig:bfs-speedup}
-\end{figure}
-
-%\section{Graph Processing Taxonomy}
-%This is in the spirit of \cite{Doekemeijer:2015:GPFSurvey}.
-Here, ``|'' means ``or'' and ``+'' means ``and.'' FOSS means Free and Open Source Software.
-The quotes around ``yes'' for HPC mean that the product claims to be amenable to high performance computing.
-Whether these actually achieve their goal is one of the purposes of this project.
-%\begin{table*}[t]
-%	\begin{minipage}{\linewidth} % So the footnotes all get printed on the same page
-%		\centering
-%		%\small
-%		\pgfplotstabletypeset[
-%			col sep=comma,
-%			string type, % Makes the .style={string type} redundant
-%			columns={Name,Type,Parallelism,Target,FOSS,Source,Notes},
-%			every head row/.style={after row=\midrule},
-%			columns/Name/.style={string type, column type={l|}},
-%			columns/Type/.style={string type},
-%			% columns/HPC/.style={string type},
-%			columns/Parallelism/.style={string type},
-%%			columns/Comm./.style={%
-%%				string type,
-%%				postproc cell content/.style={%
-%%					@cell content/.add={\footnotesize}
-%%				}
-%%			},
-%			columns/Target/.style={string type},
-%			columns/FOSS/.style={string type},
-%			columns/Source/.style={string type},
-%			columns/Notes/.style={
-%				preproc cell content/.style={@cell content=
-%					\ifx&##1&% Only make a footnote if the cell is nonempty
-%						##1
-%					\else
-%						\footnote{##1}
-%					\fi}
-%			},
-%		]{../report/platforms.csv}
-%		\caption{Tools used for graph processing}
-%		\label{tab:frameworks}
-%	\end{minipage}
-%\end{table*}
-
-\section{Conclusion} % Begin with the end in mind...
-
+\section{Future Work and Conclusion}
+The problem of choosing a system for graph processing at large scale is far from being simple.
+We make our source code open source at \url{github.com/HPCL/easy-parallel-graph} and encourage further experimentation.
 \bibliographystyle{splncs03}

### 2017-02-15 papers/publication/plot_data.R

@@ -35,23 +35,25 @@ pr_dsc$Sys <- factor(pr_dsc$Sys)
 pdf("graphics/bfs_time.pdf", width = 5.2, height = 5.2)
 boxplot(Time~Sys, bfs_time, ylab = "Time (seconds)",
 		main = "BFS Time", log = "y", col=bpc)
-mtext(paste0("Scale = ",scale), side = 3) # May want to remove subtitle later
+mtext(paste0("Scale = ",scale), side = 3)
 dev.off()
 
 pdf("graphics/bfs_dsc.pdf", width = 5.2, height = 5.2)
 boxplot(Time~Sys, bfs_dsc, ylab = "Time (seconds)",
 		main = "BFS Data Structure Construction", col=bpc, log = "y")
-mtext(paste0("Scale = ",scale), side = 3) # May want to remove subtitle later
+mtext(paste0("Scale = ",scale), side = 3)
 dev.off()
 
 pdf("graphics/sssp_time.pdf", width = 4.5, height = 4.5)
 boxplot(Time~Sys, sssp_time, ylab = "Time (seconds)",
 		main = "SSSP Time", log = "y", col=bpc)
+mtext(paste0("Scale = ",scale), side = 3)
 dev.off()
 
 pdf("graphics/sssp_dsc.pdf", width = 4.5, height = 4.5)
 boxplot(Time~Sys, sssp_time, ylab = "Time (seconds)",
 		main = "SSSP Data Structure Construction", log = "y", col=bpc)
+mtext(paste0("Scale = ",scale), side = 3)
 dev.off()
 
 pdf("graphics/pr_time.pdf", width = 4.5, height = 4.5)
@@ -164,23 +166,22 @@ bfs_cpu_nrg$Sys <- factor(bfs_cpu_nrg$Sys)
 bfs_systems <- unique(bfs_cpu_nrg$Sys)
 bfs_cpu_nrg_per_root <- numeric(length(bfs_systems))
 sleep_nrg_per_root <- numeric(length(bfs_systems))
+bfs_time_per_root <- numeric(length(bfs_systems))
 for (si in seq(length(bfs_systems))) {
 	sys <- as.character(unique(bfs_cpu_nrg$Sys)[si])
 	one_sys <- subset(x,
-			x$Algo=="BFS" & x$Metric=="Total CPU Energy (J)" & x$Sys==sys,
+			x$Algo=="BFS" & x$Metric=="Average CPU Power (W)" & x$Sys==sys,
 			Value)
[…]

### 2017-02-15 papers/publication/paper-pollard.tex
Improved bfs,sssp figures and conclusion.
Still waiting on the rest of the scale-22 results and to get
dota-league dataset working for easy-parallel-graph.

@@
 With a plugin to Graphalytics called Granula \cite{Ngai:2015:Granula} one can explicitly specify a performance model to analyze specific execution behavior such as the amount of communication or runtime of particular kernels of execution.
-The requires in-depth knowledge of the source code and execution model which is not always available.
+This requires in-depth knowledge of the source code and execution model.
 Furthermore, creating such a model requires a high level of expertise with the given system and with Granula\footnote{An example of Granula can be seen at \url{https://github.com/tudelft-atlarge/graphalytics-platforms-graphx/tree/master/granula-model-graphx}}.
@@
 	\centering
-	% Keep in mind you can do this at the beginning: string replace={s1}{s2}
+	\caption{The operating system is GNU/Linux version 4.4.0-22.
+The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
+We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.}
 	% For arya I deleted Max RAM Freq	2133MHz
@@
 	]{../report/specs.csv}
-	\caption{The operating system is GNU/Linux version 4.4.0-22.
-The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
-We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.}
 	\label{tab:specs}
@@
 
-In Table~\ref{tab:graphalytics} we show the results from running Graphalytics. For an explanation of each algorithm used, see \cite{Iosup:2016:Graphalyticstech}.
-
+In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a single dataset.
+For an explanation of each algorithm used, see \cite{Iosup:2016:Graphalyticstech}.
 Table~\ref{tab:graphalytics} shows only a broad overview of the two systems: GraphBIG and PowerGraph.
@@
 However, results such as these are preliminary and do not show the complete picture.
-It is difficult to make any claims about GraphMat from Graphalytics because of its lack of algorithm diversity.
 
@@
 
-Table~\ref{tab:epg-perf}, in contrast, gives an overview of the performance results of running graphalytics and our approach using the same dataset [TODO: GET DOTA LEAGUE WORKING FOR EPG!].
-For compactness, Table~\ref{tab:epg-perf} shows the average time across 32 roots, while Table~\ref{tab:graphalytics} shows results for a single root.
-Figures~\ref{fig:bfs-time} and \ref{fig:sssp-time} show the distribution of runtimes.
+Table~\ref{tab:epg-perf}, in contrast, gives an overview of the performance results of running graphalytics and our approach using the same dataset [TODO: Get the Dota-League Dataset working with eas-parallel-graph and replace Table~\ref{epg-perf}].
+Table~\ref{tab:epg-perf} shows the average time across 32 roots, whereas Graphalytics by default shows results for a single root.
+
+Further detail is shown in Figs.~\ref{fig:bfs-time} and \ref{fig:sssp-time}, giving box plots for the runtime distributions.
 However, even these do not give a complete picture of performance.
@@
 However, a likely explanation is that GraphMat's underlying computation model (sparse matrix operations) paired the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation is more conducive to larger-scale graphs.
-
-\begin{figure}
-	\centering
-	\caption{The $y$-axis is logarithmic. GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
-	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_time.pdf}
-	\end{minipage}
-	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_dsc.pdf}
-	\end{minipage}
-	\label{fig:bfs-time}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\caption{The $y$-axis is logarithmic. [TODO: FIX IT SO GRAPHBIG HAS NO DSC]}
-	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/sssp_time.pdf}
-	\end{minipage}
-	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/sssp_dsc.pdf}
-	\end{minipage}
-	\label{fig:sssp-time}
-\end{figure}
 
@@
 	\centering
+	\caption{The $y$-axis is logarithmic. GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_time.pdf}
+	\end{minipage}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_dsc.pdf}
+	\end{minipage}
+	\label{fig:bfs-time}
+\end{figure}
+f
+\begin{figure}
+	\centering
+	\caption{The $y$-axis is logarithmic. Both PowerGraph and GraphBIG construct their data structures at the same time as they read in the file.}
+	\begin{minipage}{0.59\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/sssp_time.pdf}
+	\end{minipage}
+	\begin{minipage}{0.365\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/sssp_dsc.pdf}
+	\end{minipage}
+	\label{fig:sssp-time}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\caption{The $y$-axis is logarithmic.
+GraphMat continues to run until none of the vertices' ranks change.
+[There is only one data point] [FIX IT SO POWERGRAPH IS INCLUDED] The $y$-axis \emph{not} logarithmic.
+GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.
+For the others, we use the stopping condition that the sum of the changes in the weights is no more than $6 \times 10^{-8}$, or approximately machine epsilon for single precision floating point numbers.
+[Note that for figure Fig.\ref{fig:pr}, mention time per iteration]}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/pr_time.pdf}
+	\end{minipage}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/pr_iters.pdf}
+	\end{minipage}
+	\label{fig:pr}
+\end{figure}
+
+\begin{figure}
+	\centering
 	\caption{Both figures use the kronecker graph with $2^{20}$ vertices with an average of 16 edges per vertex.}
@@
 \end{figure}
-
-\subsection{Power and Energy Consumption}
-We use the Performance Application Programming Interface (PAPI) \cite{Browne:2000:PAPI} to gain access to Intel's Running Average Power Limit (RAPL), which provides access to a set of hardware counters measuring energy usage.
-PAPI provides convenient access to these values and provides average energy in nanojoules for a given time interval.
-
-\begin{table}
-	\caption{The timeframe measured here is only when the BFS is performed.
-Sleeping Energy refers to the power (in Watts) consumed during the Unix \texttt{sleep} function, multiplied by time.
-Effectively, this measures the amount of energy that would have been consumed even if nothing was running.
-The increase over sleep is the ratio of the first and third columns.
-These are all averaged over the 32 roots.}
-	\centering
-	\begin{tabular}{l|c|c|c|c}
-			&	GAP  &    Graph500 & GraphMat & GraphBIG \\ \hline
-		Energy Per Root (J) &	1.184 & 1.830 & 112.213 & 111.104 \\
-		Time (s) &  0.01636 & 0.01884 & 1.600 & 1.424 \\
-		Sleeping Energy (J) & 0.4046  & 0.4660 & 39.591 &  35.234 \\
-		Increase over Sleep & 2.926 & 3.928 & 2.834 & 3.153
-	\end{tabular}
-\end{table}
 
@@
 
+\subsection{Power and Energy Consumption}
+We use the Performance Application Programming Interface (PAPI) \cite{Browne:2000:PAPI} to gain access to Intel's Running Average Power Limit (RAPL), which provides access to a set of hardware counters measuring energy usage.
+PAPI provides convenient access to these values and average energy in nanojoules for a given time interval.
+We modify the source code for each project to time only the actual BFS computation.
+In our case, the fastest code is also the most energy efficient, though with this level of granularity we could detect circumstances where one could make a tradeoff between energy and runtime.
+
+\begin{table}
+	\caption{The data are generated from a Kronecker graph with $2^{22}$ vertices.
+Sleeping Energy refers to the power (in Watts) consumed during the Unix \texttt{sleep} function, multiplied by time.
+Effectively, this measures the amount of energy that would have been consumed even if nothing was running.
+The increase over sleep is the ratio of the first and third columns.
+These are all averaged over the 32 roots.}
+	\centering
+	\begin{tabular}{l|r|r|r|r}
+			&	GAP  &    Graph500 & GraphBIG & GraphMat \\ \hline
+		Time (s) &  0.01636 & 0.01884 & 1.600 & 1.424 \\
+		Average Power per Root (W) & 72.38 & 97.17 & 78.01 & 70.12 \\
+		Energy per Root (J) &	1.184 & 1.830 & 112.213 & 111.104 \\
+		Sleeping Energy (J) & 0.4046  & 0.4660 & 39.591 &  35.234 \\
+		Increase over Sleep & 2.926 & 3.928 & 2.834 & 3.153
+	\end{tabular}
+\end{table}
+
 \section{Future Work and Conclusion}
 The problem of choosing a system for graph processing at large scale is far from being simple.
-We make our source code open source at \url{github.com/HPCL/easy-parallel-graph} and encourage further experimentation.
+We make our code available at \url{https://github.com/HPCL/easy-parallel-graph} and encourage further experimentation.
+Our results required minor changes to the original projects to add calls to PAPI sampling and those changes are also available on forked versions of the repositories.
+
+Overall, the GAP Benchmark Suite was the highest-performing system across the given datasets and the most scalable.
+However, this is only for relatively small graphs; all the graphs used here had at most $2^{22}$ vertices.
+It should be noted that GAP is also the most recent of projects.
+Thus, we recommend graph processing algorithm designers compare their implementations against the GAP Benchmark Suite to as a good test of performance.
+
+There are a vast number of directions this work could go.
+To ensure fairness, each platform must be configured to use the same graphs and the same roots.
+In the case of GraphBIG this required the file to be loaded in for each experiment.
+This was by far the most time consuming of the experiments; the file reading for GraphBIG being done serially on an uncompressed ASCII text format for graphs.
+This limited the sizes of the experiments we could run.
+Graphalytics also had circumstances with the more computationally expensive algorithms  where certain experiments fail \cite{Iosup:2016:Graphalyticstech}, so determining whether an algorithm will finish given a particular machine, input size, runtime limit, and resources is an important unanswered question.
+
 \bibliographystyle{splncs03}

### 2017-02-15 papers/publication/plot_data.R

@@ -4,7 +4,7 @@
 ###
 # Part 1: Generate the plots for a single problem size and multiple algorithms
 ###
-scale <- 21
+scale <- 22
 bpc <- "cyan"
 filename <- paste0("parsed",scale,"-32.csv")
 x <- read.csv(filename, header = FALSE)
@@ -44,16 +44,16 @@ boxplot(Time~Sys, bfs_dsc, ylab = "Time (seconds)",
 mtext(paste0("Scale = ",scale), side = 3)
 dev.off()
 
-pdf("graphics/sssp_time.pdf", width = 4.5, height = 4.5)
+pdf("graphics/sssp_time.pdf", width = 5.5, height = 4.5)
 boxplot(Time~Sys, sssp_time, ylab = "Time (seconds)",
 		main = "SSSP Time", log = "y", col=bpc)
 mtext(paste0("Scale = ",scale), side = 3)
 dev.off()
 
-pdf("graphics/sssp_dsc.pdf", width = 4.5, height = 4.5)
-boxplot(Time~Sys, sssp_time, ylab = "Time (seconds)",
-		main = "SSSP Data Structure Construction", log = "y", col=bpc)
-mtext(paste0("Scale = ",scale), side = 3)
+pdf("graphics/sssp_dsc.pdf", width = 3.5, height = 4.5)
+boxplot(Time~Sys, sssp_dsc, ylab = "Time (seconds)",
+		main = "SSSP Data Structure\nConstruction", log = "y", col=bpc)
+# mtext(paste0("Scale = ",scale), side = 3)
 dev.off()
 
 pdf("graphics/pr_time.pdf", width = 4.5, height = 4.5)
@@ -76,25 +76,30 @@ dev.off()
 ###
 # Part 2: Generate the plots for a single algorithm and multiple problem sizes
 ###
-scale <- 20
-# Read in and average the data for BFS for each thread
-# It is wasteful to reread the parsed*-1.csv but it simplifies the code
 threadcnts <- c(1,2,4,8,16,32,64,72)
-x <- read.csv(paste0("parsed",scale,"-1.csv"), header = FALSE)
-colnames(x) <- c("Sys","Algo","Metric","Time")
-systems <- unique(x$Sys)
-bfs_scale <- data.frame(
-		matrix(ncol = length(threadcnts), nrow = length(systems)),
-		row.names = systems)
[…]

### 2017-02-16 get-graphalytics.sh
Fixed issues with get-graphalytics.

@@ -24,6 +24,7 @@ set +o histexpand
 trap "exit 1" INT
 export TOP_PID=$$ 
 MAX_THREADS=64 # This is a PowerGraph limitation.
+MAX_THREADS=32 # To ensure compatibility with easy-parallel-graph
 
 OSNAME=$(uname)
 if [ "$JAVA_HOME" = "" -a "$OSNAME" = "Linux" ]; then
@@ -211,7 +212,7 @@ install_GraphX()
 # 	if [ "$EXEC_MEMORY_KB" -gt 7448928 ]; then # Yarn doesn't allow more than 8GB per executor but has overhead
 # 		EXEC_MEMORY_KB=7448928
 # 	fi
-	cp -r "$PKGDIR/config-template" "$PKGDIR/config"
+	cp -r "$BASE_DIR/ldbc_graphalytics/config" "$PKGDIR/config"
 	echo "graphx.job.num-executors = $NUM_WORKERS" > "$PKGDIR/config/graphx.properties"
 	#echo "graphx.job.executor-memory = ${EXEC_MEMORY_KB}k" >> "$PKGDIR/config/graphx.properties"	
 	echo "graphx.job.executor-memory = 6g" >> "$PKGDIR/config/graphx.properties"
@@ -263,7 +264,7 @@ install_OpenG()
  		echo "Using $NUM_THREADS threads on $NUM_CORES available cores to maintain an even comparison with PowerGraph."
  	fi
 	CONFIG=$(printf "openg.home = $GRAPHBIG_DIR\nopeng.intermediate-dir = $GRAPHBIG_DIR/intermediate\nopeng.output-dir = $GRAPHBIG_DIR/output\nopeng.num-worker-threads=$NUM_THREADS")
-	cp -r "$PKGDIR/config-template" "$PKGDIR/config"
+	cp -r "$BASE_DIR/ldbc_graphalytics/config" "$PKGDIR/config"
 	echo "$CONFIG" > $PKGDIR/config/$platform.properties
 	perl -0777 -i.original -pe "s?graphs.root-directory.*?graphs.root-directory = $DATASET_DIR?" "$PKGDIR/config/graphs.properties"
 }
@@ -357,8 +358,8 @@ install_PowerGraph()
     PKGDIR="$PLATFORM_DIR/graphalytics-$GA_VERSION-$platform-$VERSION"
 
 	# Set configuration in the packaged file
[…]

### 2017-02-16 papers/publication/paper-pollard.tex
Addded a bit to references, made plot.R better.

@@
 However, perusing the log files reveals a more complete picture: of these 5.96 seconds, 0.21 was spent computing the shortest paths while the remaining time consisted of building the necessary data structures.
-Furthermore, data such as the number of iterations run for PageRank is not easily available using Graphalytics.
+Furthermore, data such as the number of iterations run for PageRank are not easily available using Graphalytics.
 %We use GraphMat, the single source shortest paths algorithm, the dota-league dataset from the Game Trace Archive\cite{Guo:2012:GTA}, and the default settings as an example here.
@@
 Likewise, our approach provides automatic execution and performance analysis without requiring a performance and execution model for each system.
+
+Other related work includes that of \cite{Satish:2014:NavigatingGraph}, which analyzes performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} which uses six real-world datasets and focuses on the vertex-centric programming model.
 

### 2017-02-16 papers/publication/plot_data.R

@@ -193,14 +193,6 @@ bfs_ram_pwr <- subset(x, x$Algo == "BFS" & x$Metric == "Average DRAM Power (W)",
 		c("Sys","Value"))
 bfs_ram_pwr$Sys <- factor(bfs_cpu_pwr$Sys)
 
-# Make some plots
-pdf("graphics/bfs_cpu_power.pdf", width = 4.5, height = 4.5)
-boxplot(Value~Sys, bfs_cpu_pwr, ylab = "Average Power (Watts)",
-		col="yellow")
-title(main = "CPU Average Power Consumption During BFS")
-mtext(paste0("Scale = ",scale), side = 3) # May want to remove subtitle later
-dev.off()
-
 # We hope that sleeping uses less energy than running the BFS...
 stopifnot(all(sleep_nrg_per_root < bfs_cpu_nrg_per_root))
 intersperse <- function(vec, ele) { return(paste(vec, collapse = ele)) }
@@ -232,15 +224,26 @@ print(paste0("Increase over sleep & ",
 # 		fill = c("gold","orangered3"))
 # dev.off()
 
+# Make some plots
 # par(xpd = TRUE) and change inset if you want the legend to be outside the box
-pdf("graphics/bfs_ram_power.pdf", width = 4.5, height = 4.5)
+pdf("graphics/bfs_cpu_power.pdf", width = 5.2, height = 5.2)
+boxplot(Value~Sys, bfs_cpu_pwr, ylab = "Average Power (Watts)",
+		col="yellow", ylim=c(cpu_pwr_sleep$Value*0.9, max(bfs_cpu_pwr$Value)))
+title(main = "CPU Average Power Consumption During BFS")
+abline(mean(cpu_pwr_sleep$Value), 0, col = "orangered", lwd = 2)
+mtext(paste0("Scale = ",scale), side = 3) # May want to remove subtitle later
+legend(legend = c("sleep"), x = "bottomright", inset = c(0,0),
+		lty = c(1), lwd = 2, col = "orangered", bg = "white")
+dev.off()
+
+pdf("graphics/bfs_ram_power.pdf", width = 5.2, height = 5.2)
 boxplot(Value~Sys, bfs_ram_pwr, ylab = "Average Power (Watts)",
-		col="yellow")
+		col="yellow", ylim=c(ram_pwr_sleep$Value*0.9, max(bfs_ram_pwr$Value)))
 title(main = "RAM Power Consumption During BFS",
[…]

### 2017-02-16 experiment/parse-output.sh
Updated experiment scripts.
Now they work for PowerGraph, mostly and GraphBIG for RAPL.

@@ -62,3 +62,9 @@ awk -v NRT=$NRT '/Build Time:/{i++; if(i>2*NRT)print "GAP,PageRank,Data structur
 grep -B 2 'Average Time:' "$FN" | awk '/^\s*[0-9]+/{print "GAP,PageRank,Iterations," $1}'
 awk -v NRT=$NRT '/Average Time:/{i++; if(i>2*NRT)print "GAP,PageRank,Time," $3}' "$FN"
 
+
+# PowerGraph
+# The first NRT are SSSP, the next NRT PageRank.
+awk -v NRT=$NRT '/Finished Running engine/{i++; if(i<=NRT)print "PowerGraph,SSSP,Time," $5}' "$FN"
+awk -v NRT=$NRT '/Finished Running engine/{i++; if(i>NRT)print "PowerGraph,SSSP,Time," $5}' "$FN"
+

### 2017-02-16 experiment/run-experiment.sh

@@ -171,3 +171,4 @@ for ROOT in $(head -n $NRT "$DDIR/kron-${S}-roots.v"); do
 	"$POWERGRAPHDIR/release/toolkits/graph_analytics/pagerank" --graph "$DDIR/kron-${S}.el" --tol "$TOL" --format tsv
 done
 
+echo Finished experiment at $(date)

### 2017-02-16 experiment/run-power.sh

@@ -7,13 +7,15 @@
 
 # Set some parameters and environment variables
 # CHECK THAT THESE ARE CORRECT ON YOUR OWN SYSTEM!
+# NOTE: on Arya, graph500, GraphBIG, and GAP must be recompiled.
 module load intel/17
 module load papi/git
 DDIR= # Dataset directory
 GAPDIR=
 GRAPH500DIR=
 GRAPHMATDIR=
-S=16
+GRAPHBIGDIR=
+S=22
 NRT=32 # Number of roots that we did BFS on. GraphBIG has issues with >32.
 PKG=2 # The number of physical chips
 export OMP_NUM_THREADS=32
@@ -39,6 +41,11 @@ for ROOT in $(head -n $NRT "$DDIR/kron-${S}-roots.1v"); do
 	sudo bash -c "export LD_LIBRARY_PATH=$LD_LIBRARY_PATH; $GRAPHMATDIR/bin/BFS $DDIR/kron-${S}.graphmat $ROOT"
 	#"$GRAPHMATDIR/bin/BFS" "$DDIR/kron-${S}.graphmat" "$ROOT"
 done >> "$FN" 2>> "$ERRFN"
+# GraphBIG BFS
+for ROOT in $(head -n $NRT "$DDIR/kron-${S}-roots.v"); do
+	sudo "$GRAPHBIGDIR/benchmark/bench_BFS/bfs" --dataset "$DDIR/kron-${S}" --root $ROOT --threadnum $OMP_NUM_THREADS
+done >> "$FN" 2>> "$ERRFN"
+
 # Baseline (do nothing, just sleep)
 sudo "$GAPDIR"/sleep_baseline >> "$FN" 2>> "$ERRFN"
 sudo chown spollard "$FN"
@@ -57,6 +64,9 @@ grep -A 29 'RAPL on Graph500 BFS' "$FN" | awk -v PKG=$PKG '/Average.*PACKAGE_ENE
 # GraphMat
 grep -A 29 'RAPL on GraphMat BFS' "$FN" | awk -v PKG=$PKG '/Average.*PACKAGE_ENERGY:PACKAGE[0-9]+ \*/{c++;if(c%PKG==0){print "GraphMat,BFS,Average CPU Power (W)," t;t=0}else{t+=$3}}' >> "$PFN"
[…]

### 2017-02-16 get-gap.sh

@@ -1,11 +1,15 @@
 #!/bin/bash
+# Compiles and runs the experiments for power measurement
+# NOTE: This should just be used for building
+# and run-power for experiment running.
+RUN_EXPERIMENTS=false
 
 # Set and check where everything is going to be stored
 # For SamXu2
 #BASE_DIR="$HOME/uo/research/easy-parallel-graph"
 #DATASET_DIR="$HOME/uo/research/datasets"
 # For Arya
-BASE_DIR="$HOME/gap"
+BASE_DIR="$HOME/power"
 DATASET_DIR="$HOME/graphalytics/all-datasets"
 if [ -z "$BASE_DIR" ]; then
 	echo "Please set BASE_DIR to be where all the benchmarks and output will be stored."
@@ -23,12 +27,47 @@ if [ ! -d "$BASE_DIR/gapbs" ]; then
 fi
 GAPBS_DIR="$BASE_DIR/gapbs"
 cd "$GAPBS_DIR"
+sed -i 's/BUILD_RAPL = No/BUILD_RAPL = Yes/' Makefile
 make
 make test
 if [ $? -ne 0 ]; then
 	echo "Some gapbs tests didn't pass"
 	exit 1
 fi
+
+# Download and compile GraphBIG Benchmark Suite
+echo "Preparing GraphBIG..."
+cd "$BASE_DIR"
+if [ ! -d graphBIG ]; then
+	git clone "https://github.com/HPCL/graphBIG.git"
+fi
+cd graphBIG/benchmark
+sed -i 's/BUILD_RAPL = No/BUILD_RAPL = Yes/' common.mk
+make clean all
+
+# Download and compile Graph500
+echo "Preparing GraphBIG..."
+cd "$BASE_DIR"
+if [ ! -d graphBIG ]; then
+	git clone https://github.com/sampollard/graph500.git
+fi
+cd graph500
+cp make-incs/make.inc-gcc make.inc
+sed -i 's/# BUILD_OPENMP = Yes/BUILD_OPENMP = Yes/' make.inc
+sed -i 's/# CFLAGS_OPENMP = -fopenmp/CFLAGS_OPENMP = -fopenmp/' make.inc
+sed -i 's/BUILD_RAPL = No/BUILD_RAPL = Yes/' Makefile
+sed -i 's/gcc-4\.6/gcc/' make.inc
+make
+
+# Download and compile GraphMat
+echo "Preparing GraphMat..."
+cd "$BASE_DIR"
+if [ ! -d GraphMat ]; then
[…]

### 2017-02-16 experiment/parse-output.sh
Parse error file to get PowerGraph iterations.

@@ -3,9 +3,12 @@
 # Assumes the experiment is run $NRT times ($NRT roots or $NRT PageRanks)
 NRT=32
 FN=$1
+ERRFN=$2 # Needed for PowerGraph
 if [ -z "$FN" ]; then
 	echo "usage: parse-output.sh <output_file>"
 	exit 2
+elif [ -z "$ERRFN" ]; then
+	>&2 echo "WARNING: No error filename specified. May cause issues with PowerGraph."
 fi
 
 # Graph500
@@ -66,5 +69,7 @@ awk -v NRT=$NRT '/Average Time:/{i++; if(i>2*NRT)print "GAP,PageRank,Time," $3}'
 # PowerGraph
 # The first NRT are SSSP, the next NRT PageRank.
 awk -v NRT=$NRT '/Finished Running engine/{i++; if(i<=NRT)print "PowerGraph,SSSP,Time," $5}' "$FN"
-awk -v NRT=$NRT '/Finished Running engine/{i++; if(i>NRT)print "PowerGraph,SSSP,Time," $5}' "$FN"
+awk -v NRT=$NRT '/iterations completed/{i++; if(i<=NRT)print "PowerGraph,SSSP,Iterations," $(NF-2)}' "$ERRFN"
+awk -v NRT=$NRT '/Finished Running engine/{i++; if(i>NRT)print "PowerGraph,PageRank,Time," $5}' "$FN"
+awk -v NRT=$NRT '/iterations completed/{i++; if(i>NRT)print "PowerGraph,PageRank,Iterations," $(NF-2)}' "$ERRFN"
 

### 2017-02-16 papers/publication/paper-pollard.tex
Update pagerank figs, still waiting on PowerGraph
Added some more commentary in the results section.

@@
 
-\begin{table}
-	\centering
-	\caption{The above table shows times for $2^{20}$ verties and the times are in seconds.
+%\begin{table}
+%	\centering
+%	\begin{tabular}{l|c|c|c}
+%		System & Load Graph & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular} & Run BFS \\ \hline
+%		Graph500 & 0.3474 & 0.3971 & 0.003380 \\
+%		GAP      & 2.351  & 0.1935 & 0.001393 \\
+%		GraphMat & 0.1511 & 1.101  & 0.1031 \\
+%		GraphBIG & 37.22  & N/A    & 0.1528 \\
+%	\end{tabular}
+%	\caption{This table shows times for $2^{20}$ verties and the times are in seconds.
 The Graph500 generates the graph instead of loading it into a file.
@@
 These results were averaged across 32 roots.}
-	\begin{tabular}{l|c|c|c}
-		System & Load Graph & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular} & Run BFS \\ \hline
-		Graph500 & 0.3474 & 0.3971 & 0.003380 \\
-		GAP      & 2.351  & 0.1935 & 0.001393 \\
-		GraphMat & 0.1511 & 1.101  & 0.1031 \\
-		GraphBIG & 37.22  & N/A    & 0.1528 \\
-	\end{tabular}
-	\label{tab:epg-perf}
-\end{table}
+%	\label{tab:epg-perf}
+%\end{table}
 
@@
 \end{figure}
-f
+
 \begin{figure}
 	\centering
-	\caption{The $y$-axis is logarithmic. Both PowerGraph and GraphBIG construct their data structures at the same time as they read in the file.}
 	\begin{minipage}{0.59\linewidth}
@@
 	\end{minipage}
+	\caption{The $y$-axis is logarithmic. Both PowerGraph and GraphBIG construct their data structures at the same time as they read in the file.}
 	\label{fig:sssp-time}
+\end{figure}
+
+The behavior of PageRank is slightly different.
+As with SSSP and BFS, the GAP Benchmark Suite is the fastest, and it also requires the fewest iterations.
+While care was taken to ensure the same stopping criterion is met, the particular implementation methods make it impossible to ensure the same computations are taking place.
+In general, we measured $\sum_{k=1}^{n} |p_k^{(i)} - p_k^{(i-1)}|$ where $i$ is the iteration and $n$ is the number of vertices, both the vertex-centric approaches, PowerGraph and GraphMat, take more iterations.
+
+\begin{figure}
+	\centering
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/pr_time.pdf}
+	\end{minipage}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/pr_iters.pdf}
+	\end{minipage}
+	\caption{The $y$-axis is logarithmic only for the left figure.
+GraphMat continues to run until none of the vertices' ranks change.
+GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.
+For the others, we use the stopping condition that the sum of the changes in the weights is no more than $6 \times 10^{-8}$, or approximately machine epsilon for single precision floating point numbers.}
+	\label{fig:pr}
 \end{figure}
@@
 	\centering
-	\caption{The $y$-axis is logarithmic.
-GraphMat continues to run until none of the vertices' ranks change.
-[There is only one data point] [FIX IT SO POWERGRAPH IS INCLUDED] The $y$-axis \emph{not} logarithmic.
-GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.
-For the others, we use the stopping condition that the sum of the changes in the weights is no more than $6 \times 10^{-8}$, or approximately machine epsilon for single precision floating point numbers.
-[Note that for figure Fig.\ref{fig:pr}, mention time per iteration]}
 	\begin{minipage}{0.48\linewidth}
@@
 	\end{minipage}
+	\caption{[TODO: Rerun PowerGraph] The $y$-axis is logarithmic.
+GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.}
 	\label{fig:pr}
@@
 
+The challenge in comparing iteration counts for PageRank underscores an important challenge for any comparison of graph processing systems.
+The assumptions under which the various platforms operate can have a dramatic effect on the program.
+For example, the GAP Benchmark Suite stores vertex weights as 32-bit integers.
+However, other systems store them as floating point numbers.
+This may affect performance in addition to runtime behavior in cases where weights like $0.2$ are cast to $0$.
+Similarly, how a grpah is represented in the system (e.g.
+weighed or directed) may have performance and algorithmic implications but is not always readily apparent.
+
+In Fig.~\ref{fig:bfs-scaling}, we see BFS scalability\dots
 \begin{figure}
 	\centering
-	\caption{Both figures use the kronecker graph with $2^{20}$ vertices with an average of 16 edges per vertex.}
 	\begin{minipage}{0.55\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_speedup.pdf}
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/bfs_speedup.pdf}
 	\end{minipage}
 	\begin{minipage}{0.425\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ss.pdf}
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/bfs_ss.pdf}
 	\end{minipage}
+	\caption{Both figures use the kronecker graph with $2^{20}$ vertices with an average of 16 edges per vertex.}
 	\label{fig:bfs-scaling}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\caption{The $y$-axis is logarithmic.
-GraphMat continues to run until none of the vertices' ranks change.
-[There is only one data point] The $y$-axis \emph{not} logarithmic.
-GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.
-[Note that for figure Fig.\ref{fig:pr}, mention time per iteration]}
-	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/pr_time.pdf}
-	\end{minipage}
-	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/pr_iters.pdf}
-	\end{minipage}
-	\label{fig:pr}
 \end{figure}

### 2017-02-16 papers/publication/plot_data.R

@@ -56,22 +56,32 @@ boxplot(Time~Sys, sssp_dsc, ylab = "Time (seconds)",
 # mtext(paste0("Scale = ",scale), side = 3)
 dev.off()
 
-pdf("graphics/pr_time.pdf", width = 4.5, height = 4.5)
-boxplot(Time~Sys, pr_time, ylab = "Time (seconds)",
-		main = "PageRank Time", log = "y", col=bpc)
-mtext(paste0("Scale = ",scale), side = 3) # May want to remove subtitle later
+pdf("graphics/pr_iters.pdf", width = 4.5, height = 4.5)
+pr_mean_iters <- aggregate(pr_iters$Time, list(pr_iters$Sys), mean)
+pr_sys_order <- order(pr_mean_iters[[2]])
+pr_mean_iters <- pr_mean_iters[pr_sys_order,]
+bp <- barplot(pr_mean_iters[[2]], ylab = "Time (second)",
+		main = "PageRank Iterations", col=rainbow(length(pr_mean_iters[[1]])))
+text(bp, par("usr")[3], labels = pr_mean_iters[[1]], srt = 30,
+		adj = c(0.95,0.95), xpd = TRUE, cex = 1.0)
 dev.off()
 
-pdf("graphics/pr_iters.pdf", width = 4.5, height = 4.5)
-# Could be done with aggregate instead
-pr_mean_iters <- c(
-		mean(pr_iters$Time[pr_iters$Sys == "GAP"]),
-		mean(pr_iters$Time[pr_iters$Sys == "GraphBIG"]),
-		mean(pr_iters$Time[pr_iters$Sys == "GraphMat"]))
-barplot(pr_mean_iters, ylab = "Time (second)",
-		main = "PageRank Iterations", col=c("green","orange","cyan"),
-		names.arg = c("GAP","GraphBIG","GraphMat"))
+pdf("graphics/pr_time.pdf", width = 4.5, height = 4.5)
+pr_sys_labels <- pr_mean_iters[[1]] # Get the order from iterations
+pr_time$Sys <- factor(pr_time$Sys, pr_sys_labels, ordered = TRUE)
+pr_time <- pr_time[order(pr_time$Sys),]
+pr_sys <- levels(pr_time$Sys)
+boxplot(Time~Sys, pr_time, ylab = "Time (seconds)",
+		main = "PageRank Time", log = "y", col=bpc,
+		xaxt = "n", xlab = "")
[…]

### 2017-02-16 papers/publication/paper-pollard.tex
Added the RAM figure (though not in repository).

@@
 	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/pr_time.pdf}
+		\includegraphics[width=\linewidth, trim=0 24pt 18pt 0, clip]{graphics/pr_time.pdf}
 	\end{minipage}
 	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/pr_iters.pdf}
+		\includegraphics[width=\linewidth, trim=0 24pt 18pt 0, clip]{graphics/pr_iters.pdf}
 	\end{minipage}
-	\caption{[TODO: Rerun PowerGraph] The $y$-axis is logarithmic.
-GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.}
+	\caption{The $y$-axis is logarithmic. GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.}
 	\label{fig:pr}
@@
 
+RAPL also allows the measurement of DRAM power and energy, whose results are shown in Fig.~\ref{fig:ram-power}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.6\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ram_power.pdf}
+	\caption{RAM Power Consumption, averaged across all roots. [TODO: Explain more]}
+	\label{fig:ram-power}
+\end{figure}
+
 \section{Future Work and Conclusion}

### 2017-02-16 papers/publication/paper-pollard.tex
minor reorg and todo
small move
Updated paper with more exposition and todos.

@@
 \usepackage{booktabs}
+\usepackage{todonotes}
 
@@
 
+Our approach differs from the previously mentioned appproaches because it focuses on the depth of analysis rather than the breadth via diversity of datasets or graph processing systems.
+We analyze  power and energy consumption in addition to details such as the number of iterations or the time to construct the data structures.
+Additionally, our approach requires little knowledge of the inner workings of each system; the data are collected using either sampling of model specific registers (for power) or parsing log files (for timing).
+
 % The systems described in \cite{Doekemeijer:2015:GPFSurvey} operate with a wide range of parallelism paradigms and target architectures such as GPU \cite{Zhong:2014:Medusa, Kang:2009:Pegasus}, shared memory CPU \cite{Shun:2013:Ligra, kyrola:2012:Graphchi, Nguyen:2013:Galois}, a combination of CPU and GPU \cite{Gharaibeh:2012:Totem}, distributed filesystem based approaches \cite{Xin:2013:GraphX}, and distributed memory with MPI \cite{Gregor:2005:PBGL}.
@@
 \section{Experimental Setup}
+We refer to a graphs's \emph{scale} when describing the size of the graph.
+Specifically, a graph with scale $S$ has $2^S$ vertices.
+The datasets are described further in Sect.~\ref{sec:datasets}.
+We measure strong scaling and speedup by varying the number of threads from one to the total number of threads on the reasearch machine, 72.
+
+Each experiment uses 32 roots per graph. As with the Graph500, each root is selected to have a degree greater than 1. In the case of PageRank, we simply run PageRank 32 times.
+
+When possible, we measure the data set construction time as the time to translate from the unstructured file data in RAM to the graph representation on which the algorithm can be performed.
+This is not possible for PowerGraph and GraphBIG because they read in the input file and build a graph simultaneously.
+The Graph500 builds the graph once and performs the BFS on each root in turn, speeding up experimentation but only giving a single data point for the data set construction.
+
+We plot many of the results as box plots with an implied 32 datapoints per box.
+
+\subsection{Machine Specifications}
+Table~\ref{tab:specs} shows the specifications of the research computer.
 
@@
 
-We consider three algorithms though not all algorithms are available on all systems.
+We consider three algorithms: Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank, though not all algorithms are implemented on all systems.
 
-\begin{enumerate}
-	\item Breadth First Search (BFS)
-	\item Single Source Shortest Paths (SSSP)
-	\item PageRank
-\end{enumerate}
-
-The canonical performance leaderboard for parallel graph processing is the Graph500 \cite{Murphy:2010:Graph500}.
+We select BFS because the canonical performance leaderboard for parallel graph processing is the Graph500 \cite{Murphy:2010:Graph500}.
 The advantage of the Graph500 is it provides standardized measurement specifications and dataset generation.
-The primary drawback with the Graph500 is it measures a single algorithm: BFS.
+The primary drawback with the Graph500 is it measures a single algorithm.
 
@@
 The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the actual BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}.
-We run the BFS using 32 random roots.
+We run the BFS using 32 random roots with the exception of PowerGraph which doesn't provide an reference implementation of BFS in its toolkits.
 
-One straightforward extension to BFS and our second algorithm is the Single-Source Shortest Paths algorithm (SSSP). We use the same graph and the same source vertices as in BFS.
+We select SSSP because of the straightforward extension from BFS; we need not modify the graph nor the root vertices from BFS.
 
-The third algorithm is PageRank. These three algorithms are used because of their popularity; most libraries provide reference implementations.
+PageRank is selected because of its popularity; most libraries provide reference implementations.
+One challenge with using PageRank is the stopping criterion; this is addressed in Sect.~\ref{sec:perf}.
+Verification of the PageRanks is beyond the scope of this paper though this may explain some of the large performance discrepancies.
 
-\subsection{Machine Specifications}
-Table~\ref{tab:specs} shows the specifications of the research computer.
+This approach is not specific to a particular algorithm; measuring the execution time, data structure construction time, and power consumption can be easily extended to other algorithms.
 
@@
 
-\subsection{Datasets}
-We use the Graph500 synthetic graph generator which creates a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19, C = 0.19,$ and $D = 1-(A+B+C) = 0.05$.
+\subsection{Datasets}\label{sec:datasets}
+We use the Graph500 synthetic graph generator which creates a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19, C = 0.19,$ and $D = 1-(A+B+C) = 0.05$ and set the average degree of a vertex as 16.
+Hence, a Kronecker graph with scale $S$ has $2^S$ vertices and approximately $16 \times 2^S$ edges.
 
@@
 This dataset contains 61,670 vertices and 50,870,313 edges.
-This dataset is sourced from the Game Trace Archive\cite{Guo:2012:GTA} and modified for Graphalytics\footnote{This dataset is available at \url{https://atlarge.ewi.tudelft.nl/graphalytics/}.}.
+This dataset comes from the Game Trace Archive\cite{Guo:2012:GTA} and is modified for Graphalytics\footnote{This dataset is available at \url{https://atlarge.ewi.tudelft.nl/graphalytics/}.}.
+This dataset is useful because it is both weighted and more dense than the usual real-world dataset with an average out-degree of 824.
 
-\section{Performance Analysis}
+\section{Performance Analysis}\label{sec:perf}
 
@@
 
-Table~\ref{tab:epg-perf}, in contrast, gives an overview of the performance results of running graphalytics and our approach using the same dataset [TODO: Get the Dota-League Dataset working with eas-parallel-graph and replace Table~\ref{epg-perf}].
+\todo{rework this. It would be nice to compare Graphalytics and our approach on the real-world dataset.}
+Table~\ref{tab:epg-perf}, in contrast, gives an overview of the performance results of running graphalytics and our approach using the same dataset [TODO: Get the Dota-League Dataset working with easy-parallel-graph and replace Table~\ref{epg-perf}].
 Table~\ref{tab:epg-perf} shows the average time across 32 roots, whereas Graphalytics by default shows results for a single root.
 
-Further detail is shown in Figs.~\ref{fig:bfs-time} and \ref{fig:sssp-time}, giving box plots for the runtime distributions.
+Figures~\ref{fig:bfs-time} and \ref{fig:sssp-time} show more detail, with box plots for the runtime distributions.
 However, even these do not give a complete picture of performance.
 Both of these show GraphMat to not be highly performant.
-However, a likely explanation is that GraphMat's underlying computation model (sparse matrix operations) paired the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation is more conducive to larger-scale graphs.
+One explanation is GraphMat's underlying computation model (sparse matrix operations) paired the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation is more conducive to larger-scale graphs.
+There is less discrepancy between the runtimes of SSSP (between 0.1 and 1.7 seconds) compared to BFS (0.01 and 1.7 seconds) but GAP is the clear winner in both cases.
+The data structure construction times for GAP and GraphMat are consistent; in both cases the platforms create the same data structure for both algorithms.
+
+These results are consistent with \cite{Sundaram:2015:GraphMat} which lists GraphMat as more performant than PowerGraph in SSSP.
 
@@
 
+\todo{Make the text bigger for these}
 \begin{figure}
 	\centering
-	\caption{The $y$-axis is logarithmic. GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
 	\begin{minipage}{0.48\linewidth}
@@
 	\end{minipage}
+	\caption{The $y$-axes are logarithmic.
+The left box plot shows the time to compute BFS on 32 random roots while the right plot shows the times to construct the graph for each system.
+The Graph500 only constructs its graph once.
+GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
 	\label{fig:bfs-time}
@@
 	\end{minipage}
-	\caption{The $y$-axis is logarithmic. Both PowerGraph and GraphBIG construct their data structures at the same time as they read in the file.}
+	\caption{The $y$-axes is logarithmic.
+The left box plot shows the time to compute the SSSP starting at the same 32 roots as Fig.~\ref{fig:bfs-time}.
+Both PowerGraph and GraphBIG construct their data structures at the same time as they read in the file.}
 	\label{fig:sssp-time}
@@
 
-The behavior of PageRank is slightly different.
-As with SSSP and BFS, the GAP Benchmark Suite is the fastest, and it also requires the fewest iterations.
-While care was taken to ensure the same stopping criterion is met, the particular implementation methods make it impossible to ensure the same computations are taking place.
-In general, we measured $\sum_{k=1}^{n} |p_k^{(i)} - p_k^{(i-1)}|$ where $i$ is the iteration and $n$ is the number of vertices, both the vertex-centric approaches, PowerGraph and GraphMat, take more iterations.
+The behavior of PageRank is slightly different. As with SSSP and BFS, the GAP Benchmark Suite is the fastest but it also requires the fewest iterations.
+
+We attempt to use similar stopping criteria for each system, but GraphMat executes until no vertices change rank; effectively its stopping criterion requires the $\infty$-norm be less than machine epsilon.
+This could account for the increased number of iterations.
+We adjusted the other systems to use $\sum_{k=1}^{n} |p_k^{(i)} - p_k^{(i-1)}| < \epsilon $ as the stopping criteria, where $i$ is the iteration and $n$ is the number of vertices.
+We use $\epsilon = 6 \times 10^{-8}$ because this value is approximately machine epsilon for a single precision floating point number.
+This is done with the goal of making the stoppping criteria as similar as possible.
+
+Furthermore, the logarithmic scale in Figs.~\ref{fig:pr} and~\ref{fig:bfs-time} belies the actual variance in results. Each platform in Fig.~\ref{fig:pr} has a standard deviation comparable to that of
+
+%> sd(pr_time$Time[pr_time$Sys == "PowerGraph"])
+%[1] 0.4654061
+%> sd(sssp_time$Time[sssp_time$Sys == "PowerGraph"])
+%[1] 0.1475608
+%> sd(pr_time$Time[pr_time$Sys == "GAP"])
+%[1] 0.01957317
+%> sd(sssp_time$Time[sssp_time$Sys == "GAP"])
+%[1] 0.01608644
+%> sd(pr_time$Time[pr_time$Sys == "GraphMat"])
+%[1] 0.9305763
+%> sd(sssp_time$Time[sssp_time$Sys == "GraphMat"])
+%[1] 0.02191714
+%> sd(pr_time$Time[pr_time$Sys == "GraphBIG"])
+%[1] 3.308351
+%> sd(sssp_time$Time[sssp_time$Sys == "GraphBIG"])
+%[1] 0.2165655
 
@@
 GraphMat continues to run until none of the vertices' ranks change.
-GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.
 For the others, we use the stopping condition that the sum of the changes in the weights is no more than $6 \times 10^{-8}$, or approximately machine epsilon for single precision floating point numbers.}
@@
 \end{figure}
+\todo{Check this again---that's a huge discrepancy!}
 
@@
 This may affect performance in addition to runtime behavior in cases where weights like $0.2$ are cast to $0$.
-Similarly, how a grpah is represented in the system (e.g.
+Similarly, how a graph is represented in the system (e.g.
 weighed or directed) may have performance and algorithmic implications but is not always readily apparent.
@@
 In Fig.~\ref{fig:bfs-scaling}, we see BFS scalability\dots
+
 \begin{figure}
 	\centering
-	\begin{minipage}{0.55\linewidth}
+	\begin{minipage}{0.9\linewidth}
 		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/bfs_speedup.pdf}
 	\end{minipage}
-	\begin{minipage}{0.425\linewidth}
+	\begin{minipage}{0.9\linewidth}
 		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/bfs_ss.pdf}
 	\end{minipage}
-	\caption{Both figures use the kronecker graph with $2^{20}$ vertices with an average of 16 edges per vertex.}
+	\caption{The top figure has both axes logarithmic. Linear speedup is. GraphMat dips below 1 because it is slower for $2$ threads than for 1.}
 	\label{fig:bfs-scaling}
 \end{figure}
+\todo{Rerun GraphMat on scale 22 with 2 threads. Surprising that it slows down}
 

### 2017-02-16 papers/publication/plot_data.R

@@ -94,19 +94,19 @@ measure_scale <- function(algo) {
 	x <- read.csv(paste0("parsed",scale,"-1.csv"), header = FALSE)
 	colnames(x) <- c("Sys","Algo","Metric","Time")
 	systems <- unique(subset(x$Sys, x$Algo == algo, c("Sys")))
-	bfs_scale <- data.frame(
+	algo_time <- data.frame(
 			matrix(ncol = length(threadcnts), nrow = length(systems)),
 			row.names = systems)
-	colnames(bfs_scale) <- threadcnts
+	colnames(algo_time) <- threadcnts
 	for (ti in seq(length(threadcnts))) {
 		thread <- threadcnts[ti]
 		Y <- read.csv(paste0("parsed",scale,"-",thread,".csv"),
 				header = FALSE)
-		bfs_time <- subset(Y, Y[[2]] == algo & Y[[3]] == "Time",
+		ti_time <- subset(Y, Y[[2]] == algo & Y[[3]] == "Time",
 				c(V1,V4))
-		bfs_scale[ti] <- aggregate(bfs_time$V4, list(bfs_time$V1), mean)[[2]]
+		algo_time[ti] <- aggregate(ti_time$V4, list(ti_time$V1), mean)[[2]]
 	}
-	return(bfs_scale)
+	return(algo_time)
 }
 
 bfs_scale <- measure_scale("BFS")
@@ -116,33 +116,38 @@ colors <- gsub("F", "C", colors) # You want it darker
 colors <- gsub("CC$", "FF", colors) # But keep it opaque
 
 # Plot the strong scalability for BFS
-pdf("graphics/bfs_ss.pdf", width = 4.5, height = 4.5)
+pdf("graphics/bfs_ss.pdf", width = 7, height = 4)
 bfs_ss <- bfs_scale
 # Strong scaling for sequential is 1---we compute that last
 for (ti in rev(seq(length(threadcnts)))) {
 	bfs_ss[ti] <- bfs_ss[1] / (threadcnts[ti] * bfs_ss[ti])
 }
 plot(as.numeric(bfs_ss[1,]), xaxt = "n", type = "b", ylim = c(0,1),
-		ylab = "Scalability", xlab = "Threads", col = colors[1],
+		ylab = "", xlab = "Threads", col = colors[1],
 		main = "BFS Strong Scaling", lty = 1, pch = 1, lwd = 3)
 for (pli in seq(2,nrow(bfs_ss))) {
[…]

### 2017-02-17 papers/publication/paper-pollard.tex
Updated some figures and caption text.

@@
 
-Furthermore, the logarithmic scale in Figs.~\ref{fig:pr} and~\ref{fig:bfs-time} belies the actual variance in results. Each platform in Fig.~\ref{fig:pr} has a standard deviation comparable to that of
-
-%> sd(pr_time$Time[pr_time$Sys == "PowerGraph"])
-%[1] 0.4654061
-%> sd(sssp_time$Time[sssp_time$Sys == "PowerGraph"])
-%[1] 0.1475608
-%> sd(pr_time$Time[pr_time$Sys == "GAP"])
-%[1] 0.01957317
-%> sd(sssp_time$Time[sssp_time$Sys == "GAP"])
-%[1] 0.01608644
-%> sd(pr_time$Time[pr_time$Sys == "GraphMat"])
-%[1] 0.9305763
-%> sd(sssp_time$Time[sssp_time$Sys == "GraphMat"])
-%[1] 0.02191714
-%> sd(pr_time$Time[pr_time$Sys == "GraphBIG"])
-%[1] 3.308351
-%> sd(sssp_time$Time[sssp_time$Sys == "GraphBIG"])
-%[1] 0.2165655
+Furthermore, the logarithmic scale in Figs.~\ref{fig:pr} and~\ref{fig:bfs-time} compresses the apparent variance in runtime.
+Each platform in Fig.~\ref{fig:pr} has a relative standard deviation between $1/4$ and $1/2$ that of the same system executing SSSP.
 
@@
 	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/pr_time.pdf}
+		\includegraphics[width=\linewidth, trim=0pt 18pt 18pt 0pt, clip]{graphics/pr_time.pdf}
 	\end{minipage}
 	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/pr_iters.pdf}
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0pt, clip]{graphics/pr_iters.pdf}
 	\end{minipage}
@@
 
-In Fig.~\ref{fig:bfs-scaling}, we see BFS scalability\dots
-
+Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS.
+We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time and $n$ is the number of threads.
+A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}.
+These plots show in general poor scaling throughout, though GraphBIG has the best scaling.
+One explanation is GraphBIG is the slowest system so there is more room for improvement.
+Additionally, a scale of 20 generates a small graph by today's standards and thus the focus is on scalability for larger graphs.
+Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.
 \begin{figure}
@@
 	\begin{minipage}{0.9\linewidth}
-		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/bfs_speedup.pdf}
+		% trim=left lower right upper
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup.pdf}
 	\end{minipage}
 	\begin{minipage}{0.9\linewidth}
-		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/bfs_ss.pdf}
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss.pdf}
 	\end{minipage}
-	\caption{The top figure has both axes logarithmic. Linear speedup is. GraphMat dips below 1 because it is slower for $2$ threads than for 1.}
+	\caption{The top figure has both axes logarithmic.
+The exception is the extra space at 72 threads for readability.
+Linear speedup is.
+GraphMat dips below 1 because it is slower for $2$ threads than for 1.
+$T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time for $n$ threads.}
 	\label{fig:bfs-scaling}
@@
 
-RAPL also allows the measurement of DRAM power and energy, whose results are shown in Fig.~\ref{fig:ram-power}.
+RAPL also allows the measurement of DRAM power and energy, whose results are shown in Fig.~\ref{fig:power}.
 
+\todo{Make text bigger}
 \begin{figure}
 	\centering
-	\includegraphics[width=0.6\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ram_power.pdf}
-	\caption{RAM Power Consumption, averaged across all roots. [TODO: Explain more]}
-	\label{fig:ram-power}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ram_power.pdf}
+	\end{minipage}
+	\begin{minipage}{0.48\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_cpu_power.pdf}
+	\end{minipage}
+	\caption{Similar to Fig.~\ref{fig:bfs-time} we plot RAM and CPU Power Consumption for each of the 32 roots.
+Since the Graph500 runs multiple roots per execution, we only get a single datapoint.
+The baseline was computed by monitoring power consumption while a program containing only one call to the C \texttt{unistd} library function \texttt{sleep(10)} (ten seconds).}
+	\label{fig:power}
 \end{figure}
@@
 \section{Future Work and Conclusion}
-The problem of choosing a system for graph processing at large scale is far from being simple.
+We have presented an approach to analyze parallel graph processing performance at fine detail, both measuring power consumption and the two fundamental phases of execution.
+Still, the problem of choosing a system for graph processing at large scale is far from being simple.
 We make our code available at \url{https://github.com/HPCL/easy-parallel-graph} and encourage further experimentation.

### 2017-02-17 papers/publication/plot_data.R

@@ -1,6 +1,7 @@
 # Generate some plots using the data parsed by ../../misc/parse-output.sh
 # Make this section be the results from the optimal number of threads
 
+
 ###
 # Part 1: Generate the plots for a single problem size and multiple algorithms
 ###
@@ -22,6 +23,7 @@ sssp_dsc <- subset(x, x$Algo == "SSSP" & x$Metric == "Data structure build",
 		c("Sys","Time"))
 sssp_time$Sys <- factor(sssp_time$Sys)
 sssp_dsc$Sys <- factor(sssp_dsc$Sys)
+# In the paper we compare range(sssp_time$Time) and range(bfs_time$Time)
 
 pr_time <- subset(x, x$Algo == "PageRank" & x$Metric == "Time",
 		c("Sys","Time"))
@@ -60,7 +62,7 @@ pdf("graphics/pr_iters.pdf", width = 4.5, height = 4.5)
 pr_mean_iters <- aggregate(pr_iters$Time, list(pr_iters$Sys), mean)
 pr_sys_order <- order(pr_mean_iters[[2]])
 pr_mean_iters <- pr_mean_iters[pr_sys_order,]
-bp <- barplot(pr_mean_iters[[2]], ylab = "Time (second)",
+bp <- barplot(pr_mean_iters[[2]], ylab = "Time (seconds)",
 		main = "PageRank Iterations", col=rainbow(length(pr_mean_iters[[1]])))
 text(bp, par("usr")[3], labels = pr_mean_iters[[1]], srt = 30,
 		adj = c(0.95,0.95), xpd = TRUE, cex = 1.0)
@@ -114,27 +116,27 @@ bfs_scale <- measure_scale("BFS")
 colors <- rainbow(nrow(bfs_scale))
 colors <- gsub("F", "C", colors) # You want it darker
 colors <- gsub("CC$", "FF", colors) # But keep it opaque
-
-# Plot the strong scalability for BFS
-pdf("graphics/bfs_ss.pdf", width = 7, height = 4)
 bfs_ss <- bfs_scale
 # Strong scaling for sequential is 1---we compute that last
 for (ti in rev(seq(length(threadcnts)))) {
 	bfs_ss[ti] <- bfs_ss[1] / (threadcnts[ti] * bfs_ss[ti])
 }
+
+# Plot the strong scalability for BFS
[…]

### 2017-02-17 papers/publication/paper-pollard.tex
Expanded explanation in Power section.

@@
 PAPI provides convenient access to these values and average energy in nanojoules for a given time interval.
-We modify the source code for each project to time only the actual BFS computation.
-In our case, the fastest code is also the most energy efficient, though with this level of granularity we could detect circumstances where one could make a tradeoff between energy and runtime.
+We modify the source code for each project to time only the actual BFS computation and give a summary of the results in Table~\ref{tab:power}.
+In our case, the fastest code is also the most energy efficient though with this level of granularity we could detect circumstances where one could make a tradeoff between energy and runtime.
 
@@
 	\end{tabular}
+	\label{tab:power}
 \end{table}
 
-RAPL also allows the measurement of DRAM power and energy, whose results are shown in Fig.~\ref{fig:power}.
+RAPL also allows the measurement of DRAM power, the results of which are shown in Fig.~\ref{fig:power}.
+The left plot describes the second row of Table~\ref{tab:power} in more detail.
+We notice a smaller spread of RAM power consumption, but still a noticeable difference.
+GraphMat results in the lowest RAM and power consumption despite not having the largest runtime.
 

### 2017-02-17 papers/publication/paper-pollard.tex
Removed duplicate figure

@@
 \end{figure}
-\todo{Check this again---that's a huge discrepancy!}
-
-\begin{figure}
-	\centering
-	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 24pt 18pt 0, clip]{graphics/pr_time.pdf}
-	\end{minipage}
-	\begin{minipage}{0.48\linewidth}
-		\includegraphics[width=\linewidth, trim=0 24pt 18pt 0, clip]{graphics/pr_iters.pdf}
-	\end{minipage}
-	\caption{The $y$-axis is logarithmic. GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.}
-	\label{fig:pr}
-\end{figure}
 
@@
 These plots show in general poor scaling throughout, though GraphBIG has the best scaling.
-One explanation is GraphBIG is the slowest system so there is more room for improvement.
+A potential explanation is GraphBIG is the slowest system so there is more room for improvement.
 Additionally, a scale of 20 generates a small graph by today's standards and thus the focus is on scalability for larger graphs.

### 2017-02-17 papers/publication/paper-pollard.tex
some edits in 2

--- 2017-02-17 papers/publication/paper-pollard.tex
@@
 \section{Related Work}
-A related performance analysis project is Graphalytics \cite{Capota:2015:Graphalytics} which also attempts to automate the setup and execution of packages for performance analysis.
+A related project is Graphalytics \cite{Capota:2015:Graphalytics}, which also attempts to automate the setup and execution of graph packages for performance analysis.
 However, in order for Graphalytics to analyze a graph processing system---Graphalytics calls these platforms---a programmer must implement Java classes as wrappers to call the particular implementations, satisfy dependencies, and set the correct shell commands to execute.
-These all require some knowledge of the inner workings of each system in addition to familiarity with the Graphalytics API.
+These all require some knowledge of the inner workings of each system in addition to familiarity with the Graphalytics API and involve a nontrivial amount of programming.
 
-The complexity of Graphalytics obfuscates the true behavior of the program.
+Moreover, the complexity of Graphalytics can obfuscate the true behavior of the program.
 By default, Graphalytics generates an HTML report listing the runtimes for each dataset and each algorithm in seconds.
@@
 However, perusing the log files reveals a more complete picture: of these 5.96 seconds, 0.21 was spent computing the shortest paths while the remaining time consisted of building the necessary data structures.
-Furthermore, data such as the number of iterations run for PageRank are not easily available using Graphalytics.
+Furthermore, other performance-relavant details, such as the number of iterations run for PageRank, are not easily available using Graphalytics.
 %We use GraphMat, the single source shortest paths algorithm, the dota-league dataset from the Game Trace Archive\cite{Guo:2012:GTA}, and the default settings as an example here.

### 2017-02-17 papers/publication/paper-pollard.tex
more related edits (mainly rewording)

@@
 
-Other related work includes that of \cite{Satish:2014:NavigatingGraph}, which analyzes performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} which uses six real-world datasets and focuses on the vertex-centric programming model.
+Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
 
-Our approach differs from the previously mentioned appproaches because it focuses on the depth of analysis rather than the breadth via diversity of datasets or graph processing systems.
+Unlike these approaches, we focus on the \emph{depth} of analysis rather than the breadth via diversity of datasets or graph processing systems.
 We analyze  power and energy consumption in addition to details such as the number of iterations or the time to construct the data structures.
-Additionally, our approach requires little knowledge of the inner workings of each system; the data are collected using either sampling of model specific registers (for power) or parsing log files (for timing).
+Moreover, our approach requires little knowledge of the inner workings of each system; the data are collected using either hardware counter sampling of model-specific registers (for power) or through parsing log files (for timing).
 

### 2017-02-17 papers/publication/paper-pollard.tex
intro edits

@@
 \begin{abstract}
-The increasing popularity of large network analysis problems has led to the emergence of many parallel and distributed graph processing systems---one survey in 2014 identified over 80.
+The rapidly growing number of large network analysis problems has led to the emergence of many parallel and distributed graph processing systems---one survey in 2014 identified over 80.
 Since then, the landscape has evolved; some packages have become inactive while more are being developed.
@@
 Our research is motivated by the current state of parallel graph processing.
-The most comprehensive survey, released in 2014, identified and taxonomized over 80 different parallel graph processing systems without even including domain specific languages \cite{Doekemeijer:2015:GPFSurvey}.
+The most comprehensive survey, released in 2014, identified and categorized over 80 different parallel graph processing systems, without considering domain specific languages \cite{Doekemeijer:2015:GPFSurvey}.
 
@@
 
-Our contributions can be summarized as follows.
+Our research contributions in this paper can be summarized as follows.
 \begin{itemize}
@@
 Essentially, we provide a ``level playing field'' for graph algorithms.
-	\item Comparison of algorithm runtime and scalability using a variety of data sets
-	\item Comparison of our framework with the Graphalytics \cite{Capota:2015:Graphalytics} project
+	\item Comparison of algorithm runtime and scalability using a variety of data sets.
+	\item Comparison of our framework with the Graphalytics \cite{Capota:2015:Graphalytics} project.
 	\item Inspection of the source code of the surveyed parallel graph processing systems to ensure the same phases of execution are measured across differing execution and programming paradigms
-	\item Provision of a experimental framework distinguishing between the data structure generation and algorithm execution phases of each system
-	\item Analysis of energy consumption during the algorithm execution phase
+	\item Provision of a experimental framework distinguishing between the data structure generation and algorithm execution phases of each system.
+	\item Analysis of energy consumption during the algorithm execution phase.
 \end{itemize}
@@
 
-With a plugin to Graphalytics called Granula \cite{Ngai:2015:Granula} one can explicitly specify a performance model to analyze specific execution behavior such as the amount of communication or runtime of particular kernels of execution.
+With a plugin to Graphalytics called Granula \cite{Ngai:2015:Granula}, one can explicitly specify a performance model to analyze specific execution behavior such as the amount of communication or runtime of particular kernels of execution.
 This requires in-depth knowledge of the source code and execution model.

### 2017-02-17 papers/publication/paper-pollard.tex
intro analysis para

@@
 Essentially, we provide a ``level playing field'' for graph algorithms.
-	\item Comparison of algorithm runtime and scalability using a variety of data sets.
+	\item Comparison of algorithm runtime and scalability using several data sets.
 	\item Comparison of our framework with the Graphalytics \cite{Capota:2015:Graphalytics} project.
-	\item Inspection of the source code of the surveyed parallel graph processing systems to ensure the same phases of execution are measured across differing execution and programming paradigms
-	\item Provision of a experimental framework distinguishing between the data structure generation and algorithm execution phases of each system.
+	% BN: I think the bullet below is a bit too low-level for this summary, moving to later section.
+	%\item Inspection of the source code of the surveyed parallel graph processing systems to ensure the same phases of execution are measured across differing execution and programming paradigms.
+	\item An experimental framework distinguishing between the data structure generation and algorithm execution phases of each system.
 	\item Analysis of energy consumption during the algorithm execution phase.
@@
 The datasets are described further in Sect.~\ref{sec:datasets}.
-We measure strong scaling and speedup by varying the number of threads from one to the total number of threads on the reasearch machine, 72.
+We measure strong scaling and speedup by varying the number of threads from one to the total number of threads on our server, 72.
 
@@
 
-We plot many of the results as box plots with an implied 32 datapoints per box.
+We plot many of the results as box plots with an implied 32 data points per box.
 
 \subsection{Machine Specifications}
-Table~\ref{tab:specs} shows the specifications of the research computer.
+Table~\ref{tab:specs} shows the specifications of the server on which we performed all experiments.
 
@@
 
-This report explores four shared memory parallel graph processing platforms.
+This study explores four shared memory parallel graph processing platforms.
 The first three are so-called ``reference implementations'' while the remaining two are included because of their performance and popularity.
@@
 Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} are not considered here because the authors do not provide reference implementations.
-The systems are:
+We consider the following systems.
+\begin{itemize}
+	\item The Graph500\footnote{We used the most recent version from \url{https://github.com/graph500/graph500}, most similar to release 2.1.4.} \cite{Murphy:2010:Graph500}, a set of large-scale graph benchmarks.
+	\item The Graph Algorithm Platform (GAP) Benchmark Suite \cite{Beamer:2015:GAPBench}.
+	\item GraphBIG \cite{Nai:2015:Graphbig} benchmark suite.
+	\item GraphMat \cite{Sundaram:2015:GraphMat}, a high-performance graph library.
+	\item PowerGraph \cite{Gonzalez:2012:Powergraph}, a framework for distributed graph-parallel computation on natural graphs.
+\end{itemize}
 
-\begin{enumerate}
-	\item The Graph500\footnote{We used the most recent version from \url{https://github.com/graph500/graph500}, most similar to release 2.1.4.} \cite{Murphy:2010:Graph500}
-	\item The Graph Algorithm Platform (GAP) Benchmark Suite \cite{Beamer:2015:GAPBench}
-	\item GraphBIG \cite{Nai:2015:Graphbig}
-	\item GraphMat \cite{Sundaram:2015:GraphMat}
-	\item PowerGraph \cite{Gonzalez:2012:Powergraph}
-\end{enumerate}
+Our approach is not specific or limited to these graph packages and can be extended to others with relatively modest effort (building each package and if necessary, instrumenting the different phases of the computation).
 
-\subsection{Algorithms}
+\subsection{Algorithms}\label{sec:algs}
 
 We consider three algorithms: Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank, though not all algorithms are implemented on all systems.
+We inspected the source code of the surveyed parallel graph processing systems to ensure the same phases of execution are measured across differing execution and programming paradigms.
 
@@
 \section{Performance Analysis}\label{sec:perf}
+
+We analyze the performance of the algorithms described in Sec.~\ref{sec:algs} in terms of execution time, scalability over multiple threads, and power and energy consumption.
 

### 2017-02-17 papers/publication/paper-pollard.tex
Sam's update before merge. Tweaking some wording.

--- 2017-02-17 papers/publication/paper-pollard.tex
@@
 By default, Graphalytics generates an HTML report listing the runtimes for each dataset and each algorithm in seconds.
-For example, one run of a single source shortest paths algorithm took 5.6 seconds.
+For example, one run of a single source shortest paths algorithm took 5.96 seconds.
 However, perusing the log files reveals a more complete picture: of these 5.96 seconds, 0.21 was spent computing the shortest paths while the remaining time consisted of building the necessary data structures.
@@
 
-As with Graphalytics, the initial development effort is high but Granula paired with graphalytics allows automatic execution and compilation of performance results.
-Likewise, our approach provides automatic execution and performance analysis without requiring a performance and execution model for each system.
+As with Graphalytics, the initial development effort is high but Granula paired with Graphalytics allows automatic execution and compilation of performance results.
+Likewise, our approach provides automatic execution and performance analysis but without requiring a performance and execution model for each system.
 
@@
 
-Each experiment uses 32 roots per graph. As with the Graph500, each root is selected to have a degree greater than 1. In the case of PageRank, we simply run PageRank 32 times.
+Each experiment uses 32 roots per graph. As with the Graph500, each root is selected to have a degree greater than 1. For PageRank, we simply run the algorithm 32 times.
 
-When possible, we measure the data set construction time as the time to translate from the unstructured file data in RAM to the graph representation on which the algorithm can be performed.
+When possible, we measure the dataset construction time as the time to translate from the unstructured file data in RAM to the graph representation on which the algorithm can be performed.
 This is not possible for PowerGraph and GraphBIG because they read in the input file and build a graph simultaneously.
-The Graph500 builds the graph once and performs the BFS on each root in turn, speeding up experimentation but only giving a single data point for the data set construction.
+The Graph500 builds the graph once and performs the BFS on each root in turn, speeding up experimentation but only giving a single data point for the dataset construction.
 
-We plot many of the results as box plots with an implied 32 datapoints per box.
+Any box plot with the exception of the Graph500 (instead labeled as solid lines rather than boxes) is a representation of 32 datapoints per box.
 
@@
 
-We consider three algorithms: Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank, though not all algorithms are implemented on all systems.
+We consider three algorithms: Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank, though not all algorithms are implemented for all systems.
 
@@
 Our work aims to add similar rigor to other graph algorithms by borrowing heavily from the Graph500 specification.
-The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the actual BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}.
+The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an unsorted edge list stored in RAM and the actual BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}.
 We run the BFS using 32 random roots with the exception of PowerGraph which doesn't provide an reference implementation of BFS in its toolkits.
@@
 
-In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a single dataset.
-For an explanation of each algorithm used, see \cite{Iosup:2016:Graphalyticstech}.
+In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on the Dota-League dataset.
+An explanation of each algorithm is given in \cite{Iosup:2016:Graphalyticstech}.
 Table~\ref{tab:graphalytics} shows only a broad overview of the two systems: GraphBIG and PowerGraph.
@@
 
-Figures~\ref{fig:bfs-time} and \ref{fig:sssp-time} show more detail, with box plots for the runtime distributions.
-However, even these do not give a complete picture of performance.
-Both of these show GraphMat to not be highly performant.
-One explanation is GraphMat's underlying computation model (sparse matrix operations) paired the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation is more conducive to larger-scale graphs.
+Figures~\ref{fig:bfs-time} and \ref{fig:sssp-time} show performance results for a Kronecker graph with scale $22$.
+The box plots give an idea of the runtime distributions.
 There is less discrepancy between the runtimes of SSSP (between 0.1 and 1.7 seconds) compared to BFS (0.01 and 1.7 seconds) but GAP is the clear winner in both cases.
 The data structure construction times for GAP and GraphMat are consistent; in both cases the platforms create the same data structure for both algorithms.
-
 These results are consistent with \cite{Sundaram:2015:GraphMat} which lists GraphMat as more performant than PowerGraph in SSSP.
@@
 Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS.
-We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time and $n$ is the number of threads.
+We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time for $n$ threads.
 A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}.
@@
 A potential explanation is GraphBIG is the slowest system so there is more room for improvement.
-Additionally, a scale of 20 generates a small graph by today's standards and thus the focus is on scalability for larger graphs.
+Additionally, $2^{20}$ vertices is a ``small'' graph by today's standards and thus library designers focus on scalability for larger graphs.
 Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.
@@
 	\end{minipage}
-	\caption{The top figure has both axes logarithmic.
-The exception is the extra space at 72 threads for readability.
+	\caption{The top figure has both axes logarithmic with an exception at 72 threads for readability.
 Linear speedup is.
@@
 \begin{table}
-	\caption{The data are generated from a Kronecker graph with $2^{22}$ vertices.
-Sleeping Energy refers to the power (in Watts) consumed during the Unix \texttt{sleep} function, multiplied by time.
-Effectively, this measures the amount of energy that would have been consumed even if nothing was running.
+	\caption{The data are generated using a Kronecker graph with a scale of 22.
+Sleeping Energy refers to the power (in Watts) consumed during the \texttt{unistd} C \texttt{sleep} function, multiplied by the Time row.
+Essentially, this measures the amount of energy that would have been consumed even if nothing was running.
 The increase over sleep is the ratio of the first and third columns.
-These are all averaged over the 32 roots.}
+These are all averaged over 32 roots.}
 	\centering
@@
 
-Overall, the GAP Benchmark Suite was the highest-performing system across the given datasets and the most scalable.
+Overall, the GAP Benchmark Suite was the highest-performing system across the given datasets but the least scalable.
 However, this is only for relatively small graphs; all the graphs used here had at most $2^{22}$ vertices.
@@
 Thus, we recommend graph processing algorithm designers compare their implementations against the GAP Benchmark Suite to as a good test of performance.
+
+We found Graphmat to not be highly performant but relatively scalable.
+One explanation is GraphMat's underlying computation model (sparse matrix operations) paired the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation is more conducive to larger-scale graphs.
+% Furthermore, sparse matrix operations have been well studied so achieving good scalability is more realistic [cite]
 

### 2017-02-17 papers/publication/paper-pollard.tex
conclusion

--- 2017-02-17 papers/publication/paper-pollard.tex
@@
 We analyze  power and energy consumption in addition to details such as the number of iterations or the time to construct the data structures.
-Moreover, our approach requires little knowledge of the inner workings of each system; the data are collected using either hardware counter sampling of model-specific registers (for power) or through parsing log files (for timing).
+Moreover, our approach requires little knowledge of the inner workings of each system; the data are collected using either hardware counter sampling of model-specific registers (for power) or through parsing log files (for execution time).
 
@@
 We select BFS because the canonical performance leaderboard for parallel graph processing is the Graph500 \cite{Murphy:2010:Graph500}.
-The advantage of the Graph500 is it provides standardized measurement specifications and dataset generation.
+The advantage of the Graph500 is that it provides standardized measurement specifications and dataset generation.
 The primary drawback with the Graph500 is it measures a single algorithm.
@@
 One challenge with using PageRank is the stopping criterion; this is addressed in Sect.~\ref{sec:perf}.
-Verification of the PageRanks is beyond the scope of this paper though this may explain some of the large performance discrepancies.
+Verification of the PageRank results is beyond the scope of this paper, although this may explain some of the large performance discrepancies.
 
-This approach is not specific to a particular algorithm; measuring the execution time, data structure construction time, and power consumption can be easily extended to other algorithms.
+This approach is not specific to a particular algorithm; measuring the execution time, data structure construction time, and power consumption can be applied easily to other algorithms.
 
@@
 	every head row/.style={output empty row, before row=\bottomrule},
-	columns/0/.style={column type={|r|}},
-	columns/1/.style={column type={l|}},
+	columns/0/.style={column type={|p{1in}|}},
+	columns/1/.style={column type={p{3.2in}|}},
 	every last row/.style={after row=\toprule},
@@
 We use $\epsilon = 6 \times 10^{-8}$ because this value is approximately machine epsilon for a single precision floating point number.
-This is done with the goal of making the stoppping criteria as similar as possible.
+This is done with the goal of making the stopping criteria as similar as possible.
 
-Furthermore, the logarithmic scale in Figs.~\ref{fig:pr} and~\ref{fig:bfs-time} compresses the apparent variance in runtime.
+Moreover, the logarithmic scale in Figures~\ref{fig:bfs-time} and~\ref{fig:pr} compresses the apparent variance in runtime.
 Each platform in Fig.~\ref{fig:pr} has a relative standard deviation between $1/4$ and $1/2$ that of the same system executing SSSP.
@@
 \section{Future Work and Conclusion}
-We have presented an approach to analyze parallel graph processing performance at fine detail, both measuring power consumption and the two fundamental phases of execution.
-Still, the problem of choosing a system for graph processing at large scale is far from being simple.
-We make our code available at \url{https://github.com/HPCL/easy-parallel-graph} and encourage further experimentation.
+We have presented a systematic approach to analyzing parallel graph processing performance, both measuring power consumption and the two fundamental phases of execution.
+While the comparison presented here can help choose among alternatives for the selected packages and algorithms, the problem of choosing a system for graph processing at large scale is far from simple.
+Hence, we make our code available at \url{https://github.com/HPCL/easy-parallel-graph} and encourage further experimentation.
 Our results required minor changes to the original projects to add calls to PAPI sampling and those changes are also available on forked versions of the repositories.
@@
 It should be noted that GAP is also the most recent of projects.
-Thus, we recommend graph processing algorithm designers compare their implementations against the GAP Benchmark Suite to as a good test of performance.
+Thus, we recommend graph processing algorithm designers compare their implementations against the GAP Benchmark Suite as a good test of performance.
 
-There are a vast number of directions this work could go.
+Our work can continue in many directions.
 To ensure fairness, each platform must be configured to use the same graphs and the same roots.
-In the case of GraphBIG this required the file to be loaded in for each experiment.
-This was by far the most time consuming of the experiments; the file reading for GraphBIG being done serially on an uncompressed ASCII text format for graphs.
+In the case of GraphBIG, this required the file to be loaded in for each experiment.
+By far the most time consuming of the experiments was GraphBIG's serial file input of uncompressed ASCII text format for graphs.
 This limited the sizes of the experiments we could run.
-Graphalytics also had circumstances with the more computationally expensive algorithms  where certain experiments fail \cite{Iosup:2016:Graphalyticstech}, so determining whether an algorithm will finish given a particular machine, input size, runtime limit, and resources is an important unanswered question.
+Graphalytics also had circumstances with the more computationally expensive algorithms  where certain experiments fail \cite{Iosup:2016:Graphalyticstech}, so determining whether an algorithm will finish given a particular machine, input size, runtime limit, and resources is an important unanswered question we plan to pursue further.
 

### 2017-02-17 papers/publication/paper-pollard.tex
conflict merges

@@
 
-Each experiment uses 32 roots per graph. As with the Graph500, each root is selected to have a degree greater than 1. In the case of PageRank, we simply run PageRank 32 times.
+Each experiment uses 32 roots per graph.
+As with the Graph500, each root is selected to have a degree greater than 1.
+For PageRank, we simply run the algorithm 32 times.
+We plot many of the results as box plots with an implied 32 data points per box.
 
-When possible, we measure the data set construction time as the time to translate from the unstructured file data in RAM to the graph representation on which the algorithm can be performed.
+When possible, we measure the dataset construction time as the time to translate from the unstructured file data in RAM to the graph representation on which the algorithm can be performed.
 This is not possible for PowerGraph and GraphBIG because they read in the input file and build a graph simultaneously.
-The Graph500 builds the graph once and performs the BFS on each root in turn, speeding up experimentation but only giving a single data point for the data set construction.
-
-We plot many of the results as box plots with an implied 32 data points per box.
+The Graph500 builds the graph once and performs the BFS on each root in turn, speeding up experimentation but only giving a single data point for the dataset construction.
 
@@
 
-We consider three algorithms: Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank, though not all algorithms are implemented on all systems.
+We consider three algorithms: Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank, although not all algorithms are implemented on all systems.
 We inspected the source code of the surveyed parallel graph processing systems to ensure the same phases of execution are measured across differing execution and programming paradigms.
@@
 Thus, we recommend graph processing algorithm designers compare their implementations against the GAP Benchmark Suite as a good test of performance.
+
+We found Graphmat to not be highly performant but relatively scalable.
+One explanation is GraphMat's underlying computation model (sparse matrix operations) paired the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation is more conducive to larger-scale graphs.
+% Furthermore, sparse matrix operations have been well studied so achieving good scalability is more realistic [cite]
 

### 2017-02-17 papers/publication/paper-pollard.tex
conclusion

@@
 However, perusing the log files reveals a more complete picture: of these 5.96 seconds, 0.21 was spent computing the shortest paths while the remaining time consisted of building the necessary data structures.
-Furthermore, other performance-relavant details, such as the number of iterations run for PageRank, are not easily available using Graphalytics.
+Furthermore, other performance-relevant details, such as the number of iterations run for PageRank, are not easily available using Graphalytics.
 %We use GraphMat, the single source shortest paths algorithm, the dota-league dataset from the Game Trace Archive\cite{Guo:2012:GTA}, and the default settings as an example here.
@@
 
-We found Graphmat to not be highly performant but relatively scalable.
-One explanation is GraphMat's underlying computation model (sparse matrix operations) paired the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation is more conducive to larger-scale graphs.
+We found that GraphMat is not highly performant but does scale relatively well.
+One explanation is that GraphMat's underlying computation model (sparse matrix operations), paired the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation, is better suited to larger-scale graphs.
 % Furthermore, sparse matrix operations have been well studied so achieving good scalability is more realistic [cite]

### 2017-02-17 papers/publication/paper-pollard.tex
fix
misc edits

@@
 Table~\ref{tab:graphalytics} shows only a broad overview of the two systems: GraphBIG and PowerGraph.
-In general, we see GraphBIG is more performant.
-However, results such as these are preliminary and do not show the complete picture.
+In general, we see GraphBIG performs better.
+However, these preliminary results do not show the complete picture.
+\todo[inline]{Would be good to add something here on what more is needed for getting a complete picture.}
 
@@
 \todo{rework this. It would be nice to compare Graphalytics and our approach on the real-world dataset.}
-Table~\ref{tab:epg-perf}, in contrast, gives an overview of the performance results of running graphalytics and our approach using the same dataset [TODO: Get the Dota-League Dataset working with easy-parallel-graph and replace Table~\ref{epg-perf}].
+Table~\ref{tab:epg-perf}, by contrast, gives an overview of the performance results of running Graphalytics and our approach using the same dataset [TODO: Get the Dota-League Dataset working with easy-parallel-graph and replace Table~\ref{epg-perf}].
 Table~\ref{tab:epg-perf} shows the average time across 32 roots, whereas Graphalytics by default shows results for a single root.
@@
 The box plots give an idea of the runtime distributions.
-There is less discrepancy between the runtimes of SSSP (between 0.1 and 1.7 seconds) compared to BFS (0.01 and 1.7 seconds) but GAP is the clear winner in both cases.
+There is less variance in the runtimes of SSSP (between 0.1 and 1.7 seconds) compared to BFS (0.01 and 1.7 seconds) but GAP is the clear winner in both cases.
 The data structure construction times for GAP and GraphMat are consistent; in both cases the platforms create the same data structure for both algorithms.
@@
 The left box plot shows the time to compute the SSSP starting at the same 32 roots as Fig.~\ref{fig:bfs-time}.
-Both PowerGraph and GraphBIG construct their data structures at the same time as they read in the file.}
+Both PowerGraph and GraphBIG construct their data structures at the same time as they read the file.}
 	\label{fig:sssp-time}
@@
 We adjusted the other systems to use $\sum_{k=1}^{n} |p_k^{(i)} - p_k^{(i-1)}| < \epsilon $ as the stopping criteria, where $i$ is the iteration and $n$ is the number of vertices.
-We use $\epsilon = 6 \times 10^{-8}$ because this value is approximately machine epsilon for a single precision floating point number.
-This is done with the goal of making the stopping criteria as similar as possible.
+We use $\epsilon = 6 \times 10^{-8}$ because this value is approximately machine epsilon for a single precision floating-point number.
+The goal is to make the stopping criteria for all implementations as similar as possible.
 
@@
 Sleeping Energy refers to the power (in Watts) consumed during the \texttt{unistd} C \texttt{sleep} function, multiplied by the Time row.
-Essentially, this measures the amount of energy that would have been consumed even if nothing was running.
+Essentially, this measures the amount of energy that would have been consumed if nothing was running.
 The increase over sleep is the ratio of the first and third columns.
@@
 We notice a smaller spread of RAM power consumption, but still a noticeable difference.
-GraphMat results in the lowest RAM and power consumption despite not having the largest runtime.
+GraphMat results in the lowest RAM and power consumption despite not having the shortest runtime.
 

### 2017-02-17 papers/publication/paper-pollard.tex
misc edits

@@
 We have presented a systematic approach to analyzing parallel graph processing performance, both measuring power consumption and the two fundamental phases of execution.
-While the comparison presented here can help choose among alternatives for the selected packages and algorithms, the problem of choosing a system for graph processing at large scale is far from simple.
+While the comparison presented here can help one choose among alternatives for the selected packages and algorithms, the problem of selecting a graph processing framework for a given large-scale problem remains far from simple.
 Hence, we make our code available at \url{https://github.com/HPCL/easy-parallel-graph} and encourage further experimentation.

### 2017-02-17 papers/publication/paper-pollard.tex
power edits

@@
 \subsection{Power and Energy Consumption}
-We use the Performance Application Programming Interface (PAPI) \cite{Browne:2000:PAPI} to gain access to Intel's Running Average Power Limit (RAPL), which provides access to a set of hardware counters measuring energy usage.
-PAPI provides convenient access to these values and average energy in nanojoules for a given time interval.
+We use the Performance Application Programming Interface (PAPI) \cite{Browne:2000:PAPI} to gain access to Intel's Running Average Power Limit (RAPL), which provides a set of hardware counters for measuring energy usage.
+We use PAPI to obtain average energy in nanojoules for a given time interval.
 We modify the source code for each project to time only the actual BFS computation and give a summary of the results in Table~\ref{tab:power}.
-In our case, the fastest code is also the most energy efficient though with this level of granularity we could detect circumstances where one could make a tradeoff between energy and runtime.
+In our case, the fastest code is also the most energy efficient, although with this level of granularity we could detect circumstances where one could make a tradeoff between energy and runtime.
 
@@
 We notice a smaller spread of RAM power consumption, but still a noticeable difference.
-GraphMat results in the lowest RAM and power consumption despite not having the shortest runtime.
+GraphMat exhibits the lowest average RAM power consumption (but not the shortest execution time).
+This information can be useful when choosing an algorithm to use in limited power scenarios, where a slower algorithms that will not exceed the power cap is preferred to a faster one that may exceed it.
 

### 2017-02-17 papers/publication/paper-pollard.tex
fixed num cores

@@
 
-At the human level, there are optimizations for each system which may not be apparent.
-For example, the Parallel Boost Graph Library (PBGL) \cite{Gregor:2005:PBGL} provides generic implementations of their algorithms and the programmer must provide the template specializations.
+From a graph algorithm user's point of view, there are optimizations for each system which may not be apparent.
+For example, the Parallel Boost Graph Library (PBGL)~\cite{Gregor:2005:PBGL} provides generic implementations of their algorithms and the programmer must provide the template specializations.
 The optimal data structures may differ across algorithms and graphs and must be determined by the programmer.
@@
 \subsection{Machine Specifications}
-Table~\ref{tab:specs} shows the specifications of the server on which we performed all experiments.
+We performed experiments on our 32-core Intel Haswell server, whose specifications are outlined below.
+%Table~\ref{tab:specs} shows the specifications of the server on which we performed all experiments.
+%\begin{table}[h!tb]
+%	\centering
+%	\caption{
+The operating system is GNU/Linux version 4.4.0-22.
+The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
+We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.
+\begin{center}
+	% For arya I deleted Max RAM Freq	2133MHz
+	\pgfplotstabletypeset[
+	header=false,
+	col sep=tab,
+	string type,
+	every head row/.style={output empty row, before row=\bottomrule},
+	columns/0/.style={column type={|p{1in}|}},
+	columns/1/.style={column type={p{3.2in}|}},
+	every last row/.style={after row=\toprule},
+	]{../report/specs.csv}
+	%\label{tab:specs}
+\end{center}
 
@@
 This approach is not specific to a particular algorithm; measuring the execution time, data structure construction time, and power consumption can be applied easily to other algorithms.
-
-\begin{table}
-	\centering
-	\caption{The operating system is GNU/Linux version 4.4.0-22.
-The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
-We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.}
-	% For arya I deleted Max RAM Freq	2133MHz
-	\pgfplotstabletypeset[
-	header=false,
-	col sep=tab,
-	string type,
-	every head row/.style={output empty row, before row=\bottomrule},
-	columns/0/.style={column type={|p{1in}|}},
-	columns/1/.style={column type={p{3.2in}|}},
-	every last row/.style={after row=\toprule},
-	]{../report/specs.csv}
-	\label{tab:specs}
-\end{table}
 
@@
 
-We attempt to use similar stopping criteria for each system, but GraphMat executes until no vertices change rank; effectively its stopping criterion requires the $\infty$-norm be less than machine epsilon.
+We attempt to define similar stopping criteria for each system, but GraphMat executes until no vertices change rank; effectively its stopping criterion requires the $\infty$-norm be less than machine epsilon.
 This could account for the increased number of iterations.
@@
 
-The challenge in comparing iteration counts for PageRank underscores an important challenge for any comparison of graph processing systems.
+The difficulty in comparing iteration counts for PageRank underscores an important challenge for any comparison of graph processing systems.
 The assumptions under which the various platforms operate can have a dramatic effect on the program.
 For example, the GAP Benchmark Suite stores vertex weights as 32-bit integers.
-However, other systems store them as floating point numbers.
+However, other systems store them as floating-point numbers.
 This may affect performance in addition to runtime behavior in cases where weights like $0.2$ are cast to $0$.
-Similarly, how a graph is represented in the system (e.g.
-weighed or directed) may have performance and algorithmic implications but is not always readily apparent.
+Similarly, how a graph is represented in the system (e.g., weighed or directed) may have performance and algorithmic implications but is not always readily apparent.
 
@@
 A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}.
-These plots show in general poor scaling throughout, though GraphBIG has the best scaling.
-A potential explanation is GraphBIG is the slowest system so there is more room for improvement.
+These plots show in general poor scaling throughout, although GraphBIG has the best scaling.
+A potential explanation is GraphBIG is the slowest system, so there is more room for improvement.
 Additionally, $2^{20}$ vertices is a ``small'' graph by today's standards and thus library designers focus on scalability for larger graphs.

### 2017-02-17 papers/publication/paper-pollard.tex
Addressed some TODOS and grammar

@@
 \section{Experimental Setup}
-We refer to a graphs's \emph{scale} when describing the size of the graph.
+We refer to a graph's \emph{scale} when describing the size of the graph.
 Specifically, a graph with scale $S$ has $2^S$ vertices.
@@
 \subsection{Machine Specifications}
-We performed experiments on our 32-core Intel Haswell server, whose specifications are outlined below.
+We performed experiments on our 36-core Intel Haswell server whose specifications are outlined below.
 %Table~\ref{tab:specs} shows the specifications of the server on which we performed all experiments.
@@
 However, these preliminary results do not show the complete picture.
-\todo[inline]{Would be good to add something here on what more is needed for getting a complete picture.}
+For example, one may want to know the cost of performing additional computations on a graph once the data structures have been built.
+Additionally, one may desire runtime bounds to get a reasonable maximum timespan for a given problem.
 
@@
 
-\todo{Make the text bigger for these}
 \begin{figure}
@@
 
-\todo{Make text bigger}
 \begin{figure}

### 2017-02-17 papers/publication/paper-pollard.tex
minor

@@
 The left box plot shows the time to compute BFS on 32 random roots while the right plot shows the times to construct the graph for each system.
-The Graph500 only constructs its graph once.
+Graph500 only constructs its graph once.
 GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
@@
 These plots show in general poor scaling throughout, although GraphBIG has the best scaling.
-A potential explanation is GraphBIG is the slowest system, so there is more room for improvement.
+A potential explanation is that because GraphBIG is the slowest system, there is more room for improvement.
 Additionally, $2^{20}$ vertices is a ``small'' graph by today's standards and thus library designers focus on scalability for larger graphs.
@@
 GraphMat exhibits the lowest average RAM power consumption (but not the shortest execution time).
-This information can be useful when choosing an algorithm to use in limited power scenarios, where a slower algorithms that will not exceed the power cap is preferred to a faster one that may exceed it.
+This information can be useful when choosing an algorithm for limited power scenarios, where a slower algorithms that will not exceed the power cap is preferred to a faster one that may exceed it.
 

### 2017-02-17 papers/publication/paper-pollard.tex
Updated to include Boyana's fixes.

@@
 These plots show in general poor scaling throughout, although GraphBIG has the best scaling.
-A potential explanation is that because GraphBIG is the slowest system, there is more room for improvement.
+A potential explanation is GraphBIG is the slowest system, so there is more room for improvement.
 Additionally, $2^{20}$ vertices is a ``small'' graph by today's standards and thus library designers focus on scalability for larger graphs.
@@
 It should be noted that GAP is also the most recent of projects.
-Thus, we recommend graph processing algorithm designers compare their implementations against the GAP Benchmark Suite as a good test of performance.
+Thus, we recommend graph processing algorithm designers compare their implementations against the GAP Benchmark Suite as an example of good performance.
 
@@
 
-Our work can continue in many directions.
 To ensure fairness, each platform must be configured to use the same graphs and the same roots.

### 2017-02-17 papers/publication/paper-pollard.tex
Fixed bug with scaling.

--- 2017-02-17 papers/publication/paper-pollard.tex
@@
 
-\todo{rework this. It would be nice to compare Graphalytics and our approach on the real-world dataset.}
-Table~\ref{tab:epg-perf}, by contrast, gives an overview of the performance results of running Graphalytics and our approach using the same dataset [TODO: Get the Dota-League Dataset working with easy-parallel-graph and replace Table~\ref{epg-perf}].
+% Table~\ref{tab:epg-perf}, by contrast, gives an overview of the performance results of running Graphalytics and our approach using the same dataset [TODO: Get the Dota-League Dataset working with easy-parallel-graph and replace Table~\ref{epg-perf}].
 Table~\ref{tab:epg-perf} shows the average time across 32 roots, whereas Graphalytics by default shows results for a single root.
+%\begin{table}
+%	\centering
+%	\begin{tabular}{l|c|c|c}
+%		System & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular} & Run BFS & Run SSSP \\ \hline
+%		PowerGraph & N/A & 0.003380 \\
+%		GAP        & 0.412 & 0.001393 \\
+%		GraphMat   &  & 0.1031 \\
+%		GraphBIG   & N/A & 0.1528 \\
+%	\end{tabular}
+%	\caption{This table shows times for the Dota-League dataset using our method.}
+%	\label{tab:epg-perf}
+%\end{table}
 
@@
 These results are consistent with \cite{Sundaram:2015:GraphMat} which lists GraphMat as more performant than PowerGraph in SSSP.
-
-%\begin{table}
-%	\centering
-%	\begin{tabular}{l|c|c|c}
-%		System & Load Graph & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular} & Run BFS \\ \hline
-%		Graph500 & 0.3474 & 0.3971 & 0.003380 \\
-%		GAP      & 2.351  & 0.1935 & 0.001393 \\
-%		GraphMat & 0.1511 & 1.101  & 0.1031 \\
-%		GraphBIG & 37.22  & N/A    & 0.1528 \\
-%	\end{tabular}
-%	\caption{This table shows times for $2^{20}$ verties and the times are in seconds.
-The Graph500 generates the graph instead of loading it into a file.
-GraphBIG builds the graph and reads in the file simultaneously.
-These results were averaged across 32 roots.}
-%	\label{tab:epg-perf}
-%\end{table}
 
@@
 Linear speedup is.
-GraphMat dips below 1 because it is slower for $2$ threads than for 1.
+GraphBIG dips below 1 because it is slower for $2$ threads than for 1.
 $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time for $n$ threads.}
@@
 
-Overall, the GAP Benchmark Suite was the highest-performing system across the given datasets but the least scalable.
+Overall, the GAP Benchmark Suite was the highest-performing system across the given datasets though not the most scalable.
 However, this is only for relatively small graphs; all the graphs used here had at most $2^{22}$ vertices.
 It should be noted that GAP is also the most recent of projects.
-Thus, we recommend graph processing algorithm designers compare their implementations against the GAP Benchmark Suite as a good test of performance.
+Thus, we recommend graph processing algorithm designers compare their BFS implementations to the Graph500 for BFS, GAP Benchmark Suite for smaller problem sizes or a low number of cores (at most 32), and GraphMat for larger problems with more cores.
 
-We found that GraphMat is not highly performant but does scale relatively well.
-One explanation is that GraphMat's underlying computation model (sparse matrix operations), paired the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation, is better suited to larger-scale graphs.
+We found that GraphMat is relatively performant and scales better at higher cores.
+One explanation is that GraphMat's underlying computation model (sparse matrix operations), paired with the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation, is better suited to larger-scale graphs.
 % Furthermore, sparse matrix operations have been well studied so achieving good scalability is more realistic [cite]

### 2017-02-17 papers/publication/plot_data.R

@@ -95,7 +95,8 @@ measure_scale <- function(algo) {
 	# It is wasteful to reread the parsed*-1.csv but it simplifies the code
 	x <- read.csv(paste0("parsed",scale,"-1.csv"), header = FALSE)
 	colnames(x) <- c("Sys","Algo","Metric","Time")
-	systems <- unique(subset(x$Sys, x$Algo == algo, c("Sys")))
+	x$Sys <- factor(x$Sys, ordered = TRUE)
+	systems <- levels(subset(x$Sys, x$Algo == algo, c("Sys")))
 	algo_time <- data.frame(
 			matrix(ncol = length(threadcnts), nrow = length(systems)),
 			row.names = systems)
@@ -268,14 +269,12 @@ dev.off()
 ###
 # Part 4: Dota-league dataset with easy-parallel-graph
 ###
-# STILL NOT RIGHT
-dota <- read.csv("parseddota-32.csv", header = FALSE)
+dota <- read.csv("parseddota-32-redo.csv", header = FALSE)
 colnames(dota) <- c("Sys","Algo","Metric","Time")
-dota_times <- aggregate(dota, list(paste(dota$Sys, dota$Algo)), mean)
-bfs_time <- subset(dota, dota$Algo == "BFS" & dota$Metric == "Time",
-		c("Sys","Time"))
-sssp_time <- subset(x, x$Algo == "SSSP" & x$Metric == "Time",
-		c("Sys","Time"))
+dota$Sys <- factor(dota$Sys, ordered = TRUE)
+systems <- levels(subset(x$Sys, x$Algo == algo, c("Sys")))
+#dota_times <- subset(dota, dota$Metric == "Time", c("Sys","Algo","Time"))
+dota_means <- aggregate(dota$Time, list(dota$Sys, dota$Algo, dota$Metric), mean)
 
 ###
 # Part 5: Supplementary statistics used in the paper's prose.
@@ -307,5 +306,4 @@ sd(sssp_time$Time[sssp_time$Sys == "GraphMat"]) /
 #bfs_t_gap <- bfs_time[["Time"]][bfs_time[["Sys"]] == "GAP"]
 #vioplot(bfs_t_gb, bfs_t_gm, bfs_t_gap,
 #		names=c("GraphBIG", "GraphMat", "GAP"),
-#		col="bisque")
-
+#		col="bislevels

### 2017-02-17 papers/publication/paper-pollard.tex
Penultimate commit.

@@
 
-This study explores four shared memory parallel graph processing platforms.
+This study explores four shared memory parallel graph processing platforms and one distributed memory framework.
 The first three are so-called ``reference implementations'' while the remaining two are included because of their performance and popularity.
-Our target is shared memory CPU processing.
 Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} are not considered here because the authors do not provide reference implementations.
@@
 
-In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on the Dota-League dataset.
+In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$.
 An explanation of each algorithm is given in \cite{Iosup:2016:Graphalyticstech}.
-Table~\ref{tab:graphalytics} shows only a broad overview of the two systems: GraphBIG and PowerGraph.
-In general, we see GraphBIG performs better.
-However, these preliminary results do not show the complete picture.
-For example, one may want to know the cost of performing additional computations on a graph once the data structures have been built.
+Graphalytics by default does not perform SSSP on unweighted, undirected graphs.
+The discrepancy between PageRank values in Table~\ref{tab:graphalytics} and Fig.~\ref{fig:pr} is a result of the differing stopping criterion.
+
+In addition to the data presented in Table~\ref{tab:graphalytics}, one may want to know the cost of performing additional computations on a graph once the data structures have been built.
 Additionally, one may desire runtime bounds to get a reasonable maximum timespan for a given problem.
+Table~\ref{tab:epg-perf} begins to compile performance results and the rest of this section further explores these topics.
 
-%Table~\ref{tab:perf} lists performance in milliseconds of runtime according to the graphalytics output.
-Graphalytics also outputs MTEPS or millions of traversed edges pers econd.
+% Graphalytics also outputs MTEPS or millions of traversed edges per second.
 However, the graphalytics version does not make sense in all cases: for example, computing the local clustering coefficient involves traversing each edge multiple times (proportional to the sparsity of the graph), while BFS traverses each edge exactly once, and the number of edges traversed with PageRank depends on the connectivity of the graph and the number of iterations.
@@
 \begin{table}
-% ,graphmat,openg,powergraph
-%	Community Detection,0,212.5,1167
-%	PageRank,0,293.5,983
-%	Local Clustering Coefficient,0,316.5,1011
-%	Weakly Connected Components,0,89,767.5
-%	Single-Source Shortest Paths,11298.5,6378.5,34654.5
-	\caption{Performance results are in milliseconds. Community detection is performed using label propagation. At the time of this writing, Graphalytics only supports SSSP for GraphMat.}
+	\caption{The dataset used is the same Kronecker graph with scale $22$ as used in other experiments. Performance results are in seconds. Community detection is performed using label propagation.}
 	\centering
@@
 
-% Table~\ref{tab:epg-perf}, by contrast, gives an overview of the performance results of running Graphalytics and our approach using the same dataset [TODO: Get the Dota-League Dataset working with easy-parallel-graph and replace Table~\ref{epg-perf}].
-Table~\ref{tab:epg-perf} shows the average time across 32 roots, whereas Graphalytics by default shows results for a single root.
-%\begin{table}
-%	\centering
-%	\begin{tabular}{l|c|c|c}
-%		System & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular} & Run BFS & Run SSSP \\ \hline
-%		PowerGraph & N/A & 0.003380 \\
-%		GAP        & 0.412 & 0.001393 \\
-%		GraphMat   &  & 0.1031 \\
-%		GraphBIG   & N/A & 0.1528 \\
-%	\end{tabular}
-%	\caption{This table shows times for the Dota-League dataset using our method.}
-%	\label{tab:epg-perf}
-%\end{table}
+\begin{table}
+	\centering
+	\caption{This table shows averages of runtimes for the Dota-League dataset using our method.
+Each cell represents the average over 32 runs.
+The ``N/A'' for PowerGraph and GraphBIG exist because both of these systems read in the files and build a graph data structure simultaneously.
+GraphMat is not included because reading in the Dota-League dataset crashes the system.}
+	\begin{tabular}{l|c|c|c|c}
+		System & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular}
+		                   & Run BFS & Run SSSP & Run PageRank \\ \hline
+		PowerGraph & N/A   & N/A     & 0.344    & 3.606 \\
+		GAP        & 0.412 & 0.00464 & 0.267    & 33.00 \\
+%		GraphMat   & 1.013 &         &          &       \\
+		GraphBIG   & N/A   & 0.0476  & 0.162    & 21.39 \\
+	\end{tabular}
+	\label{tab:epg-perf}
+\end{table}
 
@@
 A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}.
-These plots show in general poor scaling throughout, although GraphBIG has the best scaling.
-A potential explanation is that because GraphBIG is the slowest system, there is more room for improvement.
+These plots show in general poor scaling throughout, a challenge facing most parallel graph algorithms.
 Additionally, $2^{20}$ vertices is a ``small'' graph by today's standards and thus library designers focus on scalability for larger graphs.
@@
 	\centering
-	\begin{minipage}{0.9\linewidth}
+	\begin{minipage}{\linewidth}
 		% trim=left lower right upper
@@
 	\end{minipage}
-	\begin{minipage}{0.9\linewidth}
+	\begin{minipage}{\linewidth}
 		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss.pdf}
@@
 \end{figure}
-\todo{Rerun GraphMat on scale 22 with 2 threads. Surprising that it slows down}
 
@@
 	\caption{Similar to Fig.~\ref{fig:bfs-time} we plot RAM and CPU Power Consumption for each of the 32 roots.
-Since the Graph500 runs multiple roots per execution, we only get a single datapoint.
+Since the Graph500 runs multiple roots per execution, we only get a single data point.
 The baseline was computed by monitoring power consumption while a program containing only one call to the C \texttt{unistd} library function \texttt{sleep(10)} (ten seconds).}
@@
 Thus, we recommend graph processing algorithm designers compare their BFS implementations to the Graph500 for BFS, GAP Benchmark Suite for smaller problem sizes or a low number of cores (at most 32), and GraphMat for larger problems with more cores.
+One last consideration is price: GraphMat requires the Intel compiler collection.
 

### 2017-02-17 papers/publication/paper-pollard.tex
EuroPar 2017 commit. Here goes nothing.

@@
 
-The behavior of PageRank is slightly different. As with SSSP and BFS, the GAP Benchmark Suite is the fastest but it also requires the fewest iterations.
-
+The behavior of PageRank is slightly different.
+As with SSSP and BFS, the GAP Benchmark Suite is the fastest but it also requires the fewest iterations.
 We attempt to define similar stopping criteria for each system, but GraphMat executes until no vertices change rank; effectively its stopping criterion requires the $\infty$-norm be less than machine epsilon.
 This could account for the increased number of iterations.
-We adjusted the other systems to use $\sum_{k=1}^{n} |p_k^{(i)} - p_k^{(i-1)}| < \epsilon $ as the stopping criteria, where $i$ is the iteration and $n$ is the number of vertices.
+We adjusted the other systems to use $\sum_{k=1}^{n} |p_k^{(i)} - p_k^{(i-1)}| < \epsilon $ as the stopping criterion, where $i$ is the iteration and $n$ is the number of vertices.
 We use $\epsilon = 6 \times 10^{-8}$ because this value is approximately machine epsilon for a single precision floating-point number.

### 2017-02-17 papers/publication/paper-pollard.tex
Re-assessed and resubmitted.

@@
 We analyze the performance of the algorithms described in Sec.~\ref{sec:algs} in terms of execution time, scalability over multiple threads, and power and energy consumption.
+All figures except those measuring scalability (Fig.~\ref{fig:bfs-scaling}) use 32 threads.
 
@@
 \begin{table}
-	\caption{The dataset used is the same Kronecker graph with scale $22$ as used in other experiments. Performance results are in seconds. Community detection is performed using label propagation.}
+	\caption{The dataset used is the same Kronecker graph with scale $22$ as used in other experiments.
+Performance results are in seconds running with the maximal number of threads (64 for Powergraph, 72 for the others).
+Community detection is performed using label propagation.}
 	\centering
@@
 	\centering
-	\caption{This table shows averages of runtimes for the Dota-League dataset using our method.
+	\caption{This table shows averages of runtimes for the Dota-League dataset using our method using 32 threads.
 Each cell represents the average over 32 runs.
@@
 It should be noted that GAP is also the most recent of projects.
-Thus, we recommend graph processing algorithm designers compare their BFS implementations to the Graph500 for BFS, GAP Benchmark Suite for smaller problem sizes or a low number of cores (at most 32), and GraphMat for larger problems with more cores.
+Thus, we recommend graph processing algorithm designers compare their BFS implementations to the Graph500 for BFS, GAP Benchmark Suite for smaller problem sizes or a low number of threads (at most 32), and GraphMat for larger problems with more threads.
+This is despite the fact that GraphBIG has the best scalability, because even at more than 16 threads the runtime for GraphBIG is still at least one order of magnitude slower than GAP or GraphMat for all sampled algorithms.
 One last consideration is price: GraphMat requires the Intel compiler collection.
 
-We found that GraphMat is relatively performant and scales better at higher cores.
+We found that GraphMat is only decently performant, but scales to more threads.
 One explanation is that GraphMat's underlying computation model (sparse matrix operations), paired with the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation, is better suited to larger-scale graphs.

### 2017-02-17 papers/publication/paper-pollard.tex
Made it 12, not 13 pages

@@
 
-We found that GraphMat is only decently performant, but scales to more threads.
+%We found that GraphMat is only decently performant, but scales to more threads.
 One explanation is that GraphMat's underlying computation model (sparse matrix operations), paired with the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation, is better suited to larger-scale graphs.

### 2017-05-03 autoperf/mst/plot_data.R
Fixed bug w/ gen-datasets, updated run & plot_data
Now plot_data.R is just for local stuff on arya and does not
generalize. Same for run; it shows which have been generated.

@@ -1,0 +1,46 @@
+# Given some output from ./run experiment then ./run parse
+# generate some plots of the data
+usage <- "usage: Rscript experiment_analysis.R <parsed_csv>"
+args <- commandArgs(trailingOnly = TRUE)
+if (length(args) != 1) {
+	stop(usage)
+}
+bpc <- "cyan"
+
+# Source provides scale and threads
+filename <- args[1]
+x <- read.csv(filename, header = TRUE)
+# Header:
+# algorithm,execution_phase,scale,edges_per_vertex,RMAT_type,insertion_percent,changed_vertices,threads,measurement,value
+scale <- 23
+pct_ins <- 75
+total_time <- subset(x,
+		x$execution_phase == "All" & x$scale == scale & x$insertion_percent == pct_ins,
+		c("algorithm","value"))
+time_cols <- c("algorithm","value")
+total_time$algorithm <- as.character(total_time$algorithm)
+boxplot(value~algorithm, data = total_time, ylab = "Time (seconds)",
+			main = "MST Total Time for scale 23, 75% insertions", col="green")
+
+# Assuming default edgefactor of 16
+# for (t in threads) {
+# 	print(paste0("Reading from ",filename))
+# 
+# 	colnames(x) <- c("Sys","Algo","Metric","Time")
+# 	
+# 	# Generate a figure
+# 	bfs_time <- subset(x, x$Algo == "BFS" & x$Metric == "Time",
+# 			c("Sys","Time"))
+# 	pdf(paste0("graphics/bfs_time", scale, "-", t, "t.pdf"),
+# 		width = 5.2, height = 5.2)
+# 	# The error 'adding class "factor" to an invalid object' probably means
+# 	# there was an error parsing so that extra entries were printed.
+# 	# this is caused by having too much in your log files. Just delete
+# 	# your output/kron-scale directory and start over
+# 	x <- read.csv(filename, header = FALSE)
+# 	boxplot(Time~Sys, bfs_time[,c("Time","Sys")], ylab = "Time (seconds)",
+# 			main = "BFS Time", log = "y", col=bpc)
+# 	mtext(paste0("nedges = ",nedges), side = 3)
+# 	dev.off()
+# }
+# 

### 2017-05-03 experiment/gen-datasets.sh

@@ -99,7 +99,9 @@ if [ "$FILE_PREFIX" != "kron-$S" ]; then
 	mkdir -p "$DDIR/$d"
 	echo Converting $FILE into the correct formats...
 	# If it's from SNAP then there may be some comments
-	awk '!/^#/{print}' "$FILE" > "$DDIR/$d.e"
+	if [ ! -f "$DDIR/$d.e" ]; then
+		awk '!/^#/{print}' "$FILE" > "$DDIR/$d.e"
+	fi
 	OLDPWD=$(pwd)
 	cd $DDIR
 	if [ ! -f "$d.v" ]; then
@@ -109,7 +111,7 @@ if [ "$FILE_PREFIX" != "kron-$S" ]; then
 	# nvertices is a bit of a misnomer; it should actually be "max vertex id"
 	nvertices=$(( $(sort -n "$d.v" | tail -n 1) + 1))
 	# nvertices=$(wc -l $d.v | awk '{print $1}') # Graphmat requires nvertices >= largest index
-	echo -n  "Checking whether weighted or unweighted..."
+	echo -n  "Checking whether $d.e is weighted or unweighted..."
 	if [ $(awk '{print NF; exit}' "$d.e") -eq 2 ]; then
 		echo " unweighted."
 		echo "SRC,DEST" > "$d/edge.csv"

### 2017-05-04 experiment/experiment_analysis.R
Small temporary updateds.

@@ -35,3 +35,12 @@ for (t in threads) {
 	dev.off()
 }
 
+# Weak scaling
+ws_scales <-  c(18,19,20,21,22,23,24)
+ws_threads <- c(1, 2, 4, 8, 16,32,64)
+for (t in ws_threads) {
+	for (s in wc_scales) {
+		filename <- paste0("./output/parsed-","kron-",scale,"-",t,".csv")
+	}
+}
+

### 2017-05-04 experiment/gen-datasets.sh

@@ -5,7 +5,7 @@ USAGE="usage: gen-datasets.sh [--libdir=<dir>] [--ddir=<dir>] -f=<fn>|<scale>
 		<edge1> <edge2> one per line OR
 		<edge1> <edge2> <weight> with an optional comment lines beginning with #
 	--libdir: repositories directory. Default: ./lib
-	--ddir: dataset directory. Default: ./datasets" # 2^{<scale>} = Number of vertices.
+	--ddir: dataset directory. Default: ./datasets"
 # Generate an unweighted, undirected Kronecker (RMAT) graph in the file formats
 # for graph500, GraphMat, GraphBIG, and GAP
 # for BFS, SSSP, and PageRank
@@ -122,14 +122,18 @@ if [ "$FILE_PREFIX" != "kron-$S" ]; then
 		# We don't want it to take 1 iteration.
 		echo "Getting roots."
 		"$GAPDIR/sssp" -f "$d/$d.el" -n $(($NRT*2)) > tmp.log
+		"$GAPDIR/converter" -s -f "$d/$d.el" -b "$d/$d.sg"
 		# GraphMat doesn't write out an unweighted graph. So we have output unit edge weights.
 		"$GRAPHMATDIR/bin/graph_converter" --selfloops 1 --duplicatededges 0 --bidirectional --inputformat 1 --outputformat 0 --inputheader 0 --outputheader 1 --inputedgeweights 0 --outputedgeweights 2 --nvertices $nvertices "$d/$d.el" "$d/$d.graphmat"
+		# We write a serialized graph to speed up GAP
 	elif [ $(awk '{print NF; exit}' "$d.e") -eq 3 ]; then
 		echo " weighted."
 		echo "SRC,DEST,WEIGHT" > "$d/edge.csv"
 		awk '{printf "%d %d\n", ($1+1), ($2+1) $3}' "$d.e" > "$d/$d.wel"
 		echo "Getting roots."
 		"$GAPDIR/sssp" -f "$d/$d.wel" -n $(($NRT*2)) > tmp.log
+		# TODO: Use this in real-datasets, change WeightT to float if need be and recompile GAPBS
+		"$GAPDIR/converter" -s -f "$d/$d.wel" -b "$d/$d.wsg"
 		# We make no assumptions so we output double precision edge weights
[…]

### 2017-05-04 experiment/parse-output.sh

@@ -9,11 +9,14 @@ USAGE="usage: parse-output.sh [--outdir=<dir>] -f=<fn>|<scale>
 	<scale> may be an integer between 1 and 99 inclusive.
 	--outdir: directory where the data is stored. default: ./output"
 
-OUTPUTDIR="$(pwd)/output"
+OUTDIR="$(pwd)/output"
 for arg in "$@"; do
 	case $arg in
 	--outdir=*)
-		OUTPUTDIR=${arg#*=}
+		OUTDIR=${arg#*=}
+		if [ ${OUTDIR:0:1} != '/' ]; then # Relative
+			OUTDIR="$(pwd)/$OUTDIR"
+		fi
 		shift
 	;;
 	-f=*)
@@ -47,17 +50,20 @@ if [ -z "$FILE" ]; then
 	esac
 fi
 
-OUTFN="$OUTPUTDIR/parsed-$FILE_PREFIX.csv"
-LOG_DIR="$OUTPUTDIR/$FILE_PREFIX"
+OUTFN="$OUTDIR/parsed-$FILE_PREFIX.csv"
+LOG_DIR="$OUTDIR/$FILE_PREFIX"
 # Assumes the experiment is run $NRT times ($NRT roots or $NRT PageRanks)
 NRT=32
 # Header: Sys,Algo,Metric,Time
 
 echo -n "Parsing GraphBIG for"
+echo "Clearing $OUTFN"
+rm -f "$OUTFN"
+
 for FN in $(find "$LOG_DIR" -maxdepth 1 -name '*t-GraphBIG-*.out'); do
 	f=$(basename $FN)
 	T=${f%%t[^0-9]*}
-	OUTFN="$OUTPUTDIR/parsed-$FILE_PREFIX-$T.csv"
+	OUTFN="$OUTDIR/parsed-$FILE_PREFIX-$T.csv"
 	ALGO=$(expr match "${f%.out}" '.*\(-.*\)')
 	ALGO=${ALGO#-}
 	echo -n "; $ALGO $T threads"
@@ -82,7 +88,7 @@ echo -n "Parsing Graph500 for BFS, threadcounts"
 for FN in $(find "$LOG_DIR" -maxdepth 1 -name '*t-Graph500-BFS.out'); do
 	f=$(basename $FN)
 	T=${f%%t[^0-9]*}
[…]

### 2017-05-04 experiment/real-datasets.sh

@@ -66,6 +66,8 @@ else
 fi
 echo Starting experiment at $(date)
 
+echo "Cleaning $OUTPUT_PREFIX"
+rm -f "${OUTPUT_PREFIX}-{GAP,GraphMat,PowerGraph}-{BFS,SSSP,PR}.out"
 echo "Running GAP BFS"
 # It would be nice if you could read in a file for the roots
 head -n $NRT "$DDIR/$d/${d}-roots.v" > "$DDIR/$d/${d}-${NRT}roots.v"

### 2017-05-04 experiment/run-experiment.sh

@@ -11,11 +11,13 @@
 USAGE="usage: run-experiment.sh [--libdir=<dir>] [--ddir=<dir>] <scale> <num-threads>
 	scale: 2^scale = number of vertices
 	--libdir: repositories directory. Default: ./lib
-	--ddir: dataset directory. Default: ./datasets" # 2^{<scale>} = Number of vertices.
+	--ddir: dataset directory. Default: ./datasets
+	--outdir: output directory. Default: ./output"
 
 # The edge factor (number of edges per vertex) is the default of 16.
 DDIR="$(pwd)/datasets" # Dataset directory
 LIBDIR="$(pwd)/lib"
+OUTDIR="$(pwd)/output"
 for arg in "$@"; do
 	case $arg in
 	--libdir=*)
@@ -24,6 +26,13 @@ for arg in "$@"; do
 	;;
 	--ddir=*)
 		DDIR=${arg#*=}
+		shift
+	;;
+	--outdir=*)
+		OUTDIR=${arg#*=}
+		if [ ${OUTDIR:0:1} != '/' ]; then # Relative
+			OUTDIR="$(pwd)/$OUTDIR"
+		fi
 		shift
 	;;
 	-h|--help|-help)
@@ -47,8 +56,8 @@ GRAPHBIGDIR="$LIBDIR/graphBIG"
 GRAPH500DIR="$LIBDIR/graph500"
 GRAPHMATDIR="$LIBDIR/GraphMat"
 POWERGRAPHDIR="$LIBDIR/PowerGraph"
-OUTPUT_PREFIX="$(pwd)/output/kron-$S/${OMP_NUM_THREADS}t"
-mkdir -p "$(pwd)/output/kron-$S"
+OUTPUT_PREFIX="$OUTDIR/kron-$S/${OMP_NUM_THREADS}t"
+mkdir -p "$OUTDIR/kron-$S"
 
 # Set some other variables used throughout the experiment
 # PageRank is usually represented as a 32-bit float,
@@ -62,6 +71,9 @@ export SKIP_VALIDATION=1
 
 # Load all the modules here
 module load intel/17
+
+echo "Cleaning $OUTPUT_PREFIX-*.out"
+rm -f "${OUTPUT_PREFIX}-{GAP,GraphMat,PowerGraph}-{BFS,SSSP,PR}.out"
 
[…]

### 2017-05-04 papers/publication/paper-pollard.tex
OK, final EuroPar 2017 commit.
Updated paper to IEEE format.
It's in a pretty rough transitionary period right now.

@@
-\documentclass{llncs}
+\documentclass[conference]{IEEEtran}
 \usepackage[T1]{fontenc}
@@
 \begin{document}
-\title{A Comparison of Parallel Graph Processing Benchmarks}
-\author{Samuel D. Pollard \and Boyana Norris}
-\institute{University of Oregon \\
-	Eugene OR 97403, USA \\
-	\email{spollard@cs.uoregon.edu} \\
-	\email{norris@cs.uoregon.edu}
-}
+\title{A Comparison of Parallel Graph Processing Implementations}
+\author{
+	\IEEEauthorblockN{Samuel D. Pollard}
+	\IEEEauthorblockA{Computer and Information Science\\
+		University of Oregon \\
+		Eugene, OR 97403\\
+		Email: spollard@cs.uoregon.edu}
+	\and
+	\IEEEauthorblockN{Boyana Norris}
+		\IEEEauthorblockA{Computer and Information Science\\
+		University of Oregon \\
+		Eugene, OR 97403\\
+		Email: norris@cs.uoregon.edu}}
 \maketitle
@@
 To enable easy, rigorous, and repeatable comparison of the capabilities of such systems, we present an approach and associated software for analyzing the performance and scalability of parallel, open-source graph libraries.
-We demonstrate our approach on five graph frameworks: GraphMat, the Graph500, the Graph Algorithm Platform, GraphBIG, and PowerGraph using synthetic and real-world datasets.
-We examine previously overlooked aspects of parallel graph processing performance such as phases of execution and energy usage for three algorithms: breadth first search, single source shortest paths, and PageRank and compare our results to Graphalytics.
+We demonstrate our approach on five graph frameworks: GraphMat, the Graph500, the Graph Algorithm Platform Benchmark Suite, GraphBIG, and PowerGraph using synthetic and real-world\todo{actual real-world.
+(i) Michael LeBeane et al, Supercomputing (SC 15), Data Partitioning Strategies for Graph Workloads on Heterogeneous Clusters (ii) Shuang Song et al,  ICPP 2016, Proxy-Guided Load Balancing of Graph Processing Workloads on Heterogeneous Clusters.} datasets.
+We examine previously overlooked aspects of parallel graph processing performance such as phases of execution and energy usage for three algorithms\todo{more algorithms, justify why these algorithms}: breadth first search, single source shortest paths, and PageRank and compare our results to Graphalytics.
 \end{abstract}
@@
 
-From a graph algorithm user's point of view, there are optimizations for each system which may not be apparent.
+From a graph algorithm user's point of view, optimizations for each system may not be apparent.
 For example, the Parallel Boost Graph Library (PBGL)~\cite{Gregor:2005:PBGL} provides generic implementations of their algorithms and the programmer must provide the template specializations.
-The optimal data structures may differ across algorithms and graphs and must be determined by the programmer.
+The optimal algorithms may differ across graphs and must be determined by the programmer.
+To complicate matters further, the $\Delta$-stepping implementation of the parallel Dijkstra's Algorithm for computing single source shortest paths (SSSP) is parametrized by some $\Delta > 0$ for which requires tuning for a given architecture and graph structure to achieve optimal performance \cite{Panitanarak:2014:SSSPPerf}.\todo{Will this be mentioned again?}
+
+We aim to simplify this excess of decisions by both providing results from our experiments on real world and synthetic datasets.
+We take inspiration from the Graph500 benchmark \cite{Murphy:2010:Graph500} which clearly specifies every step of the breadth first search (BFS) algorithm and how it should be timed.
+The Graph500 can be used to fairly rank parallel graph processing platforms.
+We define platforms \todo{as graphalytics does?
+Or not.} as the combination of hardware, operating system, and graph processing library.
+We refer the reader to Section~[SECTION] for the platforms used in this paper.
+Thus, a potential user of parallel graph algorithms should have no difficulty selecting the optimal library given a wide variety of graphs and performance data available on his or her system.
+
+We select a small number of graph processing platforms over which we attempt to homogenize all aspects of execution to truly compare performance.
 
 %% How do we address it?
-To mitigate the inherent bias in hand-generated implementations, all the experiments performed use the author-provided implementations; we assume the developers of each system will provide the most performant implementation.
-While this limits the scope of the experiments it mitigates the bias inherent in our programming skills.
+All the experiments performed use the author-provided implementations with modifications only to insert performance analysis hooks or to ensure homogeneous termination criteria; we assume the developers of each system will provide the best performing implementation.
+While this limits the scope of the experiments it mitigates the bias inherent in our programming skills.\todo{Should I talk about how library designers are biased?}
 
@@
 \begin{itemize}
-	\item Automation of the installation, configuration, and dataset conversion to ensure the experiments execute in the same manner.
-Essentially, we provide a ``level playing field'' for graph algorithms.
+	\item Automation of the installation\todo{This is sort of done in the repository, but not really explained in the paper}, configuration, and dataset conversion to ensure the experiments execute in the same manner.
+Essentially, we provide a ``level playing field'' for graph algorithms.\todo{put URL to github, make public}
 	\item Comparison of algorithm runtime and scalability using several data sets.
@@
 \end{itemize}
+\todo{Clarify that we're not introducing our own benchmark suite instead we're comparing existing frameworks at a greater depth.}
 
 \section{Related Work}
+\todo{Introduction sentence.
+How many others are there?
+There's a large number of qualatative.
+Most third party comparisons are qualitative and just describe the overview instead of actual runtimes.
+This isn't a level playing field.}
 A related project is Graphalytics \cite{Capota:2015:Graphalytics}, which also attempts to automate the setup and execution of graph packages for performance analysis.
@@
 These all require some knowledge of the inner workings of each system in addition to familiarity with the Graphalytics API and involve a nontrivial amount of programming.
+
+\todo{Cite the Exploring big graph computing - An empirical study from architectural perspective, but we dont do an architectural perspective.
+Don't care about distinctino between framework/usercode since we're just using the vendor provided implementations.}
+Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
 
@@
 
-Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
-
 Unlike these approaches, we focus on the \emph{depth} of analysis rather than the breadth via diversity of datasets or graph processing systems.
@@
 
-\section{Experimental Setup}
+\section{Methodology}
+
+\todo{Start with dataset subsection, write new intro to methodology, move this into inputs.
+We're comparing algorithms not applications.
+Understand those since they're the building blocks.
+Not comparing architectural details (cite that Nai paper).
+Important paragraph.}
 We refer to a graph's \emph{scale} when describing the size of the graph.
 Specifically, a graph with scale $S$ has $2^S$ vertices.
-The datasets are described further in Sect.~\ref{sec:datasets}.
+\todo{Say 2 to the S = 2351501351035923} The datasets are described further in Sect.~\ref{sec:datasets}.
 We measure strong scaling and speedup by varying the number of threads from one to the total number of threads on our server, 72.
@@
 	\item PowerGraph \cite{Gonzalez:2012:Powergraph}, a framework for distributed graph-parallel computation on natural graphs.
+	\todo{Explain each more.}
 \end{itemize}
@@
 Our approach is not specific or limited to these graph packages and can be extended to others with relatively modest effort (building each package and if necessary, instrumenting the different phases of the computation).
+\todo{Say what we automate and what we don't automate.}
+\todo{Weighted versus unweighted}
 
@@
 One challenge with using PageRank is the stopping criterion; this is addressed in Sect.~\ref{sec:perf}.
-Verification of the PageRank results is beyond the scope of this paper, although this may explain some of the large performance discrepancies.
+Verification of the PageRank results is beyond the scope of this paper, although this may explain some of the large performance discrepancies.\todo{Better explain why these were selected.
+They're the most popular of the algorithms for the papers we surveyed.}
 
@@
 
-The Graphalytics results in Table~\ref{tab:graphalytics} were performed on the Dota-League dataset.
+The Graphalytics results in Table~\ref{tab:graphalytics} were performed on the Dota-League dataset\todo{One reviewer confused this dataset with the Kronecker.}.
 This dataset contains 61,670 vertices and 50,870,313 edges.
@@
 
-We analyze the performance of the algorithms described in Sec.~\ref{sec:algs} in terms of execution time, scalability over multiple threads, and power and energy consumption.
+We analyze the performance of the algorithms described in Sec.~\ref{sec:algs} in terms of execution time, scalability over multiple threads, and power and energy consumption\todo{ A precise list and exposition of the requirements and changes	done to the application would help in understanding the methodology and its implications}.
 All figures except those measuring scalability (Fig.~\ref{fig:bfs-scaling}) use 32 threads.
@@
 
-In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$.
+In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately 33,554,432 edges)\todo{work?}.
 An explanation of each algorithm is given in \cite{Iosup:2016:Graphalyticstech}.
 Graphalytics by default does not perform SSSP on unweighted, undirected graphs.
-The discrepancy between PageRank values in Table~\ref{tab:graphalytics} and Fig.~\ref{fig:pr} is a result of the differing stopping criterion.
+The discrepancy between PageRank values in Table~\ref{tab:graphalytics} and Fig.~\ref{fig:pr} is a result of the differing stopping criterion.\todo{What are the likely culprits for performance differences---measure cache miss rates.
+Try to find something which correlates with performance.
+Resource stalls will work easily (not as interesting) Cycles per instruction?
+Automate with perfexplorer or autoperf}
 
@@
 
+\todo{The y-axis is mislabeled. Should be "iterations" for pagerank.}
 \begin{figure}
@@
 \begin{table}
-	\caption{The data are generated using a Kronecker graph with a scale of 22.
+	\caption{The data are generated using a Kronecker graph with a scale of 22 and using 32 threads.
 Sleeping Energy refers to the power (in Watts) consumed during the \texttt{unistd} C \texttt{sleep} function, multiplied by the Time row.
@@
 RAPL also allows the measurement of DRAM power, the results of which are shown in Fig.~\ref{fig:power}.
-The left plot describes the second row of Table~\ref{tab:power} in more detail.
+The right plot describes the second row of Table~\ref{tab:power} in more detail.
 We notice a smaller spread of RAM power consumption, but still a noticeable difference.
@@
 \section{Future Work and Conclusion}
+\todo{Maybe; it's still difficult to add more graphs to the dataset sinc you need to know all the metadata.}
 We have presented a systematic approach to analyzing parallel graph processing performance, both measuring power consumption and the two fundamental phases of execution.

### 2017-05-05 papers/publication/paper-pollard.tex
Updated Sections I and II.

@@
 		Eugene, OR 97403\\
-		Email: norris@cs.uoregon.edu}}
+		Email: norris@cs.uoregon.edu}
+}
 \maketitle
@@
 \section{Introduction}
-
 %% What's the problem (motivation)
 Our research is motivated by the current state of parallel graph processing.
-The most comprehensive survey, released in 2014, identified and categorized over 80 different parallel graph processing systems, without considering domain specific languages \cite{Doekemeijer:2015:GPFSurvey}.
+The most comprehensive survey, released in 2014, identified and categorized over 80 different parallel graph processing systems, not even considering domain specific languages \cite{Doekemeijer:2015:GPFSurvey}.
 
 An overarching issue among these systems is the lack of comprehensive comparisons.
-One possible reason is the considerable effort involved in getting each system to run: satisfying dependencies and ensuring data are correctly formatted are generally nontrivial tasks.
+One possible reason is the considerable effort involved in getting each system to run: satisfying dependencies and ensuring data are correctly formatted can be time consuming tasks.
 Additionally, some systems are developed with large, multi-node clusters in mind while others only work with shared memory, single node computers.
@@
 For example, the Parallel Boost Graph Library (PBGL)~\cite{Gregor:2005:PBGL} provides generic implementations of their algorithms and the programmer must provide the template specializations.
-The optimal algorithms may differ across graphs and must be determined by the programmer.
-To complicate matters further, the $\Delta$-stepping implementation of the parallel Dijkstra's Algorithm for computing single source shortest paths (SSSP) is parametrized by some $\Delta > 0$ for which requires tuning for a given architecture and graph structure to achieve optimal performance \cite{Panitanarak:2014:SSSPPerf}.\todo{Will this be mentioned again?}
+Optimal data structures may differ across graphs and must be determined by the programmer.
+To complicate matters further, the $\Delta$-stepping implementation of the parallel Dijkstra's Algorithm for computing single source shortest paths is parametrized by some $\Delta > 0$ for which requires tuning for a given architecture and graph structure to achieve optimal performance \cite{Panitanarak:2014:SSSPPerf}.
+\todo{Will this be mentioned again?}
 
-We aim to simplify this excess of decisions by both providing results from our experiments on real world and synthetic datasets.
+We aim to simplify the decision making process by providing results from our experiments on both real world and synthetic datasets.
 We take inspiration from the Graph500 benchmark \cite{Murphy:2010:Graph500} which clearly specifies every step of the breadth first search (BFS) algorithm and how it should be timed.
-The Graph500 can be used to fairly rank parallel graph processing platforms.
-We define platforms \todo{as graphalytics does?
-Or not.} as the combination of hardware, operating system, and graph processing library.
-We refer the reader to Section~[SECTION] for the platforms used in this paper.
-Thus, a potential user of parallel graph algorithms should have no difficulty selecting the optimal library given a wide variety of graphs and performance data available on his or her system.
+The Graph500 can be used to fairly rank systems on a single, well-defined benchmark.
+In this context, \emph{system} is defined as the hardware, operating system, middleware, and algorithmic implementation which accomplishes a certain task, in this case BFS.
 
-We select a small number of graph processing platforms over which we attempt to homogenize all aspects of execution to truly compare performance.
+While leaderboards can indicate computational milestones, they are not particularly useful for the average user.
+Reference implementations, however, are critical for advancement; the complexity of today's hardware and operating systems require empirical evidence to assess an algorithm's performance.
+Additionally, there is motivation to define basic building blocks for graph computations \cite{GABB16, Buluc:CombBLAS:2011}.
+However, new algorithm designers and users alike have no easy way of determining what constitutes a high-performance implementation.
 
-%% How do we address it?
-All the experiments performed use the author-provided implementations with modifications only to insert performance analysis hooks or to ensure homogeneous termination criteria; we assume the developers of each system will provide the best performing implementation.
-While this limits the scope of the experiments it mitigates the bias inherent in our programming skills.\todo{Should I talk about how library designers are biased?}
-
-%To address the challenges of getting various systems installed, we automate the process as much as possible n
-%To showcase our methods we analyze the performance of several graph processing benchmarks to facilitate the selection among such systems. Our analysis is done with minimal source code modification.
-
-Our research contributions in this paper can be summarized as follows.
-\begin{itemize}
-	\item Automation of the installation\todo{This is sort of done in the repository, but not really explained in the paper}, configuration, and dataset conversion to ensure the experiments execute in the same manner.
-Essentially, we provide a ``level playing field'' for graph algorithms.\todo{put URL to github, make public}
-	\item Comparison of algorithm runtime and scalability using several data sets.
-	\item Comparison of our framework with the Graphalytics \cite{Capota:2015:Graphalytics} project.
-	% BN: I think the bullet below is a bit too low-level for this summary, moving to later section.
-	%\item Inspection of the source code of the surveyed parallel graph processing systems to ensure the same phases of execution are measured across differing execution and programming paradigms.
-	\item An experimental framework distinguishing between the data structure generation and algorithm execution phases of each system.
-	\item Analysis of energy consumption during the algorithm execution phase.
-\end{itemize}
-\todo{Clarify that we're not introducing our own benchmark suite instead we're comparing existing frameworks at a greater depth.}
+To address these issues, we present \emph{easy-parallel-graph\mbox{-\textasteriskcentered}}
+% Here, \textasteriskcentered\ could be \emph{processing}, \emph{performance-analysis}, or \emph{algorithm-comparison})
+\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}}
+which simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
+With this tool we select a small number of graph processing libraries over which we homogenize all aspects of execution to fairly compare performance.
+Additionally, we present an example of this framework via analysis of the performance, energy consumption, and scalability results to enable algorithm designers to select high performance reference implementations with which they can compare their improvements.
+To be clear, we are not proposing a new benchmark suite or providing any reference implementations.
+We provide a framework to compare existing software packages without requiring the user to be familiar with them.
 
 \section{Related Work}
-\todo{Introduction sentence.
-How many others are there?
-There's a large number of qualatative.
-Most third party comparisons are qualitative and just describe the overview instead of actual runtimes.
-This isn't a level playing field.}
-A related project is Graphalytics \cite{Capota:2015:Graphalytics}, which also attempts to automate the setup and execution of graph packages for performance analysis.
+\todo{There's a large number of qualitative. Most third party comparisons are qualitative and just describe the overview instead of actual runtimes. This isn't a level playing field.}
+Most performance analysis of graph processing systems come from the empirical results of each new library designer's publications.
+For example, GraphMat \cite{Sundaram:2015:GraphMat}, PowerGraph \cite{Gonzalez:2012:Powergraph}, and GraphBIG \cite{Nai:2015:Graphbig} present a comparison of performance between their system and a selection of other software packages.
+
+Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
+
+Lifeng Nai et al.~\cite{Nai:2016:architectural} provide a detailed performance analysis using GraphBIG as their reference implementation, profile Intel hardware counters, and measure the bottlenecks to measure scalability, computational efficiency for a variety of datasets.
+
+Beyond this plethora of performance data, the Graph Algorithm Platform Benchmark Suite (GAP) \cite{Beamer:2015:GAPBench}, GraphBIG, and the Graph500 \cite{Murphy:2010:Graph500} consider themselves reference implementations or benchmark suites next to which other implementations can be compared.
+If no fewer than three software packages call themselves, ``reference implementations,'' which one are we to trust?
+Our belief is choosing a single reference implementation cannot capture the complexities inherent with graph processing.
+
+There do exist some performance analyses not tied to any particular library.
+The most prominent example is Graphalytics \cite{Capota:2015:Graphalytics} which like our work also automates the setup and execution of graph packages for performance analysis.
 However, in order for Graphalytics to analyze a graph processing system---Graphalytics calls these platforms---a programmer must implement Java classes as wrappers to call the particular implementations, satisfy dependencies, and set the correct shell commands to execute.
 These all require some knowledge of the inner workings of each system in addition to familiarity with the Graphalytics API and involve a nontrivial amount of programming.
-
-\todo{Cite the Exploring big graph computing - An empirical study from architectural perspective, but we dont do an architectural perspective.
-Don't care about distinctino between framework/usercode since we're just using the vendor provided implementations.}
-Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
 
@@
 
-Unlike these approaches, we focus on the \emph{depth} of analysis rather than the breadth via diversity of datasets or graph processing systems.
-We analyze  power and energy consumption in addition to details such as the number of iterations or the time to construct the data structures.
+Our contribution strikes a balance between the highly detailed analysis of a single implementation such as those presented by Nai \cite{Nai:2016:architectural} (more depth of analysis) and Graphalytics which focuses on the runtimes of several implementations and several platforms (more breadth of analysis).
+This allows us to greatly simplify deployment and allows the user to make his or her own decisions.
+
+In addition, we analyze  power and energy consumption as well as details such as the number of iterations or the time to construct the data structures.
 Moreover, our approach requires little knowledge of the inner workings of each system; the data are collected using either hardware counter sampling of model-specific registers (for power) or through parsing log files (for execution time).
+This allows arbitrary datasets to be included as long as they are in the same, simple format as those in the Stanford Network Analysis Project (SNAP) datasets \cite{snapnets}.
 
@@
 \section{Methodology}
+Automation of the installation\todo{This is sort of done in the repository, but not really explained in the paper}, configuration, and dataset conversion to ensure the experiments execute in the same manner.
+
+%% How do we address these issues?
+Our experiments use the author-provided implementations with modifications only to insert performance analysis hooks or to ensure homogeneous stopping criteria; we assume the developers of each system will provide the best performing implementation.
+While this limits the scope of the experiments it mitigates the bias inherent in our programming skills.\todo{Should I talk about how library designers are biased?}
 
@@
 
-\bibliographystyle{splncs03}
+\bibliographystyle{IEEEtran}
 \bibliography{../drp}
 \end{document}
+
+% EuroPar Reviews
+%----------------------- REVIEW 1 ---------------------
+%PAPER: 204
+%TITLE: A Comparison of Parallel Graph Processing Benchmarks
+%AUTHORS: Samuel Pollard and Boyana Norris
+%
+%Overall evaluation: 0 (borderline paper)
+%
+%----------- Overall evaluation -----------
+%Summary:
+%Due to the large amount of existing graph processing systems, comprehensive comparisons need considerable effort.
+Therefore, it is difficult for developers to select the best approach.
+This paper presents an approach and associated software to analyze the performance and scalability of parallel, open-source graph libraries.
+%
+%Strengths:
+%1. The motivation is clear, as this paper want to provide an approach to comprehensively compare the graph processing systems.
+%2. It is a well-written paper.
+%
+%Weakness:
+%1. In the contribution, you mentioned “automation of the installation”, however, I don’t think it is well demonstrated in this paper.
+%2.
+All the graph data are synthetic graphs generated via Kronecker gen.
+As far as I know, these synthetic graphs have a very different degree distribution compared to real graphs.
+Can you add some real graphs in addition to synthetic.
+Many graphs are available at SNAP (Stanford).
+%3.
+Some figures are confusing.
+For example, the right figure in “Figure 3” is named as PageRank Iterations.
+The value on y-axis is the time per iteration or the total execution time?
+If the value is the time for all iterations, then what is the difference between the left figure and right figure in “Figure 3”?
+%
+%Recommendations and Questions:
+%1. Graph algorithms on each platform are implemented in different manners. When you directly compare their performance, it should be accompanied with deeper explanations.
+%2. For strong scaling studies, there is no deep-dive to explain the poor scaling behaviors in GAP.
+%3. Why use scale 20 graph for scaling study and scale 22 graph for other study? I think demonstrating the scaling behavior, we need the relatively big graph input.
+%4.
+Most execution time for graph processing is spent on the framework ("Exploring big graph computing — An empirical study from architectural perspective", published in Journal of Parallel and Distributed Computing, 2016), focusing and comparing the framework can be another contribution that this paper can make.
+%5.
+Use graphs from SNAP, as in the following papers.
+Please cite those papers as well.
+(i) Michael LeBeane et al, Supercomputing (SC 15), Data Partitioning Strategies for Graph Workloads on Heterogeneous Clusters (ii) Shuang Song et al,  ICPP 2016, Proxy-Guided Load Balancing of Graph Processing Workloads on Heterogeneous Clusters.
+%
+%
+%----------------------- REVIEW 2 ---------------------
+%PAPER: 204
+%TITLE: A Comparison of Parallel Graph Processing Benchmarks
+%AUTHORS: Samuel Pollard and Boyana Norris
+%
+%Overall evaluation: 1 (weak accept)
+%
+%----------- Overall evaluation -----------
+%The paper performs a benchmarking study on Graph processing frameworks.
+The authors ran a large group of benchmarks to compare several frameworks in terms of the runtime and energy consumption using three different algorithm.
+%
+%While the paper is useful for audience who are looking for a starting point benchmark to select their Graph Processing framework.
+%
+%
+%The paper has a good flow of information, however adding the following sections could make help the general audience to understand the value of the study.
+%
+%--> A comparison between the 5 graph processing frame works
+%--> Justification about the selection of graph size - real world examples of graphs with these sizes of nodes/vertices.
+%--> A more fine grained  analysis of the results and reasoning behind observed performance.
+%
+%Overall,the benchamrking study is of value for the research and mainly developer communities by helping them for selecting their framework.
+%
+%
+%----------------------- REVIEW 3 ---------------------
+%PAPER: 204
+%TITLE: A Comparison of Parallel Graph Processing Benchmarks
+%AUTHORS: Samuel Pollard and Boyana Norris
+%
+%Overall evaluation: -2 (reject)
+%
+%----------- Overall evaluation -----------
+%This paper presents a comparison of runtime, scalability and energy
+%consumption of 5 graph frameworks for 3 different algorithms,
+%distinguishing the data-structure generation and algorithm execution
+%phases.  Their experiments show that the fastest runtime is obtained
+%with GAP Benchmark Suite for all of the 3 algorithms.  The most
+%scalable framework is GraphBIG. Every frameworks show poor
+%scalability beyond 32 threads.  In term of power and energy
+%consumption the GAP benchmark is the most energy efficient whereas
+%GraphMat exhibits the lowest average CPU and DRAM power consumption.
+%
+%
+%The paper felt a bit confusing. To begin with, I could not find a clear
+%exposition of the requirements and the methodology used and what are its
+%implications. A precise list and exposition of the requirements and changes
+%done to the application would help in understanding this. This is very critical
+%due to the paper positionning itself as a comparison of multiple benchmarks.
+%
+%As for the experimental part, it is as confusing as the exposition of the
+%contribution. Notably, contrarily to what the authors stated, I did not felt
+%that they actually compare themselves to the results of graphalytics, they
+%show some numbers from graphalytics but they are scarcely commented and they
+%do not compare precisely the results to their methodology, highlight the
+%differences and improvements, etc. A second problem concerns the size of the
+%benchmarks run. In the conclusion, the authors say that they could only compare
+%graphs of size 22 because of GraphBIG, but for the sake of this study why not
+%exclude GraphBIG altogether in this case in order to have more results ? A third
+%problem concerns the graph types. I understood that the graph of scale 22 is the
+%Kronecker graph, and using elimination my guess is that the graph of scale 20 is
+%the Dota-League Graph, but nowhere is this clearly stated which adds to the
+%general confusion. Overall, using two graph sizes and types is an important limitation
+%for graph algorithms. The paper could clearly benefit
+%from a more in depth study of multiple graph types with different
+%topologies, looking at the behavior of the multiple implementations with
+%different graph sizes, etc.
+%
+%
+%Details:
+%Section 4.2: The left plot describes the second row of Table 3 -> the right plot ...
+%
+%
+%----------------------- REVIEW 4 ---------------------
+%PAPER: 204
+%TITLE: A Comparison of Parallel Graph Processing Benchmarks
+%AUTHORS: Samuel Pollard and Boyana Norris
+%
+%Overall evaluation: -2 (reject)
+%
+%----------- Overall evaluation -----------
+%In this work, the authors attempt to compare parallel graph processing benchmarks.
+However, in fact, the paper proposes a benchmark of its own and pursues to demonstrate that it is superior to other benchmarks in terms of usability and performance.
+%While the importance of graph benchmarking is increasing given the large number of systems and algorithms that need performance comparison, the paper falls short in delivering a clear comparison between existing benchmarks, and enough novel contributions to propose a new benchmark.
+Besides the brief explanation regarding power consumption, which is not common in the other benchmarks, there's no originality in this work.
+%
+%I have the following comments and concerns:
+%- there's no criteria for comparing the benchmarks.
+While there's a discussion about usability and a separate discussion about failures and performance, the claim that the paper provides a rigorous comparison between these suites is not made true in the paper itself.
+%- the comparison, as well as the newly proposed benchmark, seem to also ignore the distributed aspects of some of the frameworks, which other benchmarks take into account.
+%- the newly proposed benchmark also has design flaws: the algorithms are not diverse enough, neither are the graphs.
+The authors do not give any explanation as to why such choices have been made, and therefore the proposed benchmark seems much more limited than all the previous ones.
+%- the paper claims the proposed solution does not have the complexity of Graphalytics, where users need to write Java wrappers.
+However, in their case, PAPI instrumentation is needed, which might be (a) difficult to do, (b) not feasible in a production environment for users with insufficient credentials.
+How do these challenges affect the usability of the proposed method?

### 2017-05-09 papers/publication/paper-pollard.tex
Updated paper, started rewriting Methodology.

@@
 To enable easy, rigorous, and repeatable comparison of the capabilities of such systems, we present an approach and associated software for analyzing the performance and scalability of parallel, open-source graph libraries.
-We demonstrate our approach on five graph frameworks: GraphMat, the Graph500, the Graph Algorithm Platform Benchmark Suite, GraphBIG, and PowerGraph using synthetic and real-world\todo{actual real-world.
+We demonstrate our approach on five graph processing packages: GraphMat, the Graph500, the Graph Algorithm Platform Benchmark Suite, GraphBIG, and PowerGraph using synthetic and real-world\todo{actual real-world.
 (i) Michael LeBeane et al, Supercomputing (SC 15), Data Partitioning Strategies for Graph Workloads on Heterogeneous Clusters (ii) Shuang Song et al,  ICPP 2016, Proxy-Guided Load Balancing of Graph Processing Workloads on Heterogeneous Clusters.} datasets.
@@
 % Here, \textasteriskcentered\ could be \emph{processing}, \emph{performance-analysis}, or \emph{algorithm-comparison})
-\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}}
-which simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
+\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}}, a utility which simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
 With this tool we select a small number of graph processing libraries over which we homogenize all aspects of execution to fairly compare performance.
@@
 
-Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
-
-Lifeng Nai et al.~\cite{Nai:2016:architectural} provide a detailed performance analysis using GraphBIG as their reference implementation, profile Intel hardware counters, and measure the bottlenecks to measure scalability, computational efficiency for a variety of datasets.
-
-Beyond this plethora of performance data, the Graph Algorithm Platform Benchmark Suite (GAP) \cite{Beamer:2015:GAPBench}, GraphBIG, and the Graph500 \cite{Murphy:2010:Graph500} consider themselves reference implementations or benchmark suites next to which other implementations can be compared.
-If no fewer than three software packages call themselves, ``reference implementations,'' which one are we to trust?
-Our belief is choosing a single reference implementation cannot capture the complexities inherent with graph processing.
-
-There do exist some performance analyses not tied to any particular library.
-The most prominent example is Graphalytics \cite{Capota:2015:Graphalytics} which like our work also automates the setup and execution of graph packages for performance analysis.
-However, in order for Graphalytics to analyze a graph processing system---Graphalytics calls these platforms---a programmer must implement Java classes as wrappers to call the particular implementations, satisfy dependencies, and set the correct shell commands to execute.
-These all require some knowledge of the inner workings of each system in addition to familiarity with the Graphalytics API and involve a nontrivial amount of programming.
+The most prominent example of a graph processing performance analysis tool not tied to a particular iplementation is Graphalytics \cite{Capota:2015:Graphalytics}.
+Like our work, Graphalytics also automates the setup and execution of graph packages for performance analysis.
+Graphalyics relies on Apache Maven and Java to wrap each graph processing software package---Graphalytics calls these platforms---and requires configuration files for each dataset.\todo{What am I saying with this?}
 
@@
 
-Our contribution strikes a balance between the highly detailed analysis of a single implementation such as those presented by Nai \cite{Nai:2016:architectural} (more depth of analysis) and Graphalytics which focuses on the runtimes of several implementations and several platforms (more breadth of analysis).
+Nai et al.~\cite{Nai:2016:architectural} provide a detailed performance analysis using GraphBIG as their reference implementation, profile Intel hardware counters, and measure the bottlenecks to measure scalability, computational efficiency for a variety of datasets.
+
+Beyond this plethora of performance data, the Graph Algorithm Platform Benchmark Suite (GAP) \cite{Beamer:2015:GAPBench}, GraphBIG, and the Graph500 \cite{Murphy:2010:Graph500} consider themselves reference implementations or benchmark suites next to which other implementations can be compared.
+If no fewer than three software packages call themselves, ``reference implementations,'' which one are we to trust?
+Our belief is choosing a single reference implementation cannot capture the complexities inherent with graph processing.
+
+Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
+
+Our contribution strikes a balance between the highly detailed analysis of a single implementation such as those presented by Nai \cite{Nai:2016:architectural} (greater depth of analysis) and Graphalytics which focuses on the runtimes of several implementations and several platforms (more breadth of analysis).
 This allows us to greatly simplify deployment and allows the user to make his or her own decisions.
@@
 \section{Methodology}
-Automation of the installation\todo{This is sort of done in the repository, but not really explained in the paper}, configuration, and dataset conversion to ensure the experiments execute in the same manner.
-
-%% How do we address these issues?
-Our experiments use the author-provided implementations with modifications only to insert performance analysis hooks or to ensure homogeneous stopping criteria; we assume the developers of each system will provide the best performing implementation.
-While this limits the scope of the experiments it mitigates the bias inherent in our programming skills.\todo{Should I talk about how library designers are biased?}
-
 \todo{Start with dataset subsection, write new intro to methodology, move this into inputs.
@@
 Important paragraph.}
+Our utility is broken down into five principal phases indicated in \figurename~\ref{fig:epg-overview}, each of which requires no more than a single shell command. The five phases are as follows:
+\begin{enumerate}
+	\item Installing modified, stable forks of each software package to ensure homogeneity.
+	\item Given a synthetic problem size or a real-world graph file, generate the files necessary to run each software package.
+	\item Given a graph and the number of threads, run each algorithm using each software package multiple times.
+	\item Parse through the log files to compress the output into a CSV.
+	\item Analyze the data using the provided R scripts to generate plots.
+\end{enumerate}
+
+\begin{figure*}
+	\centering
+		% trim=left lower right upper
+		\includegraphics[width=\linewidth, trim=0 144pt 0pt 156pt, clip]{graphics/overview.eps}
+	\caption{Overview of \emph{easy-parallel-graph-\textasteriskcentered}. Each orange box corresponds to a single shell script in the package while the green ellipses correspond to generated files.}
+	\label{fig:epg-overview}
+\end{figure*}
+
+\subsection{Installing Libraries}
+The libraries are stable forks of the given repositories which are configured to ensure the experiments execute in the same manner.
+Similarly, multiple copies of the datasets are stored to ensure the datasets are formatted correctly.
+For example, the roots for BFS are selected such that they are part of a nontrivial connected component.
+
+%% How do we address these issues?
+Our experiments use the author-provided implementations with modifications only to insert performance analysis hooks or to ensure homogeneous stopping criteria; we assume the developers of each system will provide the best performing implementation.
+While this limits the scope of the experiments it mitigates the bias inherent in our programming skills in addition to the bias of library designers' performance analysis; a given library designer will understand his or her source code better than any other implementation.
+
+\subsection{Datasets}
+Some software packages provide serialized file formats to speed up file IO. Homogenizing the datasets creates extra copies of the graph files.
+
 We refer to a graph's \emph{scale} when describing the size of the graph.
 Specifically, a graph with scale $S$ has $2^S$ vertices.
-\todo{Say 2 to the S = 2351501351035923} The datasets are described further in Sect.~\ref{sec:datasets}.
+For example, many of our experiments were run with $2^{22} = 4,194,304$ vertices and an average of 16 edges per vertex.
+The datasets are described further in Sect.~\ref{sec:datasets}.
 We measure strong scaling and speedup by varying the number of threads from one to the total number of threads on our server, 72.
@@
 This is not possible for PowerGraph and GraphBIG because they read in the input file and build a graph simultaneously.
-The Graph500 builds the graph once and performs the BFS on each root in turn, speeding up experimentation but only giving a single data point for the dataset construction.
 
-\subsection{Machine Specifications}
-We performed experiments on our 36-core Intel Haswell server whose specifications are outlined below.
-%Table~\ref{tab:specs} shows the specifications of the server on which we performed all experiments.
-%\begin{table}[h!tb]
-%	\centering
-%	\caption{
-The operating system is GNU/Linux version 4.4.0-22.
-The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
-We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.
-\begin{center}
-	% For arya I deleted Max RAM Freq	2133MHz
-	\pgfplotstabletypeset[
-	header=false,
-	col sep=tab,
-	string type,
-	every head row/.style={output empty row, before row=\bottomrule},
-	columns/0/.style={column type={|p{1in}|}},
-	columns/1/.style={column type={p{3.2in}|}},
-	every last row/.style={after row=\toprule},
-	]{../report/specs.csv}
-	%\label{tab:specs}
-\end{center}
+We use the Graph500 synthetic graph generator which creates a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19, C = 0.19,$ and $D = 1-(A+B+C) = 0.05$ and set the average degree of a vertex as 16.
+Hence, a Kronecker graph with scale $S$ has $2^S$ vertices and approximately $16 \times 2^S$ edges.
+
+The Graphalytics results in Table~\ref{tab:graphalytics} were performed on the Dota-League dataset\todo{One reviewer confused this dataset with the Kronecker.}.
+This dataset contains 61,670 vertices and 50,870,313 edges.
+This dataset comes from the Game Trace Archive\cite{Guo:2012:GTA} and is modified for Graphalytics\footnote{This dataset is available at \url{https://atlarge.ewi.tudelft.nl/graphalytics/}.}.
+This dataset is useful because it is both weighted and more dense than the usual real-world dataset with an average out-degree of 824.
 
@@
 
-\subsection{Datasets}\label{sec:datasets}
-We use the Graph500 synthetic graph generator which creates a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19, C = 0.19,$ and $D = 1-(A+B+C) = 0.05$ and set the average degree of a vertex as 16.
-Hence, a Kronecker graph with scale $S$ has $2^S$ vertices and approximately $16 \times 2^S$ edges.
+\subsection{Parsing and Data Analysis}
+The example workflow provided in our implementation yields the figures generated in this paper.
 
-The Graphalytics results in Table~\ref{tab:graphalytics} were performed on the Dota-League dataset\todo{One reviewer confused this dataset with the Kronecker.}.
-This dataset contains 61,670 vertices and 50,870,313 edges.
-This dataset comes from the Game Trace Archive\cite{Guo:2012:GTA} and is modified for Graphalytics\footnote{This dataset is available at \url{https://atlarge.ewi.tudelft.nl/graphalytics/}.}.
-This dataset is useful because it is both weighted and more dense than the usual real-world dataset with an average out-degree of 824.
+\subsection{Machine Specifications}
+We performed experiments on our 36-core (72 thread) Intel Haswell server with 256GB DDR4 RAM, with two Intel Xeon E5-2699 v3 CPUs. The operating system is GNU/Linux version 4.4.0-22.
+%The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
+We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.
+%\begin{center}
+%	% For arya I deleted Max RAM Freq	2133MHz
+%	\pgfplotstabletypeset[
+%	header=false,
+%	col sep=tab,
+%	string type,
+%	every head row/.style={output empty row, before row=\bottomrule},
+%	columns/0/.style={column type={|p{1in}|}},
+%	columns/1/.style={column type={p{3.2in}|}},
+%	every last row/.style={after row=\toprule},
+%	]{../report/specs.csv}
+%	%\label{tab:specs}
+%\end{center}
 

### 2017-05-09 papers/publication/paper-pollard.tex
Updated methodology section in paper.

@@
 
+Nai et al.~\cite{Nai:2016:architectural} provide a detailed performance analysis using GraphBIG as their reference implementation.
+Their approach also compares GraphBIG to the GraphLab and Pregel execution models.
+Their analysis considers architectural performance measurements such as cache miss rates to measure bottlenecks and computational efficiency for a variety of datasets.
+While incredibly thorough, these analyses focus on covering a wide range of datasets and appear difficult to apply to a new approach.
+For example, GraphMat reduce computation to sparse matrix operations which may not suffer from the same memory bottlenecks indicated in the results of Nai et al.
+
+Beyond the aforementioned performance results there are a number of existing reports on performance analysis not tied to a particular software package we presently describe.
+
+Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
+These implementations are hand tuned and provides recommendations for future improvements.
+
 The most prominent example of a graph processing performance analysis tool not tied to a particular iplementation is Graphalytics \cite{Capota:2015:Graphalytics}.
@@
 
-Nai et al.~\cite{Nai:2016:architectural} provide a detailed performance analysis using GraphBIG as their reference implementation, profile Intel hardware counters, and measure the bottlenecks to measure scalability, computational efficiency for a variety of datasets.
-
 Beyond this plethora of performance data, the Graph Algorithm Platform Benchmark Suite (GAP) \cite{Beamer:2015:GAPBench}, GraphBIG, and the Graph500 \cite{Murphy:2010:Graph500} consider themselves reference implementations or benchmark suites next to which other implementations can be compared.
@@
 Our belief is choosing a single reference implementation cannot capture the complexities inherent with graph processing.
-
-Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
 
@@
 The libraries are stable forks of the given repositories which are configured to ensure the experiments execute in the same manner.
-Similarly, multiple copies of the datasets are stored to ensure the datasets are formatted correctly.
-For example, the roots for BFS are selected such that they are part of a nontrivial connected component.
 
@@
 For example, many of our experiments were run with $2^{22} = 4,194,304$ vertices and an average of 16 edges per vertex.
-The datasets are described further in Sect.~\ref{sec:datasets}.
 We measure strong scaling and speedup by varying the number of threads from one to the total number of threads on our server, 72.
@@
 
-The Graphalytics results in Table~\ref{tab:graphalytics} were performed on the Dota-League dataset\todo{One reviewer confused this dataset with the Kronecker.}.
+\todo{More real world datasets}
+The Graphalytics results in Table~\ref{tab:graphalytics} were performed on the Dota-League datasets.
 This dataset contains 61,670 vertices and 50,870,313 edges.
@@
 PageRank is selected because of its popularity; most libraries provide reference implementations.
-One challenge with using PageRank is the stopping criterion; this is addressed in Sect.~\ref{sec:perf}.
+One challenge with using PageRank is the stopping criterion; all implementations have been modified to use $|| p_t - p_{t-1}||_1$ (the absolute sum of differences) where $p_t$ is the pagerank at step $t$.
 Verification of the PageRank results is beyond the scope of this paper, although this may explain some of the large performance discrepancies.\todo{Better explain why these were selected.
@@
 \subsection{Parsing and Data Analysis}
-The example workflow provided in our implementation yields the figures generated in this paper.
+Our example analyzes data using the R programming language and shell scripting parsers. The example workflow provided in our utility yields the figures generated in this paper.
 
@@
 
-In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately 33,554,432 edges)\todo{work?}.
+In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately $16 \times 2^22 = 33,500,000$ edges)\todo{work?}.
 An explanation of each algorithm is given in \cite{Iosup:2016:Graphalyticstech}.

### 2017-05-10 papers/publication/paper-pollard.tex
some minor fixes to get paper to build for me, also started adding notes

@@
 \usepackage{todonotes}
+
+\newcommand{\todob}[1]{\todo[inline,color=green!40]{#1}}
 
@@
 To enable easy, rigorous, and repeatable comparison of the capabilities of such systems, we present an approach and associated software for analyzing the performance and scalability of parallel, open-source graph libraries.
-We demonstrate our approach on five graph processing packages: GraphMat, the Graph500, the Graph Algorithm Platform Benchmark Suite, GraphBIG, and PowerGraph using synthetic and real-world\todo{actual real-world.
+We demonstrate our approach on five graph processing packages: GraphMat, the Graph500, the Graph Algorithm Platform Benchmark Suite, GraphBIG, and PowerGraph using synthetic and real-world\todo[inline]{actual real-world.
 (i) Michael LeBeane et al, Supercomputing (SC 15), Data Partitioning Strategies for Graph Workloads on Heterogeneous Clusters (ii) Shuang Song et al,  ICPP 2016, Proxy-Guided Load Balancing of Graph Processing Workloads on Heterogeneous Clusters.} datasets.
-We examine previously overlooked aspects of parallel graph processing performance such as phases of execution and energy usage for three algorithms\todo{more algorithms, justify why these algorithms}: breadth first search, single source shortest paths, and PageRank and compare our results to Graphalytics.
+We examine previously overlooked aspects of parallel graph processing performance such as phases of execution and energy usage for three algorithms\todo[inline]{more algorithms, justify why these algorithms}: breadth first search, single source shortest paths, and PageRank and compare our results to Graphalytics.
 \end{abstract}
@@
 %% What's the problem (motivation)
+\todob{need an introductory sentence about large graph applications and resulting proliferation of parallel frameworks}
 Our research is motivated by the current state of parallel graph processing.
@@
 One possible reason is the considerable effort involved in getting each system to run: satisfying dependencies and ensuring data are correctly formatted can be time consuming tasks.
-Additionally, some systems are developed with large, multi-node clusters in mind while others only work with shared memory, single node computers.
+Moreover, some systems are developed with large, multi-node clusters in mind while others only work with shared memory, single-node computers.
 An example of the former is GraphX \cite{Xin:2013:GraphX}, which incurs some overhead while gaining fault tolerance and an example of the latter is Ligra \cite{Shun:2013:Ligra}, a framework requiring a shared-memory architecture.
@@
 To complicate matters further, the $\Delta$-stepping implementation of the parallel Dijkstra's Algorithm for computing single source shortest paths is parametrized by some $\Delta > 0$ for which requires tuning for a given architecture and graph structure to achieve optimal performance \cite{Panitanarak:2014:SSSPPerf}.
-\todo{Will this be mentioned again?}
+\todo[inline]{Will this be mentioned again?}
 
@@
 While leaderboards can indicate computational milestones, they are not particularly useful for the average user.
-Reference implementations, however, are critical for advancement; the complexity of today's hardware and operating systems require empirical evidence to assess an algorithm's performance.
+Reference implementations, however, are critical for advancement; the complexity of today's hardware and operating systems requires empirical evidence to assess an algorithm's performance.
 Additionally, there is motivation to define basic building blocks for graph computations \cite{GABB16, Buluc:CombBLAS:2011}.
@@
 \section{Related Work}
-\todo{There's a large number of qualitative. Most third party comparisons are qualitative and just describe the overview instead of actual runtimes. This isn't a level playing field.}
+\todo[inline]{There's a large number of qualitative. Most third party comparisons are qualitative and just describe the overview instead of actual runtimes. This isn't a level playing field.}
 Most performance analysis of graph processing systems come from the empirical results of each new library designer's publications.
@@
 
-Beyond the aforementioned performance results there are a number of existing reports on performance analysis not tied to a particular software package we presently describe.
+Beyond the aforementioned performance results, a number of reports focus on performance analysis not tied to a particular software package we presently describe.
 
@@
 Like our work, Graphalytics also automates the setup and execution of graph packages for performance analysis.
-Graphalyics relies on Apache Maven and Java to wrap each graph processing software package---Graphalytics calls these platforms---and requires configuration files for each dataset.\todo{What am I saying with this?}
+Graphalyics relies on Apache Maven and Java to wrap each graph processing software package---Graphalytics calls these platforms---and requires configuration files for each dataset.\todo[inline]{What am I saying with this?}
 
@@
 \section{Methodology}
-\todo{Start with dataset subsection, write new intro to methodology, move this into inputs.
+\todo[inline]{Start with dataset subsection, write new intro to methodology, move this into inputs.
 We're comparing algorithms not applications.
@@
 
-\todo{More real world datasets}
+\todo[inline]{More real world datasets}
 The Graphalytics results in Table~\ref{tab:graphalytics} were performed on the Dota-League datasets.
@@
 	\item PowerGraph \cite{Gonzalez:2012:Powergraph}, a framework for distributed graph-parallel computation on natural graphs.
-	\todo{Explain each more.}
+	\todo[inline]{Explain each more.}
 \end{itemize}
@@
 Our approach is not specific or limited to these graph packages and can be extended to others with relatively modest effort (building each package and if necessary, instrumenting the different phases of the computation).
-\todo{Say what we automate and what we don't automate.}
-\todo{Weighted versus unweighted}
+\todo[inline]{Say what we automate and what we don't automate.}
+\todo[inline]{Weighted versus unweighted}
 
@@
 One challenge with using PageRank is the stopping criterion; all implementations have been modified to use $|| p_t - p_{t-1}||_1$ (the absolute sum of differences) where $p_t$ is the pagerank at step $t$.
-Verification of the PageRank results is beyond the scope of this paper, although this may explain some of the large performance discrepancies.\todo{Better explain why these were selected.
+Verification of the PageRank results is beyond the scope of this paper, although this may explain some of the large performance discrepancies.\todo[inline]{Better explain why these were selected.
 They're the most popular of the algorithms for the papers we surveyed.}
@@
 
-We analyze the performance of the algorithms described in Sec.~\ref{sec:algs} in terms of execution time, scalability over multiple threads, and power and energy consumption\todo{ A precise list and exposition of the requirements and changes	done to the application would help in understanding the methodology and its implications}.
+We analyze the performance of the algorithms described in Sec.~\ref{sec:algs} in terms of execution time, scalability over multiple threads, and power and energy consumption\todo[inline]{ A precise list and exposition of the requirements and changes	done to the application would help in understanding the methodology and its implications}.
 All figures except those measuring scalability (Fig.~\ref{fig:bfs-scaling}) use 32 threads.
@@
 
-In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately $16 \times 2^22 = 33,500,000$ edges)\todo{work?}.
+In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately $16 \times 2^22 = 33,500,000$ edges)\todo[inline]{work?}.
 An explanation of each algorithm is given in \cite{Iosup:2016:Graphalyticstech}.
 Graphalytics by default does not perform SSSP on unweighted, undirected graphs.
-The discrepancy between PageRank values in Table~\ref{tab:graphalytics} and Fig.~\ref{fig:pr} is a result of the differing stopping criterion.\todo{What are the likely culprits for performance differences---measure cache miss rates.
+The discrepancy between PageRank values in Table~\ref{tab:graphalytics} and Fig.~\ref{fig:pr} is a result of the differing stopping criterion.\todo[inline]{What are the likely culprits for performance differences---measure cache miss rates.
 Try to find something which correlates with performance.
@@
 
-\todo{The y-axis is mislabeled. Should be "iterations" for pagerank.}
+\todo[inline]{The y-axis is mislabeled. Should be "iterations" for pagerank.}
 \begin{figure}
@@
 \section{Future Work and Conclusion}
-\todo{Maybe; it's still difficult to add more graphs to the dataset sinc you need to know all the metadata.}
+\todo[inline]{Maybe; it's still difficult to add more graphs to the dataset sinc you need to know all the metadata.}
 We have presented a systematic approach to analyzing parallel graph processing performance, both measuring power consumption and the two fundamental phases of execution.
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{IEEEtranS}
 \bibliography{../drp}

### 2017-05-10 papers/publication/paper-pollard.tex
some edits in intro

@@
 \usepackage{hyperref}
-\usepackage{cite} % Multiple citations in one \cite command
+\usepackage{cite} % Multiple citations in one~\cite command
 \usepackage{pgfplotstable}
@@
 Our research is motivated by the current state of parallel graph processing.
-The most comprehensive survey, released in 2014, identified and categorized over 80 different parallel graph processing systems, not even considering domain specific languages \cite{Doekemeijer:2015:GPFSurvey}.
+The most comprehensive survey, released in 2014, identified and categorized over 80 different parallel graph processing systems, not even considering domain specific languages~\cite{Doekemeijer:2015:GPFSurvey}.
 
@@
 Moreover, some systems are developed with large, multi-node clusters in mind while others only work with shared memory, single-node computers.
-An example of the former is GraphX \cite{Xin:2013:GraphX}, which incurs some overhead while gaining fault tolerance and an example of the latter is Ligra \cite{Shun:2013:Ligra}, a framework requiring a shared-memory architecture.
+An example of the former is GraphX~\cite{Xin:2013:GraphX}, which incurs some overhead while gaining fault tolerance and an example of the latter is Ligra~\cite{Shun:2013:Ligra}, a framework requiring a shared-memory architecture.
 % ``The project was motivated by the fact that the largest publicly available real-world graphs all fit in shared memory''
@@
 Optimal data structures may differ across graphs and must be determined by the programmer.
-To complicate matters further, the $\Delta$-stepping implementation of the parallel Dijkstra's Algorithm for computing single source shortest paths is parametrized by some $\Delta > 0$ for which requires tuning for a given architecture and graph structure to achieve optimal performance \cite{Panitanarak:2014:SSSPPerf}.
+To complicate matters further, the $\Delta$-stepping implementation of the parallel Dijkstra's Algorithm for computing single source shortest paths is parametrized by some $\Delta > 0$ for which requires tuning for a given architecture and graph structure to achieve optimal performance~\cite{Panitanarak:2014:SSSPPerf}.
 \todo[inline]{Will this be mentioned again?}
@@
 We aim to simplify the decision making process by providing results from our experiments on both real world and synthetic datasets.
-We take inspiration from the Graph500 benchmark \cite{Murphy:2010:Graph500} which clearly specifies every step of the breadth first search (BFS) algorithm and how it should be timed.
-The Graph500 can be used to fairly rank systems on a single, well-defined benchmark.
+We take inspiration from the Graph500 benchmark~\cite{Murphy:2010:Graph500} which clearly specifies every step of the breadth first search (BFS) algorithm and how it should be timed.
+The Graph500 can be used to rank systems fairly on a single, well-defined benchmark.
 In this context, \emph{system} is defined as the hardware, operating system, middleware, and algorithmic implementation which accomplishes a certain task, in this case BFS.
@@
 Reference implementations, however, are critical for advancement; the complexity of today's hardware and operating systems requires empirical evidence to assess an algorithm's performance.
-Additionally, there is motivation to define basic building blocks for graph computations \cite{GABB16, Buluc:CombBLAS:2011}.
+Additionally, there is motivation to define basic building blocks for graph computations~\cite{GABB16, Buluc:CombBLAS:2011}.
 However, new algorithm designers and users alike have no easy way of determining what constitutes a high-performance implementation.
@@
 With this tool we select a small number of graph processing libraries over which we homogenize all aspects of execution to fairly compare performance.
-Additionally, we present an example of this framework via analysis of the performance, energy consumption, and scalability results to enable algorithm designers to select high performance reference implementations with which they can compare their improvements.
+Additionally, we present an example use of this framework to analyze the performance, energy consumption, and scalability of specific algorithm implementation.
+This enables algorithm designers to select high performance reference implementations with which they can compare their improvements.
 To be clear, we are not proposing a new benchmark suite or providing any reference implementations.
-We provide a framework to compare existing software packages without requiring the user to be familiar with them.
+We introduce a framework for detailed and fair comparison of existing graph software packages without requiring the user to be familiar with them.
 
@@
 Most performance analysis of graph processing systems come from the empirical results of each new library designer's publications.
-For example, GraphMat \cite{Sundaram:2015:GraphMat}, PowerGraph \cite{Gonzalez:2012:Powergraph}, and GraphBIG \cite{Nai:2015:Graphbig} present a comparison of performance between their system and a selection of other software packages.
+For example, GraphMat~\cite{Sundaram:2015:GraphMat}, PowerGraph~\cite{Gonzalez:2012:Powergraph}, and GraphBIG~\cite{Nai:2015:Graphbig} present a comparison of performance between their system and a selection of other software packages.
 
@@
 
-Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
+Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and~\cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
 These implementations are hand tuned and provides recommendations for future improvements.
 
-The most prominent example of a graph processing performance analysis tool not tied to a particular iplementation is Graphalytics \cite{Capota:2015:Graphalytics}.
+The most prominent example of a graph processing performance analysis tool not tied to a particular iplementation is Graphalytics~\cite{Capota:2015:Graphalytics}.
 Like our work, Graphalytics also automates the setup and execution of graph packages for performance analysis.
@@
 
-With a plugin to Graphalytics called Granula \cite{Ngai:2015:Granula}, one can explicitly specify a performance model to analyze specific execution behavior such as the amount of communication or runtime of particular kernels of execution.
+With a plugin to Graphalytics called Granula~\cite{Ngai:2015:Granula}, one can explicitly specify a performance model to analyze specific execution behavior such as the amount of communication or runtime of particular kernels of execution.
 This requires in-depth knowledge of the source code and execution model.
@@
 
-Beyond this plethora of performance data, the Graph Algorithm Platform Benchmark Suite (GAP) \cite{Beamer:2015:GAPBench}, GraphBIG, and the Graph500 \cite{Murphy:2010:Graph500} consider themselves reference implementations or benchmark suites next to which other implementations can be compared.
+Beyond this plethora of performance data, the Graph Algorithm Platform Benchmark Suite (GAP)~\cite{Beamer:2015:GAPBench}, GraphBIG~\cite{Nai:2015:Graphbig}, and the Graph500~\cite{Murphy:2010:Graph500} consider themselves reference implementations or benchmark suites next to which other implementations can be compared.
 If no fewer than three software packages call themselves, ``reference implementations,'' which one are we to trust?
@@
 
-Our contribution strikes a balance between the highly detailed analysis of a single implementation such as those presented by Nai \cite{Nai:2016:architectural} (greater depth of analysis) and Graphalytics which focuses on the runtimes of several implementations and several platforms (more breadth of analysis).
+Our contribution strikes a balance between the highly detailed analysis of a single implementation such as those presented by Nai~\cite{Nai:2016:architectural} (greater depth of analysis) and Graphalytics which focuses on the runtimes of several implementations and several platforms (more breadth of analysis).
 This allows us to greatly simplify deployment and allows the user to make his or her own decisions.
@@
 Moreover, our approach requires little knowledge of the inner workings of each system; the data are collected using either hardware counter sampling of model-specific registers (for power) or through parsing log files (for execution time).
-This allows arbitrary datasets to be included as long as they are in the same, simple format as those in the Stanford Network Analysis Project (SNAP) datasets \cite{snapnets}.
+This allows arbitrary datasets to be included as long as they are in the same, simple format as those in the Stanford Network Analysis Project (SNAP) datasets~\cite{snapnets}.
 
-% The systems described in \cite{Doekemeijer:2015:GPFSurvey} operate with a wide range of parallelism paradigms and target architectures such as GPU \cite{Zhong:2014:Medusa, Kang:2009:Pegasus}, shared memory CPU \cite{Shun:2013:Ligra, kyrola:2012:Graphchi, Nguyen:2013:Galois}, a combination of CPU and GPU \cite{Gharaibeh:2012:Totem}, distributed filesystem based approaches \cite{Xin:2013:GraphX}, and distributed memory with MPI \cite{Gregor:2005:PBGL}.
+% The systems described in~\cite{Doekemeijer:2015:GPFSurvey} operate with a wide range of parallelism paradigms and target architectures such as GPU~\cite{Zhong:2014:Medusa, Kang:2009:Pegasus}, shared memory CPU~\cite{Shun:2013:Ligra, kyrola:2012:Graphchi, Nguyen:2013:Galois}, a combination of CPU and GPU~\cite{Gharaibeh:2012:Totem}, distributed filesystem based approaches~\cite{Xin:2013:GraphX}, and distributed memory with MPI~\cite{Gregor:2005:PBGL}.
 
-% Beyond the systems described by Doekemeijer and Varbanescu, the problem has compounded with the addition of even more proprietary and open source projects such as \cite{Cheramangalath:2015:Falcon, Perez:2015:Ringo}, distributed memory approaches such as \cite{Hong:2015:PGX}.
-domain-specific languages \cite{Hong:2012:GreenMarl}, distributed database querying, \cite{Rodriguez:2015:Gremlin}, as well as novel communication schemes \cite{Edmonds:2013:ActiveMessages}.
+% Beyond the systems described by Doekemeijer and Varbanescu, the problem has compounded with the addition of even more proprietary and open source projects such as~\cite{Cheramangalath:2015:Falcon, Perez:2015:Ringo}, distributed memory approaches such as~\cite{Hong:2015:PGX}.
+domain-specific languages~\cite{Hong:2012:GreenMarl}, distributed database querying,~\cite{Rodriguez:2015:Gremlin}, as well as novel communication schemes~\cite{Edmonds:2013:ActiveMessages}.
 At the outset, this plethora of choices makes the question, ``which system is the best for my problem?'' daunting.
 
-% In addition to libraries with associated APIs there has also been a propagation of ``reference implementations'' which implement the most common graph algorithms such as \cite{Beamer:2015:GAPBench, Nai:2015:Graphbig}.
+% In addition to libraries with associated APIs there has also been a propagation of ``reference implementations'' which implement the most common graph algorithms such as~\cite{Beamer:2015:GAPBench, Nai:2015:Graphbig}.
 Thus, even selecting a standard and a benchmark over which to compare various implementations is nontrivial.
@@
 
-We use the Graph500 synthetic graph generator which creates a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19, C = 0.19,$ and $D = 1-(A+B+C) = 0.05$ and set the average degree of a vertex as 16.
+We use the Graph500 synthetic graph generator which creates a Kronecker graph~\cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19, C = 0.19,$ and $D = 1-(A+B+C) = 0.05$ and set the average degree of a vertex as 16.
 Hence, a Kronecker graph with scale $S$ has $2^S$ vertices and approximately $16 \times 2^S$ edges.
@@
 The first three are so-called ``reference implementations'' while the remaining two are included because of their performance and popularity.
-Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} are not considered here because the authors do not provide reference implementations.
+Other popular libraries such as the Parallel Boost Graph Library~\cite{Gregor:2005:PBGL} are not considered here because the authors do not provide reference implementations.
 We consider the following systems.
 \begin{itemize}
-	\item The Graph500\footnote{We used the most recent version from \url{https://github.com/graph500/graph500}, most similar to release 2.1.4.} \cite{Murphy:2010:Graph500}, a set of large-scale graph benchmarks.
-	\item The Graph Algorithm Platform (GAP) Benchmark Suite \cite{Beamer:2015:GAPBench}.
-	\item GraphBIG \cite{Nai:2015:Graphbig} benchmark suite.
-	\item GraphMat \cite{Sundaram:2015:GraphMat}, a high-performance graph library.
-	\item PowerGraph \cite{Gonzalez:2012:Powergraph}, a framework for distributed graph-parallel computation on natural graphs.
+	\item The Graph500\footnote{We used the most recent version from \url{https://github.com/graph500/graph500}, most similar to release 2.1.4.}~\cite{Murphy:2010:Graph500}, a set of large-scale graph benchmarks.
+	\item The Graph Algorithm Platform (GAP) Benchmark Suite~\cite{Beamer:2015:GAPBench}.
+	\item GraphBIG~\cite{Nai:2015:Graphbig} benchmark suite.
+	\item GraphMat~\cite{Sundaram:2015:GraphMat}, a high-performance graph library.
+	\item PowerGraph~\cite{Gonzalez:2012:Powergraph}, a framework for distributed graph-parallel computation on natural graphs.
 	\todo[inline]{Explain each more.}
@@
 
-We select BFS because the canonical performance leaderboard for parallel graph processing is the Graph500 \cite{Murphy:2010:Graph500}.
+We select BFS because the canonical performance leaderboard for parallel graph processing is the Graph500~\cite{Murphy:2010:Graph500}.
 The advantage of the Graph500 is that it provides standardized measurement specifications and dataset generation.
@@
 In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately $16 \times 2^22 = 33,500,000$ edges)\todo[inline]{work?}.
-An explanation of each algorithm is given in \cite{Iosup:2016:Graphalyticstech}.
+An explanation of each algorithm is given in~\cite{Iosup:2016:Graphalyticstech}.
 Graphalytics by default does not perform SSSP on unweighted, undirected graphs.
@@
 The data structure construction times for GAP and GraphMat are consistent; in both cases the platforms create the same data structure for both algorithms.
-These results are consistent with \cite{Sundaram:2015:GraphMat} which lists GraphMat as more performant than PowerGraph in SSSP.
+These results are consistent with~\cite{Sundaram:2015:GraphMat} which lists GraphMat as more performant than PowerGraph in SSSP.
 
@@
 \subsection{Power and Energy Consumption}
-We use the Performance Application Programming Interface (PAPI) \cite{Browne:2000:PAPI} to gain access to Intel's Running Average Power Limit (RAPL), which provides a set of hardware counters for measuring energy usage.
+We use the Performance Application Programming Interface (PAPI)~\cite{Browne:2000:PAPI} to gain access to Intel's Running Average Power Limit (RAPL), which provides a set of hardware counters for measuring energy usage.
 We use PAPI to obtain average energy in nanojoules for a given time interval.
@@
 This limited the sizes of the experiments we could run.
-Graphalytics also had circumstances with the more computationally expensive algorithms  where certain experiments fail \cite{Iosup:2016:Graphalyticstech}, so determining whether an algorithm will finish given a particular machine, input size, runtime limit, and resources is an important unanswered question we plan to pursue further.
+Graphalytics also had circumstances with the more computationally expensive algorithms  where certain experiments fail~\cite{Iosup:2016:Graphalyticstech}, so determining whether an algorithm will finish given a particular machine, input size, runtime limit, and resources is an important unanswered question we plan to pursue further.
 

### 2017-05-11 experiment/gen-datasets.sh
Fixed bug with weighted edge list

@@ -88,7 +88,7 @@ if [ -z "$FILE" ]; then
 fi
 
 mkdir -p "$DDIR"
-export OMP_NUM_THREADS=32
+export OMP_NUM_THREADS=64
 NRT=64
 GAPDIR="$LIBDIR/gapbs"
 GRAPH500DIR="$LIBDIR/graph500"
@@ -129,7 +129,7 @@ if [ "$FILE_PREFIX" != "kron-$S" ]; then
 	elif [ $(awk '{print NF; exit}' "$d.e") -eq 3 ]; then
 		echo " weighted."
 		echo "SRC,DEST,WEIGHT" > "$d/edge.csv"
-		awk '{printf "%d %d\n", ($1+1), ($2+1) $3}' "$d.e" > "$d/$d.wel"
+		awk '{printf "%d %d %s\n", ($1+1), ($2+1), $3}' "$d.e" > "$d/$d.wel"
 		echo "Getting roots."
 		"$GAPDIR/sssp" -f "$d/$d.wel" -n $(($NRT*2)) > tmp.log
 		# TODO: Use this in real-datasets, change WeightT to float if need be and recompile GAPBS

### 2017-05-11 experiment/real-datasets.sh

@@ -66,7 +66,7 @@ else
 fi
 echo Starting experiment at $(date)
 
-echo "Cleaning $OUTPUT_PREFIX"
+echo "Cleaning $OUTPUT_PREFIX-*"
 rm -f "${OUTPUT_PREFIX}-{GAP,GraphMat,PowerGraph}-{BFS,SSSP,PR}.out"
 echo "Running GAP BFS"
 # It would be nice if you could read in a file for the roots

### 2017-05-11 experiment/example-workflow.sh
Fixed bug with real world datasets for real.

@@ -5,7 +5,7 @@
 # the experiments will just append to existing log files.
 # IF YOU STOP AN EXPERIMENT HALFWAY DELETE THE LOG FILES
 # they are stored in output/kron-$S
-THREADS="1 2 4 8 16 32"
+THREADS="1 2 4 8 16"
 S=13
 ./get-libraries.sh
 ./gen-datasets.sh $S

### 2017-05-11 experiment/gen-datasets.sh

@@ -127,15 +127,17 @@ if [ "$FILE_PREFIX" != "kron-$S" ]; then
 		"$GRAPHMATDIR/bin/graph_converter" --selfloops 1 --duplicatededges 0 --bidirectional --inputformat 1 --outputformat 0 --inputheader 0 --outputheader 1 --inputedgeweights 0 --outputedgeweights 2 --nvertices $nvertices "$d/$d.el" "$d/$d.graphmat"
 		# We write a serialized graph to speed up GAP
 	elif [ $(awk '{print NF; exit}' "$d.e") -eq 3 ]; then
-		echo " weighted."
+		echo " weighted. Weighted graphs are currently not supported because of a bug in GraphMat."
 		echo "SRC,DEST,WEIGHT" > "$d/edge.csv"
 		awk '{printf "%d %d %s\n", ($1+1), ($2+1), $3}' "$d.e" > "$d/$d.wel"
+		awk '{printf "%d %d\n", ($1+1), ($2+1)}' "$d.e" > "$d/$d.el" # For GraphMat
 		echo "Getting roots."
-		"$GAPDIR/sssp" -f "$d/$d.wel" -n $(($NRT*2)) > tmp.log
+		"$GAPDIR/sssp" -f "$d/$d.el" -n $(( $NRT * 2 )) > tmp.log
 		# TODO: Use this in real-datasets, change WeightT to float if need be and recompile GAPBS
 		"$GAPDIR/converter" -s -f "$d/$d.wel" -b "$d/$d.wsg"
 		# We make no assumptions so we output double precision edge weights
-		"$GRAPHMATDIR/bin/graph_converter" --selfloops 1 --duplicatededges 0 --bidirectional --inputformat 1 --outputformat 0 --inputheader 0 --outputheader 1 --inputedgeweights 1 --outputedgeweights 1 --edgeweighttype 1 --nvertices $nvertices "$d/$d.wel" "$d/$d.graphmat"
+		# "$GRAPHMATDIR/bin/graph_converter" --selfloops 1 --duplicatededges 0 --bidirectional --inputformat 1 --outputformat 0 --inputheader 0 --outputheader 1 --inputedgeweights 1 --outputedgeweights 0 --edgeweighttype 1 --nvertices $nvertices "$d/$d.wel" "$d/$d.graphmat"
[…]

### 2017-05-11 papers/publication/paper-pollard.tex
minor edits

@@
 % Here, \textasteriskcentered\ could be \emph{processing}, \emph{performance-analysis}, or \emph{algorithm-comparison})
-\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}}, a utility which simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
+\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}}, a framework which simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
 With this tool we select a small number of graph processing libraries over which we homogenize all aspects of execution to fairly compare performance.
@@
 Important paragraph.}
-Our utility is broken down into five principal phases indicated in \figurename~\ref{fig:epg-overview}, each of which requires no more than a single shell command. The five phases are as follows:
+Our framework breaks the process of characterizing performance into five principal phases shown in \figurename~\ref{fig:epg-overview}, each of which requires no more than a single shell command.
+The five phases are as follows.
 \begin{enumerate}
@@
 \subsection{Parsing and Data Analysis}
-Our example analyzes data using the R programming language and shell scripting parsers. The example workflow provided in our utility yields the figures generated in this paper.
+Our example analyzes data using the R programming language and shell scripting parsers. The example workflow provided in our framework yields the figures generated in this paper.
 

### 2017-05-12 autoperf/mst/experiment_analysis.R
Added an R analysis script, updated MST real.
MST still not working for livejournal.

@@ -1,0 +1,101 @@
+library(ggplot2)
+
+time_boxplot_with_batches <- function(filename, scale, ins_pct, cverts, num_threads) {
+	x <- read.csv(filename, header = TRUE)
+	# algorithm,execution_phase,scale,edges_per_vertex,RMAT_type,insertion_percent,changed_vertices,threads,measurement,value
+	# execution_phase: All, rooting tree, first pass, insertion, deletion
+	epv <- x$edges_per_vertex[1] # Assume the same throughout experiments
+	nedges <- epv * 2^scale
+	# Generate a figure
+	salient_cols <- c("algorithm", "execution_phase", "value")
+	method_time <- subset(x,
+			x$scale==scale &
+			x$edges_per_vertex == epv &
+			x$changed_vertices==cverts &
+			x$insertion_percent==ins_pct &
+			x$threads == num_threads,
+			salient_cols)
+	# method_time$value <- as.numeric(as.character(algo_time$Time)) # May not be necessary
+	# Remove zero rows---they're invalid and don't work with the log plot
+	# If some factors were coerced into NAs then there was some issue parsing
+	#method_time <- algo_time[!algo_time$Time == 0.0, ]
+	method_time$algo_and_phase <- interaction(method_time$execution_phase, method_time$algorithm)
+	
+	# If you want ugly base graphics but also want it to work on arya
+	pdf(paste0("plot",scale,"_",ins_pct,"i_",cverts,"_10b.pdf"), width = 3.3, height = 3.3)
+	#boxplot(value~algo_and_phase, data = method_time,
+	#		ylab = "Time (seconds)" )
+	# Prettier
+	box <- ggplot(aes(y = value, x = algo_and_phase), data = method_time) +
+		ylab("Time (seconds)") +
+		xlab("Execution Phase.Algorithm") +
+		geom_boxplot()
+	box + theme(axis.text.x = element_text(angle = 90, hjust = 1))
+	dev.off()
+}
+
+measure_scalability <- function(filename, one_scale) {
+    # Read in and average the data for BFS for each thread
+    # It is wasteful to reread the parsed*-1t.csv but it simplifies the code
+    x <- read.csv(filename, header = TRUE)
+	#x$algo_and_phase <- interaction(x$execution_phase, x$algorithm)
[…]

### 2017-05-12 papers/publication/paper-pollard.tex
algorithms edits
% insertion R plot for MST. Updated related work.
Added some reviewer's comments for more graph performance analysis.

@@
 \usepackage{todonotes}
+\usepackage{listings}
 
@@
 We demonstrate our approach on five graph processing packages: GraphMat, the Graph500, the Graph Algorithm Platform Benchmark Suite, GraphBIG, and PowerGraph using synthetic and real-world\todo[inline]{actual real-world.
-(i) Michael LeBeane et al, Supercomputing (SC 15), Data Partitioning Strategies for Graph Workloads on Heterogeneous Clusters (ii) Shuang Song et al,  ICPP 2016, Proxy-Guided Load Balancing of Graph Processing Workloads on Heterogeneous Clusters.} datasets.
+(i) Michael LeBeane et al, Supercomputing (SC 15), Data Partitioning Strategies for Graph Workloads on Heterogeneous Clusters (ii) Shuang Song et al (say it's distributed and we're shared),  ICPP 2016, Proxy-Guided Load Balancing of Graph Processing Workloads on Heterogeneous Clusters.
+(partitioning, but yet again, we're shared)} datasets.
 We examine previously overlooked aspects of parallel graph processing performance such as phases of execution and energy usage for three algorithms\todo[inline]{more algorithms, justify why these algorithms}: breadth first search, single source shortest paths, and PageRank and compare our results to Graphalytics.
@@
 Most performance analysis of graph processing systems come from the empirical results of each new library designer's publications.
-For example, GraphMat~\cite{Sundaram:2015:GraphMat}, PowerGraph~\cite{Gonzalez:2012:Powergraph}, and GraphBIG~\cite{Nai:2015:Graphbig} present a comparison of performance between their system and a selection of other software packages.
+For example, GraphMat~\cite{Sundaram:2015:GraphMat}, PowerGraph~\cite{Gonzalez:2012:Powergraph}, and GraphBIG~\cite{Nai:2015:Graphbig} present a comparison of performance between their system and a selection of other software packages and in general any new approach will compare itself to existing implementations.
 
@@
 
-Beyond the aforementioned performance results, a number of reports focus on performance analysis not tied to a particular software package we presently describe.
+Research by Song et al.~\cite{Song:2016:LoadBalancing} and LeBeane et al.~\cite{LeBeane:2015:GraphHetero} considers the important problem of partitioning on heterogeneous architectures and provide profiling to reduce runtime, improve load balance, and reduce energy consumption.
+Our framework instead focuses on shared memory packagess.
 
+Beyond the aforementioned performance results, a number of reports focus on performance analysis not tied to a particular software package.
 Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and~\cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
@@
 Like our work, Graphalytics also automates the setup and execution of graph packages for performance analysis.
-Graphalyics relies on Apache Maven and Java to wrap each graph processing software package---Graphalytics calls these platforms---and requires configuration files for each dataset.\todo[inline]{What am I saying with this?}
+Graphalyics relies on Apache Maven and Java to wrap the execution of each graph processing software package.
+%---Graphalytics calls these platforms---and requires configuration files for each dataset.\todo[inline]{What am I saying with this?}
+%Moreover, the complexity of Graphalytics can obfuscate the true behavior of the program.
 
-Moreover, the complexity of Graphalytics can obfuscate the true behavior of the program.
 By default, Graphalytics generates an HTML report listing the runtimes for each dataset and each algorithm in seconds.
@@
 
-We consider three algorithms: Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank, although not all algorithms are implemented on all systems.
+We consider three algorithms: Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank, though not all algorithms are implemented on all systems.
 We inspected the source code of the surveyed parallel graph processing systems to ensure the same phases of execution are measured across differing execution and programming paradigms.
@@
 We select BFS because the canonical performance leaderboard for parallel graph processing is the Graph500~\cite{Murphy:2010:Graph500}.
-The advantage of the Graph500 is that it provides standardized measurement specifications and dataset generation.
+One strength of the Graph500 is it provides standardized measurement specifications and dataset generation.
 The primary drawback with the Graph500 is it measures a single algorithm.
@@
 In our case, the fastest code is also the most energy efficient, although with this level of granularity we could detect circumstances where one could make a tradeoff between energy and runtime.
+\begin{figure}
+\begin{lstlisting}[language=C, frame=single]
+#ifdef POWER_PROFILING
+power_rapl_t ps;
+power_rapl_init(&ps);
+power_rapl_start(&ps);
+#endif
+	<region of code to monitor>
+#ifdef POWER_PROFILING
+power_rapl_end(&ps);
+power_rapl_print(&ps);
+#endif
+\end{lstlisting}
+\begin{lstlisting}[language=make,  basicstyle=\footnotesize]
+CFLAGS += -I$(PAPI)/include -DPOWER_PROFILING=1
+LDLIBS += -L$(PAPI)/lib
+LDLIBS += -Wl,-rpath,$(PAPI)/lib -lpapi
+\end{lstlisting}
+\caption{The changes required to monitor energy usage for a C or C++ file and its corresponding Makefile, given the papi header file is included and PAPI is the directory where PAPI is installed.}\todob{Is this too much detail?}
+\label{lst:RAPL}
+\end{figure}
+
+Listing~\ref{lst:RAPL} shows the simplicity of adding RAPL profiling to existing code.
+Once the source code is modified, the implementation may be recompiled and our provided header file linked in.
+Alternatively, there may be scenarios where this approach is not feasible.
+For example, if the user does not have root privileges s/he may not be able to access the model specific registers storing the energy usage.
+One alternative is to use TAU paired with PAPI to instrument the source code, and if the counters are available on the host CPU then the power and energy can still be measured.
 
@@
 \todo[inline]{Maybe; it's still difficult to add more graphs to the dataset sinc you need to know all the metadata.}
+
+\todo[inline]{Difficulty in running each dataset; GraphMat requires vertices to be 1 indexed and graph dataset conversion tools are limited.}
 We have presented a systematic approach to analyzing parallel graph processing performance, both measuring power consumption and the two fundamental phases of execution.

### 2017-05-12 autoperf/mst/experiment_analysis.R

@@ -34,6 +34,43 @@ time_boxplot_with_batches <- function(filename, scale, ins_pct, cverts, num_thre
 	dev.off()
 }
 
+percent_insertions <- function(filename, scale_num, cvert,
+		RMAT = "ER", num_threads = 72) {
+	x <- read.csv(filename, header = TRUE)
+	# algorithm,execution_phase,scale,edges_per_vertex,RMAT_type,insertion_percent,changed_vertices,threads,measurement,value
+	# execution_phase: All, rooting tree, first pass, insertion, deletion
+	# where rooting tree is only done for the first batch
+	epv <- x$edges_per_vertex[1] # Assume the same throughout experiments
+	nedges <- epv * 2^scale
+	# Generate a figure
+	salient_cols <- c("insertion_percent", "value")
+	method_time <- subset(x,
+			x$scale==scale_num &
+			x$edges_per_vertex==epv &
+			x$RMAT_type==RMAT &
+			x$threads==num_threads &
+			x$execution_phase=="All" &
+			x$changed_vertices==cvert &
+			x$algorithm=="MST",
+			salient_cols)
+	# method_time$value <- as.numeric(as.character(algo_time$Time)) # May not be necessary
+	# Remove zero rows---they're invalid and don't work with the log plot
+	# If some factors were coerced into NAs then there was some issue parsing
+	#method_time <- algo_time[!algo_time$Time == 0.0, ]
+	#method_time$algo_and_phase <- interaction(method_time$execution_phase, method_time$algorithm)
+	
+	pdf(paste0("insertion_pct_",scale_num,"_",cvert,"_time.pdf"), width = 2.5, height = 2.5)
+	ggplot(aes(y = value, x = insertion_percent, group = insertion_percent),
+			data = method_time) +
+		ylab("Time (seconds)") +
+		xlab("Insertion Percent") +
+		scale_x_continuous(breaks = as.numeric(method_time$insertion_percent)) +
+		geom_boxplot() +
+		labs(title = "Update vs. Insertion %") +
+		theme(axis.text.x = element_text(angle = 90, hjust = 1))
+	dev.off()
+}
+
 measure_scalability <- function(filename, one_scale) {
     # Read in and average the data for BFS for each thread
     # It is wasteful to reread the parsed*-1t.csv but it simplifies the code
[…]

### 2017-05-13 autoperf/mst/experiment_analysis.R
Updated for livejournal dataset.

@@ -5,9 +5,21 @@ time_boxplot_with_batches <- function(filename, scale, ins_pct, cverts, num_thre
 	# algorithm,execution_phase,scale,edges_per_vertex,RMAT_type,insertion_percent,changed_vertices,threads,measurement,value
 	# execution_phase: All, rooting tree, first pass, insertion, deletion
 	epv <- x$edges_per_vertex[1] # Assume the same throughout experiments
-	nedges <- epv * 2^scale
+	if (scale > 99) {
+		if (x$RMAT_type[1] == "com-lj.ungraph") {
+			plot_title <- "Livejournal Dataset"
+			nedges <- 34681189
+			plot_subtitle <- paste0(nedges, " edges, ", scale, " vertices")
+		} else {
+			nedges <- epv * scale # Just an estimation
+		}
+		plotfilename <- paste0("plot-",x$RMAT_type[1],"_",ins_pct,"i_",cverts,"_10b.pdf")
+	} else {
+		nedges <- epv * 2^scale
+		plotfilename <- paste0("plot",scale,"_",ins_pct,"i_",cverts,"_10b.pdf")
+	}
 	# Generate a figure
-	salient_cols <- c("algorithm", "execution_phase", "value")
+	salient_cols <- c("algorithm", "execution_phase", "value", "batch")
 	method_time <- subset(x,
 			x$scale==scale &
 			x$edges_per_vertex == epv &
@@ -21,15 +33,12 @@ time_boxplot_with_batches <- function(filename, scale, ins_pct, cverts, num_thre
 	#method_time <- algo_time[!algo_time$Time == 0.0, ]
 	method_time$algo_and_phase <- interaction(method_time$execution_phase, method_time$algorithm)
 	
-	# If you want ugly base graphics but also want it to work on arya
-	pdf(paste0("plot",scale,"_",ins_pct,"i_",cverts,"_10b.pdf"), width = 3.3, height = 3.3)
-	#boxplot(value~algo_and_phase, data = method_time,
-	#		ylab = "Time (seconds)" )
-	# Prettier
+	pdf(plotfilename, width = 3.3, height = 3.3)
 	box <- ggplot(aes(y = value, x = algo_and_phase), data = method_time) +
 		ylab("Time (seconds)") +
 		xlab("Execution Phase.Algorithm") +
-		geom_boxplot()
+		geom_boxplot() +
[…]

### 2017-05-13 papers/publication/paper-pollard.tex
Touched up figures and tables.

@@
 To enable easy, rigorous, and repeatable comparison of the capabilities of such systems, we present an approach and associated software for analyzing the performance and scalability of parallel, open-source graph libraries.
-We demonstrate our approach on five graph processing packages: GraphMat, the Graph500, the Graph Algorithm Platform Benchmark Suite, GraphBIG, and PowerGraph using synthetic and real-world\todo[inline]{actual real-world.
-(i) Michael LeBeane et al, Supercomputing (SC 15), Data Partitioning Strategies for Graph Workloads on Heterogeneous Clusters (ii) Shuang Song et al (say it's distributed and we're shared),  ICPP 2016, Proxy-Guided Load Balancing of Graph Processing Workloads on Heterogeneous Clusters.
-(partitioning, but yet again, we're shared)} datasets.
+We demonstrate our approach on five graph processing packages: GraphMat, the Graph500, the Graph Algorithm Platform Benchmark Suite, GraphBIG, and PowerGraph using synthetic and real-world datasets.
 We examine previously overlooked aspects of parallel graph processing performance such as phases of execution and energy usage for three algorithms\todo[inline]{more algorithms, justify why these algorithms}: breadth first search, single source shortest paths, and PageRank and compare our results to Graphalytics.
@@
 
-\todo[inline]{The y-axis is mislabeled. Should be "iterations" for pagerank.}
 \begin{figure}
@@
 
-\begin{table}
-	\caption{The data are generated using a Kronecker graph with a scale of 22 and using 32 threads.
+\begin{table*}
+	\caption{The data are generated using a Kronecker graph with a scale of 22.
+The program is executed with 32 threads.
 Sleeping Energy refers to the power (in Watts) consumed during the \texttt{unistd} C \texttt{sleep} function, multiplied by the Time row.
@@
 	\label{tab:power}
-\end{table}
+\end{table*}
 

### 2017-05-14 autoperf/mst/experiment_analysis.R
Added parsing and plot generation for power on MST

@@ -1,22 +1,25 @@
 library(ggplot2)
 
-time_boxplot_with_batches <- function(filename, scale, ins_pct, cverts, num_threads) {
+time_boxplot_with_batches <- function(filename, scale_num, ins_pct, cverts, num_threads)
+{
 	x <- read.csv(filename, header = TRUE)
 	# algorithm,execution_phase,scale,edges_per_vertex,RMAT_type,insertion_percent,changed_vertices,threads,measurement,value
 	# execution_phase: All, rooting tree, first pass, insertion, deletion
 	epv <- x$edges_per_vertex[1] # Assume the same throughout experiments
-	if (scale > 99) {
+	if (scale_num > 99) {
 		if (x$RMAT_type[1] == "com-lj.ungraph") {
 			plot_title <- "Livejournal Dataset"
 			nedges <- 34681189
-			plot_subtitle <- paste0(nedges, " edges, ", scale, " vertices")
+			plot_subtitle <- paste0(nedges, " edges, ", scale_num, " vertices")
 		} else {
-			nedges <- epv * scale # Just an estimation
+			nedges <- epv * scale_num # Just an estimation
 		}
 		plotfilename <- paste0("plot-",x$RMAT_type[1],"_",ins_pct,"i_",cverts,"_10b.pdf")
 	} else {
-		nedges <- epv * 2^scale
-		plotfilename <- paste0("plot",scale,"_",ins_pct,"i_",cverts,"_10b.pdf")
+		nedges <- epv * 2^scale_num
+		plotfilename <- paste0("plot",scale_num,"_",ins_pct,"i_",cverts,"_10b.pdf")
+		plot_title <- paste0("RMAT with Scale ",scale)
+		plot_subtitle <- paste0(nedges, "edges, ", 2^scale_num, " vertices")
 	}
 	# Generate a figure
 	salient_cols <- c("algorithm", "execution_phase", "value", "batch")
@@ -26,6 +29,7 @@ time_boxplot_with_batches <- function(filename, scale, ins_pct, cverts, num_thre
 			x$changed_vertices==cverts &
 			x$insertion_percent==ins_pct &
 			x$threads == num_threads,
+			x$measurement == "Time (s)",
 			salient_cols)
 	# method_time$value <- as.numeric(as.character(algo_time$Time)) # May not be necessary
 	# Remove zero rows---they're invalid and don't work with the log plot
[…]

### 2017-05-14 experiment/real-datasets.sh

@@ -28,7 +28,7 @@ for arg in "$@"; do
 done
 if [ "$#" -lt 1 ]; then
 	echo 'Please provide <filename> or the prefix of the filename'
-	echo $USAGE
+	echo "$USAGE"
 	exit 2
 fi
 FILE="$1"

### 2017-05-14 papers/publication/paper-pollard.tex

@@
 We demonstrate our approach on five graph processing packages: GraphMat, the Graph500, the Graph Algorithm Platform Benchmark Suite, GraphBIG, and PowerGraph using synthetic and real-world datasets.
-We examine previously overlooked aspects of parallel graph processing performance such as phases of execution and energy usage for three algorithms\todo[inline]{more algorithms, justify why these algorithms}: breadth first search, single source shortest paths, and PageRank and compare our results to Graphalytics.
+We examine previously overlooked aspects of parallel graph processing performance such as phases of execution and energy usage for three algorithms: breadth first search, single source shortest paths, and PageRank and compare our results to Graphalytics.
 \end{abstract}
@@
 \section{Introduction}
-%% What's the problem (motivation)
-\todob{need an introductory sentence about large graph applications and resulting proliferation of parallel frameworks}
 Our research is motivated by the current state of parallel graph processing.
+Fields such as social network analysis, \cite{Kang:2011:BillionGraph}, and computational biology \cite{Paylopoulos:2011:biograph} require the analysis of ever-increasing graph sizes.
+The wide variety of problem domains is resulting in the proliferation of parallel graph processing frameworks.
 The most comprehensive survey, released in 2014, identified and categorized over 80 different parallel graph processing systems, not even considering domain specific languages~\cite{Doekemeijer:2015:GPFSurvey}.
@@
 Optimal data structures may differ across graphs and must be determined by the programmer.
-To complicate matters further, the $\Delta$-stepping implementation of the parallel Dijkstra's Algorithm for computing single source shortest paths is parametrized by some $\Delta > 0$ for which requires tuning for a given architecture and graph structure to achieve optimal performance~\cite{Panitanarak:2014:SSSPPerf}.
-\todo[inline]{Will this be mentioned again?}
 
@@
 
-To address these issues, we present \emph{easy-parallel-graph\mbox{-\textasteriskcentered}}
-% Here, \textasteriskcentered\ could be \emph{processing}, \emph{performance-analysis}, or \emph{algorithm-comparison})
-\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}}, a framework which simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
+To address these issues, we present our framework called \mbox{\emph{easy-parallel-graph-\textasteriskcentered}}\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}}, a framework which simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
 With this tool we select a small number of graph processing libraries over which we homogenize all aspects of execution to fairly compare performance.
-Additionally, we present an example use of this framework to analyze the performance, energy consumption, and scalability of specific algorithm implementation.
+Additionally, we present results from a use case of this framework to analyze the performance, energy consumption, and scalability of specific algorithm implementation.
 This enables algorithm designers to select high performance reference implementations with which they can compare their improvements.
@@
 \section{Related Work}
-\todo[inline]{There's a large number of qualitative. Most third party comparisons are qualitative and just describe the overview instead of actual runtimes. This isn't a level playing field.}
 Most performance analysis of graph processing systems come from the empirical results of each new library designer's publications.
 For example, GraphMat~\cite{Sundaram:2015:GraphMat}, PowerGraph~\cite{Gonzalez:2012:Powergraph}, and GraphBIG~\cite{Nai:2015:Graphbig} present a comparison of performance between their system and a selection of other software packages and in general any new approach will compare itself to existing implementations.
+These performance results are inherently biased because the authors will know their own package better than the competition and as such are aware of more optimizations.
 
 Nai et al.~\cite{Nai:2016:architectural} provide a detailed performance analysis using GraphBIG as their reference implementation.
-Their approach also compares GraphBIG to the GraphLab and Pregel execution models.
+Their approach also compares GraphBIG to the GraphLab and Pregel execution models (Gather Apply Scatter).
 Their analysis considers architectural performance measurements such as cache miss rates to measure bottlenecks and computational efficiency for a variety of datasets.
 While incredibly thorough, these analyses focus on covering a wide range of datasets and appear difficult to apply to a new approach.
-For example, GraphMat reduce computation to sparse matrix operations which may not suffer from the same memory bottlenecks indicated in the results of Nai et al.
+For example, GraphMat reduces computation to sparse matrix operations which may not suffer from the same memory bottlenecks indicated in the results of Nai et al.
 
@@
 
-By default, Graphalytics generates an HTML report listing the runtimes for each dataset and each algorithm in seconds.
+\todo[inline]{Put real world figure here}
+
+Graphalytics generates an HTML report listing the runtimes for each dataset and each algorithm.
 For example, one run of a single source shortest paths algorithm took 5.96 seconds.
-However, perusing the log files reveals a more complete picture: of these 5.96 seconds, 0.21 was spent computing the shortest paths while the remaining time consisted of building the necessary data structures.
-Furthermore, other performance-relevant details, such as the number of iterations run for PageRank, are not easily available using Graphalytics.
+However, perusing the log files reveals a more complete picture: of these \todo[inline]{Make this from table I!}5.96 seconds, 0.21 was spent computing the shortest paths while the remaining time consisted of building the necessary data structures.
+Table~\ref{tab:graphalytics-realworld} shows the results from a typical experiment run by Graphalytics.
+
 %We use GraphMat, the single source shortest paths algorithm, the dota-league dataset from the Game Trace Archive\cite{Guo:2012:GTA}, and the default settings as an example here.
@@
 With a plugin to Graphalytics called Granula~\cite{Ngai:2015:Granula}, one can explicitly specify a performance model to analyze specific execution behavior such as the amount of communication or runtime of particular kernels of execution.
-This requires in-depth knowledge of the source code and execution model.
-Furthermore, creating such a model requires a high level of expertise with the given system and with Granula\footnote{An example of Granula can be seen at \url{https://github.com/tudelft-atlarge/graphalytics-platforms-graphx/tree/master/granula-model-graphx}}.
-
-As with Graphalytics, the initial development effort is high but Granula paired with Graphalytics allows automatic execution and compilation of performance results.
-Likewise, our approach provides automatic execution and performance analysis but without requiring a performance and execution model for each system.
+This requires in-depth knowledge of the source code and execution model in addition to expertise with the Granula API but allows detailed performance analysis and automatic execution and compilation of performance results.\footnote{An example of Granula can be seen at \url{https://github.com/tudelft-atlarge/graphalytics-platforms-graphx/tree/master/granula-model-graphx}}.
 
@@
 		\includegraphics[width=\linewidth, trim=0 144pt 0pt 156pt, clip]{graphics/overview.eps}
-	\caption{Overview of \emph{easy-parallel-graph-\textasteriskcentered}. Each orange box corresponds to a single shell script in the package while the green ellipses correspond to generated files.}
+	\caption{Overview of \emph{easy-parallel-graph-\textasteriskcentered}. Each cyan box corresponds to a single shell script in the package while the green ellipses correspond to generated files.}
 	\label{fig:epg-overview}
@@
 For example, many of our experiments were run with $2^{22} = 4,194,304$ vertices and an average of 16 edges per vertex.
-We measure strong scaling and speedup by varying the number of threads from one to the total number of threads on our server, 72.
+We measure strong scaling and speedup by varying the number of threads from one to the total number of threads on our research computer, 72.
 
@@
 
-\todo[inline]{More real world datasets}
 The Graphalytics results in Table~\ref{tab:graphalytics} were performed on the Dota-League datasets.
@@
 This dataset is useful because it is both weighted and more dense than the usual real-world dataset with an average out-degree of 824.
+
+We also present results for the \verb|cit-Patents| dataset from SNAP \cite{snap-cit-patents}.
+This is a widely-used network of citations from the National Bureau of Economic Research (NBER) and is less dense than Dota-League.
+We stress that though these two datasets are presented, any network in the SNAP data format\footnote{A file in the SNAP format consists of one edge per line, with vertices separated by whitespace and lines which begin with \# are comments.} can be used in \mbox{\emph{easy-parallel-graph-\textasteriskcentered}}.
 
@@
 \begin{itemize}
-	\item The Graph500\footnote{We used the most recent version from \url{https://github.com/graph500/graph500}, most similar to release 2.1.4.}~\cite{Murphy:2010:Graph500}, a set of large-scale graph benchmarks.
-	\item The Graph Algorithm Platform (GAP) Benchmark Suite~\cite{Beamer:2015:GAPBench}.
-	\item GraphBIG~\cite{Nai:2015:Graphbig} benchmark suite.
-	\item GraphMat~\cite{Sundaram:2015:GraphMat}, a high-performance graph library.
-	\item PowerGraph~\cite{Gonzalez:2012:Powergraph}, a framework for distributed graph-parallel computation on natural graphs.
-	\todo[inline]{Explain each more.}
+	\item The Graph500~\cite{Murphy:2010:Graph500}, the canonical BFS benchmark which consists of a specification and reference implementation. We use a modified version most similar to 2.1.4.
+	\item The Graph Algorithm Platform (GAP) Benchmark Suite~\cite{Beamer:2015:GAPBench}, a set of reference implementations for shared memory graph processing.
+	\item GraphBIG~\cite{Nai:2015:Graphbig} benchmark suite. We consider only the shared memory solutions but there are also GPU benchmarks.
+	\item GraphMat~\cite{Sundaram:2015:GraphMat}, a library and programming model along with reference implementations of common algorithms.
+	\item PowerGraph~\cite{Gonzalez:2012:Powergraph}, a library and programming model for distributed (and shared memory) graph-parallel computation.
+PowerGraph includes additional toolkits for tasks such as clustering and computer vision.
 \end{itemize}
 
-Our approach is not specific or limited to these graph packages and can be extended to others with relatively modest effort (building each package and if necessary, instrumenting the different phases of the computation).
-\todo[inline]{Say what we automate and what we don't automate.}
+Our approach is not specific or limited to these graph packages and can be extended to others.
 \todo[inline]{Weighted versus unweighted}
@@
 \subsection{Parsing and Data Analysis}
-Our example analyzes data using the R programming language and shell scripting parsers. The example workflow provided in our framework yields the figures generated in this paper.
+Our example analyzes data using the R programming language and shell scripting parsers. The example workflow provided in our framework yields the figures and tables generated in this paper.
 
 \subsection{Machine Specifications}
-We performed experiments on our 36-core (72 thread) Intel Haswell server with 256GB DDR4 RAM, with two Intel Xeon E5-2699 v3 CPUs. The operating system is GNU/Linux version 4.4.0-22.
-%The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit.
-We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.
+We performed experiments on our 36-core (72 thread) Intel Haswell server with 256GB DDR4 RAM, with two Intel Xeon E5-2699 v3 CPUs.
+The operating system is GNU/Linux version 4.4.0-22.
+Our code was compiled with GCC version 4.8.5 with the exception of GraphMat which was compiled with ICPC version 17.0.0.
 %\begin{center}
@@
 
-\subsection{Runtime}
-
-In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately $16 \times 2^22 = 33,500,000$ edges)\todo[inline]{work?}.
-An explanation of each algorithm is given in~\cite{Iosup:2016:Graphalyticstech}.
-Graphalytics by default does not perform SSSP on unweighted, undirected graphs.
-The discrepancy between PageRank values in Table~\ref{tab:graphalytics} and Fig.~\ref{fig:pr} is a result of the differing stopping criterion.\todo[inline]{What are the likely culprits for performance differences---measure cache miss rates.
-Try to find something which correlates with performance.
-Resource stalls will work easily (not as interesting) Cycles per instruction?
-Automate with perfexplorer or autoperf}
-
-In addition to the data presented in Table~\ref{tab:graphalytics}, one may want to know the cost of performing additional computations on a graph once the data structures have been built.
-Additionally, one may desire runtime bounds to get a reasonable maximum timespan for a given problem.
-Table~\ref{tab:epg-perf} begins to compile performance results and the rest of this section further explores these topics.
-
-% Graphalytics also outputs MTEPS or millions of traversed edges per second.
-However, the graphalytics version does not make sense in all cases: for example, computing the local clustering coefficient involves traversing each edge multiple times (proportional to the sparsity of the graph), while BFS traverses each edge exactly once, and the number of edges traversed with PageRank depends on the connectivity of the graph and the number of iterations.
+\subsection{Runtime for Synthetic Graphs}
 
@@
 	\caption{The dataset used is the same Kronecker graph with scale $22$ as used in other experiments.
-Performance results are in seconds running with the maximal number of threads (64 for Powergraph, 72 for the others).
+Performance results are in seconds running with the maximal number of threads (64 for PowerGraph, 72 for the others).
 Community detection is performed using label propagation.}
@@
 
-\begin{table}
-	\centering
-	\caption{This table shows averages of runtimes for the Dota-League dataset using our method using 32 threads.
-Each cell represents the average over 32 runs.
-The ``N/A'' for PowerGraph and GraphBIG exist because both of these systems read in the files and build a graph data structure simultaneously.
-GraphMat is not included because reading in the Dota-League dataset crashes the system.}
-	\begin{tabular}{l|c|c|c|c}
-		System & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular}
-		                   & Run BFS & Run SSSP & Run PageRank \\ \hline
-		PowerGraph & N/A   & N/A     & 0.344    & 3.606 \\
-		GAP        & 0.412 & 0.00464 & 0.267    & 33.00 \\
-%		GraphMat   & 1.013 &         &          &       \\
-		GraphBIG   & N/A   & 0.0476  & 0.162    & 21.39 \\
-	\end{tabular}
-	\label{tab:epg-perf}
-\end{table}
+In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately $16 \times 2^{22} = 33,500,000$ edges)\todo[inline]{work?}.
+An explanation of each algorithm is given in~\cite{Iosup:2016:Graphalyticstech}.
+Graphalytics by default does not perform SSSP on unweighted, undirected graphs.
+The discrepancy between PageRank values in Table~\ref{tab:graphalytics} and Fig.~\ref{fig:pr} is a result of the differing stopping criterion.
+
+In contrast, our results for the Kronecker graph of scale $22$ are presented in Figs.~\ref{fig:bfs-time},~\ref{fig:sssp-time}, and~\ref{fig:pr}.
+
+\todo[inline]{What are the likely culprits for performance differences---measure cache miss rates.
+Try to find something which correlates with performance.
+Resource stalls will work easily (not as interesting) Cycles per instruction?
+Automate with perfexplorer or autoperf}
+
+% Graphalytics also outputs MTEPS or millions of traversed edges per second.
+However, the graphalytics version does not make sense in all cases: for example, computing the local clustering coefficient involves traversing each edge multiple times (proportional to the sparsity of the graph), while BFS traverses each edge exactly once, and the number of edges traversed with PageRank depends on the connectivity of the graph and the number of iterations.
 
@@
 The data structure construction times for GAP and GraphMat are consistent; in both cases the platforms create the same data structure for both algorithms.
-These results are consistent with~\cite{Sundaram:2015:GraphMat} which lists GraphMat as more performant than PowerGraph in SSSP.
+These results are consistent with~\cite{Sundaram:2015:GraphMat} which lists GraphMat as more higher performing than PowerGraph in SSSP.
 
@@
 The assumptions under which the various platforms operate can have a dramatic effect on the program.
-For example, the GAP Benchmark Suite stores vertex weights as 32-bit integers.
+\todo[inline]{Rewrite this portion} For example, the GAP Benchmark Suite stores vertex weights as 32-bit integers.
 However, other systems store them as floating-point numbers.
@@
 
+\subsection{Scalability}
 Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS.
@@
 Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.
+GCC version 4.8.5 supports OpenMP version 3.1 but the most recent release of OpenMP is 4.5.
 \begin{figure}
@@
 \end{figure}
+
+\subsection{Real World Datasets}
+
+\todo[inline]{Rerun this}
+\begin{table}
+	\centering
+	\caption{This table shows averages of runtimes for the Dota-League dataset using our method using 32 threads.
+Each cell represents the average over 32 runs.
+The ``N/A'' for PowerGraph and GraphBIG exist because both of these systems read in the files and build a graph data structure simultaneously.
+GraphMat is not included because reading in the Dota-League dataset crashes the system.}
+	\begin{tabular}{l|c|c|c|c}
+		System & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular}
+		& \begin{tabular}[x]{@{}c@{}}Run \\ BFS\end{tabular}
+		& \begin{tabular}[x]{@{}c@{}}Run \\ SSSP\end{tabular}
+		& \begin{tabular}[x]{@{}c@{}}Run \\ PageRank\end{tabular}
+		\\ \hline
+		PowerGraph & N/A   & N/A     & 0.344    & 3.606 \\
+		GAP        & 0.412 & 0.00464 & 0.267    & 33.00 \\
+		%		GraphMat   & 1.013 &         &          &       \\
+		GraphBIG   & N/A   & 0.0476  & 0.162    & 21.39 \\
+	\end{tabular}
+	\label{tab:epg-perf}
+\end{table}
+
+In addition to the data presented in Table~\ref{tab:graphalytics}, one may want to know the cost of performing additional computations on a graph once the data structures have been built.
+Additionally, one may desire runtime bounds to get a reasonable maximum timespan for a given problem.
+Table~\ref{tab:epg-perf} begins to compile performance results and the rest of this section further explores these topics.
 
@@
 \end{lstlisting}
-\caption{The changes required to monitor energy usage for a C or C++ file and its corresponding Makefile, given the papi header file is included and PAPI is the directory where PAPI is installed.}\todob{Is this too much detail?}
+\caption{The changes required to monitor energy usage for a C or C++ file and its corresponding Makefile, given the papi header file is included and \$PAPI is the directory where PAPI is installed.}\todob{Is this too much detail?}
 \label{lst:RAPL}
@@
 
-\section{Future Work and Conclusion}
-\todo[inline]{Maybe; it's still difficult to add more graphs to the dataset sinc you need to know all the metadata.}
+\section{Future Work}
+%We found that GraphMat is only decently performant, but scales to more threads.
+One explanation is that GraphMat's underlying computation model (sparse matrix operations), paired with the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation, is better suited to larger-scale graphs.
+% Furthermore, sparse matrix operations have been well studied so achieving good scalability is more realistic [cite]
 
-\todo[inline]{Difficulty in running each dataset; GraphMat requires vertices to be 1 indexed and graph dataset conversion tools are limited.}
-We have presented a systematic approach to analyzing parallel graph processing performance, both measuring power consumption and the two fundamental phases of execution.
-While the comparison presented here can help one choose among alternatives for the selected packages and algorithms, the problem of selecting a graph processing framework for a given large-scale problem remains far from simple.
-Hence, we make our code available at \url{https://github.com/HPCL/easy-parallel-graph} and encourage further experimentation.
-Our results required minor changes to the original projects to add calls to PAPI sampling and those changes are also available on forked versions of the repositories.
+Our selection of algorithms comes from a common subset of algorithms analyzed in prior work and implemented in the selected software packages.
+The standardization of graph algorithm building blocks (graph kernels) is being developed by the Graph BLAS Forum~\cite{Mattson:2013:graphblas}).
+Once this standardization is finalized there is motivation from both library designers and performance analyzers to implement and profile each kernel.
+Regardless, algorithms like triangle counting and betweenness centrality are widely implemented but not supported by either Graphalytics nor \emph{easy-parallel-graph\mbox{-\textasteriskcentered}}.
 
-Overall, the GAP Benchmark Suite was the highest-performing system across the given datasets though not the most scalable.
+We hope to add some level of parameter tuning to the next iteration of our software.
+Recent advances in parallel SSSP and BFS algorithms make use of a concept called $\Delta$-stepping.
+This consists of the algorithm being parametrized by some $\Delta > 0$ for which requires tuning for a given architecture and graph structure to achieve optimal performance~\cite{Panitanarak:2014:SSSPPerf}.
+This is provided in the GAP benchmark suite as well as the Parallel Boost Graph Library.
+
+Graphalytics also had circumstances with the more computationally expensive algorithms  where certain experiments fail~\cite{Iosup:2016:Graphalyticstech}, so determining whether an algorithm will finish given a particular machine, input size, runtime limit, and resources is an important unanswered question we plan to pursue further.
+
+Graph data structure representation by nature varies across software packages.
+There may be significant performance differences among the various packages for using directed or undirected, or weighted and unweighted graphs.
+We hope to investigate this issue further.
+% Our framework abstracts away much of the data formatting problem, but issues such as whether a framework supports a given dataset is a complex task.
+For example, GraphMat currently cannot convert weighted graphs to unweighted and does not support zero-indexed graphs.
+
+One overarching issue with these software packages is the speed at which they change.
+On one end of the spectrum, the code base of PowerGraph was made closed-source in 2015.
+Simple fixes such as increasing the maximum number of threads per process and updating dependency URLs are not merged and as a result there is much duplicated effort among the over than 400 forks of PowerGraph.
+On the other end, GraphMat's update to version 2.0 has not yet been adopted by Graphalytics nor our framework.
+One potential solution would be to add each framework to a package management system such as Spack \cite{Gamblin:2015:Spack} which focuses on High Performance Computing.
+
+\section{Conclusion}
+Overall, the GAP Benchmark Suite was the best-performing system across the given datasets though not the most scalable.
 However, this is only for relatively small graphs; all the graphs used here had at most $2^{22}$ vertices.
+Additionally, GAP is the most recent of projects and also the most robust in terms of the input graphs accepted.
 It should be noted that GAP is also the most recent of projects.
@@
 
-%We found that GraphMat is only decently performant, but scales to more threads.
-One explanation is that GraphMat's underlying computation model (sparse matrix operations), paired with the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation, is better suited to larger-scale graphs.
-% Furthermore, sparse matrix operations have been well studied so achieving good scalability is more realistic [cite]
-
-To ensure fairness, each platform must be configured to use the same graphs and the same roots.
-In the case of GraphBIG, this required the file to be loaded in for each experiment.
-By far the most time consuming of the experiments was GraphBIG's serial file input of uncompressed ASCII text format for graphs.
-This limited the sizes of the experiments we could run.
-Graphalytics also had circumstances with the more computationally expensive algorithms  where certain experiments fail~\cite{Iosup:2016:Graphalyticstech}, so determining whether an algorithm will finish given a particular machine, input size, runtime limit, and resources is an important unanswered question we plan to pursue further.
+We have presented a tool called \emph{easy-parallel-graph\mbox{-\textasteriskcentered}}, where \textasteriskcentered $\ = $ installer, comparator, performance analyzer, dataset homogenizer, or performance visualizer.
+This framework can analyze both power consumption and the two fundamental phases of execution: data structure construction and algorithm execution.
+While the comparison presented here can help one choose among alternatives for the selected packages and algorithms, the problem of selecting a graph processing framework for a given large-scale problem remains far from simple.
+Furthermore, increasing hardware heterogeneity demands performance analysis be easily repeatable on the target architecture.
+Because of this, our framework is designed to be easy to use and we make our code freely available at \url{https://github.com/HPCL/easy-parallel-graph} to encourage further experimentation.
 

### 2017-05-14 experiment/example-workflow.sh
Updated example workflow for real-world datsets.

@@ -1,12 +1,10 @@
 #!/bin/bash
 # An example workflow, going from building the libraries to
 # generating some figures.
-# NOTE: This is in the early stages of development so currently
-# the experiments will just append to existing log files.
-# IF YOU STOP AN EXPERIMENT HALFWAY DELETE THE LOG FILES
-# they are stored in output/kron-$S
-THREADS="1 2 4 8 16"
-S=13
+# Synthetic datasets are stored in output/kron-$S
+mkdir -p output
+THREADS="1 2 4 8 16 32 64 72"
+S=22
 ./get-libraries.sh
 ./gen-datasets.sh $S
 for T in $THREADS; do
@@ -14,9 +12,27 @@ for T in $THREADS; do
 done
 ./parse-output.sh $S
 echo "# Config file for experiment_analysis.R. threads a vector, scale an int.
+prefix <- './output/'
 threads <- c(${THREADS// /,})
 scale <- $S
 " > example_config.R # Warning: this file is sourced in experiment_analysis.R
 mkdir -p graphics
 Rscript experiment_analysis.R example_config.R
+
+# Dota-league dataset
+DSET=dota-league
+curl -o datasets/$DSET.zip https://atlarge.ewi.tudelft.nl/graphalytics/zip/$DSET.zip
+unzip -d datasets datasets/$DSET.zip
+mv datasets/$DSET/$DSET.e datasets/$DSET.e
+mv datasets/$DSET/$DSET.v datasets/$DSET.v
+./gen-datasets.sh -f=datasets/$DSET.e
+./run-experiment.sh -f=$DSET 32
+./parse-output.sh -f=$DSET
+
+echo "# Config file for experiment_analysis.R. threads a vector, scale an int.
+prefix <- './output/'
+threads <- 32
+dataset <- '$DSET'
+" > ${DSET}_config.R # Warning: this file is sourced in experiment_analysis.R
+Rscript realworld_analysis.R ${DSET}_config.R
 

### 2017-05-14 experiment/experiment_analysis.R

@@ -5,7 +5,7 @@ if (length(args) != 1) {
 	stop(usage)
 }
 bpc <- "cyan"
-prefix <- "./output/"
+prefix <- "./output/" # The default.
 
 # source sets scale and threads
 source(args[1]) # This is a security vulnerability... just be careful
@@ -21,7 +21,7 @@ time_boxplot <- function(scale, thr, algo) {
 	nedges <- 16 * 2^scale
 	filename <- paste0(prefix,"parsed-","kron-",scale,"-",thr,".csv")
 	x <- read.csv(filename, header = FALSE)
-	colnames(x) <- c("Sys","Algo","Metric","Time")	
+	colnames(x) <- c("Sys","Algo","Metric","Time")
 	# Generate a figure
 	algo_time <- subset(x, x$Algo == algo & x$Metric == "Time",
 			c("Sys","Time"))

### 2017-05-14 experiment/experiment_analysis.R
Fixed bug with removing previous log files.
Updated experiment analysis to reflect the paper's workflow.

@@ -37,19 +37,6 @@ time_boxplot <- function(scale, thr, algo) {
 	dev.off()
 }
 
-for (thr in threads) {
-	time_boxplot(scale, thr, "BFS")
-}
-
-
-# UNIMPLEMENTED
-# weak_scaling <- function(threads, algo, all_scales) {
-# 	for (thr in threads) {
-# 		for (s in all_scales) {
-# 		}
-# 	}
-# }
-
 ###
 # Part 2: Generate the plots for a single algorithm and multiple problem sizes
 ###
@@ -116,4 +103,8 @@ plot_strong_scaling <- function(scaling_data, scale, threads, algo) {
 
 bfs_scale <- measure_scale(scale, threads, "BFS") # Possiblities: BFS, SSSP, PageRank
 plot_strong_scaling(bfs_scale, scale, threads, "BFS")
+
+for (thr in threads) {
+	time_boxplot(scale, thr, "BFS")
+}
 

### 2017-05-14 experiment/run-experiment.sh

@@ -72,14 +72,15 @@ export SKIP_VALIDATION=1
 # Load all the modules here
 module load intel/17
 
-echo "Cleaning $OUTPUT_PREFIX-*.out"
-rm -f "${OUTPUT_PREFIX}-{GAP,GraphMat,PowerGraph}-{BFS,SSSP,PR}.out"
+echo Note: Files of the form
+echo "${OUTPUT_PREFIX}-{GAP,GraphMat,PowerGraph}-{BFS,SSSP,PR}.out"
+echo get overwritten.
 
 echo Starting experiment at $(date)
 
 echo -n "Running Graph500 BFS"
-# This isn't working, so just regenerate the data
 #omp-csr/omp-csr -s $S -o "$DDIR/kron-$S/kron-${S}.graph500" -r "$DDIR/kron-$S/kron-${S}.roots"
+# ^ This isn't working, so just regenerate the data ^
 if [ "$OMP_NUM_THREADS" -gt 1 ]; then
 	echo " with OpenMP"
 	"$GRAPH500DIR/omp-csr/omp-csr" -s $S > "${OUTPUT_PREFIX}-Graph500-BFS.out"
@@ -89,7 +90,7 @@ else
 fi
 
 # GAP
-rm -f "${OUTPUT_PREFIX}-GAP-{BFS,SSSP,PR}.out"
+rm -f "${OUTPUT_PREFIX}"-GAP-{BFS,SSSP,PR}.out
 echo "Running GAP BFS"
 # It would be nice if you could read in a file for the roots
 # Just do one trial to be the same as the rest of the experiments
@@ -111,7 +112,7 @@ for ROOT in $(head -n $NRT "$DDIR/kron-$S/kron-${S}-roots.v"); do
 done
 
 # PowerGraph
-rm -f "${OUTPUT_PREFIX}-PowerGraph-{SSSP,PR}.{out,err}"
+rm -f "${OUTPUT_PREFIX}"-PowerGraph-{SSSP,PR}.{out,err}
 echo "Running PowerGraph SSSP"
 # Note that PowerGraph also sends diagnostic output to stderr so we redirect that too.
 if [ "$OMP_NUM_THREADS" -gt 128 ]; then
[…]

### 2017-05-14 papers/publication/paper-pollard.tex

@@
 \usepackage{hyperref}
+\usepackage{url} 
 \usepackage{cite} % Multiple citations in one~\cite command
@@
 
-To address these issues, we present our framework called \mbox{\emph{easy-parallel-graph-\textasteriskcentered}}\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}}, a framework which simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
+To address these issues, we present our framework called \mbox{\emph{easy-parallel-graph-\textasteriskcentered}},\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}} a framework which simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
 With this tool we select a small number of graph processing libraries over which we homogenize all aspects of execution to fairly compare performance.
@@
 
-\section{Related Work}
+\section{Related Work}\label{sec:relatedwork}
 Most performance analysis of graph processing systems come from the empirical results of each new library designer's publications.
@@
 Research by Song et al.~\cite{Song:2016:LoadBalancing} and LeBeane et al.~\cite{LeBeane:2015:GraphHetero} considers the important problem of partitioning on heterogeneous architectures and provide profiling to reduce runtime, improve load balance, and reduce energy consumption.
-Our framework instead focuses on shared memory packagess.
+Our framework instead focuses on shared memory packages.
 
 Beyond the aforementioned performance results, a number of reports focus on performance analysis not tied to a particular software package.
-Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and~\cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
+Satish et al.~\mbox{\cite{Satish:2014:NavigatingGraph}} analyze the performance of several systems on datasets on the order of $30$ billion edges and~\cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
 These implementations are hand tuned and provides recommendations for future improvements.
@@
 
-\todo[inline]{Put real world figure here}
+\begin{table}
+	\caption{A tabulated sample run of Graphalytics with 32 threads.
+Just one run per experiment is performed and these times do not take into account the different phases of execution.
+Runtimes are in seconds.}
+	\centering
+	% cit-Patents: (3 774 768 vertices, 16 518 948 edges)
+	% dota-league (61 170 vertices, 50 870 313 edges)
+	\begin{tabular}{l|c|c|c|c|c|c}
+	GraphBIG & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
+	cit-Patents & 1.0 & 11.1 & 13.0 &   5.0 & N/A  & 1.3 \\
+	dota-league & 1.0 & 3.8 & 1104.0 & 2.5 & \textbf{2.9} & 0.9 \\
+	\multicolumn{1}{l}{}\\
+	PowerGraph & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
+	\multicolumn{1}{l}{}\\
+	GraphMat & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
+	cit-Patents & 7.8 & 20.6 & 11.2 &  8.4 & N/A &  7.0 \\
+	dota-league & 2.2 & 19.7 & 150.1 & 2.8 & \textbf{3.8} & 3.1
+	\end{tabular}
 
-Graphalytics generates an HTML report listing the runtimes for each dataset and each algorithm.
-For example, one run of a single source shortest paths algorithm took 5.96 seconds.
-However, perusing the log files reveals a more complete picture: of these \todo[inline]{Make this from table I!}5.96 seconds, 0.21 was spent computing the shortest paths while the remaining time consisted of building the necessary data structures.
-Table~\ref{tab:graphalytics-realworld} shows the results from a typical experiment run by Graphalytics.
+	\begin{lstlisting}
+Timing results (for GraphMat SSSP on dota-league)
+- Finished file read of dota-league. time: 0.9495
+- load graph: 3.48167 sec
+- initialize engine: 3.60012e-05 sec
+- run algorithm: 0.168526 sec
+- print output: 0.105846 sec
+- deinitialize engine: 6.00815e-05 sec
+	\end{lstlisting}
+	\label{tab:graphalytics-realworld}
+\end{table}
+
+Graphalytics generates an HTML report listing the runtimes for each dataset and each algorithm. Table~\ref{tab:graphalytics-realworld} shows the results from a typical experiment run by Graphalytics.
+
+However, perusal of the log files reveals a different story.
+The listing contained below Table~\ref{tab:graphalytics-realworld} is the output from GraphMat itself and the execution timing (3.8 seconds) includes the time to read in the input file.
+However, the GraphBIG timing (2.9 seconds) does not include the time to read the dota-league file.
+\textbf{If the time to read in the text file was computed correctly, GraphMat would be faster than GraphBIG.} To call this a fair comparison is dubious at best when phases of execution are omitted.
 
 %We use GraphMat, the single source shortest paths algorithm, the dota-league dataset from the Game Trace Archive\cite{Guo:2012:GTA}, and the default settings as an example here.
-% The full logs:
-%Finished setting ids, time: 0.157009
+%source vertex: 55713
+%num. threads: 64
+%Starting file read of /home/users/spollard/graphalytics/graphalytics-platforms-graphmat/intermediate/dota-league.e_weight..mtx
+%Got graph with m=61170  n=61170 nnz=101740626
+%Finished file read of /home/users/spollard/graphalytics/graphalytics-platforms-graphmat/intermediate/dota-league.e_weight..mtx, time: 0.94948
+%Finished setting ids, time: 0.077502
 %Starting sort
-%Finished sort, time: 1.39571
-%Finished setting edge pointers, time: 0.000133
+%Finished sort, time: 0.767248
+%Finished setting edge pointers, time: 0.00017
 %Starting build_dcsc
-%Finished build_dcsc, time: 0.235948
-%Finished setting ids, time: 0.154693
+%Finished build_dcsc, time: 0.162558
+%Finished setting ids, time: 0.07698
 %Starting sort
-%Finished sort, time: 1.75676
-%Finished setting edge pointers, time: 0.000142
+%Finished sort, time: 1.20438
+%Finished setting edge pointers, time: 0.000169
 %Starting build_dcsc
-%Finished build_dcsc, time: 0.238386
-%Completed reading A from memory in 4.089144 seconds.
-%Completed reading A from file in 5.114708 seconds
+%Finished build_dcsc, time: 0.156858
+%Completed reading A from memory in 2.522761 seconds.
+%Completed reading A from file in 3.481659 seconds.
 %Completed 19 iterations
 %Timing results:
-%- load graph: 5.11472 sec
-%- initialize engine: 2.88486e-05 sec
-%- run algorithm: 0.212451 sec
-%- print output: 0.0799532 sec
-%- deinitialize engine: 2.88486e-05 sec
-%17:29:05.575 [INFO ] Benchmarking algorithm "Single source shortest paths" on graph "dota-league" took 5955 ms.
-
+%- load graph: 3.48167 sec
+%- initialize engine: 3.60012e-05 sec
+%- run algorithm: 0.168526 sec
+%- print output: 0.105846 sec
+%- deinitialize engine: 6.00815e-05 sec
+%03:17:45.114 [INFO ] Benchmarked algorithm "Single source shortest paths" on graph "dota-league".
+%03:17:45.114 [INFO ] Benchmarking algorithm "Single source shortest paths" on graph "dota-league" succeed.
+%03:17:45.114 [INFO ] Benchmarking algorithm "Single source shortest paths" on graph "dota-league" took 3830 ms.
 With a plugin to Graphalytics called Granula~\cite{Ngai:2015:Granula}, one can explicitly specify a performance model to analyze specific execution behavior such as the amount of communication or runtime of particular kernels of execution.
@@
 Our belief is choosing a single reference implementation cannot capture the complexities inherent with graph processing.
-
-Our contribution strikes a balance between the highly detailed analysis of a single implementation such as those presented by Nai~\cite{Nai:2016:architectural} (greater depth of analysis) and Graphalytics which focuses on the runtimes of several implementations and several platforms (more breadth of analysis).
-This allows us to greatly simplify deployment and allows the user to make his or her own decisions.
-
-In addition, we analyze  power and energy consumption as well as details such as the number of iterations or the time to construct the data structures.
-Moreover, our approach requires little knowledge of the inner workings of each system; the data are collected using either hardware counter sampling of model-specific registers (for power) or through parsing log files (for execution time).
-This allows arbitrary datasets to be included as long as they are in the same, simple format as those in the Stanford Network Analysis Project (SNAP) datasets~\cite{snapnets}.
 
@@
 \section{Methodology}
-\todo[inline]{Start with dataset subsection, write new intro to methodology, move this into inputs.
-We're comparing algorithms not applications.
-Understand those since they're the building blocks.
-Not comparing architectural details (cite that Nai paper).
-Important paragraph.}
+Our contribution strikes a balance between the highly detailed analysis such as those presented by Nai~\cite{Nai:2016:architectural} (greater depth of analysis) and Graphalytics which compares only the runtimes of a greater number of software packages and algorithms (greater breadth of analysis).
+This simplifies our interface which in turn allows users to perform their own analysis on their own datasets.
+
+In addition, we analyze  power usage, energy consumption, and details such as the number of iterations for PageRank or the time to construct the graph data structures.
+Moreover, our approach requires little knowledge of the inner workings of each system; the data are collected by either parsing log files (for execution time) or hardware counter sampling of model-specific registers (for power).
+This allows arbitrary datasets to be included as long as they are in the same, simple format as those in the Stanford Network Analysis Project (SNAP) datasets~\cite{snapnets}.
+
 Our framework breaks the process of characterizing performance into five principal phases shown in \figurename~\ref{fig:epg-overview}, each of which requires no more than a single shell command.
@@
 
-The Graphalytics results in Table~\ref{tab:graphalytics} were performed on the Dota-League datasets.
-This dataset contains 61,670 vertices and 50,870,313 edges.
-This dataset comes from the Game Trace Archive\cite{Guo:2012:GTA} and is modified for Graphalytics\footnote{This dataset is available at \url{https://atlarge.ewi.tudelft.nl/graphalytics/}.}.
+The Graphalytics results in Table~\ref{tab:graphalytics-realworld} were performed on the Dota-League dataset.
+This dataset contains 61,670 vertices and 50,870,313 edges and measures interactions between players in the online video game Defense of the Ancients.
+This was sourced from the Game Trace Archive\cite{Guo:2012:GTA} and is modified for Graphalytics\footnote{This dataset is available at \url{https://atlarge.ewi.tudelft.nl/graphalytics/}.}.
 This dataset is useful because it is both weighted and more dense than the usual real-world dataset with an average out-degree of 824.
@@
 We also present results for the \verb|cit-Patents| dataset from SNAP \cite{snap-cit-patents}.
-This is a widely-used network of citations from the National Bureau of Economic Research (NBER) and is less dense than Dota-League.
+This is a widely-used network of citations from the National Bureau of Economic Research (NBER) and is less dense than Dota-League with 3,774,768 vertices and 16,518,948 edges.
 We stress that though these two datasets are presented, any network in the SNAP data format\footnote{A file in the SNAP format consists of one edge per line, with vertices separated by whitespace and lines which begin with \# are comments.} can be used in \mbox{\emph{easy-parallel-graph-\textasteriskcentered}}.
@@
 Our approach is not specific or limited to these graph packages and can be extended to others.
-\todo[inline]{Weighted versus unweighted}
 
@@
 We select SSSP because of the straightforward extension from BFS; we need not modify the graph nor the root vertices from BFS.
+Furthermore, SSSP is used as a building block for other graph algorithms such as Betweenness Centrality.
 
@@
 One challenge with using PageRank is the stopping criterion; all implementations have been modified to use $|| p_t - p_{t-1}||_1$ (the absolute sum of differences) where $p_t$ is the pagerank at step $t$.
-Verification of the PageRank results is beyond the scope of this paper, although this may explain some of the large performance discrepancies.\todo[inline]{Better explain why these were selected.
-They're the most popular of the algorithms for the papers we surveyed.}
+Verification of the PageRank results is beyond the scope of this paper, although this may explain some of the large performance discrepancies.
 
-This approach is not specific to a particular algorithm; measuring the execution time, data structure construction time, and power consumption can be applied easily to other algorithms.
+As with the graph processing systems, this approach is not specific to a particular algorithm; measuring the execution time, data structure construction time, and power consumption can be applied easily to other algorithms.
 
@@
 
-We analyze the performance of the algorithms described in Sec.~\ref{sec:algs} in terms of execution time, scalability over multiple threads, and power and energy consumption\todo[inline]{ A precise list and exposition of the requirements and changes	done to the application would help in understanding the methodology and its implications}.
+We analyze the performance of the algorithms described in Sec.~\ref{sec:algs} in terms of execution time, scalability over multiple threads, and power and energy consumption\todo{ A precise list and exposition of the requirements and changes	done to the application would help in understanding the methodology and its implications}.
 All figures except those measuring scalability (Fig.~\ref{fig:bfs-scaling}) use 32 threads.
@@
 
-In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately $16 \times 2^{22} = 33,500,000$ edges)\todo[inline]{work?}.
+In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately $16 \times 2^{22} = 33,500,000$ edges).
 An explanation of each algorithm is given in~\cite{Iosup:2016:Graphalyticstech}.
@@
 
-\todo[inline]{What are the likely culprits for performance differences---measure cache miss rates.
+\todo{What are the likely culprits for performance differences---measure cache miss rates.
 Try to find something which correlates with performance.
@@
 Automate with perfexplorer or autoperf}
-
 % Graphalytics also outputs MTEPS or millions of traversed edges per second.
@@
 
+\todo{Rotate the x-tick text so you can make all of the text larger}
 \begin{figure}
@@
 We adjusted the other systems to use $\sum_{k=1}^{n} |p_k^{(i)} - p_k^{(i-1)}| < \epsilon $ as the stopping criterion, where $i$ is the iteration and $n$ is the number of vertices.
-We use $\epsilon = 6 \times 10^{-8}$ because this value is approximately machine epsilon for a single precision floating-point number.
-The goal is to make the stopping criteria for all implementations as similar as possible.
+We use $\epsilon = 6 \times 10^{-8}$ because this value is approximately machine epsilon for a single precision floating-point number to make the stopping criteria for all implementations as similar as possible.
+However, with GraphMat there is no computation of $|p_k^{(i)} - p_k^{(i-1)}|$.
 
@@
 The assumptions under which the various platforms operate can have a dramatic effect on the program.
-\todo[inline]{Rewrite this portion} For example, the GAP Benchmark Suite stores vertex weights as 32-bit integers.
-However, other systems store them as floating-point numbers.
+For example, the GAP Benchmark Suite can be recompiled to store weights as integers or floating point values.
 This may affect performance in addition to runtime behavior in cases where weights like $0.2$ are cast to $0$.
@@
 
-\todo[inline]{Rerun this}
+\todo{Plot this}
 \begin{table}
@@
 Each cell represents the average over 32 runs.
-The ``N/A'' for PowerGraph and GraphBIG exist because both of these systems read in the files and build a graph data structure simultaneously.
-GraphMat is not included because reading in the Dota-League dataset crashes the system.}
+The ``N/A'' for PowerGraph and GraphBIG exist because both of these systems read in the files and build a graph data structure simultaneously.}
 	\begin{tabular}{l|c|c|c|c}
@@
 		\\ \hline
-		PowerGraph & N/A   & N/A     & 0.344    & 3.606 \\
-		GAP        & 0.412 & 0.00464 & 0.267    & 33.00 \\
-		%		GraphMat   & 1.013 &         &          &       \\
 		GraphBIG   & N/A   & 0.0476  & 0.162    & 21.39  \\
+		PowerGraph & N/A   & 0.36874 & 0.344    & 3.441  \\
+		GAP        & 0.412 & 0.00464 & 0.267    & 0.2464 \\
+		GraphMat   & 1.013 &         &          & 2.1587 \\
 	\end{tabular}
@@
 \end{table}
+%1         GAP      BFS Data structure build  0.25732000
+%2    GraphMat      BFS Data structure build  2.10878359
+%3         GAP PageRank Data structure build  0.32198688
+%4    GraphMat PageRank Data structure build  2.15872441
+%5         GAP     SSSP Data structure build  0.47479500
+%6    GraphMat     SSSP Data structure build  2.35433178
+%7         GAP      BFS                 Time  0.07449250
+%8    GraphBIG      BFS                 Time  0.01132847
+%9    GraphMat      BFS                 Time  0.02272806
+%10        GAP PageRank                 Time  0.24646406
+%11   GraphBIG PageRank                 Time 19.53251875
+%12   GraphMat PageRank                 Time  1.69520531
+%13 PowerGraph PageRank                 Time  3.44062500
+%14        GAP     SSSP                 Time  0.07461594
+%15   GraphBIG     SSSP                 Time  0.03791720
+%16   GraphMat     SSSP                 Time  0.03936912
+%17 PowerGraph     SSSP                 Time  0.36874997
 
+\todo{This paragraph is out of place}
 In addition to the data presented in Table~\ref{tab:graphalytics}, one may want to know the cost of performing additional computations on a graph once the data structures have been built.
@@
 #endif
-	<region of code to monitor>
+	<region of code to profile>
 #ifdef POWER_PROFILING
@@
 	\centering
-	\begin{minipage}{0.48\linewidth}
+	\begin{minipage}{0.78\linewidth}
 		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ram_power.pdf}
 	\end{minipage}
-	\begin{minipage}{0.48\linewidth}
+	\begin{minipage}{0.78\linewidth}
 		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_cpu_power.pdf}

### 2017-05-14 papers/publication/paper-pollard.tex
Added more todos, expanded explanation of systems.

@@
 Our research is motivated by the current state of parallel graph processing.
-Fields such as social network analysis, \cite{Kang:2011:BillionGraph}, and computational biology \cite{Paylopoulos:2011:biograph} require the analysis of ever-increasing graph sizes.
+Fields such as social network analysis \cite{Kang:2011:BillionGraph} and computational biology \cite{Paylopoulos:2011:biograph} require the analysis of ever-increasing graph sizes.
 The wide variety of problem domains is resulting in the proliferation of parallel graph processing frameworks.
-The most comprehensive survey, released in 2014, identified and categorized over 80 different parallel graph processing systems, not even considering domain specific languages~\cite{Doekemeijer:2015:GPFSurvey}.
+The most comprehensive survey, released in 2014, identified and categorized over 80 different parallel graph processing systems not even considering domain specific languages~\cite{Doekemeijer:2015:GPFSurvey}.
 
@@
 To address these issues, we present our framework called \mbox{\emph{easy-parallel-graph-\textasteriskcentered}},\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}} a framework which simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
-With this tool we select a small number of graph processing libraries over which we homogenize all aspects of execution to fairly compare performance.
+With this framework we select a small number of graph processing libraries over which we homogenize all aspects of execution to fairly compare performance.
 Additionally, we present results from a use case of this framework to analyze the performance, energy consumption, and scalability of specific algorithm implementation.
 This enables algorithm designers to select high performance reference implementations with which they can compare their improvements.
-To be clear, we are not proposing a new benchmark suite or providing any reference implementations.
-We introduce a framework for detailed and fair comparison of existing graph software packages without requiring the user to be familiar with them.
+To be clear, we are not proposing a new benchmark suite or providing any new reference implementations.
+Instead, we introduce a method for detailed and fair comparison of existing graph software packages without requiring the user to be familiar with them.
 
@@
 However, perusal of the log files reveals a different story.
-The listing contained below Table~\ref{tab:graphalytics-realworld} is the output from GraphMat itself and the execution timing (3.8 seconds) includes the time to read in the input file.
+The listing contained below Table~\ref{tab:graphalytics-realworld} is the output from GraphMat itself.
+Graphalytics reports a 3.8 second runtime but 0.94 seconds of which GraphMat is simply reading the input file from disk.
 However, the GraphBIG timing (2.9 seconds) does not include the time to read the dota-league file.
@@
 \section{Methodology}
+
+\begin{figure*}
+	\centering
+	% trim=left lower right upper
+	\includegraphics[width=\linewidth, trim=0 144pt 0pt 156pt, clip]{graphics/overview.eps}
+	\caption{Overview of \emph{easy-parallel-graph-\textasteriskcentered}. Each cyan box corresponds to a single shell script in the package while the green ellipses correspond to generated files.}
+	\label{fig:epg-overview}
+\end{figure*}
+
 Our contribution strikes a balance between the highly detailed analysis such as those presented by Nai~\cite{Nai:2016:architectural} (greater depth of analysis) and Graphalytics which compares only the runtimes of a greater number of software packages and algorithms (greater breadth of analysis).
-This simplifies our interface which in turn allows users to perform their own analysis on their own datasets.
+This middle-ground simplifies our interface which in turn allows users to perform their own analysis on their own datasets with greater ease than existing frameworks.
 
@@
 \begin{enumerate}
-	\item Installing modified, stable forks of each software package to ensure homogeneity.
+	\item Installing homogenized, stable forks of each software package to ensure homogeneity.
 	\item Given a synthetic problem size or a real-world graph file, generate the files necessary to run each software package.
@@
 \end{enumerate}
-
-\begin{figure*}
-	\centering
-		% trim=left lower right upper
-		\includegraphics[width=\linewidth, trim=0 144pt 0pt 156pt, clip]{graphics/overview.eps}
-	\caption{Overview of \emph{easy-parallel-graph-\textasteriskcentered}. Each cyan box corresponds to a single shell script in the package while the green ellipses correspond to generated files.}
-	\label{fig:epg-overview}
-\end{figure*}
 
@@
 Hence, a Kronecker graph with scale $S$ has $2^S$ vertices and approximately $16 \times 2^S$ edges.
+Kronecker graphs are a generalization of RMAT graphs.
 
 The Graphalytics results in Table~\ref{tab:graphalytics-realworld} were performed on the Dota-League dataset.
-This dataset contains 61,670 vertices and 50,870,313 edges and measures interactions between players in the online video game Defense of the Ancients.
+This dataset contains 61,670 vertices and 50,870,313 edges and models interactions between players in the online video game Defense of the Ancients.
 This was sourced from the Game Trace Archive\cite{Guo:2012:GTA} and is modified for Graphalytics\footnote{This dataset is available at \url{https://atlarge.ewi.tudelft.nl/graphalytics/}.}.
@@
 We consider the following systems.
-\begin{itemize}
-	\item The Graph500~\cite{Murphy:2010:Graph500}, the canonical BFS benchmark which consists of a specification and reference implementation. We use a modified version most similar to 2.1.4.
+\begin{enumerate}
+	\item The Graph500~\cite{Murphy:2010:Graph500}, the canonical BFS benchmark which consists of a specification and reference implementation.
+We use a modified version most similar to 2.1.4.
+The Graph500 provides reference implementations for MPI and OpenMP but since our framework focuses on shared memory we use only the OpenMP version.
+The Graph500 uses a compressed sparse row (CSR) representation.
 	\item The Graph Algorithm Platform (GAP) Benchmark Suite~\cite{Beamer:2015:GAPBench}, a set of reference implementations for shared memory graph processing.
-	\item GraphBIG~\cite{Nai:2015:Graphbig} benchmark suite. We consider only the shared memory solutions but there are also GPU benchmarks.
+The author of GAP has contributed to the Graph500 so the BFS implementations are similar.
+Additionally, GAP uses OpenMP to achieve parallelism and uses a CSR representation.
+	\item GraphBIG~\cite{Nai:2015:Graphbig} benchmark suite.
+We consider only the shared memory solutions but GraphBIG also provides GPU benchmarks.
+GraphBIG uses a CSR representation for graphs and OpenMP for parallelism.
 	\item GraphMat~\cite{Sundaram:2015:GraphMat}, a library and programming model along with reference implementations of common algorithms.
+GraphMat uses a doubly-compressed sparse row representation and OpenMP for parallelism.
 	\item PowerGraph~\cite{Gonzalez:2012:Powergraph}, a library and programming model for distributed (and shared memory) graph-parallel computation.
 PowerGraph includes additional toolkits for tasks such as clustering and computer vision.
-\end{itemize}
+Parallelism is achieved via a combination of OpenMP and light-weight, user-level threads called fibers.
+Powergraph uses a novel storage scheme on top of CSR.
+\end{enumerate}
 
@@
 
-We analyze the performance of the algorithms described in Sec.~\ref{sec:algs} in terms of execution time, scalability over multiple threads, and power and energy consumption\todo{ A precise list and exposition of the requirements and changes	done to the application would help in understanding the methodology and its implications}.
+We analyze the performance of the algorithms described in Sec.~\ref{sec:algs} in terms of execution time, scalability over multiple threads, and power and energy consumption.
 All figures except those measuring scalability (Fig.~\ref{fig:bfs-scaling}) use 32 threads.
@@
 
+\todo{redo this table}
 \begin{table}
@@
 
-In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately $16 \times 2^{22} = 33,500,000$ edges).
+In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately $16 \times 2^{22} \approx 33,500,000$ edges).
 An explanation of each algorithm is given in~\cite{Iosup:2016:Graphalyticstech}.
@@
 \subsection{Scalability}
-Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS.
+\todo{People want more explanation of this}Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS.
 We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time for $n$ threads.
@@
 \subsection{Real World Datasets}
-
 \todo{Plot this}
@@
 \end{lstlisting}
-\caption{The changes required to monitor energy usage for a C or C++ file and its corresponding Makefile, given the papi header file is included and \$PAPI is the directory where PAPI is installed.}\todob{Is this too much detail?}
+\caption{The changes required to monitor energy usage for a C or C++ file and its corresponding Makefile, given the papi header file is included and \$PAPI is the directory where PAPI is installed.}
 \label{lst:RAPL}

### 2017-05-14 autoperf/mst/experiment_analysis.R
Added more energy/power EPV measurements.
Fixed bug where the logs wouldn't get deleted upon a new run.

@@ -1,5 +1,35 @@
 library(ggplot2)
 
+# This takes a filename, reads it in, then filters it down by the parameters
+# and adds a column for their interations.
+# E.g. Galois.All, MST.insertion, MST.rooting tree, etc.
+# Expects the header to be
+# algorithm,execution_phase,scale,edges_per_vertex,RMAT_type,insertion_percent,changed_vertices,threads,measurement,value
+get_method_time <- function(filename, scale_num, ins_pct, cverts, num_threads,
+							rmat_type = "ER", measurement_type = "Time (s)")
+{
+	x <- read.csv(filename, header = TRUE)
+	epv <- x$edges_per_vertex[1] # Assume the same throughout experiments
+	salient_cols <- c("algorithm", "execution_phase", "batch", "measurement", "value")
+	method_time <- subset(x,
+			x$scale==scale &
+			x$RMAT_type==rmat &
+			x$edges_per_vertex==epv &
+			x$changed_vertices==cverts &
+			x$insertion_percent==ins_pct &
+			x$threads==num_threads &
+			x$measurement==measurement_type,
+			salient_cols)
+	# This is necessary if the data is noisy; then $value becomes a factor
+	# method_time$value <- as.numeric(as.character(method_time$value))
+	# Remove zero rows---they're invalid and don't work with the log plot
+	# If some factors were coerced into NAs then there was some issue parsing
+	#method_time <- algo_time[!algo_time$Time == 0.0, ]
+	method_time$algo_and_phase <-
+			interaction(method_time$execution_phase, method_time$algorithm)
+	return(method_time)
+}
+
 time_boxplot_with_batches <- function(filename, scale_num, ins_pct, cverts, num_threads)
 {
 	x <- read.csv(filename, header = TRUE)
@@ -144,22 +174,24 @@ plot_strong_scaling <- function(scaling_data, scale)
 	dev.off()
 }
 
+# Can also not do power if you pick "Time (s)". Generates a box
+# for each execution phase:
+# execution_phase: All, rooting tree, first pass, insertion, deletion
[…]

### 2017-05-15 papers/publication/paper-pollard.tex
reordered paragraphs in conclusion, it's better to start by summarizing the contribution of the paper rather than more specific things

@@
 
-The most prominent example of a graph processing performance analysis tool not tied to a particular iplementation is Graphalytics~\cite{Capota:2015:Graphalytics}.
-Like our work, Graphalytics also automates the setup and execution of graph packages for performance analysis.
+The most prominent example of a graph processing performance analysis tool not tied to a particular implementation is Graphalytics~\cite{Capota:2015:Graphalytics}.
+Similar to our work, Graphalytics also automates the setup and execution of graph packages for performance analysis.
 Graphalyics relies on Apache Maven and Java to wrap the execution of each graph processing software package.
@@
 PageRank is selected because of its popularity; most libraries provide reference implementations.
-One challenge with using PageRank is the stopping criterion; all implementations have been modified to use $|| p_t - p_{t-1}||_1$ (the absolute sum of differences) where $p_t$ is the pagerank at step $t$.
+One challenge with using PageRank is the stopping criterion; all implementations have been modified to use $|| p_t - p_{t-1}||_1$ (the absolute sum of differences) where $p_t$ is the page rank at step $t$.
 Verification of the PageRank results is beyond the scope of this paper, although this may explain some of the large performance discrepancies.
@@
 \section{Conclusion}
+We have presented a tool called \emph{easy-parallel-graph\mbox{-\textasteriskcentered}}, where \textasteriskcentered $\ = $ installer, comparator, performance analyzer, dataset homogenizer, or performance visualizer.
+This framework can analyze both power consumption and the two fundamental phases of execution: data structure construction and algorithm execution.
+While the comparison presented here can help one choose among alternatives for the selected packages and algorithms, the problem of selecting a graph processing framework for a given large-scale problem remains far from simple.
+Furthermore, increasing hardware heterogeneity demands performance analysis be easily repeatable on the target architecture.
+Because of this, our framework is designed to be easy to use and we make our code freely available at \url{https://github.com/HPCL/easy-parallel-graph} to encourage further experimentation.
+
 Overall, the GAP Benchmark Suite was the best-performing system across the given datasets though not the most scalable.
@@
 One last consideration is price: GraphMat requires the Intel compiler collection.
-
-We have presented a tool called \emph{easy-parallel-graph\mbox{-\textasteriskcentered}}, where \textasteriskcentered $\ = $ installer, comparator, performance analyzer, dataset homogenizer, or performance visualizer.
-This framework can analyze both power consumption and the two fundamental phases of execution: data structure construction and algorithm execution.
-While the comparison presented here can help one choose among alternatives for the selected packages and algorithms, the problem of selecting a graph processing framework for a given large-scale problem remains far from simple.
-Furthermore, increasing hardware heterogeneity demands performance analysis be easily repeatable on the target architecture.
-Because of this, our framework is designed to be easy to use and we make our code freely available at \url{https://github.com/HPCL/easy-parallel-graph} to encourage further experimentation.
 

### 2017-05-15 papers/publication/paper-pollard.tex
Updated some figures and reworked the tables.

@@
 	PowerGraph & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
+	cit-Patents & 13.8 & 30.1 & 23.9 & 18.8 & N/A & 22.1 \\
+	dota-league & 25.6 & 31.2 & 458.1 & 26.7 & 28.9 & 22.9 \\
+	% graph500-22 43.0 s 55.6 s 299.8 s 46.4 s N/A    40.5 s
 	\multicolumn{1}{l}{}\\
@@
 	\caption{The dataset used is the same Kronecker graph with scale $22$ as used in other experiments.
-Performance results are in seconds running with the maximal number of threads (64 for PowerGraph, 72 for the others).
+Performance results are in seconds with 32 threads.
 Community detection is performed using label propagation.}
@@
 		% trim=left lower right upper
-		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup.pdf}
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup20.pdf}
 	\end{minipage}
 	\begin{minipage}{\linewidth}
-		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss.pdf}
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss20.pdf}
 	\end{minipage}
@@
 \subsection{Real World Datasets}
-\todo{Plot this}
 \begin{table}
 	\centering
-	\caption{This table shows averages of runtimes for the Dota-League dataset using our method using 32 threads.
-Each cell represents the average over 32 runs.
-The ``N/A'' for PowerGraph and GraphBIG exist because both of these systems read in the files and build a graph data structure simultaneously.}
+	\caption{This table shows averages of runtimes for the Dota-League dataset using our method with 32 threads.
+We average over 32 runs and since the data structure construction times do not differ significantly we average across all algorithms.
+Both PowerGraph and GraphBIG read in the files and build a graph data structure simultaneously so we cannot measure data structure construction time.}
 	\begin{tabular}{l|c|c|c|c}
@@
 		\\ \hline
-		GraphBIG   & N/A   & 0.0476  & 0.162    & 21.39  \\
-		PowerGraph & N/A   & 0.36874 & 0.344    & 3.441  \\
-		GAP        & 0.412 & 0.00464 & 0.267    & 0.2464 \\
-		GraphMat   & 1.013 &         &          & 2.1587 \\
+		GraphBIG   & N/A   & 0.01133        & {\bf 0.03791} & 19.5325  \\
+		PowerGraph & N/A   & N/A            & 0.3687        & 3.4406 \\
+		GAP        & 0.351 & 0.07449        & 0.0746        & {\bf 0.2465} \\
+		GraphMat   & 2.207 & {\bf 0.02273 } & 0.0397        & 1.6952 \\
 	\end{tabular}

### 2017-05-15 papers/publication/paper-pollard.tex
Clarified tables by adding Graphalytics or e-p-g*.
Added overview.graphml.

@@
 	\begin{tabular}{l|c|c|c|c|c|c}
+	\multicolumn{5}{l}{\large{\emph{Graphalytics}}} & \\
 	GraphBIG & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
@@
 	every head row/.style={after row=\midrule},
-	columns/0/.style={string type, column type={l|}, column name={}},
+	columns/0/.style={string type, column type={l|}, column name={\large{\emph{Graphalytics}}}},
 	columns/graphmat/.style={
@@
 	\begin{tabular}{l|c|c|c|c}
+		\multicolumn{2}{l}{\large{\emph{easy-parallel-graph\mbox{-\textasteriskcentered}}}} & & &\\
 		System & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular}
@@
 	\begin{tabular}{l|r|r|r|r}
-			&	GAP  &    Graph500 & GraphBIG & GraphMat \\ \hline
+		\large{\emph{easy-parallel-graph\mbox{-\textasteriskcentered}}}	&	GAP  &    Graph500 & GraphBIG & GraphMat \\ \hline
 		Time (s) &  0.01636 & 0.01884 & 1.600 & 1.424 \\
@@
 
-We hope to add some level of parameter tuning to the next iteration of our software.
 Recent advances in parallel SSSP and BFS algorithms make use of a concept called $\Delta$-stepping.
 This consists of the algorithm being parametrized by some $\Delta > 0$ for which requires tuning for a given architecture and graph structure to achieve optimal performance~\cite{Panitanarak:2014:SSSPPerf}.
-This is provided in the GAP benchmark suite as well as the Parallel Boost Graph Library.
+This is provided in the GAP benchmark suite as well as the Parallel Boost Graph Library and we plan to add some level of parameter tuning to the next iteration of our framework.
 
-Graphalytics also had circumstances with the more computationally expensive algorithms  where certain experiments fail~\cite{Iosup:2016:Graphalyticstech}, so determining whether an algorithm will finish given a particular machine, input size, runtime limit, and resources is an important unanswered question we plan to pursue further.
+Graphalytics encountered circumstances with the more computationally expensive algorithms fail~\cite{Iosup:2016:Graphalyticstech}, so determining whether an algorithm will finish given a particular machine, input size, runtime limit, and resources is an important unanswered question we plan to pursue further.
 
-Graph data structure representation by nature varies across software packages.
-There may be significant performance differences among the various packages for using directed or undirected, or weighted and unweighted graphs.
-We hope to investigate this issue further.
+Even though most software packages represent graphs using CSR format, the implementation details differ across packages.
+There may be significant performance differences among the various packages between using directed or undirected, or weighted and unweighted graphs.
 % Our framework abstracts away much of the data formatting problem, but issues such as whether a framework supports a given dataset is a complex task.
@@
 On one end of the spectrum, the code base of PowerGraph was made closed-source in 2015.
-Simple fixes such as increasing the maximum number of threads per process and updating dependency URLs are not merged and as a result there is much duplicated effort among the over than 400 forks of PowerGraph.
-On the other end, GraphMat's update to version 2.0 has not yet been adopted by Graphalytics nor our framework.
+Simple fixes such as increasing the maximum number of threads per process and updating dependency URLs are not merged and as a result there is much duplicated effort among more than 400 forks of PowerGraph.
+On the other end, GraphMat's update to version 2.0 is in general incompatible with version 1.0 and as such has not yet been adopted by Graphalytics nor our framework.
 One potential solution would be to add each framework to a package management system such as Spack \cite{Gamblin:2015:Spack} which focuses on High Performance Computing.

### 2017-05-15 experiment/example-workflow.sh
Updated workflow, added realworld analysis.
Still an issue with realworld analysis plotting.

@@ -19,20 +19,21 @@ scale <- $S
 mkdir -p graphics
 Rscript experiment_analysis.R example_config.R
 
-# Dota-league dataset
-DSET=dota-league
-curl -o datasets/$DSET.zip https://atlarge.ewi.tudelft.nl/graphalytics/zip/$DSET.zip
-unzip -d datasets datasets/$DSET.zip
-mv datasets/$DSET/$DSET.e datasets/$DSET.e
-mv datasets/$DSET/$DSET.v datasets/$DSET.v
-./gen-datasets.sh -f=datasets/$DSET.e
-./run-experiment.sh -f=$DSET 32
-./parse-output.sh -f=$DSET
+DATASETS="dota-league cit-Patents"
+for DSET in $DATASETS; do
+	curl -o datasets/$DSET.zip https://atlarge.ewi.tudelft.nl/graphalytics/zip/$DSET.zip
+	unzip -d datasets datasets/$DSET.zip
+	mv datasets/$DSET/$DSET.e datasets/$DSET.e
+	mv datasets/$DSET/$DSET.v datasets/$DSET.v
+	./gen-datasets.sh -f=datasets/$DSET.e
+	./run-experiment.sh -f=$DSET 32
+	./parse-output.sh -f=$DSET
+done
 
 echo "# Config file for experiment_analysis.R. threads a vector, scale an int.
 prefix <- './output/'
 threads <- 32
-dataset <- '$DSET'
+dataset_list <- c(${DATASETS// /,})
 " > ${DSET}_config.R # Warning: this file is sourced in experiment_analysis.R
 Rscript realworld_analysis.R ${DSET}_config.R
 

### 2017-05-15 experiment/experiment_analysis.R

@@ -4,7 +4,6 @@ args <- commandArgs(trailingOnly = TRUE)
 if (length(args) != 1) {
 	stop(usage)
 }
-bpc <- "cyan"
 prefix <- "./output/" # The default.
 
 # source sets scale and threads
@@ -12,34 +11,39 @@ source(args[1]) # This is a security vulnerability... just be careful
 
 # Assuming default edgefactor of 16
 stopifnot(length(scale) == 1) # Not supported yet
-algo <- "BFS"
-threadcnt <- 2
 
-# given a scale, algorithm, (BFS, SSSP, or PageRank) and a number of threads,
-# generate a pdf boxplot, one box per library, of the runtimes.
-time_boxplot <- function(scale, thr, algo) {
+###
+# Define some functions
+###
+# given a scale, algorithm, (BFS, SSSP, or PageRank), a number of threads,
+# and a timing metric where timing_metric %in%
+# {"Time", "File reading", "Iterations", "Data structure build"},
+# generate a pdf boxplot, one box per System, of the runtimes.
+time_boxplot <- function(scale, thr, algo, timing_metric = "Time") {
 	nedges <- 16 * 2^scale
 	filename <- paste0(prefix,"parsed-","kron-",scale,"-",thr,".csv")
 	x <- read.csv(filename, header = FALSE)
 	colnames(x) <- c("Sys","Algo","Metric","Time")
 	# Generate a figure
-	algo_time <- subset(x, x$Algo == algo & x$Metric == "Time",
+	algo_time <- subset(x, x$Algo == algo & x$Metric == timing_metric,
 			c("Sys","Time"))
 	algo_time$Time <- as.numeric(as.character(algo_time$Time))
 	# Remove zero rows---they're invalid and don't work with the log plot
 	# If some factors were coerced into NAs then there was some issue parsing
 	algo_time <- algo_time[!algo_time$Time == 0.0, ]
-	pdf(paste0("graphics/bfs_time", scale, "-", thr, "t.pdf"),
+	pdf(paste0("graphics/bfs_",timing_metric,"_",scale,"-", thr,"t.pdf"),
 		width = 5.2, height = 5.2)
-	boxplot(Time~Sys, data = algo_time, ylab = "Time (seconds)",
-			main = "BFS Time", log = "y", col=bpc)
+	ylabel <- timing_metric
[…]

### 2017-05-15 experiment/realworld_analysis.R

@@ -1,0 +1,44 @@
+# Given the results from parse-output.sh,
+# tabulate and generate some plots.
+usage <- "usage: Rscript experiment_analysis.R <config_file>"
+args <- commandArgs(trailingOnly = TRUE)
+if (length(args) != 1) {
+	stop(usage)
+}
+prefix <- "./output/" # The default.
+
+# source should set dataset_list and threads
+source(args[1]) # This is a security vulnerability... only run trusted files.
+
+# Tabulate the average runtime and data set
+avgs <- tabulate_data <- function(thr, dataset)
+{
+	filename <- paste0(prefix,"parsed-",dataset,"-",thr,".csv")
+	x <- read.csv(filename, header = FALSE)
+	colnames(x) <- c("Sys","Algo","Metric","Time")
+	avgs <- aggregate(x$Time, list(x$Sys, x$Algo, x$Metric), mean)
+	return(avgs)
+}
+
+# Compare datasets using one metric where
+# metric %in% c("Data structure build", "File reading", "Iterations", "Time")
+# (Iterations only makes sense for PageRank)
+# Supported options for algo: "BFS", "SSSP", "PageRank"
+compare_datasets <- function(all_avgs, dataset_list, algo, metric = "Time")
+{
+	# TODO: Currently not working
+	avgs.m.a <- subset(all_avgs, all_avgs$Metric==metric & all_avgs$Algo==algo)
+	ggplot(avgs.m.a, aes(Dataset, Sys)) +
+			geom_col(aes(fill = Time))
+}
+
+all_avgs <- data.frame()
+thr <- threads[1] # Just start with one thread count
+for (dset in dataset_list) {
+	avgs <- tabulate_data(thr, dset)
+	all_avgs <- rbind(all_avgs, cbind(dset, avgs))
+}
+colnames(all_avgs) <- c("Dataset","Sys","Algo","Metric","Time")
+
+compare_datasets(avgs_list, dataset_list, "SSSP")
+

### 2017-05-15 experiment/run-experiment.sh

@@ -86,7 +86,9 @@ if [ "$OMP_NUM_THREADS" -gt 1 ]; then
 	"$GRAPH500DIR/omp-csr/omp-csr" -s $S > "${OUTPUT_PREFIX}-Graph500-BFS.out"
 else
 	echo " sequentially"
-	"$GRAPH500DIR/seq-csr/seq-csr" -s $S > "${OUTPUT_PREFIX}-Graph500-BFS.out"
+	# seq-csr is slower than omp-csr with one thread
+	# "$GRAPH500DIR/seq-csr/seq-csr" -s $S > "${OUTPUT_PREFIX}-Graph500-BFS.out"
+	"$GRAPH500DIR/omp-csr/omp-csr" -s $S > "${OUTPUT_PREFIX}-Graph500-BFS.out"
 fi
 
 # GAP

### 2017-05-15 experiment/realworld_analysis.R
ggplot real world dataset and algorithm comparison

@@ -1,3 +1,4 @@
+library(ggplot2)
 # Given the results from parse-output.sh,
 # tabulate and generate some plots.
 usage <- "usage: Rscript experiment_analysis.R <config_file>"
@@ -26,10 +27,17 @@ avgs <- tabulate_data <- function(thr, dataset)
 # Supported options for algo: "BFS", "SSSP", "PageRank"
 compare_datasets <- function(all_avgs, dataset_list, algo, metric = "Time")
 {
-	# TODO: Currently not working
-	avgs.m.a <- subset(all_avgs, all_avgs$Metric==metric & all_avgs$Algo==algo)
-	ggplot(avgs.m.a, aes(Dataset, Sys)) +
-			geom_col(aes(fill = Time))
+	# If you just want a single algorithm, remove the "facet_wrap" line and use
+	# avgs.m.a <- subset(all_avgs, all_avgs$Metric==metric & all_avgs$Algo==algo)
+	avgs.m <- subset(all_avgs, all_avgs$Metric==metric)
+	p <- ggplot(avgs.m, aes(Dataset, Time, fill = Sys)) +
+			geom_bar(stat = "identity", position = "dodge") +
+			facet_wrap(~Algo, scales = "free") +
+			theme(axis.text.x = element_text(angle = 15, hjust = 1))
+	pdf(paste0("graphics/compare-",metric,".pdf"),
+		width = 5, height = 3)
+	p
+	dev.off()
 }
 
 all_avgs <- data.frame()

### 2017-05-15 autoperf/mst/experiment_analysis.R
Debugged to regenerate some figures.

@@ -48,18 +48,18 @@ time_boxplot_with_batches <- function(filename, scale_num, ins_pct, cverts, num_
 	} else {
 		nedges <- epv * 2^scale_num
 		plotfilename <- paste0("plot",scale_num,"_",ins_pct,"i_",cverts,"_10b.pdf")
-		plot_title <- paste0("RMAT with Scale ",scale)
+		plot_title <- paste0("RMAT with Scale ",scale_num)
 		plot_subtitle <- paste0(nedges, "edges, ", 2^scale_num, " vertices")
 	}
 	# Generate a figure
 	salient_cols <- c("algorithm", "execution_phase", "value", "batch")
 	method_time <- subset(x,
-			x$scale==scale &
-			x$edges_per_vertex == epv &
+			x$scale==scale_num &
+			x$edges_per_vertex==epv &
 			x$changed_vertices==cverts &
 			x$insertion_percent==ins_pct &
-			x$threads == num_threads,
-			x$measurement == "Time (s)",
+			x$threads==num_threads &
+			x$measurement=="Time (s)",
 			salient_cols)
 	# method_time$value <- as.numeric(as.character(algo_time$Time)) # May not be necessary
 	# Remove zero rows---they're invalid and don't work with the log plot
@@ -73,7 +73,7 @@ time_boxplot_with_batches <- function(filename, scale_num, ins_pct, cverts, num_
 		xlab("Execution Phase.Algorithm") +
 		geom_boxplot() +
 		labs(title = plot_title, subtitle = plot_subtitle)
-	box + theme(axis.text.x = element_text(angle = 90, hjust = 1))
+	box + theme(axis.text.x = element_text(angle = 30, hjust = 1))
 	dev.off()
 }
 
@@ -85,7 +85,7 @@ percent_insertions <- function(filename, scale_num, cvert,
 	# execution_phase: All, rooting tree, first pass, insertion, deletion
 	# where rooting tree is only done for the first batch
 	epv <- x$edges_per_vertex[1] # Assume the same throughout experiments
-	nedges <- epv * 2^scale
+	nedges <- epv * 2^scale_num
 	# Generate a figure
 	salient_cols <- c("insertion_percent", "value")
 	method_time <- subset(x,
@@ -103,15 +103,16 @@ percent_insertions <- function(filename, scale_num, cvert,
[…]

### 2017-05-15 papers/publication/paper-pollard.tex
ggplot real world dataset and algorithm comparison
Updated figures and screenshots, added todos.

@@
 \begin{table}
-	\caption{A tabulated sample run of Graphalytics with 32 threads.
-Just one run per experiment is performed and these times do not take into account the different phases of execution.
-Runtimes are in seconds.}
+	\caption{A tabulated sample run of Graphalytics with 32 threads. Just one run per experiment is performed. Runtimes are in seconds. Below the table is an excerpt from the GraphMat log file.}
 	\centering
@@
 	GraphBIG & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
-	cit-Patents & 1.0 & 11.1 & 13.0 &   5.0 & N/A  & 1.3 \\
-	dota-league & 1.0 & 3.8 & 1104.0 & 2.5 & \textbf{2.9} & 0.9 \\
+	cit-Patents & 0.8 & 11.8 & 15.5  &  4.5 & N/A  & 1.3 \\
+	dota-league & 1.1 & 3.9 & 1073.7 &  2.6 & \textbf{3.0} & 1.0 \\
 	\multicolumn{1}{l}{}\\
@@
 	dota-league & 2.2 & 19.7 & 150.1 & 2.8 & \textbf{3.8} & 3.1
-	\end{tabular}
-
-	\begin{lstlisting}
+	\end{tabular} \\
+	\vspace{12pt}
 	Timing results (for GraphMat SSSP on dota-league)
-- Finished file read of dota-league. time: 0.9495
-- load graph: 3.48167 sec
-- initialize engine: 3.60012e-05 sec
-- run algorithm: 0.168526 sec
-- print output: 0.105846 sec
-- deinitialize engine: 6.00815e-05 sec
-	\end{lstlisting}
+	\begin{itemize}
+		\item Finished file read of dota-league. time: 0.9495
+		\item load graph: 3.48167 sec
+		\item initialize engine: 3.60012e-05 sec
+		\item run algorithm: 0.168526 sec
+		\item print output: 0.105846 sec
+		\item deinitialize engine: 6.00815e-05 sec
+	\end{itemize}
 	\label{tab:graphalytics-realworld}
@@
 Graphalytics reports a 3.8 second runtime but 0.94 seconds of which GraphMat is simply reading the input file from disk.
-However, the GraphBIG timing (2.9 seconds) does not include the time to read the dota-league file.
+However, the GraphBIG timing (3.0 seconds) does not include the time to read the dota-league file.
 \textbf{If the time to read in the text file was computed correctly, GraphMat would be faster than GraphBIG.} To call this a fair comparison is dubious at best when phases of execution are omitted.
@@
 
-Our approach is not specific or limited to these graph packages and can be extended to others.
+Our approach is not specific or limited to these graph packages and can be extended to others. When comparing to Graphalytics, we use the most recent stable release, v0.2.
 
@@
 \begin{table}
-	\caption{The dataset used is the same Kronecker graph with scale $22$ as used in other experiments.
-Performance results are in seconds with 32 threads.
-Community detection is performed using label propagation.}
+	\caption{Graphalytics on the same Kronecker graph with scale $22$ as used in other experiments. Performance results are in seconds with 32 threads. Community detection uses label propagation.}
 	\centering
@@
 \subsection{Real World Datasets}
-\begin{table}
+\begin{figure}
 	\centering
-	\caption{This table shows averages of runtimes for the Dota-League dataset using our method with 32 threads.
+		% trim=left lower right upper
+		\includegraphics[width=\linewidth, trim=0 18pt 6pt 0pt, clip]{graphics/compare-realworld.pdf}
+	\caption{Real world experiments using \emph{easy-parallel-graph-\textasteriskcentered}. The leftmost plot does is missing Powergraph because Powergraph does not provide BFS.}
+	\label{fig:epg-realworld}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\linewidth, trim=0 18pt 6pt 0pt, clip]{graphics/graphalytics-GraphBIG.png}
+	\caption{Real world and synthetic experiments using Graphalytics on GraphBIG using 32 threads. Graphalytics outputs one HTML page per software package.}
+	\label{fig:graphalytics-html}
+\end{figure}
+%\begin{table}
+%	\centering
+%	\caption{This table shows averages of runtimes for the Dota-League dataset using our method with 32 threads.
 We average over 32 runs and since the data structure construction times do not differ significantly we average across all algorithms.
 Both PowerGraph and GraphBIG read in the files and build a graph data structure simultaneously so we cannot measure data structure construction time.}
-	\begin{tabular}{l|c|c|c|c}
-		\multicolumn{2}{l}{\large{\emph{easy-parallel-graph\mbox{-\textasteriskcentered}}}} & & &\\
-		System & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular}
-		& \begin{tabular}[x]{@{}c@{}}Run \\ BFS\end{tabular}
-		& \begin{tabular}[x]{@{}c@{}}Run \\ SSSP\end{tabular}
-		& \begin{tabular}[x]{@{}c@{}}Run \\ PageRank\end{tabular}
-		\\ \hline
-		GraphBIG   & N/A   & 0.01133        & {\bf 0.03791} & 19.5325  \\
-		PowerGraph & N/A   & N/A            & 0.3687        & 3.4406 \\
-		GAP        & 0.351 & 0.07449        & 0.0746        & {\bf 0.2465} \\
-		GraphMat   & 2.207 & {\bf 0.02273 } & 0.0397        & 1.6952 \\
-	\end{tabular}
-	\label{tab:epg-perf}
-\end{table}
+%	\begin{tabular}{l|c|c|c|c}
+%		\multicolumn{2}{l}{\large{\emph{easy-parallel-graph\mbox{-\textasteriskcentered}}}} & & &\\
+%		System & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular}
+%		& \begin{tabular}[x]{@{}c@{}}Run \\ BFS\end{tabular}
+%		& \begin{tabular}[x]{@{}c@{}}Run \\ SSSP\end{tabular}
+%		& \begin{tabular}[x]{@{}c@{}}Run \\ PageRank\end{tabular}
+%		\\ \hline
+%		GraphBIG   & N/A   & 0.01133        & {\bf 0.03791} & 19.5325  \\
+%		PowerGraph & N/A   & N/A            & 0.3687        & 3.4406 \\
+%		GAP        & 0.351 & 0.07449        & 0.0746        & {\bf 0.2465} \\
+%		GraphMat   & 2.207 & {\bf 0.02273 } & 0.0397        & 1.6952 \\
+%	\end{tabular}
+%	\label{tab:epg-perf}
+%\end{table}
 %1         GAP      BFS Data structure build  0.25732000
@@
 
-\todo{This paragraph is out of place}
 In addition to the data presented in Table~\ref{tab:graphalytics}, one may want to know the cost of performing additional computations on a graph once the data structures have been built.
-Additionally, one may desire runtime bounds to get a reasonable maximum timespan for a given problem.
-Table~\ref{tab:epg-perf} begins to compile performance results and the rest of this section further explores these topics.
+Figure~\ref{fig:epg-realworld} compiles average performance results for \emph{easy-parallel-graph-\textasteriskcentered}.
+We contrast this with Graphalytics which outputs an HTML page resulting from a single trial.
+We display a partial screenshot in \figurename~\ref{fig:graphalytics-html}.
+\todo{Update figure to include "Algorithm run" on the y axis. Then add in data structure construction time and comment on it; Why does dota league or patents take more time?}
 

### 2017-05-15 papers/publication/paper-pollard.tex

@@
 
-As with the graph processing systems, this approach is not specific to a particular algorithm; measuring the execution time, data structure construction time, and power consumption can be applied easily to other algorithms.
+This approach is not specific to a particular algorithm; measuring the execution time, data structure construction time, and power consumption can be applied easily to other algorithms.
+We select some representative algorithms here to demonstrate the process, which we hope will motivate others to use
+the framework to add more algorithms.
 

### 2017-05-15 papers/publication/paper-pollard.tex
mostly results discussion edits

@@
 \begin{table}
-	\caption{A tabulated sample run of Graphalytics with 32 threads. Just one run per experiment is performed. Runtimes are in seconds. Below the table is an excerpt from the GraphMat log file.}
-	\centering
+	\caption{Graphalytics: tabulated sample run times (seconds) with 32 threads. Just one run per experiment is performed. Below the table is an excerpt from the GraphMat log file.}
+		\label{tab:graphalytics-realworld}
+	\vspace{-10pt}
+	\begin{center}
 	% cit-Patents: (3 774 768 vertices, 16 518 948 edges)
@@
 	\begin{tabular}{l|c|c|c|c|c|c}
-	\multicolumn{5}{l}{\large{\emph{Graphalytics}}} & \\
+	%% BN: Commented out because it's redundant with the caption
+	%\multicolumn{5}{l}{\large{\emph{Graphalytics}}}  \\
 	GraphBIG & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
@@
 	cit-Patents & 7.8 & 20.6 & 11.2 &  8.4 & N/A &  7.0 \\
-	dota-league & 2.2 & 19.7 & 150.1 & 2.8 & \textbf{3.8} & 3.1
-	\end{tabular} \\
-	\vspace{12pt}
-	Timing results (for GraphMat SSSP on dota-league)
+	dota-league & 2.2 & 19.7 & 150.1 & 2.8 & \textbf{3.8} & 3.1\\ \hline
+	\end{tabular}
+	\end{center}
+	%\vspace{12pt}
+\noindent Timing results (for GraphMat SSSP on dota-league)
 	\begin{itemize}
@@
 	\end{itemize}
-	\label{tab:graphalytics-realworld}
 \end{table}
@@
 
-Beyond this plethora of performance data, the Graph Algorithm Platform Benchmark Suite (GAP)~\cite{Beamer:2015:GAPBench}, GraphBIG~\cite{Nai:2015:Graphbig}, and the Graph500~\cite{Murphy:2010:Graph500} consider themselves reference implementations or benchmark suites next to which other implementations can be compared.
+Beyond this plethora of performance data, the Graph Algorithm Platform Benchmark Suite (GAP)~\cite{Beamer:2015:GAPBench}, GraphBIG~\cite{Nai:2015:Graphbig}, and the Graph500~\cite{Murphy:2010:Graph500} consider themselves reference implementations or benchmark suites with which other implementations can be compared.
 If no fewer than three software packages call themselves, ``reference implementations,'' which one are we to trust?
-Our belief is choosing a single reference implementation cannot capture the complexities inherent with graph processing.
+Our belief is that choosing a single reference implementation cannot capture the complexities inherent in graph processing.
 
@@
 Specifically, a graph with scale $S$ has $2^S$ vertices.
-For example, many of our experiments were run with $2^{22} = 4,194,304$ vertices and an average of 16 edges per vertex.
-We measure strong scaling and speedup by varying the number of threads from one to the total number of threads on our research computer, 72.
+For example, many of our experiments were performed on graphs with $2^{22} = 4,194,304$ vertices and an average of 16 edges per vertex.
+We measure strong scaling and speedup by varying the number of threads from one to the total number of threads available on our research server, 72.
 
@@
 \subsection{Scalability}
-\todo{People want more explanation of this}Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS.
-We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time for $n$ threads.
-A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}.
-These plots show in general poor scaling throughout, a challenge facing most parallel graph algorithms.
-Additionally, $2^{20}$ vertices is a ``small'' graph by today's standards and thus library designers focus on scalability for larger graphs.
+Figures~\ref{fig:bfs-speedup} and~\ref{fig:bfs-efficiency} illustrate the parallel strong scaling of the BFS algorithm in different packages on a Kronecker graph of scale $20$ .
+The parallel speedup shown in Fig.~\ref{fig:bfs-speedup} is computed as $T_1 / T_n$ where $T_1$ is the sequential time and $T_n$ is the execution time on $n$ threads.
+\begin{figure}[htb]
+	\centering
+		% trim=left lower right upper
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup20.pdf}
+		\caption{Speedup of BFS for a scale-20 graph, black solid line represents ideal speedup. Both axes are logarithmic with an exception at 72 threads for readability. }
+	\label{fig:bfs-speedup}
+\end{figure}
+
+We define the parallel efficiency as $T_1 / (n T_n)$ where $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time for $n$ threads.
+Ideal efficiency is defined as $T_n = T_1/n$ and is the horizontal line near the top of Fig.~\ref{fig:bfs-efficiency}.
+\begin{figure}[htb]
+	\centering
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss20.pdf}
+	\caption{Parallel efficiency for a scale-20 graph.
+GraphBIG dips below 1 because it is slower for $2$ threads than for 1.
+$T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time with $n$ threads.}
+	\label{fig:bfs-efficiency}
+\end{figure}
+
+These plots show generally poor scaling for this size problem, a challenge facing most parallel graph algorithms. 
+\todo[inline]{If time permits, replace with a larger graph to show scaling.}
+Additionally, $2^{20}$ vertices is a relatively ``small'' graph resulting for by today's standards and thus library designers focus on scalability for larger graphs.
 Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.
 GCC version 4.8.5 supports OpenMP version 3.1 but the most recent release of OpenMP is 4.5.
-\begin{figure}
-	\centering
-	\begin{minipage}{\linewidth}
-		% trim=left lower right upper
-		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup20.pdf}
-	\end{minipage}
-	\begin{minipage}{\linewidth}
-		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss20.pdf}
-	\end{minipage}
-	\caption{The top figure has both axes logarithmic with an exception at 72 threads for readability.
-Linear speedup is.
-GraphBIG dips below 1 because it is slower for $2$ threads than for 1.
-$T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time for $n$ threads.}
-	\label{fig:bfs-scaling}
-\end{figure}
 
@@
 On the other end, GraphMat's update to version 2.0 is in general incompatible with version 1.0 and as such has not yet been adopted by Graphalytics nor our framework.
-One potential solution would be to add each framework to a package management system such as Spack \cite{Gamblin:2015:Spack} which focuses on High Performance Computing.
+One potential solution would be to add each framework to a package management system such as Spack \cite{Gamblin:2015:Spack} which supports packaging of high performance computing software.
 

### 2017-05-15 papers/publication/paper-pollard.tex
added some future work

@@
 
-Overall, the GAP Benchmark Suite was the best-performing system across the given datasets though not the most scalable.
-However, this is only for relatively small graphs; all the graphs used here had at most $2^{22}$ vertices.
-Additionally, GAP is the most recent of projects and also the most robust in terms of the input graphs accepted.
-It should be noted that GAP is also the most recent of projects.
+Overall, the GAP Benchmark Suite was the best-performing system across the given datasets though not the most scalable for the graphs used in this study (at most $2^{22}$ vertices).
+In addition, GAP is the most recent project and also the most robust in terms of the input graphs accepted.
 Thus, we recommend graph processing algorithm designers compare their BFS implementations to the Graph500 for BFS, GAP Benchmark Suite for smaller problem sizes or a low number of threads (at most 32), and GraphMat for larger problems with more threads.
 This is despite the fact that GraphBIG has the best scalability, because even at more than 16 threads the runtime for GraphBIG is still at least one order of magnitude slower than GAP or GraphMat for all sampled algorithms.
-One last consideration is price: GraphMat requires the Intel compiler collection.
+One last consideration is price: GraphMat requires the Intel compiler collection, which is not free.
+
+Future extensions to this work include providing additional automation, for example, for graph format conversions, which would make the use of our framework even easier.
+Depending on the granularity of experiments, one can also automatically determine the number of repetitions required to get meaningful timing results in a potentially noisy system environment.
 

### 2017-05-15 papers/publication/paper-pollard.tex
intro edits

@@
 
-We aim to simplify the decision making process by providing results from our experiments on both real world and synthetic datasets.
-We take inspiration from the Graph500 benchmark~\cite{Murphy:2010:Graph500} which clearly specifies every step of the breadth first search (BFS) algorithm and how it should be timed.
+We aim to simplify the decision making process by providing an easy-to-use framework for consistently and fairly evaluating the performance of different algorithms.
+To illustrate the use of this framework, we include results from our experiments on both real world and synthetic datasets.
+We take inspiration from the Graph500 benchmark~\cite{Murphy:2010:Graph500}, which clearly specifies every step of the breadth first search (BFS) algorithm and how it should be timed.
 The Graph500 can be used to rank systems fairly on a single, well-defined benchmark.
@@
 Reference implementations, however, are critical for advancement; the complexity of today's hardware and operating systems requires empirical evidence to assess an algorithm's performance.
-Additionally, there is motivation to define basic building blocks for graph computations~\cite{GABB16, Buluc:CombBLAS:2011}.
-However, new algorithm designers and users alike have no easy way of determining what constitutes a high-performance implementation.
+In addition, there is motivation to define basic building blocks for graph computations~\cite{GABB16, Buluc:CombBLAS:2011}.
+However, new algorithm designers and users alike have no easy way to accurately determine what constitutes a high-performance implementation.
 
-To address these issues, we present our framework called \mbox{\emph{easy-parallel-graph-\textasteriskcentered}},\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}} a framework which simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
-With this framework we select a small number of graph processing libraries over which we homogenize all aspects of execution to fairly compare performance.
-Additionally, we present results from a use case of this framework to analyze the performance, energy consumption, and scalability of specific algorithm implementation.
+To address these issues, we introduce \mbox{\emph{easy-parallel-graph-\textasteriskcentered}},\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}}---a framework that simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
+With this framework we select a small number of graph processing libraries over which we homogenize all aspects of execution to compare performance fairly.
+Another contribution of this paper is a demonstration of the use of this framework to analyze the performance, energy consumption, and scalability of specific algorithm implementation.
 This enables algorithm designers to select high performance reference implementations with which they can compare their improvements.

### 2017-05-15 papers/publication/paper-pollard.tex
related edits

@@
 For example, GraphMat~\cite{Sundaram:2015:GraphMat}, PowerGraph~\cite{Gonzalez:2012:Powergraph}, and GraphBIG~\cite{Nai:2015:Graphbig} present a comparison of performance between their system and a selection of other software packages and in general any new approach will compare itself to existing implementations.
-These performance results are inherently biased because the authors will know their own package better than the competition and as such are aware of more optimizations.
+These performance results are inherently biased because the authors typically know their own package better than the competition and as such are aware of more optimizations.
 
@@
 However, perusal of the log files reveals a different story.
-The listing contained below Table~\ref{tab:graphalytics-realworld} is the output from GraphMat itself.
+The listing shown below Table~\ref{tab:graphalytics-realworld} is the output from GraphMat itself.
 Graphalytics reports a 3.8 second runtime but 0.94 seconds of which GraphMat is simply reading the input file from disk.
 However, the GraphBIG timing (3.0 seconds) does not include the time to read the dota-league file.
-\textbf{If the time to read in the text file was computed correctly, GraphMat would be faster than GraphBIG.} To call this a fair comparison is dubious at best when phases of execution are omitted.
+\textbf{If the time to read in the text file was computed correctly, GraphMat would be faster than GraphBIG.} To call this a fair comparison is dubious at best when certain phases of execution are omitted for some packages but not others.
 

### 2017-05-15 papers/publication/paper-pollard.tex
efficiency text fix

@@
 With a plugin to Graphalytics called Granula~\cite{Ngai:2015:Granula}, one can explicitly specify a performance model to analyze specific execution behavior such as the amount of communication or runtime of particular kernels of execution.
-This requires in-depth knowledge of the source code and execution model in addition to expertise with the Granula API but allows detailed performance analysis and automatic execution and compilation of performance results.\footnote{An example of Granula can be seen at \url{https://github.com/tudelft-atlarge/graphalytics-platforms-graphx/tree/master/granula-model-graphx}}.
+This requires in-depth knowledge of the source code and execution model in addition to expertise with the Granula API but allows detailed performance analysis and automatic execution and compilation of performance results.\footnote{An example of Granula can be seen at \url{http://bit.ly/granula-example}}.
+%\url{https://github.com/tudelft-atlarge/graphalytics-platforms-graphx/tree/master/granula-model-graphx}}.
 
@@
 We analyze the performance of the algorithms described in Sec.~\ref{sec:algs} in terms of execution time, scalability over multiple threads, and power and energy consumption.
-All figures except those measuring scalability (Fig.~\ref{fig:bfs-scaling}) use 32 threads.
+All figures except those measuring scalability (Figs.~\ref{fig:bfs-speedup} and~\ref{fig:bfs-efficiency}) use 32 threads.
 
@@
 
-We define the parallel efficiency as $T_1 / (n T_n)$ where $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time for $n$ threads.
+Figure~\ref{fig:bfs-efficiency} shows the parallel efficiency,  $T_1 / (nT_n)$ for different implementations of BFS.
 Ideal efficiency is defined as $T_n = T_1/n$ and is the horizontal line near the top of Fig.~\ref{fig:bfs-efficiency}.

### 2017-05-15 papers/publication/paper-pollard.tex
conclusion edits

@@
 This is despite the fact that GraphBIG has the best scalability, because even at more than 16 threads the runtime for GraphBIG is still at least one order of magnitude slower than GAP or GraphMat for all sampled algorithms.
-One last consideration is price: GraphMat requires the Intel compiler collection, which is not free.
-
-Future extensions to this work include providing additional automation, for example, for graph format conversions, which would make the use of our framework even easier.
-Depending on the granularity of experiments, one can also automatically determine the number of repetitions required to get meaningful timing results in a potentially noisy system environment.
+Last, but not least, one must consider cost and portability: GraphMat requires the Intel compiler collection, which is not free or universally available for all platforms.
 

### 2017-05-15 papers/publication/paper-pollard.tex
minor

@@
 Alternatively, there may be scenarios where this approach is not feasible.
-For example, if the user does not have root privileges s/he may not be able to access the model specific registers storing the energy usage.
+For example, in some cases, if the user does not have root privileges s/he may not be able to access the model specific registers storing the energy usage.
 One alternative is to use TAU paired with PAPI to instrument the source code, and if the counters are available on the host CPU then the power and energy can still be measured.
@@
 Sleeping Energy refers to the power (in Watts) consumed during the \texttt{unistd} C \texttt{sleep} function, multiplied by the Time row.
-Essentially, this measures the amount of energy that would have been consumed if nothing was running.
+Essentially, this measures the energy that would have been consumed when the CPU and memory are (nearly) idle.
 The increase over sleep is the ratio of the first and third columns.
@@
 
-Our selection of algorithms comes from a common subset of algorithms analyzed in prior work and implemented in the selected software packages.
+Our choice of algorithms is based on the common subset of methods analyzed in prior work and implemented in the selected software packages.
 The standardization of graph algorithm building blocks (graph kernels) is being developed by the Graph BLAS Forum~\cite{Mattson:2013:graphblas}).

### 2017-05-15 papers/publication/paper-pollard.tex
added power to future

@@
 Listing~\ref{lst:RAPL} shows the simplicity of adding RAPL profiling to existing code.
-Once the source code is modified, the implementation may be recompiled and our provided header file linked in.
+Once the source code is modified, it is compiled with our provided header file and linked with a small library.
 Alternatively, there may be scenarios where this approach is not feasible.
@@
 One potential solution would be to add each framework to a package management system such as Spack \cite{Gamblin:2015:Spack} which supports packaging of high performance computing software.
+
+With respect to power and energy profiling, while our current implementation supports measurements based on PAPI's interface to RAPL, which is only available on Intel platforms, the interface is simple and easy to adapt to other platforms in future without requiring PAPI support.
+In particular, fine-grained measurements provided through potentially available custom hardware~\cite{naecon15} can be enabled through the same interface.
 

### 2017-05-15 papers/publication/paper-pollard.tex
minor

@@
 We aim to simplify the decision making process by providing an easy-to-use framework for consistently and fairly evaluating the performance of different algorithms.
-To illustrate the use of this framework, we include results from our experiments on both real world and synthetic datasets.
+To illustrate the use of this framework, we include results from our experiments on both real-world and synthetic datasets.
 We take inspiration from the Graph500 benchmark~\cite{Murphy:2010:Graph500}, which clearly specifies every step of the breadth first search (BFS) algorithm and how it should be timed.
@@
 Moreover, our approach requires little knowledge of the inner workings of each system; the data are collected by either parsing log files (for execution time) or hardware counter sampling of model-specific registers (for power).
-This allows arbitrary datasets to be included as long as they are in the same, simple format as those in the Stanford Network Analysis Project (SNAP) datasets~\cite{snapnets}.
+This allows arbitrary datasets to be included as long as they are in the same format as those in the Stanford Network Analysis Project (SNAP) datasets~\cite{snapnets}.
 
@@
 The assumptions under which the various platforms operate can have a dramatic effect on the program.
-For example, the GAP Benchmark Suite can be recompiled to store weights as integers or floating point values.
+For example, the GAP Benchmark Suite can be recompiled to store weights as integers or floating-point values.
 This may affect performance in addition to runtime behavior in cases where weights like $0.2$ are cast to $0$.
@@
 \todo[inline]{If time permits, replace with a larger graph to show scaling.}
-Additionally, $2^{20}$ vertices is a relatively ``small'' graph resulting for by today's standards and thus library designers focus on scalability for larger graphs.
+In addition, $2^{20}$ vertices is a relatively ``small'' graph resulting for by today's standards and thus library designers focus on scalability for larger graphs.
 Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.
@@
 
-\subsection{Real World Datasets}
+\subsection{Real-World Datasets}
 \begin{figure}
@@
 	\includegraphics[width=\linewidth, trim=0 18pt 6pt 0pt, clip]{graphics/graphalytics-GraphBIG.png}
-	\caption{Real world and synthetic experiments using Graphalytics on GraphBIG using 32 threads. Graphalytics outputs one HTML page per software package.}
+	\caption{Real-world and synthetic experiments using Graphalytics on GraphBIG using 32 threads. Graphalytics outputs one HTML page per software package.}
 	\label{fig:graphalytics-html}
@@
 Simple fixes such as increasing the maximum number of threads per process and updating dependency URLs are not merged and as a result there is much duplicated effort among more than 400 forks of PowerGraph.
-On the other end, GraphMat's update to version 2.0 is in general incompatible with version 1.0 and as such has not yet been adopted by Graphalytics nor our framework.
+On the other end, GraphMat's update to version 2.0 is in general incompatible with version 1.0 and as such has not yet been adopted by Graphalytics or our framework.
 One potential solution would be to add each framework to a package management system such as Spack \cite{Gamblin:2015:Spack} which supports packaging of high performance computing software.
@@
 
-Overall, the GAP Benchmark Suite was the best-performing system across the given datasets though not the most scalable for the graphs used in this study (at most $2^{22}$ vertices).
+Overall, the GAP Benchmark Suite was the best-performing system across all chosen datasets, although not the most scalable for the graphs used in this study (at most $2^{22}$ vertices).
 In addition, GAP is the most recent project and also the most robust in terms of the input graphs accepted.

### 2017-05-16 experiment/gen-datasets.sh
fixed gen-datasets to not use Graph500 generator.

@@ -153,17 +153,26 @@ else
 	d="$FILE_PREFIX"
 	mkdir -p "$DDIR/$d"
 	# FIXME: make-edgelist only generates files at most 2.0Gb
-	"$GRAPH500DIR/make-edgelist" -s $S -o "$DDIR/$d/$d.graph500" -r "$DDIR/$d/$d.roots"
-	# Convert to GAP format (edgelist)
-	"$GRAPH500DIR/graph5002el" "$DDIR/$d/$d.graph500" "$DDIR/$d/$d.roots" "$DDIR/$d/$d.el" "$DDIR/$d/${d}-roots.v"
+	# "$GRAPH500DIR/make-edgelist" -s $S -o "$DDIR/$d/$d.graph500" -r "$DDIR/$d/$d.roots"
+
+	# Convert to edgelist
+	#"$GRAPH500DIR/graph5002el" "$DDIR/$d/$d.graph500" "$DDIR/$d/$d.roots" "$DDIR/$d/$d.el" "$DDIR/$d/${d}-roots.v" #FIXME
+	"$GAPDIR/converter" -g $S -e "$DDIR/$d/$d.el"
+
+	# Symmetrize (make undirected)
+	"$GAPDIR/converter" -g $S -s -e "$DDIR/$d/${d}-undir.el"
+
+	# Convert to GAP serialized format
+	"$GAPDIR/converter" -g $S -s -b "$DDIR/$d/$d.sg"
+
+	# Generate roots (usually graph500 does this but it doesn't work for scale > 22)
+	"$GAPDIR/bfs" -n $NRT -f "$DDIR/$d/$d.sg" | awk '/Source/{print $2}' > "$DDIR/$d/${d}-roots.v"
+
 	# Sort the roots (mitigate that weird issue with GraphBIG not working for root > # vertices read)
 	cat "$DDIR/$d/${d}-roots.v" | sort -n > tmp.txt
 	cp tmp.txt "$DDIR/$d/${d}-roots.v"
 	rm tmp.txt
 
-	# Symmetrize (make undirected)
-	"$GAPDIR/converter" -f "$DDIR/$d/${d}.el" -s -e "$DDIR/$d/${d}-undir.el"
-	"$GAPDIR/converter" -s -f "$DDIR/$d/$d.el" -b "$DDIR/$d/$d.sg"
 	# Convert to GraphBIG format
[…]

### 2017-05-16 experiment/realworld_analysis.R

@@ -33,7 +33,9 @@ compare_datasets <- function(all_avgs, dataset_list, algo, metric = "Time")
 	p <- ggplot(avgs.m, aes(Dataset, Time, fill = Sys)) +
 			geom_bar(stat = "identity", position = "dodge") +
 			facet_wrap(~Algo, scales = "free") +
-			theme(axis.text.x = element_text(angle = 15, hjust = 1))
+			# If your datasets are too long you can rename them
+			scale_x_discrete(labels = c("dota","Patents")) +
+			theme(axis.text.x = element_text(angle = 30, hjust = 1))
 	pdf(paste0("graphics/compare-",metric,".pdf"),
 		width = 5, height = 3)
 	p
@@ -48,5 +50,5 @@ for (dset in dataset_list) {
 }
 colnames(all_avgs) <- c("Dataset","Sys","Algo","Metric","Time")
 
-compare_datasets(avgs_list, dataset_list, "SSSP")
+compare_datasets(all_avgs, dataset_list, "SSSP")
 

### 2017-05-16 papers/publication/paper-pollard.tex
Updated BFS strong scaling title, misc fixes.

@@
 	GraphMat & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
-	cit-Patents & 7.8 & 20.6 & 11.2 &  8.4 & N/A &  7.0 \\
-	dota-league & 2.2 & 19.7 & 150.1 & 2.8 & \textbf{3.8} & 3.1\\ \hline
+	% This was with 64 threads
+	%cit-Patents & 7.8 & 20.6 & 11.2 &  8.4 & N/A &  7.0 \\
+	%dota-league & 2.2 & 19.7 & 150.1 & 2.8 & \textbf{3.8} & 3.1\\ \hline
+	cit-Patents 8.3 & 21.7 & 11.4 & 9.1 & N/A & 7.8 \\
+	dota-league 3.8 & 19.7 & 236.4 & 3.8 & 7.4 & 3.9 \\
 	\end{tabular}
@@
 \todo[inline]{If time permits, replace with a larger graph to show scaling.}
-In addition, $2^{20}$ vertices is a relatively ``small'' graph resulting for by today's standards and thus library designers focus on scalability for larger graphs.
+In addition, $2^{20}$ vertices is a relatively small graph by today's standards and thus library designers focus on scalability for larger graphs.
 Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.

### 2017-05-16 papers/publication/plot_data.R

@@ -7,7 +7,7 @@
 ###
 scale <- 22
 bpc <- "cyan"
-filename <- paste0("parsed",scale,"-32.csv")
+filename <- paste0("parsed-kron-",scale,"-32.csv")
 x <- read.csv(filename, header = FALSE)
 colnames(x) <- c("Sys","Algo","Metric","Time")
 
@@ -93,7 +93,7 @@ scale <- 20
 measure_scale <- function(algo) {
 	# Read in and average the data for BFS for each thread
 	# It is wasteful to reread the parsed*-1.csv but it simplifies the code
-	x <- read.csv(paste0("parsed",scale,"-1.csv"), header = FALSE)
+	x <- read.csv(paste0("parsed-kron-",scale,"-1.csv"), header = FALSE)
 	colnames(x) <- c("Sys","Algo","Metric","Time")
 	x$Sys <- factor(x$Sys, ordered = TRUE)
 	systems <- levels(subset(x$Sys, x$Algo == algo, c("Sys")))
@@ -103,7 +103,7 @@ measure_scale <- function(algo) {
 	colnames(algo_time) <- threadcnts
 	for (ti in seq(length(threadcnts))) {
 		thread <- threadcnts[ti]
-		Y <- read.csv(paste0("parsed",scale,"-",thread,".csv"),
+		Y <- read.csv(paste0("parsed-kron-",scale,"-",thread,".csv"),
 				header = FALSE)
 		ti_time <- subset(Y, Y[[2]] == algo & Y[[3]] == "Time",
 				c(V1,V4))
@@ -125,9 +125,10 @@ for (ti in rev(seq(length(threadcnts)))) {
 
 # Plot the strong scalability for BFS
 pdf("graphics/bfs_ss.pdf", width = 7, height = 4)
+systems <- rownames(bfs_ss)
 plot(as.numeric(bfs_ss[1,]), xaxt = "n", type = "b", ylim = c(0,1),
 		ylab = "", xlab = "Threads", col = colors[1],
-		main = "BFS Strong Scaling", cex.main = 1.4, lty = 1, pch = 1, lwd = 3)
+		main = "BFS Parallel Efficiency", cex.main = 1.4, lty = 1, pch = 1, lwd = 3)
 for (pli in seq(2,nrow(bfs_ss))) {
 	lines(as.numeric(bfs_ss[pli,]), col = colors[pli], type = "b",
[…]

### 2017-05-16 papers/publication/paper-pollard.tex
Added todo for redoing graphalytics table.
First submission to Cluster. Rewrote, reorganized.
Still waiting on scale 23 data.

--- 2017-05-15 papers/publication/paper-pollard.tex
@@
 An example of the former is GraphX~\cite{Xin:2013:GraphX}, which incurs some overhead while gaining fault tolerance and an example of the latter is Ligra~\cite{Shun:2013:Ligra}, a framework requiring a shared-memory architecture.
-% ``The project was motivated by the fact that the largest publicly available real-world graphs all fit in shared memory''
+Additionally, there is a growing number of systems designed to run on GPUs \cite{Zhong:2014:Medusa}.
+% Ligra's motivation for shared memory is the claim that `` the largest publicly available real-world graphs all fit in shared memory.''
 
@@
 
-To address these issues, we introduce \mbox{\emph{easy-parallel-graph-\textasteriskcentered}},\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}}---a framework that simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
+To address these issues, we introduce \emph{easy-parallel-\mbox{graph-\textasteriskcentered}},\footnote{Available at \url{https://github.com/HPCL/easy-parallel-graph}} a framework which simplifies the installation, comparison, and performance analysis of the three most widely implemented graph algorithm building blocks: breadth first search (BFS), single source shortest paths (SSSP), and PageRank (PR).
 With this framework we select a small number of graph processing libraries over which we homogenize all aspects of execution to compare performance fairly.
-Another contribution of this paper is a demonstration of the use of this framework to analyze the performance, energy consumption, and scalability of specific algorithm implementation.
+Another contribution of this paper is a demonstration of the use of this framework to analyze the performance, energy consumption, and scalability of specific algorithm implementations.
 This enables algorithm designers to select high performance reference implementations with which they can compare their improvements.
@@
 
+Section~\ref{sec:relatedwork} describes previous performance analysis of parallel graph algorithms.
+Section~\ref{sec:meth} describes the architecture of \emph{easy-parallel-\mbox{graph-\textasteriskcentered}} and the algorithms, datasets, and graph processing implementations it uses.
+Section~\ref{sec:perf} presents performance analysis and discusses possible explanations for the results.
+Section~\ref{sec:fw} describes future directions for this research and is followed by the conclusion.
+
 \section{Related Work}\label{sec:relatedwork}
@@
 For example, GraphMat~\cite{Sundaram:2015:GraphMat}, PowerGraph~\cite{Gonzalez:2012:Powergraph}, and GraphBIG~\cite{Nai:2015:Graphbig} present a comparison of performance between their system and a selection of other software packages and in general any new approach will compare itself to existing implementations.
-These performance results are inherently biased because the authors typically know their own package better than the competition and as such are aware of more optimizations.
+We assume the performance results which accompany a new library are at risk of bias because the authors typically know their own package better than the competition and as such are aware of more optimizations.
+For a collection of references to the original papers of each new library or framework, see \cite{Doekemeijer:2015:GPFSurvey}.
+Instead, we focus on related research which provide analysis of existing software.
 
@@
 While incredibly thorough, these analyses focus on covering a wide range of datasets and appear difficult to apply to a new approach.
-For example, GraphMat reduces computation to sparse matrix operations which may not suffer from the same memory bottlenecks indicated in the results of Nai et al.
+For example, GraphMat reduces computation to sparse matrix operations which may not suffer from the same memory bottlenecks indicated in~\cite{Nai:2016:architectural}.
 
-Research by Song et al.~\cite{Song:2016:LoadBalancing} and LeBeane et al.~\cite{LeBeane:2015:GraphHetero} considers the important problem of partitioning on heterogeneous architectures and provide profiling to reduce runtime, improve load balance, and reduce energy consumption.
+Research by Song et al.~\cite{Song:2016:LoadBalancing} and LeBeane et al.~\cite{LeBeane:2015:GraphHetero} considers the important problem of partitioning on heterogeneous architectures and provides profiling to reduce runtime, improve load balance, and reduce energy consumption.
 Our framework instead focuses on shared memory packages.
@@
 Satish et al.~\mbox{\cite{Satish:2014:NavigatingGraph}} analyze the performance of several systems on datasets on the order of $30$ billion edges and~\cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model.
-These implementations are hand tuned and provides recommendations for future improvements.
-
-The most prominent example of a graph processing performance analysis tool not tied to a particular implementation is Graphalytics~\cite{Capota:2015:Graphalytics}.
-Similar to our work, Graphalytics also automates the setup and execution of graph packages for performance analysis.
-Graphalyics relies on Apache Maven and Java to wrap the execution of each graph processing software package.
-%---Graphalytics calls these platforms---and requires configuration files for each dataset.\todo[inline]{What am I saying with this?}
-%Moreover, the complexity of Graphalytics can obfuscate the true behavior of the program.
+These implementations are hand tuned and provide recommendations for future improvements.
 
@@
 			cit-Patents & 0.8 & 11.8 & 15.5  &  4.5 & N/A  & 1.3 \\
-	dota-league & 1.1 & 3.9 & 1073.7 &  2.6 & \textbf{3.0} & 1.0 \\
-	\multicolumn{1}{l}{}\\
+			dota-league & 1.1 & 3.9 & 1073.7 &  {\bf 2.6} & 3.0 & 1.0 \\
+			\multicolumn{7}{l}{} \\
 			PowerGraph & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
@@
 			% graph500-22 43.0 s 55.6 s 299.8 s 46.4 s N/A    40.5 s
-	\multicolumn{1}{l}{}\\
+			\multicolumn{7}{l}{} \\
 			GraphMat & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
-	cit-Patents & 7.8 & 20.6 & 11.2 &  8.4 & N/A &  7.0 \\
-	dota-league & 2.2 & 19.7 & 150.1 & 2.8 & \textbf{3.8} & 3.1\\ \hline
+			cit-Patents &  7.5 &  20.1 & 9.8 &  8.1 & N/A  & 6.6 \\
+			dota-league &  2.7 &  21.2 & 239.7 & {\bf 6.3} & 9.4 & 6.9 \\
 		\end{tabular}
 	\end{center}
-	%\vspace{12pt}
-\noindent Timing results (for GraphMat SSSP on dota-league)
+
+	\noindent Timing results (for GraphMat PageRank on dota-league)
 	\begin{itemize}
-		\item Finished file read of dota-league. time: 0.9495
-		\item load graph: 3.48167 sec
-		\item initialize engine: 3.60012e-05 sec
-		\item run algorithm: 0.168526 sec
-		\item print output: 0.105846 sec
-		\item deinitialize engine: 6.00815e-05 sec
+		\item Finished file read of dota-league. time: 2.65211
+		\item load graph: 5.91229 sec
+		\item initialize engine: 8.32081e-05 sec
+		\item run algorithm 1 (count degree): 0.0555639 sec
+		\item run algorithm 2 (compute PageRank): 0.149445 sec
+		\item print output: 0.0641179 sec
+		\item deinitialize engine: 0.00022006 sec
 	\end{itemize}
 \end{table}
+
+The most prominent example of a graph processing performance analysis tool not tied to a particular implementation is Graphalytics~\cite{Capota:2015:Graphalytics}.
+Similar to our work, Graphalytics also automates the setup and execution of graph packages for performance analysis.
+Graphalyics relies on Apache Maven and Java to wrap the execution of each graph processing software package.
+%---Graphalytics calls these platforms---and requires configuration files for each dataset.\todo[inline]{What am I saying with this?}
+%Moreover, the complexity of Graphalytics can obfuscate the true behavior of the program.
 
@@
 However, perusal of the log files reveals a different story.
-The listing shown below Table~\ref{tab:graphalytics-realworld} is the output from GraphMat itself.
-Graphalytics reports a 3.8 second runtime but 0.94 seconds of which GraphMat is simply reading the input file from disk.
-However, the GraphBIG timing (3.0 seconds) does not include the time to read the dota-league file.
-\textbf{If the time to read in the text file was computed correctly, GraphMat would be faster than GraphBIG.} To call this a fair comparison is dubious at best when certain phases of execution are omitted for some packages but not others.
+The bullets shown below Table~\ref{tab:graphalytics-realworld} summarize the output from GraphMat itself.
+Graphalytics reports a 6.3 second runtime but 2.7 seconds of that time GraphMat is simply reading the input file from disk.
+However, the GraphBIG timing (2.6 seconds) does not include the time to read the dota-league file.
+\textbf{If the time to read in the text file was ignored then GraphMat would complete nearly twice as quickly.} To call this a fair comparison is dubious at best when certain phases of execution are omitted for some packages but not others.
 
@@
 
-\section{Methodology}
+\section{Methodology}\label{sec:meth}
 
@@
 \begin{enumerate}
-	\item Installing homogenized, stable forks of each software package to ensure homogeneity.
-	\item Given a synthetic problem size or a real-world graph file, generate the files necessary to run each software package.
+	\item Installing modified, stable forks of each software package to ensure homogeneity.
+	\item Given a synthetic graph size or a real-world graph file, generate the files necessary to run each software package.
 	\item Given a graph and the number of threads, run each algorithm using each software package multiple times.
@@
 \subsection{Datasets}
-Some software packages provide serialized file formats to speed up file IO. Homogenizing the datasets creates extra copies of the graph files.
+Homogenizing the datasets creates copies of the graph files and auxiliary files in various formats.
+This is both to ensure they are correctly formatted for each system and to speed up file I/O whenever possible by using the library designer's serialized data structure file formats.
 
@@
 For example, many of our experiments were performed on graphs with $2^{22} = 4,194,304$ vertices and an average of 16 edges per vertex.
-We measure strong scaling and speedup by varying the number of threads from one to the total number of threads available on our research server, 72.
+We measure parallel efficiency and speedup by varying the number of threads from one to the total number of threads available on our research server, 72.
 
@@
 
-The Graphalytics results in Table~\ref{tab:graphalytics-realworld} were performed on the Dota-League dataset.
-This dataset contains 61,670 vertices and 50,870,313 edges and models interactions between players in the online video game Defense of the Ancients.
+Part of the Graphalytics results in Table~\ref{tab:graphalytics-realworld} were performed on the Dota-League dataset.
+We use this dataset for other experiments and as well.
+It contains 61,670 vertices and 50,870,313 edges and models interactions between players in the online video game Defense of the Ancients.
 This was sourced from the Game Trace Archive\cite{Guo:2012:GTA} and is modified for Graphalytics\footnote{This dataset is available at \url{https://atlarge.ewi.tudelft.nl/graphalytics/}.}.
@@
 
-This study explores four shared memory parallel graph processing platforms and one distributed memory framework.
+This study explores four shared memory parallel graph processing frameworks and one distributed memory framework operating on a single node (Powergraph).
 The first three are so-called ``reference implementations'' while the remaining two are included because of their performance and popularity.
 Other popular libraries such as the Parallel Boost Graph Library~\cite{Gregor:2005:PBGL} are not considered here because the authors do not provide reference implementations.
-We consider the following systems.
+The frameworks are as follows:
 \begin{enumerate}
@@
 
-Our approach is not specific or limited to these graph packages and can be extended to others. When comparing to Graphalytics, we use the most recent stable release, v0.2.
+Our approach is not specific or limited to these graph packages and can be extended to others. When comparing to Graphalytics, we use the most recent stable release, v0.3.
 
@@
 
-We consider three algorithms: Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank, though not all algorithms are implemented on all systems.
+We consider three parallel algorithms: Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank, though not all algorithms are implemented on all systems.
 We inspected the source code of the surveyed parallel graph processing systems to ensure the same phases of execution are measured across differing execution and programming paradigms.
@@
 
-We select SSSP because of the straightforward extension from BFS; we need not modify the graph nor the root vertices from BFS.
+We select SSSP because of the straightforward extension from BFS; we need not modify the graph and can use the same the root vertices from BFS.
 Furthermore, SSSP is used as a building block for other graph algorithms such as Betweenness Centrality.
@@
 \subsection{Parsing and Data Analysis}
-Our example analyzes data using the R programming language and shell scripting parsers. The example workflow provided in our framework yields the figures and tables generated in this paper.
+Our example analyzes data using the R programming language and shell (Bash and AWK) parsers. The example workflow provided in our framework yields the figures and tables generated in this paper.
 
@@
 \subsection{Runtime for Synthetic Graphs}
-
-\todo{redo this table}
 \begin{table}
@@
 Graphalytics by default does not perform SSSP on unweighted, undirected graphs.
-The discrepancy between PageRank values in Table~\ref{tab:graphalytics} and Fig.~\ref{fig:pr} is a result of the differing stopping criterion.
+The discrepancy between PageRank values in Table~\ref{tab:graphalytics} and Fig.~\ref{fig:pr} is a result of the differing stopping criterion and the aforementioned inconsistency of Graphalytics's performance collection scheme.
 
@@
 
-\todo{What are the likely culprits for performance differences---measure cache miss rates.
+% \todo{What are the likely culprits for performance differences---measure cache miss rates.
 Try to find something which correlates with performance.
@@
 
-\todo{Rotate the x-tick text so you can make all of the text larger}
 \begin{figure}
@@
 \subsection{Scalability}
-Figures~\ref{fig:bfs-speedup} and~\ref{fig:bfs-efficiency} illustrate the parallel strong scaling of the BFS algorithm in different packages on a Kronecker graph of scale $20$ .
+Figures~\ref{fig:bfs-speedup} and~\ref{fig:bfs-efficiency} illustrate the parallel strong efficiency of the BFS algorithm in different packages on a Kronecker graph of scale $20$ .
 The parallel speedup shown in Fig.~\ref{fig:bfs-speedup} is computed as $T_1 / T_n$ where $T_1$ is the sequential time and $T_n$ is the execution time on $n$ threads.
@@
 These plots show generally poor scaling for this size problem, a challenge facing most parallel graph algorithms.
-\todo[inline]{If time permits, replace with a larger graph to show scaling.}
-In addition, $2^{20}$ vertices is a relatively ``small'' graph resulting for by today's standards and thus library designers focus on scalability for larger graphs.
+In addition, $2^{20}$ vertices is a relatively small graph by today's standards and thus library designers focus on scalability for larger graphs.
 Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.
@@
 \subsection{Real-World Datasets}
-\begin{figure}
-	\centering
-		% trim=left lower right upper
-		\includegraphics[width=\linewidth, trim=0 18pt 6pt 0pt, clip]{graphics/compare-realworld.pdf}
-	\caption{Real world experiments using \emph{easy-parallel-graph-\textasteriskcentered}. The leftmost plot does is missing Powergraph because Powergraph does not provide BFS.}
-	\label{fig:epg-realworld}
-\end{figure}
-
 \begin{figure}
@@
 We display a partial screenshot in \figurename~\ref{fig:graphalytics-html}.
-\todo{Update figure to include "Algorithm run" on the y axis. Then add in data structure construction time and comment on it; Why does dota league or patents take more time?}
+
+\begin{figure}
+	\centering
+	% trim=left lower right upper
+	\includegraphics[width=\linewidth, trim=0 18pt 6pt 0pt, clip]{graphics/compare-realworld.pdf}
+	\caption{Real world experiments using \emph{easy-parallel-graph-\textasteriskcentered}. The leftmost plot does is missing Powergraph because Powergraph does not provide BFS.}
+	\label{fig:epg-realworld}
+\end{figure}
+
+Comparison of these two datasets in \figurename~\ref{fig:epg-realworld} brings up some surprising variations in the data.
+For example, PowerGraph is faster for SSSP.
+This could be because of the efficient edge-cut partitioning scheme in place on PowerGraph which can more efficiently deal with the high degree vertices present on the denser Dota-League graph.
+However, this comes with a significant overhead; PowerGraph is slower for SSSP than the other platforms.
+GraphBIG seems to have the widest variation; it is by far the slowest for PageRank but is the fastest for the Dota-League BFS, beating out even GAP which uses a more modern algorithm for BFS called $\Delta$-stepping \cite{Meyer:2003:delta}.
+One reason for this lack of performance from GAP is our lack of tuning; we use the default $\Delta = 1$ which is known not to be optimal.
+One possible explanation for the poor performance of GraphMat compared is the overhead of the sparse matrix operations.
+These may pay off for larger datasets and we see good performance on the denser Dota-League dataset for GraphMat across all algorithms.
 
@@
 In our case, the fastest code is also the most energy efficient, although with this level of granularity we could detect circumstances where one could make a tradeoff between energy and runtime.
+
+\begin{table*}
+	\caption{The data are generated using a Kronecker graph with a scale of 22.
+The program is executed with 32 threads.
+Sleeping Energy refers to the power (in Watts) consumed during the \texttt{unistd} C \texttt{sleep} function, multiplied by the Time row.
+Essentially, this measures the energy that would have been consumed when the CPU and memory are (nearly) idle.
+The increase over sleep is the ratio of the first and third columns.
+These are all averaged over 32 roots.}
+	\centering
+	\begin{tabular}{l|r|r|r|r}
+		\large{\emph{easy-parallel-graph\mbox{-\textasteriskcentered}}}	&	GAP  &    Graph500 & GraphBIG & GraphMat \\ \hline
+		Time (s) &  0.01636 & 0.01884 & 1.600 & 1.424 \\
+		Average Power per Root (W) & 72.38 & 97.17 & 78.01 & 70.12 \\
+		Energy per Root (J) &	1.184 & 1.830 & 112.213 & 111.104 \\
+		Sleeping Energy (J) & 0.4046  & 0.4660 & 39.591 &  35.234 \\
+		Increase over Sleep & 2.926 & 3.928 & 2.834 & 3.153
+	\end{tabular}
+	\label{tab:power}
+\end{table*}
+
+\begin{figure}
+	\centering
+	\begin{minipage}{0.78\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ram_power.pdf}
+	\end{minipage}
+	\begin{minipage}{0.78\linewidth}
+		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_cpu_power.pdf}
+	\end{minipage}
+	\caption{Similar to Fig.~\ref{fig:bfs-time} we plot RAM and CPU Power Consumption for each of the 32 roots.
+Since the Graph500 runs multiple roots per execution, we only get a single data point.
+The baseline was computed by monitoring power consumption while a program containing only one call to the C \texttt{unistd} library function \texttt{sleep(10)} (ten seconds).}
+	\label{fig:power}
+\end{figure}
+
+RAPL also allows the measurement of DRAM power, the results of which are shown in Fig.~\ref{fig:power}.
+The right plot describes the second row of Table~\ref{tab:power} in more detail.
+We notice a smaller spread of RAM power consumption, but still a noticeable difference.
+GraphMat exhibits the lowest average RAM power consumption (but not the shortest execution time).
+This information can be useful when choosing an algorithm to use in limited power scenarios, where a slower algorithms that will not exceed the power cap is preferred to a faster one that may exceed it.
+
 \begin{figure}
@@
 
-\begin{table*}
-	\caption{The data are generated using a Kronecker graph with a scale of 22.
-The program is executed with 32 threads.
-Sleeping Energy refers to the power (in Watts) consumed during the \texttt{unistd} C \texttt{sleep} function, multiplied by the Time row.
-Essentially, this measures the energy that would have been consumed when the CPU and memory are (nearly) idle.
-The increase over sleep is the ratio of the first and third columns.
-These are all averaged over 32 roots.}
-	\centering
-	\begin{tabular}{l|r|r|r|r}
-		\large{\emph{easy-parallel-graph\mbox{-\textasteriskcentered}}}	&	GAP  &    Graph500 & GraphBIG & GraphMat \\ \hline
-		Time (s) &  0.01636 & 0.01884 & 1.600 & 1.424 \\
-		Average Power per Root (W) & 72.38 & 97.17 & 78.01 & 70.12 \\
-		Energy per Root (J) &	1.184 & 1.830 & 112.213 & 111.104 \\
-		Sleeping Energy (J) & 0.4046  & 0.4660 & 39.591 &  35.234 \\
-		Increase over Sleep & 2.926 & 3.928 & 2.834 & 3.153
-	\end{tabular}
-	\label{tab:power}
-\end{table*}
-
-RAPL also allows the measurement of DRAM power, the results of which are shown in Fig.~\ref{fig:power}.
-The right plot describes the second row of Table~\ref{tab:power} in more detail.
-We notice a smaller spread of RAM power consumption, but still a noticeable difference.
-GraphMat exhibits the lowest average RAM power consumption (but not the shortest execution time).
-This information can be useful when choosing an algorithm to use in limited power scenarios, where a slower algorithms that will not exceed the power cap is preferred to a faster one that may exceed it.
-
-\begin{figure}
-	\centering
-	\begin{minipage}{0.78\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ram_power.pdf}
-	\end{minipage}
-	\begin{minipage}{0.78\linewidth}
-		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_cpu_power.pdf}
-	\end{minipage}
-	\caption{Similar to Fig.~\ref{fig:bfs-time} we plot RAM and CPU Power Consumption for each of the 32 roots.
-Since the Graph500 runs multiple roots per execution, we only get a single data point.
-The baseline was computed by monitoring power consumption while a program containing only one call to the C \texttt{unistd} library function \texttt{sleep(10)} (ten seconds).}
-	\label{fig:power}
-\end{figure}
-
-\section{Future Work}
+\section{Future Work}\label{sec:fw}
 %We found that GraphMat is only decently performant, but scales to more threads.
@@
 
-Recent advances in parallel SSSP and BFS algorithms make use of a concept called $\Delta$-stepping.
+Recent advances in parallel SSSP and BFS algorithms make use of $\Delta$-stepping.
 This consists of the algorithm being parametrized by some $\Delta > 0$ for which requires tuning for a given architecture and graph structure to achieve optimal performance~\cite{Panitanarak:2014:SSSPPerf}.

### 2017-05-16 papers/publication/plot_data.R

@@ -90,13 +90,13 @@ dev.off()
 ###
 threadcnts <- c(1,2,4,8,16,32,64,72)
 scale <- 20
-measure_scale <- function(algo) {
+measure_scale <- function(algo, scale) {
 	# Read in and average the data for BFS for each thread
 	# It is wasteful to reread the parsed*-1.csv but it simplifies the code
 	x <- read.csv(paste0("parsed-kron-",scale,"-1.csv"), header = FALSE)
 	colnames(x) <- c("Sys","Algo","Metric","Time")
 	x$Sys <- factor(x$Sys, ordered = TRUE)
-	systems <- levels(subset(x$Sys, x$Algo == algo, c("Sys")))
+	systems <- unique(subset(x$Sys, x$Algo == algo, c("Sys")))
 	algo_time <- data.frame(
 			matrix(ncol = length(threadcnts), nrow = length(systems)),
 			row.names = systems)
@@ -112,7 +112,7 @@ measure_scale <- function(algo) {
 	return(algo_time)
 }
 
-bfs_scale <- measure_scale("BFS")
+bfs_scale <- measure_scale("BFS", scale)
 
 colors <- rainbow(nrow(bfs_scale))
 colors <- gsub("F", "C", colors) # You want it darker
@@ -124,6 +124,7 @@ for (ti in rev(seq(length(threadcnts)))) {
 }
 
 # Plot the strong scalability for BFS
+# Actually, this is called Parallel Efficiency!
 pdf("graphics/bfs_ss.pdf", width = 7, height = 4)
 systems <- rownames(bfs_ss)
 plot(as.numeric(bfs_ss[1,]), xaxt = "n", type = "b", ylim = c(0,1),

### 2017-05-16 papers/publication/paper-pollard.tex
Added scale 23. Turning in again.

@@
 \subsection{Scalability}
-Figures~\ref{fig:bfs-speedup} and~\ref{fig:bfs-efficiency} illustrate the parallel strong efficiency of the BFS algorithm in different packages on a Kronecker graph of scale $20$ .
+Figures~\ref{fig:bfs-speedup} and~\ref{fig:bfs-efficiency} illustrate the parallel strong efficiency of the BFS algorithm in different packages on a Kronecker graph of scale $23$.
 The parallel speedup shown in Fig.~\ref{fig:bfs-speedup} is computed as $T_1 / T_n$ where $T_1$ is the sequential time and $T_n$ is the execution time on $n$ threads.
+Because of timing considerations, only four trials were run for these experiments.
 \begin{figure}[htb]
@@
 		% trim=left lower right upper
-		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup20.pdf}
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup23.pdf}
 		\caption{Speedup of BFS for a scale-20 graph, black solid line represents ideal speedup. Both axes are logarithmic with an exception at 72 threads for readability. }
@@
 	\centering
-		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss20.pdf}
-	\caption{Parallel efficiency for a scale-20 graph.
-GraphBIG dips below 1 because it is slower for $2$ threads than for 1.
+		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss23.pdf}
+	\caption{Parallel efficiency for a scale-23 graph.
+Graph500 dips below 1 because it is slower for $2$ threads than for 1.
 $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time with $n$ threads.}
@@
 These plots show generally poor scaling for this size problem, a challenge facing most parallel graph algorithms.
-In addition, $2^{20}$ vertices is a relatively small graph by today's standards and thus library designers focus on scalability for larger graphs.
+In addition, $2^{23}$ vertices is a relatively small graph by today's standards and thus library designers focus on scalability for larger graphs.
 Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.
 GCC version 4.8.5 supports OpenMP version 3.1 but the most recent release of OpenMP is 4.5.
+
+Overall, GAP is the most scalable with GraphMat close behind for larger threads and even slightly beating GAP at 72 threads.
+While care was taken to ensure there was no other load on the system when the experiments were performed, the efficiency for two threads is surprisingly low for the Graph500.
+Because the Graph500 spends a shorter amount of time executing in general (there is no file I/O performs multiple BFS passes one right after another), it is more sensitive to spikes in CPU usage.
 
@@
 
-Overall, the GAP Benchmark Suite was the best-performing system across all chosen datasets, although not the most scalable for the graphs used in this study (at most $2^{22}$ vertices).
+Overall, the GAP Benchmark Suite is the best-performing system across all chosen datasets and in general the most scalable for the graphs used in this study (at most $2^{23}$ vertices).
 In addition, GAP is the most recent project and also the most robust in terms of the input graphs accepted.
-Thus, we recommend graph processing algorithm designers compare their BFS implementations to the Graph500 for BFS, GAP Benchmark Suite for smaller problem sizes or a low number of threads (at most 32), and GraphMat for larger problems with more threads.
-This is despite the fact that GraphBIG has the best scalability, because even at more than 16 threads the runtime for GraphBIG is still at least one order of magnitude slower than GAP or GraphMat for all sampled algorithms.
-Last, but not least, one must consider cost and portability: GraphMat requires the Intel compiler collection, which is not free or universally available for all platforms.
+Powergraph and Graphmat are designed for larger graphs (it is possible to run in distributed mode for both) so the overhead of these frameworks may dominate for smaller problem sizes.
+Two more potentially important considerations are cost and portability: GraphMat requires the Intel compiler collection which may be cost-prohibitive for some users.
+Despite the GAP Benchmark Suite being the overall best performing benchmark, it was not the best performing in every case so we unsurprisingly recommend the more comprehensive comparison using all five benchmarks offered by \emph{easy-parallel-\mbox{graph-\textasteriskcentered}}.
 

### 2017-05-16 papers/publication/plot_data.R

@@ -151,7 +151,7 @@ dev.off()
 pdf("graphics/bfs_speedup.pdf", width = 7, height = 4)
 bfs_spd <- data.frame(t(apply(bfs_ss, 1, function(x){x*threadcnts})))
 colnames(bfs_spd) <- threadcnts
-plot(as.numeric(bfs_spd[1,]), xaxt = "n", type = "b", ylim = c(1,10),
+plot(as.numeric(bfs_spd[1,]), xaxt = "n", type = "b", ylim = c(1,18),
 		ylab = "Speedup", xlab = "Threads", col = colors[1], log = "y",
 		main = "BFS Speedup", cex.main=1.4, lty = 1, pch = 1, lwd = 3)
 for (pli in seq(2,nrow(bfs_ss))) {

### 2017-07-30 papers/poster/extended-abstract-pollard.tex
Added Cluster '17 extended abstract for poster.

\documentclass[conference]{IEEEtran}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[draft]{hyperref}
\usepackage{url}
\usepackage{cite} % Multiple citations in one~\cite command
\usepackage{pgfplotstable}
\usepackage{array} % Required for pgfplotstable dec sep align
\pgfplotsset{compat=1.12}
\usepackage{booktabs}
\usepackage{todonotes}
\usepackage{listings}

%\newcommand{\todob}[1]{\todo[inline,color=green!40]{#1}}

\usepackage{enumitem}
\setlist{parsep=0pt, listparindent=0.5cm}
\pdfminorversion=7

\begin{document}
\title{A Comparison of Parallel Graph Processing Implementations}
\author{
	\IEEEauthorblockN{Samuel D. Pollard}
	\IEEEauthorblockA{Computer and Information Science\\
		University of Oregon \\
		Eugene, OR 97403\\
		Email: spollard@cs.uoregon.edu}
	\and
	\IEEEauthorblockN{Boyana Norris}
	\IEEEauthorblockA{Computer and Information Science\\
		University of Oregon \\
		Eugene, OR 97403\\
		Email: norris@cs.uoregon.edu}
}
\maketitle

\begin{abstract}
The rapidly growing number of large network analysis problems has led to the emergence of many parallel and distributed graph processing systems---one survey in 2014 identified over 80. Determining the best approach for a given problem is infeasible for most developers. We present an approach and associated software for analyzing the performance and scalability of parallel, open-source graph libraries. We demonstrate our approach on five graph processing packages: GraphMat, Graph500, Graph Algorithm Platform Benchmark Suite, GraphBIG, and PowerGraph using synthetic and real-world datasets. We examine previously overlooked aspects of parallel graph processing performance such as phases of execution and energy usage for three algorithms: breadth first search, single source shortest paths, and PageRank.
\end{abstract}

\section{Extended Abstract}

\subsection{Motivation}
Any user wishing to perform graph analytics faces the daunting task of selecting which software package to use given a problem. Installing, satisfying dependencies, and determining which algorithms are supported is nontrivial. Input file formats, and stopping criteria vary across packages.

\subsection{Architectural Overview}
Our framework, easy-parallel-graph-\textasteriskcentered, comprises Bash shell scripts which automate each step of the experiment. Our framework breaks the process of characterizing performance into five principal phases as follows.
\begin{enumerate}
	\item Installing modified, stable forks of each software package to ensure homogeneity.
	\item Given a synthetic graph size or a real-world graph file, generate the files necessary to run each software package.
	\item Given a graph and the number of threads, run each algorithm using each software package multiple times.
	\item Parse through the log files to compress the output into a CSV.
	\item Analyze the data using the provided R scripts to generate plots.
\end{enumerate}
The source code is freely available at \url{https://github.com/HPCL/easy-parallel-graph}.

%\begin{figure*}
%	\centering
%	% trim=left lower right upper
%	\includegraphics[width=0.8\linewidth, trim=0 8pt 0pt 8pt, clip]{graphics/overview.eps}
%	\caption{Overview of \emph{easy-parallel-graph-\textasteriskcentered}. Each cyan box corresponds to a single shell script in the package while the green ellipses correspond to generated files.}
%	\label{fig:epg-overview}
%\end{figure*}

\subsection{Algorithms and Packges}
We provide analysis for three algorithms and five packages, though not all packages implement all algorithms. The algorithms are Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank. The packages are listed below:

\begin{enumerate}
	\item The Graph500~\cite{Murphy:2010:Graph500} consists of a specification and reference implementation of BFS.
	\item The Graph Algorithm Platform (GAP) Benchmark Suite~\cite{Beamer:2015:GAPBench}, a set of reference implementations for shared memory graph processing. GAP implements all three algorithms.
	\item GraphBIG~\cite{Nai:2015:Graphbig} benchmark suite, all three algorithms.
	\item GraphMat~\cite{Sundaram:2015:GraphMat}, a library and programming model, all three algorithms.
	\item PowerGraph~\cite{Gonzalez:2012:Powergraph}, a library for distributed and shared memory graph-parallel computation. Powergraph provides SSSP and PageRank reference implementations.
\end{enumerate}

\subsection{Related Work}\label{sec:relatedwork}
Graphalytics~\cite{Capota:2015:Graphalytics} is the most prominent benchmark suite presented and is still active. Other benchmark suites which to the best of our knowledge do not have associated publications are GraphBench\footnote{\url{https://github.com/uwsampa/graphbench}} and Graph Package Testing\footnote{\url{ https://github.com/robmccoll/graphdb-testing}}. Additionally, each graph processing package typically presents its own performance analysis.

\subsection{Datasets}
Our framework supports synthetic datasets consisting of Kronecker Graphs~\cite{Leskovec:2010:Kronecker}, a generalization of the RMAT graphs and are used in the Graph500.

Additionally, any dataset in the format of the Stanford Network Analysis Project~\cite{snapnets} (SNAP) may be used\footnote{This format is one edge per line with \# as comment lines.}. For our examples we use the Dota-League dataset~\cite{Guo:2012:GTA} and the cit-Patents dataset~\cite{snap-cit-patents}.

Each graph is made undirected and unweighted, then converted to the formats necessary for each implementations

\subsection{Results}

\begin{figure}
	\centering
	\begin{minipage}{0.7\linewidth}
		\includegraphics[width=\linewidth, trim=0 40pt 18pt 24pt, clip]{graphics/bfs_cpu_power.pdf}
	\end{minipage}
	\caption{We plot CPU Power Consumption for each of the 32 roots. Since Graph500 runs multiple roots per execution, we only get a single data point. The baseline monitors power consumption during the execution of the C \texttt{unistd} function \texttt{sleep(10)} (ten seconds).}
	\label{fig:power}
\end{figure}

We use the Performance Application Programming Interface (PAPI)~\cite{Browne:2000:PAPI} to access Intel's Running Average Power Limit (RAPL), which provides a set of hardware counters for measuring energy usage. Results from this are shown in Fig.~\ref{fig:power}.

\begin{figure}[htb]
	\centering
		% trim=left lower right upper
		\includegraphics[width=0.95\linewidth, trim=0 18pt 24pt 18pt, clip]{graphics/bfs_ss23.pdf}
	\caption{Parallel efficiency for a scale-23 graph. $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time with $n$ threads.}
	\label{fig:bfs-efficiency}
\end{figure}
Figure~\ref{fig:bfs-efficiency} shows the parallel efficiency, $T_1 / (nT_n)$ for different implementations of BFS. Ideal efficiency is defined as $T_n = T_1/n$ and is the horizontal line near the top of Fig.~\ref{fig:bfs-efficiency}.

\begin{figure}
	\centering
	% trim=left lower right upper
	\includegraphics[width=0.95\linewidth, trim=0 18pt 6pt 6pt, clip]{graphics/compare-realworld.pdf}
	\caption{Real world experiments using \emph{easy-parallel-graph-\textasteriskcentered} averaged across 32 runs.}
	\label{fig:epg-realworld}
\end{figure}

Fig.~\ref{fig:epg-realworld} shows results for real-world experiments.

\subsection{Conclusion and Future Work}
A comparison of implementations requires tedious engineering effort. The newest framework, GAP, is the best performing framework in most cases.

Future improvements will add more algorithms and packages; support for triangle counting and Galois \cite{Nguyen:2013:Galois} is nearly complete.

Searching for optimal parameters for SSSP ($\Delta$-stepping) and BFS (direction-optimizing parameters $\alpha$ and $\beta$) via search would also increase performance for graph package users.

\bibliographystyle{IEEEtranS}
\bibliography{../drp}
\end{document}

### 2017-07-30 experiment/get-libraries.sh
Tuned up documentation and fixed R plotting bugs.

@@ -2,6 +2,8 @@
 # Download and build all the required graph processing libraries
 USAGE="usage: get-libraries.sh <libdir>
 	<libdir> repositories directory. Default: ./lib"
+# TODO: Add a --power option which checks if $PAPI is defined
+#       and by default builds in $(pwd)/powerlib
 
 LIBDIR="$(pwd)/lib"
 if [ -n "$1" ]; then
@@ -71,7 +73,7 @@ cd build
 mkdir -p default
 cd default
 cmake -DCMAKE_CXX_COMPILER=g++-4.8 -DCMAKE_C_COMPILER=gcc-4.8 ../..
-make # TODO: This fails right now for DESorderedHandNB on sansa. I think it has to do with boost or gcc5 ABI or gcc5 being 32 bit only...
+make
 cd "$LIBDIR/.."
 if [ "$?" -ne 0 ]; then FAILED="$FAILED Galois"; fi
 

### 2017-07-30 experiment/run-multi-thread.sh

@@ -2,15 +2,14 @@
 # Runs run-experiment.sh for each thread number.
 # Must have been parsed beforehand.
 S=$1
-if [ -z "$S" ]; then
-	echo "usage: run-multi-thread.sh <scale>"
+THREADS="1 2 4 8 12 16 20 24 28 32"
+if [ -z "$S" ] || [ "$S" = '--help' ] || [ "$S" = '-h' ]; then
+	echo "usage: run-multi-thread.sh <scale>
+	You can also change the THREADS variable in this script"
 	exit 2
 fi
-for T in $(echo 1 2 4 8 16 32 64 72); do
-	./run-experiment --ddir=$HOME/graphalytics/all-datasets/gabb17 --libdir=$HOME/graphalytics $S $T >> output/out${S}-${T}.log 2>> output/out${S}-${T}.err
+mkdir -p output/out-kron-$S
+for T in $(echo $THREADS); do
+	./run-experiment.sh $S $T >> output/kron-${S}-logs/${T}t.log 2>> output/kron-${S}-logs/${T}t.err
 done
 
-# Other experiment examples on arya:
-# ./gen-datasets.sh --libdir=$LIBDIR --ddir=$DDIR $S
-# ./run-experiment.sh $S $T --libdir=$LIBDIR --ddir=$DDIR > output/out${S}-${T}.log 2> output/out${S}-${T}v2.err &
-# ./run-power.sh --ddir=$DDIR $S $T

### 2017-07-30 experiment/run-multi-thread.sh
Overwrite log file in run-multi-thread.

@@ -2,7 +2,7 @@
 # Runs run-experiment.sh for each thread number.
 # Must have been parsed beforehand.
 S=$1
-THREADS="1 2 4 8 12 16 20 24 28 32"
+THREADS="12 16 20 24 28 32"
 if [ -z "$S" ] || [ "$S" = '--help' ] || [ "$S" = '-h' ]; then
 	echo "usage: run-multi-thread.sh <scale>
 	You can also change the THREADS variable in this script"
@@ -10,6 +10,6 @@ if [ -z "$S" ] || [ "$S" = '--help' ] || [ "$S" = '-h' ]; then
 fi
 mkdir -p output/out-kron-$S
 for T in $(echo $THREADS); do
-	./run-experiment.sh $S $T >> output/kron-${S}-logs/${T}t.log 2>> output/kron-${S}-logs/${T}t.err
+	./run-experiment.sh $S $T > output/kron-${S}-logs/${T}t.log 2> output/kron-${S}-logs/${T}t.err
 done
 

### 2017-08-14 papers/publication/paper-pollard.tex
This is the vesion I turned into Cluster'17

@@
 		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup23.pdf}
-		\caption{Speedup of BFS for a scale-20 graph, black solid line represents ideal speedup. Both axes are logarithmic with an exception at 72 threads for readability. }
+		\caption{Speedup of BFS for a scale-23 graph, black solid line represents ideal speedup. Both axes are logarithmic with an exception at 72 threads for readability. }
 	\label{fig:bfs-speedup}
@@
 However, this comes with a significant overhead; PowerGraph is slower for SSSP than the other platforms.
-GraphBIG seems to have the widest variation; it is by far the slowest for PageRank but is the fastest for the Dota-League BFS, beating out even GAP which uses a more modern algorithm for BFS called $\Delta$-stepping \cite{Meyer:2003:delta}.
-One reason for this lack of performance from GAP is our lack of tuning; we use the default $\Delta = 1$ which is known not to be optimal.
+GraphBIG seems to have the widest variation; it is by far the slowest for PageRank but is the fastest for the Dota-League BFS, beating out even GAP which uses a more modern algorithm for BFS called Direction-optiming BFS \cite{Beamer:2012:DOBFS}.
+One reason for this lack of performance from GAP is our lack of tuning; we use the default parameterization of $\alpha = 15$ and $\beta = 18$, which may not be optimal for all graphs.
 One possible explanation for the poor performance of GraphMat compared is the overhead of the sparse matrix operations.
@@
 
-Recent advances in parallel SSSP and BFS algorithms make use of $\Delta$-stepping.
-This consists of the algorithm being parametrized by some $\Delta > 0$ for which requires tuning for a given architecture and graph structure to achieve optimal performance~\cite{Panitanarak:2014:SSSPPerf}.
-This is provided in the GAP benchmark suite as well as the Parallel Boost Graph Library and we plan to add some level of parameter tuning to the next iteration of our framework.
+Advances in parallel SSSP and BFS contain parameterizations ($\Delta$ for SSSP and $\alpha$ and $\beta$ for BFS) which affects performance depending on graph structure.
+These are provided in GAP.
+We plan to add some level of heuristic parameter tuning as performed in \cite{Beamer:2012:DOBFS} to the next iteration of our framework to take advantage of these algorithmic advances.
 

### 2017-08-19 autoperf/mst/experiment_analysis.R
Added time and strong scaling

@@ -16,6 +16,7 @@ cverts <- as.numeric(args[4])
 num_threads <- as.numeric(args[5])
 rmat_type <- args[6]
 num_batches <- 1
+epv <- 8
 if (num_batches <= 1) {
 	with_batches <- FALSE
 } else {
@@ -99,7 +100,7 @@ time_boxplot_with_batches <- function(filename, scale_num, ins_pct, cverts, num_
 	dev.off()
 }
 
-percent_insertions <- function(filename, scale_num, cvert,
+percent_insertions <- function(filename, scale_num, cverts,
 		RMAT = "ER", num_threads = 72)
 {
 	x <- read.csv(filename, header = TRUE)
@@ -116,7 +117,7 @@ percent_insertions <- function(filename, scale_num, cvert,
 			x$RMAT_type==RMAT &
 			x$threads==num_threads &
 			x$execution_phase=="All" &
-			x$changed_vertices==cvert &
+			x$changed_vertices==cverts &
 			x$algorithm=="MST",
 			salient_cols)
 	# method_time$value <- as.numeric(as.character(algo_time$Time)) # May not be necessary
@@ -133,21 +134,21 @@ percent_insertions <- function(filename, scale_num, cvert,
 		geom_boxplot() +
 		labs(title = "Update vs. Insertion %") +
 		theme(axis.text.x = element_text(angle = 90, hjust = 1))
-	pdf(paste0("insertion_pct_",scale_num,"_",cvert,"_time.pdf"), width = 2.5, height = 2.5)
+	pdf(paste0("insertion_pct_",scale_num,"_",cverts,"_time.pdf"), width = 2.5, height = 2.5)
 	p
 	dev.off()
 }
 
-measure_scalability <- function(filename, one_scale)
+measure_scalability <- function(filename, scale_num)
 {
     x <- read.csv(filename, header = TRUE)
 	yy <- subset(x,
-			x$scale == one_scale & x$changed_vertices == 5000 &
-			x$insertion_percent == 75 & x$RMAT_type == "ER")
+			x$scale == scale_num & x$changed_vertices == cverts &
+			x$insertion_percent == ins_pct & x$RMAT_type == rmat_type)
 	threads <- unique(yy$threads)
 	# TEST
 	# subset(x,
[…]

### 2018-04-13 papers/publication/paper-pollard.tex
Updated figures with embedded fonts
Added an extra paragraph and resized fig. tiny changes from a while ago

@@
 %Starting build_dcsc
-%Finished build_dcsc, time: 0.156858
+%Finished build_dcsc, time: 0.858
 %Completed reading A from memory in 2.522761 seconds.
@@
 	% trim=left lower right upper
-	\includegraphics[width=\linewidth, trim=0 144pt 0pt 156pt, clip]{graphics/overview.eps}
+	\includegraphics[width=0.8\linewidth, trim=0 8pt 0pt 8pt, clip]{graphics/overview.eps}
 	\caption{Overview of \emph{easy-parallel-graph-\textasteriskcentered}. Each cyan box corresponds to a single shell script in the package while the green ellipses correspond to generated files.}
@@
 
+Ultimately, the automated aspect of \emph{easy-parallel-graph\mbox{-\textasteriskcentered}} allows us to generate a large amount of performance data which can in turn be fed into machine learning models.
+These models can then be used to predict performance given a new dataset.
+This involves feature selection, model selection, and training, but once this is complete this will allow a model to be used in lieu of expensive experimental runs.
+
 \section{Conclusion}

### 2018-04-13 experiment/experiment_analysis.R
Added more features to experiment analysis, renamed to preprocess

@@ -225,9 +225,11 @@ if (coalesce == TRUE) {
 							gsub("(parsed-)|(-[0-9]+\\.csv)", "", basename(perf_fn)),
 							"features.csv")
 					if (file.exists(feature_fn)) {
+						message("Reading ", feature_fn)
 						feature_df <- read.csv(feature_fn)
 					} else {
-						error("Could not find feature file ", feature_fn)
+						message("Could not find feature file ", feature_fn)
+						feature_df <- data.frame()
 					}
 					x <- read.csv(perf_fn, header = FALSE)
 					# TODO: Get the actual number of edges and vertices
@@ -241,7 +243,7 @@ if (coalesce == TRUE) {
 					combo <- avg_x[avg_x[[3]] == "Time", c(1,2,4) ]
 					# This is the actual order of the parsed data
 					combo <- cbind(combo, 2^scl, nvertices * 2^scl, thr)
-					if (!ignore_extra_features) {
+					if (!ignore_extra_features && length(feature_df) > 0) {
 						combo <- cbind(combo, feature_df)
 						colnames(combo) <- c(synth_colnames, colnames(feature_df))
 						# it must be reordered to have time at the end
@@ -257,6 +259,7 @@ if (coalesce == TRUE) {
 					combo <- cbind(
 							packagename = as.character(combo$package),
 							algorithmname = as.character(combo$algorithm),
+							datasetname = basename(perf_fn),
 							combo)
 					package_mappings <- rbind(package_mappings,
 							data.frame(as.character(combo$package), as.numeric(combo$package)))
@@ -273,6 +276,7 @@ if (coalesce == TRUE) {
 			message("Collecting data for ", rw)
 			# Also want to include the features downloaded from SNAP
 			feature_fn <- file.path(data_dir, rw, "features.csv")
+			message("Reading ", feature_fn)
 			feature_df <- read.csv(feature_fn)
 			for (nthreads in threads) {
 				perf_fn <- paste0(prefix, "parsed-", rw, "-", nthreads, ".csv")
@@ -301,7 +305,12 @@ if (coalesce == TRUE) {
 				} else {
 					combo <- cbind(combo, nthreads, feature_df)
 				}
-				colnames(combo)[colnames(combo) == "Nodes"] <- "nvertices"
+				combo <- cbind(
+						packagename = as.character(combo$package),
+						algorithmname = as.character(combo$algorithm),
[…]
